```{r chain-setup, include=FALSE}
# This is a CHILD document. It assumes the parent master .Rmd has already:
#   - sourced config/endpoints.R and R/serology_helpers.R
#   - created `cfg`  (via get_endpoint_config(params$endpoint))
#   - created `data_raw` (recoded long antibody data)
#   - created `nt`   (the participant N-tracker)
roles <- chain_visit_roles(cfg)   # $early $matbirth $cord $infant (recoded names)
```

# Maternal → Cord → Infant Antibody Chain {#chain}

This part traces how the infant's **`r cfg$outcome_label` at `r roles$infant`**
(2 months of age, before first vaccination) is shaped by the chain of antibody
events running from maternal vaccination, through placental transfer, into early
infancy. Six sequential predictor blocks each interrogate one mechanistic step.

**Primary stratification is the maternal vaccination arm** (TdaP vs TT). Within
each maternal arm we report (a) all infant arms pooled, then (b) each infant
priming level (aP, wP) separately — i.e. the maternal margins plus the
maternal × infant cross. *(This re-centres the analysis on the maternal arms;
the earlier infant-arm-primary layout is available via `build_infant_strata()`.)*

```{r chain-framework, echo=FALSE}
fw <- data.frame(
  Block = paste0("Block ", 1:6),
  Period = c(
    paste0(roles$early, " & ", roles$matbirth, " (maternal)"),
    paste0(roles$early, " \u2192 ", roles$matbirth, " (maternal)"),
    paste0(roles$matbirth, " \u2192 ", roles$cord, " (maternal \u2192 cord)"),
    paste0(roles$matbirth, " \u2192 ", roles$cord, " (maternal \u2192 cord)"),
    paste0(roles$cord, " \u2192 ", roles$infant, " (cord \u2192 infant)"),
    paste0(roles$cord, " \u2192 ", roles$infant, " (cord \u2192 infant)")),
  Type = c("Antibody quantity", "Class switching", "Antibody quantity",
           "Class switching", "Antibody quantity", "Class switching"),
  Question = c(
    "Do maternal IgG levels before and at delivery predict the infant outcome?",
    "Does gestational subclass polarisation predict it?",
    "Does the quantity transferred across the placenta predict it?",
    "Does subclass-selective placental transfer predict it?",
    "Does the cord\u2192infant quantity change predict it?",
    "Does subclass-selective cord\u2192infant change predict it?"),
  stringsAsFactors = FALSE)
knitr::kable(fw, col.names = c("Block", "Period", "Predictor type", "Biological question"),
             caption = "Table C1. Six-block predictor framework (visit names recoded).")
```

Pro-inflammatory subclasses: IgG1, IgG3. Tolerogenic: IgG2, IgG4.
Ratio = log10[(IgG1 + IgG3) / (IgG2 + IgG4)]; positive = pro-inflammatory polarisation.

```{r chain-mapnote-1, echo=FALSE, results='asis'}
mapnote(
  "This whole part corresponds to the **dual-stratified interpretation summary** ",
  "for ", cfg$outcome_label, " (the \"six-block maternal-to-infant antibody chain\").",
  feeds = c(
    "Block 1 results \u2192 summary \u00a72 (variance landscape) and \u00a73 (dominant predictor / antigen identity); synthesis Q1, Q2",
    "Maternal-arm contrast (TdaP vs TT) \u2192 \u00a74 (booster dependence); synthesis Q5",
    "Ratio Blocks 2/4/6 \u2192 \u00a75 (qualitative layer); synthesis Q3",
    "Blocks 3 & 5 \u2192 \u00a78 (mechanistic synthesis; Block-5 positive-median caveat); synthesis Q6",
    "Cross-stratum comparisons \u2192 \u00a76 (what each stratification isolates) and \u00a77 (cross-outcome); synthesis Q4",
    "R\u00b2 / reduced-model inputs \u2192 \"Proposed Reduced Model\" section"))
```

---

## Data Preparation

```{r chain-data, results='asis'}
chain_feats <- c(cfg$outcome_feature, PREDICTOR_FEATURES)

data_long_chain <- data_raw %>%
  dplyr::filter(feature    %in% chain_feats,
                visit_name %in% cfg$chain_visits,
                arm_name   %in% cfg$maternal_arms) %>%
  dplyr::select(subject_accession, visit_name, arm_name, infant_arm,
                feature, log_assay_value) %>%
  distinct(subject_accession, visit_name, feature, .keep_all = TRUE)

data_wide_chain <- data_long_chain %>%
  pivot_wider(id_cols     = c(subject_accession, arm_name, infant_arm),
              names_from  = c(feature, visit_name),
              values_from = log_assay_value,
              names_sep   = "_") %>%
  add_chain_derived(cfg)

cat(sprintf("Wide chain dataset: **%d subjects \u00d7 %d columns**  \n",
            nrow(data_wide_chain), ncol(data_wide_chain)))
cat(sprintf("Outcome column `%s`: **%d non-missing**\n\n",
            cfg$outcome_col, sum(!is.na(data_wide_chain[[cfg$outcome_col]]))))

ft <- as.data.frame(table(data_long_chain$feature, data_long_chain$visit_name))
colnames(ft) <- c("Feature", "Visit", "n")
ft <- ft %>% dplyr::filter(n > 0) %>%
  pivot_wider(names_from = Visit, values_from = n, values_fill = 0)
knitr::kable(ft, caption = "Table C2. Observations per feature \u00d7 visit (after recode & filter).") %>% print()
```

```{r chain-strata, results='asis'}
# Maternal-arm-primary strata: each maternal arm, pooled + per infant arm.
strata <- build_maternal_strata(data_wide_chain)

# Per-stratum headcount BEFORE any predictor-driven attrition.
nt_outcome_counts(data_wide_chain, strata, cfg$outcome_col,
                  caption = "Table C3. Subjects per stratum and number with a non-missing outcome.")
mapnote("This headcount is the starting N for each stratum; subsequent ",
        "per-model attrition (driven by missing subclass/cord measurements) ",
        "is tracked in the participant-accounting tables and explains the small ",
        "subclass-model samples flagged as a caveat in the interpretation summary.")
```

```{r chain-run, results='hide'}
# Fit all six blocks for every stratum. Every model is logged to `nt`.
chain <- lapply(names(strata), function(k)
  run_chain_blocks(strata[[k]]$data, cfg, tracker = nt,
                   part = "Maternal chain", stratum = strata[[k]]$label))
names(chain) <- names(strata)

# Display order + which models to show for pooled margins vs infant splits.
mat_primary_order <- c("TdaP", "TdaP_aP", "TdaP_wP", "TT", "TT_aP", "TT_wP")
mat_primary_order <- intersect(mat_primary_order, names(chain))

# Infant splits show a reduced key set (antigen-level + full-12) to keep the
# document readable; set infant_split_full <- TRUE to show every model.
infant_split_full <- FALSE
split_ids <- list("1" = c("1a","1b","1c","1k"),
                  "3" = c("3a","3e"),
                  "5" = c("5a","5e"))

display_stratum <- function(key, ids, show_relimp = TRUE) {
  cat("####", strata[[key]]$label, "\n\n")
  for (id in ids) if (!is.null(chain[[key]]$br[[id]]))
    show_model(chain[[key]]$br[[id]], show_relimp = show_relimp)
}

display_block <- function(full_ids, split_key, show_relimp = TRUE) {
  sids <- if (isTRUE(infant_split_full) || is.null(split_key)) full_ids
          else split_ids[[split_key]]
  for (m in cfg$maternal_arms) {
    cat("###", m, "mothers\n\n")
    display_stratum(m, full_ids, show_relimp)                 # pooled margin
    for (i in cfg$infant_arms) {
      key <- paste0(m, "_", i)
      if (!is.null(chain[[key]])) display_stratum(key, sids, show_relimp)
    }
  }
}
```

---

## Block 1 — Maternal Antibody Levels (`r roles$early` and `r roles$matbirth`)

**Biological context.** Maternal vaccination during pregnancy raises antigen-specific
IgG; the levels in maternal circulation at **`r roles$matbirth` (delivery)** are the
pool available for placental transfer, while **`r roles$early`** captures the
pre-vaccination baseline. Entering both visits separates pre-existing background from
the vaccine-induced rise, and asks whether it is the absolute level at delivery or the
gestational *change* that matters for the infant outcome. All models use listwise
deletion, so the effective n varies with subclass/visit missingness (tracked below).

```{r chain-b1-mapnote, echo=FALSE, results='asis'}
mapnote("Block 1 is the part the interpretation leans on most heavily.",
        feeds = c(
          paste0("Which antigen dominates the maternal supply (\u00a73 / Q2 of the ",
                 cfg$outcome_label, " summary)"),
          "How much variance the maternal supply explains (\u00a72 / Q1)",
          "The TdaP-vs-TT contrast that diagnoses booster dependence (\u00a74 / Q5)"))
```

```{r chain-b1, results='asis'}
display_block(full_ids = paste0("1", letters[1:12]), split_key = "1")
```

### Maternal subclass slopes at `r roles$matbirth`: TdaP vs TT

```{r chain-b1-slopes, fig.cap=paste0("Figure C1. Maternal subclass IgG at ", roles$matbirth, " vs infant ", cfg$outcome_label, " at ", roles$infant, ", by antigen and maternal arm."), fig.height=6}
vm  <- roles$matbirth
pat <- paste0("(PT|FHA|PRN)_(IgG[1-4])_(", vm, ")")
dl_b1 <- data_wide_chain %>%
  dplyr::filter(arm_name %in% cfg$maternal_arms) %>%
  dplyr::select(arm_name, all_of(cfg$outcome_col), matches(paste0("IgG[1-4]_", vm, "$"))) %>%
  pivot_longer(matches(paste0("IgG[1-4]_", vm, "$")),
               names_to = c("antigen","subclass","visit"),
               names_pattern = pat,
               values_to = "subclass_value") %>%
  tidyr::drop_na(tidyselect::all_of(c("subclass_value", cfg$outcome_col)))

ggplot(dl_b1, aes(x = subclass_value, y = .data[[cfg$outcome_col]],
                  color = subclass, linetype = arm_name)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ arm_name) +
  scale_color_brewer(palette = "Set1") +
  labs(x = paste0("Log10 maternal subclass IgG at ", vm),
       y = cfg$outcome_y_label, color = "Subclass", linetype = "Maternal arm",
       title = paste0("Block 1 (", vm, "): antigen \u00d7 subclass \u2192 ",
                      cfg$outcome_label, " at ", roles$infant)) +
  theme_bw(base_size = 11)
```

---

## Block 2 — Gestational Subclass Switching (`r roles$early` → `r roles$matbirth`)

**Biological context.** Beyond absolute levels, the *direction of class switching*
during pregnancy may carry independent information. A positive Δ ratio
(`r roles$matbirth` − `r roles$early`) indicates a shift toward pro-inflammatory
(IgG1+IgG3) dominance; a positive coefficient means that shift predicts a higher
infant outcome.

```{r chain-b2-mapnote, echo=FALSE, results='asis'}
mapnote("Ratio block \u2014 tests the *qualitative* (subclass-balance) layer.",
        feeds = paste0("\u00a75 (is there a reproducible qualitative layer?) and synthesis Q3 of the ",
                       cfg$outcome_label, " summary"))
```

```{r chain-b2, results='asis'}
display_block(full_ids = "2", split_key = NULL, show_relimp = FALSE)
```

---

## Block 3 — Placental Transfer: Quantity (`r roles$matbirth` → `r roles$cord`)

**Biological context.** Placental IgG transfer is an active FcRn-mediated process.
The transferred quantity is Δ = `r roles$cord` − `r roles$matbirth` (log10). A positive
coefficient means more efficient transfer of that antigen/subclass predicts a higher
infant outcome.

```{r chain-b3-mapnote, echo=FALSE, results='asis'}
mapnote("Transfer-quantity block.",
        feeds = paste0("\u00a78 (mechanistic synthesis \u2014 transfer step) of the ",
                       cfg$outcome_label, " summary; synthesis Q6"))
```

```{r chain-b3, results='asis'}
display_block(full_ids = paste0("3", letters[1:5]), split_key = "3")
```

---

## Block 4 — Placental Transfer: Selectivity (`r roles$matbirth` → `r roles$cord`)

**Biological context.** FcRn transfers subclasses with differing efficiency. The
change in pro-/tolerogenic ratio across transfer tests whether *selective* transfer
(IgG1 enrichment or depletion) predicts the infant outcome.

```{r chain-b4-mapnote, echo=FALSE, results='asis'}
mapnote("Ratio block (transfer selectivity).",
        feeds = paste0("\u00a75 (qualitative layer) and synthesis Q3 of the ",
                       cfg$outcome_label, " summary"))
```

```{r chain-b4, results='asis'}
display_block(full_ids = "4", split_key = NULL, show_relimp = FALSE)
```

---

## Block 5 — Neonatal Change: Quantity (`r roles$cord` → `r roles$infant`)

**Biological context.** Between birth and 2 months, measured IgG changes. The
quantity change is Δ = `r roles$infant` − `r roles$cord` (log10). **Caveat (carried
in the interpretation summary):** in these data the medians are typically *positive*
— measured IgG is higher at `r roles$infant` than in cord blood, which is backwards
for passively transferred antibody and most likely reflects a cord-vs-infant
assay/normalisation difference. Block 5 should be read as "`r roles$cord`→`r roles$infant`
change," **not** catabolic decay.

```{r chain-b5-check, results='asis'}
b5_ant <- paste0("delta_", cfg$antigen_totals, "_", roles$cord, "_", roles$infant)
for (m in cfg$maternal_arms) {
  d <- strata[[m]]$data
  present <- intersect(b5_ant, names(d))
  if (length(present)) {
    meds <- round(apply(d[, present, drop = FALSE], 2, median, na.rm = TRUE), 3)
    knitr::kable(data.frame(Antigen = names(meds), median_delta = unname(meds)),
                 col.names = c("Antigen", "Median \u0394 log10"),
                 caption = paste0("Median ", roles$cord, "\u2192", roles$infant,
                                  " change (", m, " mothers) \u2014 positive = measured rise (artifact, see caveat)."),
                 row.names = FALSE) %>% print()
    cat("\n")
  }
}
mapnote("The positive-median check above is the quantitative basis for the ",
        "Block-5 caveat.",
        feeds = paste0("\u00a71 (caveats) and \u00a78 of the ", cfg$outcome_label, " summary"))
```

```{r chain-b5, results='asis'}
display_block(full_ids = paste0("5", letters[1:5]), split_key = "5")
```

### Subclass change slopes: TdaP vs TT

```{r chain-b5-slopes, fig.cap=paste0("Figure C2. Subclass ", roles$cord, "\u2192", roles$infant, " change vs infant ", cfg$outcome_label, ", by antigen and maternal arm."), fig.height=6}
vc <- roles$cord; vi <- roles$infant
patd <- paste0("^delta_(PT|FHA|PRN)_(IgG[1-4])_", vc, "_", vi, "$")
dl_b5 <- data_wide_chain %>%
  dplyr::filter(arm_name %in% cfg$maternal_arms) %>%
  dplyr::select(arm_name, all_of(cfg$outcome_col),
                matches(paste0("^delta_(PT|FHA|PRN)_IgG[1-4]_", vc, "_", vi, "$"))) %>%
  pivot_longer(matches(paste0("^delta_(PT|FHA|PRN)_IgG[1-4]_", vc, "_", vi, "$")),
               names_to = c("antigen","subclass"),
               names_pattern = patd, values_to = "delta_value") %>%
  tidyr::drop_na(tidyselect::all_of(c("delta_value", cfg$outcome_col)))

ggplot(dl_b5, aes(x = delta_value, y = .data[[cfg$outcome_col]],
                  color = subclass, linetype = arm_name)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ arm_name) +
  scale_color_brewer(palette = "Set1") +
  labs(x = paste0("\u0394 log10 subclass IgG (", vi, " \u2212 ", vc, ")"),
       y = cfg$outcome_y_label, color = "Subclass", linetype = "Maternal arm",
       title = paste0("Block 5: antigen \u00d7 subclass change \u2192 ", cfg$outcome_label)) +
  theme_bw(base_size = 11)
```

---

## Block 6 — Neonatal Change: Selectivity (`r roles$cord` → `r roles$infant`)

**Biological context.** If subclasses change at different rates from cord to 2 months,
the ratio shifts. A positive Δ ratio means pro-inflammatory subclasses are relatively
better preserved.

```{r chain-b6-mapnote, echo=FALSE, results='asis'}
mapnote("Ratio block (neonatal-change selectivity).",
        feeds = paste0("\u00a75 (qualitative layer) and synthesis Q3 of the ",
                       cfg$outcome_label, " summary"))
```

```{r chain-b6, results='asis'}
display_block(full_ids = "6", split_key = NULL, show_relimp = FALSE)
```

---

## Cross-Stratum Comparisons

```{r chain-r2-mapnote, echo=FALSE, results='asis'}
mapnote("The R\u00b2 landscape and the TdaP-vs-TT contrast below are the direct inputs ",
        "to the summary's variance-landscape and booster-dependence sections.",
        feeds = c(
          paste0("\u00a72 (variance landscape) / Q1 of the ", cfg$outcome_label, " summary"),
          "\u00a74 (booster dependence) / Q5",
          "\u00a76 (what each stratification isolates) / Q4",
          "Reduced-model R\u00b2 benchmarks"))
```

### R² Across All Blocks and Strata

```{r chain-r2-table, results='asis'}
r2_all <- bind_rows(lapply(names(chain), function(k)
  chain[[k]]$r2 %>% mutate(stratum = strata[[k]]$label,
                           marm = strata[[k]]$marm, iarm = strata[[k]]$iarm))) %>%
  arrange(block, model)
r2_disp <- r2_all %>%
  mutate(R2     = ifelse(saturated, "\u2014", sprintf("%.3f", R2)),
         adj_R2 = ifelse(saturated, "\u2014", sprintf("%.3f", adj_R2))) %>%
  dplyr::select(stratum, block, model, n_obs, R2, adj_R2)
knitr::kable(r2_disp,
             caption = "Table C4. R\u00b2 for every model across all blocks and strata (saturated models, n \u2264 predictors + 1, shown as \u2014).",
             align = c("l","l","l","r","r","r")) %>% print()
```

```{r chain-r2-wide, results='asis'}
# Primary contrast: pooled maternal margins TdaP vs TT.
r2_wide <- r2_all %>%
  dplyr::filter(iarm == "all") %>%
  mutate(R2 = ifelse(saturated, NA_real_, R2)) %>%
  dplyr::select(marm, block, model, R2) %>%
  pivot_wider(names_from = marm, values_from = R2) %>%
  mutate(delta_R2_TdaP_minus_TT = round(.data[["TdaP"]] - .data[["TT"]], 4)) %>%
  arrange(block, model) %>%
  mutate(across(c("TdaP", "TT", "delta_R2_TdaP_minus_TT"),
                ~ ifelse(is.na(.), "\u2014", sprintf("%.3f", .))))
knitr::kable(r2_wide,
             caption = "Table C5. R\u00b2 side-by-side, pooled maternal margins: TdaP vs TT (saturated shown as \u2014).",
             align = c("l","l","r","r","r")) %>% print()
```

```{r chain-r2-barplot, fig.cap="Figure C3. R\u00b2 of antigen-total models by block, maternal arm (facets) and infant level (fill; 'all' = pooled).", fig.height=5}
iarm_fill <- c(all = "grey45", cfg$iarm_colors)
r2_ant <- r2_all %>%
  dplyr::filter(grepl("antigen", model)) %>%
  mutate(block_clean = gsub("_qty|_quantity|_switch|_ratio|_change", "", block))
ggplot(r2_ant, aes(x = block_clean, y = R2, fill = iarm)) +
  geom_col(position = position_dodge(0.7), width = 0.6) +
  facet_wrap(~ marm) +
  scale_fill_manual(values = iarm_fill, name = "Infant level") +
  labs(x = "Block", y = "R\u00b2",
       title = paste0("Explained variance (antigen-total models) \u2014 ", cfg$outcome_label)) +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

### Ratio-Model Coefficients: TdaP vs TT (Blocks 2, 4, 6)

```{r chain-ratio-coef, fig.cap="Figures C4-C6. Ratio-model coefficients for the pooled maternal margins. Error bars = 95% CI.", fig.height=5}
blk_lab <- c(B2 = paste0("Block 2: \u0394 ratio ", roles$early, "\u2192", roles$matbirth),
             B4 = paste0("Block 4: \u0394 ratio ", roles$matbirth, "\u2192", roles$cord),
             B6 = paste0("Block 6: \u0394 ratio ", roles$cord, "\u2192", roles$infant))
for (blk in c("B2","B4","B6")) {
  rows <- lapply(cfg$maternal_arms, function(m) {
    mod <- chain[[m]]$ratio_models[[blk]]
    if (is.null(mod)) return(NULL)
    broom::tidy(mod) %>% dplyr::filter(term != "(Intercept)") %>% mutate(maternal_arm = m)
  }) %>% bind_rows()
  if (nrow(rows) == 0) next
  p <- ggplot(rows, aes(x = term, y = estimate, fill = maternal_arm,
                        ymin = estimate - 1.96*std.error, ymax = estimate + 1.96*std.error)) +
    geom_col(position = position_dodge(0.6), width = 0.5) +
    geom_errorbar(position = position_dodge(0.6), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_fill_manual(values = cfg$marm_colors) +
    labs(x = "Pro-inflammatory / tolerogenic ratio",
         y = paste0("\u03b2 (\u2192 ", cfg$outcome_label, " ", roles$infant, ")"),
         title = blk_lab[blk], fill = "Maternal arm") +
    theme_bw(base_size = 11)
  print(p)
}
```

### Relative Importance: TdaP vs TT (full 12-subclass models)

```{r chain-relimp, results='asis'}
relimp_blocks <- c("B1_birth" = paste0("Block 1 full-12 @ ", roles$matbirth),
                   "B3_transfer" = "Block 3 full-12 (transfer)",
                   "B5_change"   = "Block 5 full-12 (change)")
for (rb in names(relimp_blocks)) {
  rTdaP <- chain[["TdaP"]]$relimp[[rb]]; rTT <- chain[["TT"]]$relimp[[rb]]
  if (is.null(rTdaP) || is.null(rTT)) next
  cmp <- merge(
    data.frame(feature = names(rTdaP@lmg), lmg_TdaP = round(rTdaP@lmg*100, 2)),
    data.frame(feature = names(rTT@lmg),   lmg_TT   = round(rTT@lmg*100, 2)),
    by = "feature") %>%
    mutate(delta_lmg = round(lmg_TdaP - lmg_TT, 2)) %>%
    arrange(desc(abs(delta_lmg)))
  knitr::kable(cmp, caption = paste0("Relative importance (lmg %) \u2014 ",
               relimp_blocks[rb], ". delta_lmg = TdaP \u2212 TT."),
               align = c("l","r","r","r")) %>% print()
  cat("\n")
}
```

---

## Participant Accounting (Maternal Chain) {#chain-naccounting}

Every model above is listed here with how many subjects entered its stratum, how many
were actually used after listwise deletion, and which predictor(s) drove the drop. This
is what lets a reader reconcile the differing sample sizes across sections — in
particular the steep attrition from antigen-total models (large n) to subclass/full-12
models (small n) flagged as a caveat in the interpretation summary.

```{r chain-naccount, results='asis'}
nt_render(nt, part = "Maternal chain",
          caption = "Table C7. Participant accounting for every maternal-chain model.")
```
