<!--
=====================================================================
parts/01_maternal_chain.Rmd   (REFACTORED — Result #2 only)
TITLE: Result #2 — Absolute subclass ratios at every upstream stage vs infant SBA
---------------------------------------------------------------------
FOCUSED REFACTOR of the original six-block maternal chain. The original
emitted absolute ratios only implicitly (Blocks 2/4/6 modelled the *delta*
ratios across transitions). This version consolidates the ABSOLUTE
pro-/tolerogenic subclass ratios at each upstream visit
(PregEarly, MatBirth, CordBlood) as predictors of the infant outcome
(SBA at InfMon2), all in one place, as requested.

Ratio definition (identical to the original):
  pro-inflammatory = IgG1 + IgG3 ; tolerogenic = IgG2 + IgG4
  ratio = log10[(IgG1 + IgG3) / (IgG2 + IgG4)]   (positive = pro-inflammatory)
The ratio columns are built by add_chain_derived() -> `<ANT>_ratio_<visit>`.

Runs as a CHILD (master provides cfg / data_raw / nt / roles) or STANDALONE
(bootstrap chunk sources the pipeline).
=====================================================================
-->

```{r ratio-setup, include=FALSE}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(broom)
})
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

if (!exists("cfg") || !exists("data_raw")) {
  .stub <- function(rel, optional = FALSE) {
    cand <- character(0)
    if (requireNamespace("here", quietly = TRUE))
      cand <- c(cand, here::here(rel), here::here("analysis", rel))
    rmd_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                        error = function(e) NA_character_)
    if (!is.na(rmd_dir)) cand <- c(cand, file.path(rmd_dir, rel),
                                   file.path(dirname(rmd_dir), rel))
    cand <- unique(c(cand, file.path(getwd(), rel), rel))
    hit  <- cand[file.exists(cand)]
    if (length(hit)) return(normalizePath(hit[1], winslash = "/", mustWork = FALSE))
    if (optional) return(NULL)
    stop("Cannot locate '", rel, "'. Run from project root.", call. = FALSE)
  }
  if (!exists("find_proj_file")) source(.stub("R/find_proj_file.R"))
  source(find_proj_file("config/endpoints.R"),  encoding = "UTF-8")
  source(find_proj_file("R/serology_helpers.R"), encoding = "UTF-8")
  cfg      <- get_endpoint_config("SBA")
  data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))
  nt       <- new_ntracker()
}
if (!exists("roles")) roles <- chain_visit_roles(cfg)
```

# Absolute subclass ratios (all upstream stages) → `r cfg$outcome_label` at `r roles$infant` {#ratio-stages}

Pro-inflammatory subclasses: IgG1, IgG3. Tolerogenic: IgG2, IgG4.
Ratio = log10[(IgG1 + IgG3) / (IgG2 + IgG4)]; positive = pro-inflammatory
polarisation. This part regresses the **absolute** PT / FHA / PRN ratios
measured at **`r paste(cfg$ratio_stages, collapse = ", ")`** on the infant
**`r cfg$outcome_label`** at `r roles$infant`, adjusting for maternal arm —
first one ratio at a time (consolidated table + forest plot), then all
three antigen ratios per stage, then every stage together.

```{r ratio-data, results='asis'}
chain_feats <- c(cfg$outcome_feature, PREDICTOR_FEATURES)

data_wide_chain <- data_raw %>%
  dplyr::filter(feature    %in% chain_feats,
                visit_name %in% cfg$chain_visits,
                arm_name   %in% cfg$maternal_arms) %>%
  dplyr::select(subject_accession, arm_name, infant_arm,
                feature, visit_name, log_assay_value) %>%
  dplyr::distinct(subject_accession, visit_name, feature, .keep_all = TRUE) %>%
  tidyr::pivot_wider(id_cols     = c(subject_accession, arm_name, infant_arm),
                     names_from  = c(feature, visit_name),
                     values_from = log_assay_value,
                     names_sep   = "_") %>%
  dplyr::mutate(arm_name = factor(arm_name, levels = cfg$maternal_arms)) %>%
  add_chain_derived(cfg)                    # builds <ANT>_ratio_<visit>

oc     <- cfg$outcome_col                   # WHOLE_SBA_InfMon2
stages <- cfg$ratio_stages                  # PregEarly, MatBirth, CordBlood
ants   <- cfg$antigens                      # PT, FHA, PRN

# every absolute-ratio column we expect, keeping only those actually built
ratio_grid <- expand.grid(antigen = ants, stage = stages,
                          stringsAsFactors = FALSE)
ratio_grid$col <- paste0(ratio_grid$antigen, "_ratio_", ratio_grid$stage)
ratio_grid     <- ratio_grid[ratio_grid$col %in% names(data_wide_chain), ]

cat(sprintf("Wide chain frame: **%d subjects**; outcome `%s` non-missing = %d.  \n",
            nrow(data_wide_chain), oc, sum(!is.na(data_wide_chain[[oc]]))))
cat(sprintf("Absolute-ratio predictors available: **%d of %d** (antigen \u00d7 stage).\n\n",
            nrow(ratio_grid), length(ants) * length(stages)))
if (nrow(ratio_grid) == 0)
  cat("> No ratio columns were built — check that the IgG1-4 subclasses exist ",
      "at these visits in the data.\n\n", sep = "")
```

## 1 · Consolidated univariate table (one ratio at a time, arm-adjusted)

Each row is a single model `r cfg$outcome_col` ~ ratio + maternal arm. The
coefficient (beta) is the slope on that ratio; a positive value means a more
pro-inflammatory profile at that stage predicts higher infant
`r cfg$outcome_label`.

```{r ratio-univariate, results='asis'}
uni_rows <- lapply(seq_len(nrow(ratio_grid)), function(i) {
  col <- ratio_grid$col[i]
  res <- run_lm(oc, c(col, "arm_name"), data_wide_chain,
                label = paste0(col, " + arm"),
                show_relimp_calc = FALSE,
                tracker = nt, part = "Ratio stages",
                section = "Univariate (arm-adjusted)",
                stratum = "TdaP+TT pooled")
  if (is.null(res$model)) return(NULL)
  ci  <- tryCatch(stats::confint(res$model, col), error = function(e) c(NA, NA))
  cf  <- summary(res$model)$coefficients
  data.frame(
    Antigen  = ratio_grid$antigen[i],
    Stage    = factor(ratio_grid$stage[i], levels = stages),
    n        = res$n_obs,
    beta     = if (col %in% rownames(cf)) unname(cf[col, 1]) else NA,
    lo       = ci[1], hi = ci[2],
    p        = if (col %in% rownames(cf)) unname(cf[col, 4]) else NA,
    stringsAsFactors = FALSE)
})
uni <- dplyr::bind_rows(uni_rows)

if (nrow(uni)) {
  uni_disp <- uni %>%
    dplyr::arrange(Stage, Antigen) %>%
    dplyr::mutate(`95% CI` = sprintf("[%.3f, %.3f]", lo, hi),
                  beta = round(beta, 3), p = signif(p, 3),
                  sig = pvalue_to_label(p)) %>%
    dplyr::select(Antigen, Stage, n, beta, `95% CI`, p, sig)
  print(knitr::kable(
    uni_disp, row.names = FALSE,
    caption = paste0("Absolute pro-/tolerogenic ratio at each stage vs ",
                     cfg$outcome_label, " at ", roles$infant,
                     " (each row: outcome ~ ratio + maternal arm; pooled arms)."),
    align = c("l", "l", "r", "r", "r", "r", "c")))
  cat("\n")
}
```

```{r ratio-forest, fig.width=8, fig.height=4.8, fig.cap=paste0("Univariate arm-adjusted slope of each absolute subclass ratio on ", cfg$outcome_label, " at ", roles$infant, " (beta with 95% CI), by stage and antigen.")}
if (exists("uni") && nrow(uni)) {
  ggplot(uni, aes(x = beta, y = Antigen, colour = Antigen)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_point(size = 2.4) +
    geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.25, linewidth = 0.7) +
    facet_wrap(~ Stage, ncol = 1, strip.position = "right") +
    scale_colour_brewer(palette = "Set1", guide = "none") +
    labs(x = paste0("beta (ratio -> ", cfg$outcome_label, " at ", roles$infant, ")"),
         y = NULL,
         title = paste0("Absolute subclass ratio -> ", cfg$outcome_label,
                        " by upstream stage")) +
    theme_bw(base_size = 11)
}
```

## 2 · Per-stage multivariable models (three antigen ratios together, arm-adjusted)

```{r ratio-perstage, results='asis'}
r2_perstage <- list()
for (st in stages) {
  cols_st <- paste0(ants, "_ratio_", st)
  cols_st <- cols_st[cols_st %in% names(data_wide_chain)]
  if (length(cols_st) == 0) next
  cat("###", st, "\n\n")
  res <- run_lm(oc, c(cols_st, "arm_name"), data_wide_chain,
                label = paste0("Ratios @ ", st, " (PT+FHA+PRN) + arm"),
                show_relimp_calc = FALSE,
                tracker = nt, part = "Ratio stages",
                section = "Per-stage multivariable", stratum = st)
  show_model(res, show_relimp = FALSE)
  r2_perstage[[st]] <- data.frame(Stage = st, n = res$n_obs,
                                  R2 = res$R2, adj_R2 = res$adj_R2)
}
if (length(r2_perstage))
  print(knitr::kable(dplyr::bind_rows(r2_perstage), row.names = FALSE,
        caption = "Per-stage multivariable fit (all three antigen ratios + arm).",
        align = c("l", "r", "r", "r")))
cat("\n")
```

## 3 · All stages together

A single model with every available absolute ratio (antigen x stage) plus
maternal arm. Complete-case deletion across all predictors makes this the
smallest-n model; the n is reported so the table can be read against the
univariate rows above.

```{r ratio-allstages, results='asis'}
all_cols <- ratio_grid$col
if (length(all_cols) >= 2) {
  res_all <- run_lm(oc, c(all_cols, "arm_name"), data_wide_chain,
                    label = "All absolute ratios (antigen x stage) + arm",
                    show_relimp_calc = FALSE,
                    tracker = nt, part = "Ratio stages",
                    section = "All stages combined", stratum = "TdaP+TT pooled")
  show_model(res_all, show_relimp = FALSE)

  # Optional collinearity diagnostic
  if (!is.null(res_all$model) && requireNamespace("car", quietly = TRUE)) {
    vif <- tryCatch(car::vif(res_all$model), error = function(e) NULL)
    if (!is.null(vif)) {
      vdf <- data.frame(term = names(vif), VIF = round(as.numeric(vif), 2))
      print(knitr::kable(vdf, row.names = FALSE,
            caption = "Variance inflation factors (all-stages model).",
            align = c("l", "r")))
      cat("\n")
    }
  }
} else {
  cat("> Fewer than two ratio predictors available — the combined model is skipped.\n\n")
}
```

## 4 · Interpretation aid

```{r ratio-interpret, results='asis'}
mapnote(
  "Reading this block: \u00a71 is the requested single consolidated view — every ",
  "absolute PT/FHA/PRN ratio at ", paste(cfg$ratio_stages, collapse = ", "),
  " against infant ", cfg$outcome_label,
  ". \u00a72 asks whether the three antigen ratios at a given stage jointly add ",
  "signal, and \u00a73 whether ratios across stages carry independent information ",
  "(watch the n and the VIFs, as the stage ratios are correlated within antigen).",
  feeds = "Result #2 of the request")
```
