```{r conc-setup, include=FALSE}
# CHILD document. Assumes cfg, data_raw, nt, find_proj_file() exist.
# Heavy network/NCT packages are optional: if absent, those sections are
# skipped with a note (the regression + mixed-model layers still run).
set.seed(20240614)
NCT_IT <- 1000L   # NCT permutations (slow; lower for a quick pass)

have_net <- all(vapply(c("bootnet", "qgraph", "networktools", "NetworkComparisonTest"),
                       requireNamespace, logical(1), quietly = TRUE))
have_lmer <- requireNamespace("lme4", quietly = TRUE)
have_corr <- requireNamespace("corrplot", quietly = TRUE)

# Network nodes: outcome + pertussis subclasses + FIM subclasses (PATCH).
# Uses cfg$network_subclasses (NETWORK_SUBCLASSES in config/endpoints.R),
# which adds FIM — estimable at the infant visits used here. ALL_SUBCLASSES
# is intentionally NOT used for nodes anymore so the chain stays unaffected.
net_vars   <- c(cfg$outcome_feature, cfg$network_subclasses)   # PATCH
node_names <- net_vars
groups_idx <- list(
  OUT = which(node_names == cfg$outcome_feature),
  PT  = which(node_names %in% PT_SUBCLASSES),
  FHA = which(node_names %in% FHA_SUBCLASSES),
  PRN = which(node_names %in% PRN_SUBCLASSES),
  FIM = which(node_names %in% FIM_SUBCLASSES))                  # PATCH: +FIM
group_colors <- c("gold", "tomato", "steelblue", "seagreen", "mediumpurple")  # PATCH: +FIM colour
# Build communities from membership (robust to node count / order) instead
# of a hard-coded positional vector. PATCH.
communities <- integer(length(node_names))
communities[groups_idx$OUT] <- 1L
communities[groups_idx$PT]  <- 2L
communities[groups_idx$FHA] <- 3L
communities[groups_idx$PRN] <- 4L
communities[groups_idx$FIM] <- 5L

# reference visit for single-visit networks / cross-arm NCT
ref_visit <- if ("InfMon5" %in% cfg$concurrent_visits) "InfMon5" else
             tail(cfg$concurrent_visits, 1)
```

# Concurrent Within-Visit Architecture {#concurrent}

This part models the infant's **`r cfg$outcome_label`** against pertussis-antigen IgG
**measured at the same infant visit** (visits: `r paste(cfg$concurrent_visits, collapse=", ")`),
rather than tracing the upstream maternal chain. It reproduces the network /
mixed-model architecture: relative-importance decomposition, the
pro-/tolerogenic ratio models, the antigen × subclass × level interaction, partial-correlation
(GGM) networks with bridge centrality, and network-comparison tests (NCT).

**Stratification is maternal-arm primary** (TdaP / TT), with infant splits (aP, wP)
nested. Network and NCT steps need adequate n (≥ `r NETWORK_MIN` complete cases at a
visit); underpowered cells are skipped and recorded in the participant accounting.

```{r conc-mapnote, echo=FALSE, results='asis'}
mapnote("This part corresponds to the **stratified within-visit interpretation summary** ",
        "(the \"", cfg$outcome_label, " aP vs wP\" narrative), now re-centred on the maternal arm.",
        feeds = c(
          "Antigen / subclass relative importance \u2192 summary \u00a71 (PRN dominance) and \u00a73",
          "FHA negative association \u2192 \u00a72",
          "Ratio models (overall + by visit) \u2192 \u00a74",
          "Antigen \u00d7 subclass \u00d7 level interaction \u2192 \u00a75",
          "GGM networks + bridge centrality \u2192 \u00a76",
          "NCT (visit-to-visit, arm contrasts) \u2192 \u00a77/\u00a78"))
if (!have_net)
  cat("\n> **Note:** network packages (bootnet/qgraph/networktools/NetworkComparisonTest) ",
      "not installed; network and NCT sections are skipped. Install them to enable \u00a76\u2013\u00a78.\n\n")
```

---

## Data Preparation

```{r conc-data, results='asis'}
conc_feats <- c(cfg$outcome_feature, PREDICTOR_FEATURES, cfg$network_subclasses)  # PATCH: +FIM nodes
data_long_c <- data_raw %>%
  dplyr::filter(feature %in% conc_feats,
                visit_name %in% cfg$concurrent_visits,
                arm_name %in% cfg$maternal_arms) %>%
  dplyr::select(subject_accession, visit_name, arm_name, infant_arm,
                feature, log_assay_value) %>%
  distinct(subject_accession, visit_name, feature, .keep_all = TRUE)

data_wide_c <- data_long_c %>%
  pivot_wider(id_cols = c(subject_accession, visit_name, arm_name, infant_arm),
              names_from = feature, values_from = log_assay_value) %>%
  add_concurrent_ratios()

cat(sprintf("Concurrent dataset: **%d subject\u00d7visit rows \u00d7 %d columns**  \n",
            nrow(data_wide_c), ncol(data_wide_c)))

cnt <- as.data.frame(table(data_wide_c$arm_name, data_wide_c$visit_name))
colnames(cnt) <- c("Maternal arm", "Visit", "n_rows")
cnt <- cnt %>% pivot_wider(names_from = Visit, values_from = n_rows, values_fill = 0)
knitr::kable(cnt, caption = "Table N1. Subject\u00d7visit rows by maternal arm and visit.") %>% print()
```

```{r conc-strata, results='hide'}
strata_x <- build_maternal_strata(data_wide_c)
conc <- lapply(names(strata_x), function(k)
  run_concurrent_models(strata_x[[k]]$data, cfg, tracker = nt,
                        part = "Concurrent", stratum = strata_x[[k]]$label))
names(conc) <- names(strata_x)

display_conc_stratum <- function(key, full = TRUE) {
  cat("####", strata_x[[key]]$label, "\n\n")
  cm <- conc[[key]]
  show_model(cm$br[["antigen"]])
  if (full) { show_model(cm$br[["PT"]]); show_model(cm$br[["FHA"]]); show_model(cm$br[["PRN"]]) }
  show_model(cm$br[["full12"]])
  show_model(cm$br[["ratio"]], show_relimp = FALSE)
}
```

---

## Regression Layer: Antigen, Subclass, and Ratio Models

```{r conc-reg-mapnote, echo=FALSE, results='asis'}
mapnote("Relative-importance and ratio models \u2014 the basis for the PRN-dominance and ",
        "FHA-negative findings.",
        feeds = c("\u00a71 (PRN dominance) and \u00a73 (arm-specific signatures)",
                  "\u00a72 (FHA negative association)",
                  "\u00a74 (ratio model)"))
```

```{r conc-reg, results='asis'}
for (m in cfg$maternal_arms) {
  cat("###", m, "mothers\n\n")
  display_conc_stratum(m, full = TRUE)                       # pooled margin
  for (i in cfg$infant_arms) {
    key <- paste0(m, "_", i)
    if (!is.null(conc[[key]])) display_conc_stratum(key, full = FALSE)
  }
}
```

### Ratio Model by Visit (pooled maternal margins)

```{r conc-ratio-by-visit, results='asis'}
for (m in cfg$maternal_arms) {
  d <- strata_x[[m]]$data
  rbv <- d %>% group_by(visit_name) %>%
    group_modify(~ broom::tidy(lm(reformulate(c("PT_ratio","FHA_ratio","PRN_ratio"),
                                              response = cfg$outcome_feature), data = .x))) %>%
    dplyr::filter(term != "(Intercept)") %>%
    mutate(across(c(estimate, std.error, statistic), ~round(., 3)),
           p.value = signif(p.value, 3), sig = pvalue_to_label(p.value)) %>%
    ungroup()
  knitr::kable(rbv, caption = paste0("Ratio model by visit \u2014 ", m, " mothers."),
               align = c("l","l","r","r","r","r","c")) %>% print()
  cat("\n")
}
```

---

## Antigen × Subclass × Level Interaction (mixed model)

```{r conc-lmer-mapnote, echo=FALSE, results='asis'}
mapnote("The three-way interaction tests whether which subclass matters depends on ",
        "antigen and on binding level.",
        feeds = "\u00a75 (three-way antigen \u00d7 subclass interaction)")
```

```{r conc-lmer, results='asis'}
if (have_lmer) {
  for (m in cfg$maternal_arms) {
    d <- strata_x[[m]]$data
    dl <- d %>%
      dplyr::select(subject_accession, visit_name, arm_name,
                    all_of(cfg$outcome_feature), matches("IgG[1-4]$")) %>%
      pivot_longer(matches("IgG[1-4]$"), names_to = c("antigen","subclass"),
                   names_pattern = "(PT|FHA|PRN)_(IgG[1-4])", values_to = "subclass_value") %>%
      tidyr::drop_na(tidyselect::all_of(c("subclass_value", cfg$outcome_feature)))
    fml <- as.formula(paste0(cfg$outcome_feature,
                             " ~ antigen * subclass * subclass_value + (1 | subject_accession)"))
    mm <- tryCatch(lme4::lmer(fml, data = dl), error = function(e) NULL)
    if (is.null(mm)) { cat(sprintf("*%s: mixed model could not be fit.*\n\n", m)); next }
    av <- as.data.frame(anova(mm))
    av <- cbind(Term = rownames(av), round(av, 3)); rownames(av) <- NULL
    knitr::kable(av, caption = paste0("Type-III ANOVA of fixed effects \u2014 ", m, " mothers (lmer)."),
                 row.names = FALSE) %>% print()
    cat("\n")
  }
} else cat("*lme4 not installed; mixed-model interaction skipped.*\n\n")
```

### Antigen × Subclass Slopes (pooled maternal margins)

```{r conc-slopes, fig.cap=paste0("Figure N1. Antigen \u00d7 subclass slopes \u2192 ", cfg$outcome_label, ", by maternal arm."), fig.height=6}
dl_all <- bind_rows(lapply(cfg$maternal_arms, function(m)
  strata_x[[m]]$data %>%
    dplyr::select(arm_name, all_of(cfg$outcome_feature), matches("IgG[1-4]$")))) %>%
  pivot_longer(matches("IgG[1-4]$"), names_to = c("antigen","subclass"),
               names_pattern = "(PT|FHA|PRN)_(IgG[1-4])", values_to = "subclass_value") %>%
  tidyr::drop_na(tidyselect::all_of(c("subclass_value", cfg$outcome_feature)))
ggplot(dl_all, aes(x = subclass_value, y = .data[[cfg$outcome_feature]],
                   color = subclass, linetype = arm_name)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ arm_name) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Log10 subclass-specific IgG (concurrent)", y = cfg$outcome_y_label,
       color = "Subclass", linetype = "Maternal arm",
       title = paste0("Antigen \u00d7 subclass \u2192 ", cfg$outcome_label, " (within-visit)")) +
  theme_bw(base_size = 11)
```

---

## GGM Networks and Bridge Centrality

```{r conc-net-mapnote, echo=FALSE, results='asis'}
mapnote("Partial-correlation networks and bridge centrality at visit ", ref_visit, ".",
        feeds = "\u00a76 (network architecture, bridge nodes)")
```

```{r conc-networks, results='asis', fig.height=6, fig.width=7}
bridge_by_marm <- list(); nets_by_marm <- list()
if (have_net) {
  for (m in cfg$maternal_arms) {
    dnet <- strata_x[[m]]$data %>%
      dplyr::filter(visit_name == ref_visit) %>%
      dplyr::select(all_of(net_vars)) %>% tidyr::drop_na()
    nt_log(nt, "Concurrent", paste0("Network @ ", ref_visit), strata_x[[m]]$label,
           paste0("GGM network @ ", ref_visit), net_vars, strata_x[[m]]$data %>% dplyr::filter(visit_name == ref_visit))
    cat(sprintf("\n#### %s mothers \u2014 network at %s (n = %d)\n\n", m, ref_visit, nrow(dnet)))
    if (nrow(dnet) < NETWORK_MIN) { cat(sprintf("*n < %d; network skipped.*\n\n", NETWORK_MIN)); next }
    net <- tryCatch(bootnet::estimateNetwork(dnet, default = "EBICglasso", threshold = TRUE),
                    error = function(e) NULL)
    if (is.null(net)) { cat("*Network estimation failed.*\n\n"); next }
    nets_by_marm[[m]] <- net
    qgraph::qgraph(net$graph, groups = groups_idx, color = group_colors,
                   layout = "spring", labels = node_names, legend = TRUE,
                   legend.cex = 0.35, vsize = 8,
                   title = paste0(m, " \u2014 ", cfg$outcome_label, " network @ ", ref_visit))
    br <- networktools::bridge(net$graph, communities = communities)
    bdf <- data.frame(Variable = names(br$`Bridge Strength`),
                      Bridge_Strength = round(br$`Bridge Strength`, 3),
                      Bridge_EI = round(br$`Bridge Expected Influence (1-step)`, 3),
                      row.names = NULL) %>% arrange(desc(Bridge_Strength))
    bridge_by_marm[[m]] <- bdf
    print(knitr::kable(bdf, caption = paste0("Bridge centrality \u2014 ", m, " @ ", ref_visit)))
    cat("\n")
    if (have_corr) {
      adj <- net$graph; rownames(adj) <- colnames(adj) <- net_vars
      corrplot::corrplot(adj, method = "color", type = "lower", tl.col = "black",
                         tl.cex = 0.7, order = "hclust", addrect = 4, mar = c(0,0,2,0),
                         col = colorRampPalette(c("blue","white","red"))(200),
                         title = paste0("Partial correlations \u2014 ", m, " @ ", ref_visit))
    }
  }
} else cat("*Network packages not installed; GGM / bridge sections skipped.*\n\n")
```

```{r conc-bridge-compare, results='asis'}
if (length(bridge_by_marm) == 2) {
  cmp <- merge(bridge_by_marm[["TdaP"]] %>% dplyr::select(Variable, Bridge_Strength) %>%
                 rename(BS_TdaP = Bridge_Strength),
               bridge_by_marm[["TT"]] %>% dplyr::select(Variable, Bridge_Strength) %>%
                 rename(BS_TT = Bridge_Strength), by = "Variable") %>%
    mutate(delta_BS = round(BS_TdaP - BS_TT, 3)) %>% arrange(desc(abs(delta_BS)))
  knitr::kable(cmp, caption = paste0("Table N2. Bridge-strength comparison TdaP vs TT @ ", ref_visit, ".")) %>% print()
}
```

---

## Network Comparison Tests (NCT)

```{r conc-nct-mapnote, echo=FALSE, results='asis'}
mapnote("NCT contrasts: maternal arm (TdaP vs TT), infant arm within maternal arm, ",
        "and first-vs-last visit within maternal arm.",
        feeds = "\u00a77 (temporal dynamics) and \u00a78 (cross-arm network comparison)")
```

```{r conc-nct, results='asis'}
nct_subset <- function(stratum_data, visit) stratum_data %>%
  dplyr::filter(visit_name == visit) %>% dplyr::select(all_of(net_vars)) %>% tidyr::drop_na()

run_nct <- function(d1, d2, lab) {
  cat(sprintf("\n**%s** \u2014 n1 = %d, n2 = %d\n\n", lab, nrow(d1), nrow(d2)))
  if (nrow(d1) < NETWORK_MIN || nrow(d2) < NETWORK_MIN) {
    cat(sprintf("*One or both groups n < %d; NCT skipped.*\n\n", NETWORK_MIN)); return(invisible())
  }
  res <- tryCatch(NetworkComparisonTest::NCT(d1, d2, it = NCT_IT, test.edges = TRUE,
                  test.centrality = TRUE, centrality = c("strength","expectedInfluence")),
                  error = function(e) NULL)
  if (is.null(res)) { cat("*NCT failed.*\n\n"); return(invisible()) }
  cat(sprintf("- Network-structure invariance (M) p = %.3f\n", res$nwinv.pval))
  cat(sprintf("- Global-strength invariance (S) p = %.3f\n\n", res$glstrinv.pval))
}

if (have_net) {
  # (a) TdaP vs TT at reference visit (pooled infants)
  run_nct(nct_subset(strata_x[["TdaP"]]$data, ref_visit),
          nct_subset(strata_x[["TT"]]$data,   ref_visit),
          paste0("Maternal arm: TdaP vs TT @ ", ref_visit))
  # (b) aP vs wP within each maternal arm at reference visit
  for (m in cfg$maternal_arms) {
    ka <- paste0(m, "_aP"); kw <- paste0(m, "_wP")
    if (!is.null(strata_x[[ka]]) && !is.null(strata_x[[kw]]))
      run_nct(nct_subset(strata_x[[ka]]$data, ref_visit),
              nct_subset(strata_x[[kw]]$data, ref_visit),
              paste0("Infant arm within ", m, ": aP vs wP @ ", ref_visit))
  }
  # (c) first vs last visit within each maternal arm
  v1 <- cfg$concurrent_visits[1]; vL <- tail(cfg$concurrent_visits, 1)
  for (m in cfg$maternal_arms)
    run_nct(nct_subset(strata_x[[m]]$data, v1),
            nct_subset(strata_x[[m]]$data, vL),
            paste0("Temporal within ", m, ": ", v1, " vs ", vL))
} else cat("*Network packages not installed; NCT skipped.*\n\n")
```

---

## Cross-Stratum Comparisons

```{r conc-relimp, results='asis'}
rT <- conc[["TdaP"]]$relimp_full; rN <- conc[["TT"]]$relimp_full
if (!is.null(rT) && !is.null(rN)) {
  cmp <- merge(data.frame(feature = names(rT@lmg), lmg_TdaP = round(rT@lmg*100, 2)),
               data.frame(feature = names(rN@lmg), lmg_TT   = round(rN@lmg*100, 2)),
               by = "feature") %>%
    mutate(delta_lmg = round(lmg_TdaP - lmg_TT, 2)) %>% arrange(desc(abs(delta_lmg)))
  knitr::kable(cmp, caption = "Table N3. Full-12 relative importance (lmg %): TdaP vs TT (concurrent).",
               align = c("l","r","r","r")) %>% print()
}
```

```{r conc-ratio-coef, fig.cap="Figure N2. Concurrent ratio-model coefficients: TdaP vs TT. Error bars = 95% CI.", fig.height=5}
rows <- bind_rows(lapply(cfg$maternal_arms, function(m) {
  mod <- conc[[m]]$ratio_model
  if (is.null(mod)) return(NULL)
  broom::tidy(mod) %>% dplyr::filter(term != "(Intercept)") %>% mutate(maternal_arm = m)
}))
if (nrow(rows) > 0)
  ggplot(rows, aes(x = term, y = estimate, fill = maternal_arm,
                   ymin = estimate - 1.96*std.error, ymax = estimate + 1.96*std.error)) +
    geom_col(position = position_dodge(0.6), width = 0.5) +
    geom_errorbar(position = position_dodge(0.6), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_fill_manual(values = cfg$marm_colors) +
    labs(x = "Pro-/tolerogenic ratio", y = paste0("\u03b2 (\u2192 ", cfg$outcome_label, ")"),
         title = paste0("Concurrent ratio model: TdaP vs TT (", cfg$outcome_label, ")"),
         fill = "Maternal arm") +
    theme_bw(base_size = 11)
```

---

## Participant Accounting (Concurrent) {#concurrent-naccounting}

```{r conc-naccount, results='asis'}
nt_render(nt, part = "Concurrent",
          caption = "Table N4. Participant accounting for every concurrent model / network.")
```
