# Step 5 — FcγR Collinearity Diagnostics {#step5}

In the earlier partial-correlation networks every FcγR3b node dropped to zero — the signature of
FcγR2a and FcγR3b being near-collinear, so conditioning on one erases its twin. That is a modelling
artifact, not evidence FcγR3b is unimportant. This step recovers the **FcγR3b-specific** signal you
care about by reporting, per antigen and arm: the receptor–receptor correlation, the variance
inflation, and — crucially — FcγR3b's **zero-order** associations (with the outcome, with IgG1, with
ADCD) beside its **partial** association after adjusting for FcγR2a.

```{r step5-helpers, include=FALSE}
pcor_resid <- function(d, x, y, z) {  # partial cor of x,y given z, via residuals
  rx <- resid(lm(reformulate(z, response = x), data = d))
  ry <- resid(lm(reformulate(z, response = y), data = d))
  suppressWarnings(cor(rx, ry))
}
vif2 <- function(d, a, b, ctrl) {  # VIF of a given b + ctrl
  r2 <- summary(lm(reformulate(c(b, ctrl), response = a), data = d))$r.squared
  1 / (1 - r2)
}
```

```{r step5-run, results='asis'}
.stage(5, "FcγR collinearity diagnostics")
rows <- list()
for (A in PATHWAY_ANTIGENS) {
  r2a <- paste0(A, "_FcgR2a"); r3b <- paste0(A, "_FcgR3b")
  igg1 <- paste0(A, "_IgG1"); adcd <- paste0(A, "_ADCD")
  if (!all(c(r2a, r3b, igg1) %in% names(path_wide))) next
  for (m in cfg$maternal_arms) {
    need <- keep_present(c(OUTCOME, r2a, r3b, igg1, adcd))
    d <- arm_data(m) %>% dplyr::select(dplyr::all_of(need)) %>% tidyr::drop_na()
    if (nrow(d) < 20) next
    .info("step5 %s %s n=%d", A, m, nrow(d))
    rows[[length(rows) + 1]] <- data.frame(
      antigen = A, arm = m, n = nrow(d),
      cor_2a_3b     = round(cor(d[[r2a]], d[[r3b]]), 3),
      VIF_3b        = round(vif2(d, r3b, r2a, igg1), 2),
      r_3b_OUT      = round(cor(d[[r3b]], d[[OUTCOME]]), 3),
      pr_3b_OUT_2a  = round(pcor_resid(d, r3b, OUTCOME, r2a), 3),
      r_3b_IgG1     = round(cor(d[[r3b]], d[[igg1]]), 3),
      r_3b_ADCD     = if (adcd %in% names(d)) round(cor(d[[r3b]], d[[adcd]]), 3) else NA_real_,
      stringsAsFactors = FALSE)
  }
}
diag_tbl <- do.call(rbind, rows)
if (is.null(diag_tbl)) {
  cat("*No antigen had a complete FcγR pair; step 5 skipped.*\n\n")
} else {
  print(knitr::kable(diag_tbl, row.names = FALSE,
    caption = "FcγR collinearity and FcγR3b-specific associations (zero-order r vs partial pr | FcγR2a).",
    col.names = c("Antigen", "Arm", "n", "r(2a,3b)", "VIF 3b", "r(3b,OUT)",
                  "pr(3b,OUT|2a)", "r(3b,IgG1)", "r(3b,ADCD)")))
  cat("\n")
}
```

**How to read this.** A high `r(2a,3b)` (say > 0.8) and `VIF 3b` well above ~5 confirm the receptors
are redundant, explaining why the joint network zeroed FcγR3b. The informative columns are the
zero-order associations: a solid `r(3b,IgG1)` supports IgG1↔FcγR3b co-regulation, a solid
`r(3b,ADCD)` supports the FcγR3b→ADCD step, and `r(3b,OUT)` its link to killing — none of which the
partialled network could show. If `pr(3b,OUT|2a)` collapses toward 0 while `r(3b,OUT)` is large, the
two receptors are interchangeable carriers of the same signal, and FcγR3b-specific claims should
rest on the zero-order / single-receptor evidence rather than the joint model. Consistent patterns
across arms indicate the redundancy (and the underlying signal) is arm-invariant.
