# Step 7 — FcγR3b Unique Role {#step7}

The joint networks zeroed FcγR3b because it is collinear with FcγR2a. This step separates the two
receptors and asks whether **FcγR3b carries information independent of FcγR2a** — for ADCD (the
activation step) and for `r cfg$outcome_label` (the functional endpoint). Two decouplings are used:
the **residual of FcγR3b on FcγR2a** (the part of 3b orthogonal to 2a) and the **2a/3b log-ratio**.
Run per antigen, per arm, and in each data mode.

```{r step7-helpers, include=FALSE}
# unique contribution of `focal` beyond `other`: residualize focal on other,
# then measure the residual's incremental R² and coefficient on outcome y.
unique_contrib <- function(d, y, focal, other) {
  d <- d[stats::complete.cases(d[, c(y, focal, other)]), ]
  if (nrow(d) < 20) return(NULL)
  d$.fres <- resid(lm(reformulate(other, response = focal), data = d))
  base <- lm(reformulate(other, response = y), data = d)
  full <- lm(reformulate(c(other, ".fres"), response = y), data = d)
  co <- summary(full)$coefficients[".fres", ]
  list(n = nrow(d), dR2 = summary(full)$r.squared - summary(base)$r.squared,
       beta = unname(co[1]), p = unname(co[4]),
       r0 = suppressWarnings(cor(d[[focal]], d[[y]])))
}
```

```{r step7-run, results='asis'}
.stage(7, "FcγR3b unique role (residual-on-2a and 2a/3b ratio)")
if (!RUN_FCGR3B) {
  cat("*Step 7 skipped (run_fcgr3b = FALSE).*\n\n")
} else {
  .modes <- if (RUN_STD_RERUNS) MODES else "raw"
  for (mode in .modes) {
    cat("## Data mode: `", mode, "`", mode_desc(mode), "\n\n", sep = "")
    wd <- wide_for(mode)
    rows <- list()
    for (A in PATHWAY_ANTIGENS) {
      r2a <- paste0(A, "_FcgR2a"); r3b <- paste0(A, "_FcgR3b"); adcd <- paste0(A, "_ADCD")
      if (!all(c(r2a, r3b) %in% names(wd))) next
      for (m in cfg$maternal_arms) {
        d <- arm_data_m(mode, m)
        # FcγR3b unique contribution to ADCD (activation step)
        uc_adcd <- if (adcd %in% names(wd)) unique_contrib(d, adcd, r3b, r2a) else NULL
        # FcγR3b unique contribution to the whole-cell outcome
        uc_out  <- if (OUTCOME %in% names(wd)) unique_contrib(d, OUTCOME, r3b, r2a) else NULL
        dd <- d[stats::complete.cases(d[, c(r2a, r3b)]), ]
        ratio <- dd[[r3b]] - dd[[r2a]]                       # log(3b/2a)
        r_ratio_out <- if (OUTCOME %in% names(wd))
          suppressWarnings(cor(ratio, dd[[OUTCOME]], use = "complete.obs")) else NA
        rows[[length(rows) + 1]] <- data.frame(
          antigen = A, arm = m,
          cor_2a_3b = round(suppressWarnings(cor(dd[[r2a]], dd[[r3b]])), 3),
          "3b|2a→ADCD_dR2" = if (is.null(uc_adcd)) NA else round(uc_adcd$dR2, 3),
          "3b|2a→ADCD_p"   = if (is.null(uc_adcd)) NA else signif(uc_adcd$p, 3),
          "3b|2a→OUT_dR2"  = if (is.null(uc_out))  NA else round(uc_out$dR2, 3),
          "3b|2a→OUT_p"    = if (is.null(uc_out))  NA else signif(uc_out$p, 3),
          "r(3b/2a,OUT)"   = round(r_ratio_out, 3),
          check.names = FALSE, stringsAsFactors = FALSE)
        .info("step7 [%s] %s %s done", mode, A, m)
      }
    }
    tb <- do.call(rbind, rows)
    if (is.null(tb)) { cat("*No complete FcγR pair in this mode.*\n\n"); next }
    print(knitr::kable(tb, row.names = FALSE,
      caption = paste0("FcγR3b unique contribution beyond FcγR2a [", mode,
                       "]: incremental R² and coefficient p of the 3b-residual on ADCD and on the outcome; last column is the 2a/3b ratio's correlation with the outcome."))); cat("\n")
  }
}
```

**How to read this.** `3b|2a→ADCD` columns test the activation claim *net of FcγR2a*: a significant
positive incremental R² means FcγR3b engages complement independently of FcγR2a (supporting the
"FcγR3b essential for ADCD" premise); near-zero means the two receptors are interchangeable for ADCD.
`3b|2a→OUT` does the same for killing. The ratio column asks whether the *balance* between the two
receptors (rather than either level) tracks the outcome. A consistent, significant 3b-unique
contribution in one arm but not the other localises the FcγR3b dependence to that arm.
