---
title: "Figure 4 (revised) — Three-panel determinants of maternal responder status"
subtitle: |
  Left: predictor AUC ladder (arm / +pre-vaccination / +all-maternal).
  Middle (NEW): pre-vaccination (early-pregnancy) serology signature.
  Right: all-maternal serology signature.
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
params:
  responder_features: !r c("PT_IgG", "FHA_IgG", "PRN_IgG")
  maternal_visits:    !r c("PregEarly", "MatBirth")
  baseline_visit:     "PregEarly"     # the pre-vaccination draw
  breadth_thresh:     0.6
  en_alpha:           0.5
  en_folds:           10
  min_coverage:       0.6
  seed:               1
  top_n_mid:          12               # pre-vaccination features to show (middle panel)
  top_n_right:        12               # all-maternal features to show (right panel)
  # Right panel now defaults to the top `top_n_right` features by AUC (curated =
  # NULL). To instead pin an exact curated set, list feature@visit strings here
  # (best first), e.g. c("PRN_IgG1@MatBirth", "PT_IgG1@MatBirth", ...).
  curated_right: !r NULL
  fig_out_dir:        "figures"
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
```

```{r bootstrap, include=FALSE}
library(here); library(dplyr); library(tidyr); library(ggplot2)

# ---- self-locating resolver (identical contract to the Phase-08 doc) ----
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R", "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE); rm(.ff_cand, .ff_rmd, .ff_hit)
}

# ---- shared config + helpers (read-only, exactly as Phase 08) ----
source(find_proj_file("config/endpoints.R"),             local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"),           local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/within_visit_helpers.R"),       local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/maternal_responder_helpers.R"), local = FALSE, encoding = "UTF-8")

data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))
has_glmnet <- requireNamespace("glmnet", quietly = TRUE)
has_cluster <- requireNamespace("cluster", quietly = TRUE)
```

# Rebuild the responder class, arm, and serology matrices

Everything here re-uses the Phase-08 engine so the figure is guaranteed to be the
same construct as the rest of the analysis. The only *new* computation is the
univariate AUC of the **pre-vaccination** (`r params$baseline_visit`) matrix, which
drives the middle panel.

```{r rebuild}
RESP_FEATS <- params$responder_features
MAT_VIS    <- params$maternal_visits

mat_cls <- build_maternal_responder_class(
  data_raw, MAT_VIS, RESP_FEATS,
  breadth_thresh = params$breadth_thresh, has_cluster = has_cluster)

arm_tbl <- subject_maternal_arm(data_raw)

# never let ACT / PENTAMER in (protocol), and drop the 3 score-defining features
.excl_anti <- { av <- unique(as.character(data_raw$feature))
                av[sub("_.*$", "", toupper(av)) %in% c("ACT", "PENTAMER")] }
.excl <- union(RESP_FEATS, .excl_anti)

W_all  <- build_maternal_serology_matrix(data_raw, MAT_VIS, exclude_feats = .excl,
                                         min_coverage = params$min_coverage)
W_base <- build_maternal_serology_matrix(data_raw, params$baseline_visit,
                                         exclude_feats = .excl,
                                         min_coverage = params$min_coverage)

# univariate AUC of every feature against the Low/High class
su_all  <- serology_univariate_auc(W_all,  mat_cls)   # right panel source
su_base <- serology_univariate_auc(W_base, mat_cls)   # middle panel source (NEW)
```

```{r left-panel-aucs, results='asis'}
# Left panel: the AUC ladder. arm-only is the rank statistic (auc_binary of the
# arm indicator); the two serology rows are cross-validated elastic-net AUCs.
data_lr <- merge(mat_cls[, c("subject_accession", "responder")], arm_tbl,
                 by = "subject_accession")
data_lr <- data_lr[data_lr$arm_name %in% c("TdaP", "TT"), , drop = FALSE]
arm_auc <- auc_binary(as.numeric(data_lr$arm_name == "TdaP"), data_lr$responder)

cv_base <- cv_all <- NA_real_
if (has_glmnet) {
  en_base <- elastic_net_signature(W_base, mat_cls, arm_tbl = arm_tbl,
                                   alpha = params$en_alpha, nfolds = params$en_folds,
                                   seed = params$seed)
  en_all  <- elastic_net_signature(W_all,  mat_cls, arm_tbl = arm_tbl,
                                   alpha = params$en_alpha, nfolds = params$en_folds,
                                   seed = params$seed)
  if (isTRUE(en_base$ok)) cv_base <- en_base$cv_auc
  if (isTRUE(en_all$ok))  cv_all  <- en_all$cv_auc
} else {
  # fall back to the values reported in Table C2 if glmnet is unavailable
  cv_base <- 0.883; cv_all <- 0.918
  cat("> *glmnet not installed — using the Table C2 CV-AUC values (0.88 / 0.92) ",
      "for the left panel. Install `glmnet` to recompute live.*\n\n")
}

left_df <- data.frame(
  label = c("Maternal arm only\n(rank statistic)",
            "Arm + pre-vaccination\nserology (CV)",
            "Arm + all maternal\nserology (CV)"),
  auc   = c(arm_auc, cv_base, cv_all),
  kind  = c("arm", "base", "all"),
  stringsAsFactors = FALSE)
left_df$label <- factor(left_df$label, levels = rev(left_df$label))
knitr::kable(left_df[, c("label", "auc")], digits = 3,
             caption = "Left-panel AUC ladder (arm-only = rank statistic; serology rows = CV-AUC).")
```

# Shared styling helpers

```{r style-helpers}
# antigen palette. IPV uses the orange from the published figure; FIM/TT get
# their own distinct hues so they are legible in the signatures.
ANTIGEN_COL <- c(PRN = "#206090", PT = "#d04050", FHA = "#308040",
                 FIM = "#0f9b9b", IPV = "#e08000", DT = "#7050a0",
                 TT = "#8a6d3b", DEFAULT = "#9a9a9a")
LEFT_COL <- c(arm = "#9a9a9a", base = "#5b8fc0", all = "#206090")

# pretty label: "PRN_IgG1@MatBirth" -> "PRN IgG1 @ birth"
VISIT_PRETTY <- c(PregEarly = "early", MatBirth = "birth")
pretty_feature <- function(f) {
  base  <- sub("@.*$", "", f)
  visit <- sub("^.*@", "", f)
  vp <- ifelse(visit %in% names(VISIT_PRETTY), VISIT_PRETTY[visit], visit)
  paste0(gsub("_", " ", base), " @ ", vp)
}

# build a plot-ready frame from a univariate-AUC table (adds antigen + prevax flag)
prep_features <- function(su, keep = NULL, top_n = NULL, baseline_visit = params$baseline_visit) {
  su$base_feature <- sub("@.*$", "", su$feature)
  su$visit        <- sub("^.*@", "", su$feature)
  cls <- classify_feature(unique(su$base_feature))          # -> feature, antigen, ...
  su  <- dplyr::left_join(su, cls, by = c("base_feature" = "feature"))
  if (!is.null(keep)) {
    su <- su[match(keep, su$feature), , drop = FALSE]       # keep + preserve order
    su <- su[!is.na(su$feature), , drop = FALSE]
  } else {
    su <- su[order(-su$auc), , drop = FALSE]
    if (!is.null(top_n)) su <- utils::head(su, top_n)
  }
  # colour poliovirus serotypes (IPV1/IPV2/IPV3) as one IPV family, then fall
  # back to grey for anything outside the named palette.
  su$antigen  <- sub("^IPV[0-9]+$", "IPV", su$antigen)
  su$antigen  <- ifelse(su$antigen %in% names(ANTIGEN_COL), su$antigen, "DEFAULT")
  su$prevax   <- su$visit == baseline_visit
  su$plabel   <- pretty_feature(su$feature)
  su$plabel   <- factor(su$plabel, levels = rev(su$plabel)) # top bar on top
  su
}

# is a diagonal-hatch pattern available? (nicer, matches the original)
use_pattern <- requireNamespace("ggpattern", quietly = TRUE)

# One horizontal-bar signature panel: antigen colour + a pre-vaccination marker
# (diagonal hatch via ggpattern when available, else a dark outline). Value
# labels are drawn in a separate geom_text with a fixed colour so they never get
# recoloured by the fill/outline/pattern scales. Guides are emitted identically
# in every panel (all antigens, drop = FALSE) so patchwork can collect them into
# one shared legend at the bottom of the figure.
signature_panel <- function(df, title, show_legend = FALSE) {
  ANT_BREAKS <- c("PRN","PT","FHA","FIM","IPV","DT","TT")
  fill_guide <- if (show_legend) guide_legend(order = 1, ncol = 2) else "none"
  mark_guide <- if (show_legend) guide_legend(order = 2, ncol = 1) else "none"
  p <- ggplot(df, aes(auc, plabel, fill = antigen))
  if (use_pattern) {
    p <- p + ggpattern::geom_col_pattern(
      aes(pattern = prevax), width = 0.72, colour = NA,
      pattern_fill = "white", pattern_angle = 45,
      pattern_density = 0.12, pattern_spacing = 0.03, pattern_size = 0.05) +
      ggpattern::scale_pattern_manual(
        values = c(`TRUE` = "stripe", `FALSE` = "none"),
        breaks = TRUE, labels = c(`TRUE` = "pre-vaccination"),
        name = NULL, guide = mark_guide)
  } else {
    p <- p + geom_col(aes(colour = prevax), width = 0.72, linewidth = 0.6) +
      scale_colour_manual(values = c(`TRUE` = "grey15", `FALSE` = NA),
                          breaks = TRUE, labels = c(`TRUE` = "pre-vaccination"),
                          name = NULL, guide = mark_guide)
  }
  # value labels: inside (white) for long bars, outside (dark) for short bars
  df$inside <- df$auc >= 0.72
  p +
    geom_text(data = df[df$inside, , drop = FALSE],
              aes(label = sprintf("%.2f", auc)),
              hjust = 1.15, colour = "white", fontface = "bold", size = 3.1) +
    geom_text(data = df[!df$inside, , drop = FALSE],
              aes(label = sprintf("%.2f", auc)),
              hjust = -0.20, colour = "grey20", fontface = "bold", size = 3.1) +
    scale_fill_manual(values = ANTIGEN_COL, breaks = ANT_BREAKS, drop = FALSE,
                      name = NULL, guide = fill_guide) +
    scale_x_continuous(breaks = seq(0.5, 1.0, 0.1), expand = c(0, 0)) +
    # zoom the view to 0.5-1.0 WITHOUT censoring the bars (bars are drawn from 0;
    # scale limits would drop them as out-of-bounds — coord_cartesian just clips).
    coord_cartesian(xlim = c(0.5, 1.0)) +
    labs(x = "Univariate AUC (Low vs High responder)", y = NULL, title = title) +
    theme_minimal(base_size = 11) +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
          plot.title = element_text(face = "bold", size = 12),
          axis.ticks = element_blank(),
          # boxed legend tucked into the bottom-right corner (like the preview)
          legend.position = if (show_legend) c(0.99, 0.02) else "none",
          legend.justification = c(1, 0),
          legend.box = "vertical", legend.box.just = "right",
          legend.spacing.y = grid::unit(1, "pt"),
          legend.key.size = grid::unit(11, "pt"),
          legend.text = element_text(size = 8.5),
          legend.margin = margin(1, 1, 1, 1),
          legend.box.margin = margin(2, 3, 2, 3),
          legend.background = element_blank(),
          legend.box.background = element_rect(fill = "white", colour = "grey70", linewidth = 0.4))
}
```

# Assemble the three panels

```{r build-panels}
## --- LEFT: predicting responder status ---
p_left <- ggplot(left_df, aes(auc, label, fill = kind)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = sprintf("%.2f", auc)), hjust = 1.2,
            colour = "white", fontface = "bold", size = 3.6) +
  scale_fill_manual(values = LEFT_COL, guide = "none") +
  scale_x_continuous(breaks = seq(0.5, 1.0, 0.1), expand = c(0, 0)) +
  coord_cartesian(xlim = c(0.5, 1.0)) +
  labs(x = "AUC (Low vs High responder)", y = NULL, title = "Predicting responder status") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 12), axis.ticks = element_blank())

## --- MIDDLE: pre-vaccination signature (NEW) — pre-vaccination features only ---
mid_df <- prep_features(su_base, top_n = params$top_n_mid)
p_mid  <- signature_panel(mid_df, "Pre-vaccination serology signature", show_legend = FALSE)

## --- RIGHT: all-maternal signature (kept) ---
right_keep <- params$curated_right
if (!is.null(right_keep)) right_keep <- right_keep[right_keep %in% su_all$feature]
right_df <- if (length(right_keep))
  prep_features(su_all, keep = right_keep) else
  prep_features(su_all, top_n = params$top_n_right)
p_right  <- signature_panel(right_df, "Maternal serology signature (all visits)",
                            show_legend = TRUE)
```

```{r figure4, fig.width=15.5, fig.height=5.0, fig.cap="Figure 4 (revised). Determinants of maternal responder status. Left: AUC ladder (arm-only rank statistic; +pre-vaccination and +all-maternal serology are cross-validated elastic-net AUCs). Middle: top-12 pre-vaccination (early-pregnancy) serology features by univariate AUC. Right: top-12 all-maternal serology features by univariate AUC; hatched/outlined bars are the pre-vaccination draw. Bars coloured by antigen family (poliovirus serotypes IPV1/IPV2/IPV3 collapsed to IPV); the three score-defining features (PT/FHA/PRN total IgG) are excluded from both signatures."}
if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  fig4 <- p_left + p_mid + p_right + plot_layout(widths = c(1.05, 1.05, 1.25))
} else {
  stop("Please install `patchwork` to assemble the three panels: install.packages('patchwork').")
}
fig4
```

```{r save-figure}
dir.create(params$fig_out_dir, showWarnings = FALSE, recursive = TRUE)
png_path <- file.path(params$fig_out_dir, "figure4_three_panel.png")
pdf_path <- file.path(params$fig_out_dir, "figure4_three_panel.pdf")
ggsave(png_path, fig4, width = 15.5, height = 5.0, dpi = 300, bg = "white")
ggsave(pdf_path, fig4, width = 15.5, height = 5.0, bg = "white")
cat("Saved:\n- ", png_path, "\n- ", pdf_path, "\n", sep = "")
```

---

*Revised Figure 4. Both signature panels now show the top 12 features by
univariate AUC. Poliovirus serotypes (IPV1/IPV2/IPV3) share one IPV colour, and
FIM/TT have their own hues, so every antigen family is legible; the boxed legend
in the right panel carries all seven families plus the pre-vaccination hatch. The
middle panel is computed on the pre-vaccination (`r params$baseline_visit`) matrix
`W_base`; the right panel on the all-maternal matrix `W_all`. To pin an exact
curated set for the right panel instead of the automatic top 12, set
`params$curated_right` to a vector of feature@visit strings.*
