---
title: "GaPs infant post-priming serum IgG vs de Graaf CHIM 'resistant' adults"
subtitle: "Is post-vaccination serum PT IgG (and PRN/FHA/FIM) in GaPs infants above the tentative CHIM correlate-of-protection concentrations? Anchored to de Graaf et al. Lancet Microbe 2026, Table 1."
author: "Prepared for co-author review"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
    code_folding: hide
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
options(scipen = 999)
```

```{r libs, include=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
```

# What this document does

The de Graaf *et al.* controlled human infection (CHIM) study (Lancet Microbe 2026)
identified a **sub-group of adults who resisted colonisation** after intranasal *B pertussis*
challenge, and this resistant phenotype was associated with **higher pre-inoculation serum
anti-pertussis-toxin (PT) IgG** (among other markers). PT IgG is therefore proposed — very
tentatively — as a possible correlate of protection against colonisation.

The question from the co-author: **how do post-primary-vaccination serum PT IgG
concentrations in our GaPs infants (especially the blunted wP / maternal-Tdap infants)
compare with the CHIM adult "resistant" concentration?** If the infant concentrations sit
well *above* the resistant-adult anchor, that is a substantive point.

The comparison rests on the units being directly comparable. The GaPs total-IgG assay is the
**RIVM (Bilthoven) multiplex/ELISA in IU/mL**, and the de Graaf CHIM PT/PRN/FHA IgG values are
reported in the **same RIVM IU/mL scale** (de Graaf Table 1 footnote; fimbriae 2/3 are in
AU/mL in *both* studies). So PT, PRN and FHA are compared IU-for-IU; FIM is compared AU-for-AU.
This is the assumption that makes the numeric comparison legitimate and it is stated explicitly
in the output so a reader can judge it.

---

# CHIM anchors (from de Graaf Table 1, NOT the digitized Figure 3)

These are the **published geometric means (95% CI)** for pre-inoculation serum IgG, taken
directly from Table 1 of de Graaf *et al.* 2026. They are hard-coded here so the comparison is
anchored to the paper's own reported statistics rather than to points digitized off Figure 3.

```{r chim-anchors}
# --- de Graaf 2026, Table 1: pre-inoculation serum IgG, geometric mean (95% CI) ---
#   PT, PRN, FHA reported in IU/mL ; Fimbriae 2/3 (FIM) reported in AU/mL.
#   colonised   = became colonised after primary challenge (n = 20)  -> "susceptible"
#   resistant   = non-colonised after primary challenge   (n = 30)  -> "protected"
chim <- tibble::tribble(
  ~antigen, ~unit,   ~n_col, ~gm_col, ~col_lo, ~col_hi, ~p_table1, ~n_res, ~gm_res, ~res_lo, ~res_hi,
  "PT",     "IU/mL",  20,     1.9,     0.9,     4.0,     0.028,      30,     6.2,     3.0,    12.8,
  "PRN",    "IU/mL",  20,     7.3,     2.4,    22.3,     0.037,      30,    17.6,     8.1,    38.2,
  "FHA",    "IU/mL",  20,     8.5,     2.5,    29.2,     0.010,      30,    30.1,    13.5,    67.3,
  "FIM",    "AU/mL",  20,    28.2,     7.4,   108.0,     0.97,       30,    25.4,     8.6,    75.5
)

chim_disp <- chim %>%
  transmute(
    Antigen = antigen,
    Unit = unit,
    `CHIM colonised (susceptible) GM (95% CI)` =
      sprintf("%.1f (%.1f\u2013%.1f)", gm_col, col_lo, col_hi),
    `CHIM resistant (protected) GM (95% CI)` =
      sprintf("%.1f (%.1f\u2013%.1f)", gm_res, res_lo, res_hi),
    `p (colonised vs resistant, Table 1)` = p_table1)

kable(chim_disp,
      caption = "CHIM adult pre-inoculation serum IgG anchors (de Graaf 2026, Table 1). PT/PRN/FHA in IU/mL; FIM in AU/mL. p-values are the Table 1 colonised-vs-non-colonised comparisons (Mann-Whitney; FIM n.s.).")
```

The single most relevant anchors for the co-author's argument are the **PT IgG** values:
**resistant adults GM = 6.2 IU/mL (3.0\u201312.8)** vs **colonised adults GM = 1.9 IU/mL
(0.9\u20134.0)**. The two thresholds carried through every infant comparison below are the
**resistant GM (6.2)** and the **colonised GM (1.9)**.

---

# Load GaPs infant data (`d_set` / `ebaa_extra2`)

```{r load-dset}
# ---------------------------------------------------------------------------
# d_set path: put d_set.RData beside this Rmd (or edit the path below).
# The object inside is named `ebaa_extra2` (per the project pipeline). We load
# it defensively so this file also works if the object name changes.
# ---------------------------------------------------------------------------
DSET_PATH <- if (file.exists("d_set.RData")) "d_set.RData" else
             if (file.exists("data/d_set.RData")) "data/d_set.RData" else "d_set.RData"

stopifnot(file.exists(DSET_PATH))
.e <- new.env()
.loaded <- load(DSET_PATH, envir = .e)

# Prefer the known object name; otherwise take the first data.frame loaded.
if ("ebaa_extra2" %in% .loaded) {
  d_set <- get("ebaa_extra2", envir = .e)
} else {
  .dfs <- .loaded[vapply(.loaded, function(o) is.data.frame(get(o, envir = .e)), logical(1))]
  stopifnot(length(.dfs) >= 1)
  d_set <- get(.dfs[1], envir = .e)
}
d_set <- as.data.frame(d_set)
cat(sprintf("Loaded object(s): %s\nUsing a data frame with %d rows and %d columns.\n",
            paste(.loaded, collapse = ", "), nrow(d_set), ncol(d_set)))
```

### Structure discovery (so column/label choices are auditable)

I cannot assume the exact spelling of every column, so the chunk below **prints what is
actually in `d_set`** — the columns, the visit labels, the antigen and analyte levels, and the
infant/maternal group labels. Skim this once; if any of the auto-detected choices in the next
section are wrong, they are trivially editable there.

```{r discover, results='asis'}
show_levels <- function(df, col, max_n = 40) {
  if (!col %in% names(df)) return(sprintf("- **%s**: (absent)\n", col))
  u <- sort(unique(as.character(df[[col]])))
  sprintf("- **%s** (%d levels): %s\n", col, length(u),
          paste(utils::head(u, max_n), collapse = ", "))
}
cat("**Columns present:**\n\n")
cat(paste0("`", names(d_set), "`", collapse = ", "), "\n\n")

cat("**Key level sets:**\n\n")
for (cc in c("visit_name", "visit", "timepoint",
             "antigen", "analyte", "feature", "isotype", "subclass",
             "infant_arm", "arm_name", "maternal_arm", "mat_arm",
             "group", "study_group", "arm"))
  cat(show_levels(d_set, cc))
```

---

# Focal selection: 5-month post-priming visit, total IgG, PT/PRN/FHA/FIM

```{r focal-config}
# ---- EDIT HERE if the discovery output above shows different spellings -----

# 1) Visit column and the 5-month (InfMon5) post-priming label.
VISIT_COL   <- if ("visit_name" %in% names(d_set)) "visit_name" else
               if ("visit" %in% names(d_set)) "visit" else "timepoint"
FOCAL_VISIT <- "InfMon5"   # 5 months after the infant priming series

# 2) Antigen column + the four antigens of interest.
ANTIGEN_COL <- if ("antigen" %in% names(d_set)) "antigen" else "antigen"
ANTIGENS    <- c("PT", "PRN", "FHA", "FIM")

# 3) Analyte column + the value that denotes TOTAL IgG (not a subclass).
ANALYTE_COL   <- if ("analyte" %in% names(d_set)) "analyte" else "analyte"
TOTAL_IGG_VAL <- "IgG"     # total IgG; subclasses are IgG1..IgG4 in this dataset

# 4) The native concentration column (IU/mL or AU/mL), NOT the log10 column.
#    In this pipeline the native value is `value_reported` and the log10 is
#    `log_assay_value`. We use the native value and compute GM ourselves.
CONC_COL <- if ("value_reported" %in% names(d_set)) "value_reported" else
            if ("concentration"  %in% names(d_set)) "concentration"  else
            if ("value"          %in% names(d_set)) "value"          else NA_character_
stopifnot(!is.na(CONC_COL))

# 5) Infant priming arm (wP vs aP) and maternal arm (Tdap vs TT).
INFANT_ARM_COL <- if ("infant_arm" %in% names(d_set)) "infant_arm" else NA_character_

# Maternal arm: try common names; else try to derive from arm_name.
MAT_ARM_COL <- {
  cand <- c("maternal_arm", "mat_arm", "maternal_vaccine", "mat_vaccine")
  hit  <- cand[cand %in% names(d_set)]
  if (length(hit)) hit[1] else NA_character_
}

cat(sprintf("Using: visit='%s' (focal='%s'), antigen='%s', analyte='%s' (total='%s'), conc='%s', infant_arm='%s', maternal_arm='%s'\n",
            VISIT_COL, FOCAL_VISIT, ANTIGEN_COL, ANALYTE_COL, TOTAL_IGG_VAL, CONC_COL,
            ifelse(is.na(INFANT_ARM_COL), "(none)", INFANT_ARM_COL),
            ifelse(is.na(MAT_ARM_COL), "(derive/none)", MAT_ARM_COL)))
```

```{r derive-maternal, results='asis'}
# If there is no explicit maternal-arm column, try to derive one from arm_name.
# GaPs maternal randomisation: Tdap-IPV vs tetanus-toxoid (TT). We look for a
# text signature in arm_name; anything matching Tdap/aP-IPV -> "Tdap", TT -> "TT".
d <- d_set
if (is.na(MAT_ARM_COL) && "arm_name" %in% names(d)) {
  an <- tolower(as.character(d$arm_name))
  mat <- ifelse(grepl("tdap|t-dap|ipv|pertussis", an) & !grepl("^tt|tetanus[- ]?toxoid", an), "Tdap",
         ifelse(grepl("\\btt\\b|tetanus", an), "TT", NA_character_))
  if (any(!is.na(mat))) {
    d$maternal_arm_derived <- mat
    MAT_ARM_COL <- "maternal_arm_derived"
    cat("> Derived a maternal arm from `arm_name` by text match. **Verify** the mapping below ",
        "against the discovery output; edit the `grepl()` rules if it is wrong.\n\n")
    print(kable(as.data.frame(table(arm_name = d$arm_name, maternal = d$maternal_arm_derived)),
                caption = "Derived maternal-arm mapping (check me)"))
    cat("\n\n")
  }
}
if (is.na(MAT_ARM_COL)) {
  cat("> **No maternal-arm column found or derivable.** The maternal-split rows in the ",
      "comparison table will be skipped; the infant wP/aP and all-infant rows will still be ",
      "produced. Add/point `MAT_ARM_COL` to the correct column to enable the maternal split.\n\n")
}
```

```{r build-focal}
# Coerce helper columns to character for robust filtering.
d$.visit   <- as.character(d[[VISIT_COL]])
d$.antigen <- as.character(d[[ANTIGEN_COL]])
d$.analyte <- as.character(d[[ANALYTE_COL]])
d$.conc    <- suppressWarnings(as.numeric(d[[CONC_COL]]))
d$.infarm  <- if (!is.na(INFANT_ARM_COL)) as.character(d[[INFANT_ARM_COL]]) else NA_character_
d$.matarm  <- if (!is.na(MAT_ARM_COL))    as.character(d[[MAT_ARM_COL]])    else NA_character_

focal <- d %>%
  filter(.visit == FOCAL_VISIT,
         .analyte == TOTAL_IGG_VAL,
         .antigen %in% ANTIGENS,
         is.finite(.conc), .conc > 0)

# One value per subject x antigen (guard against duplicate assay rows).
subj_col <- if ("subject_accession" %in% names(focal)) "subject_accession" else
            if ("subject_id" %in% names(focal)) "subject_id" else names(focal)[1]
focal <- focal %>%
  distinct(.data[[subj_col]], .antigen, .infarm, .matarm, .keep_all = TRUE)

cat(sprintf("Focal rows (visit=%s, total IgG, 4 antigens): %d across %d subjects.\n",
            FOCAL_VISIT, nrow(focal), dplyr::n_distinct(focal[[subj_col]])))
```

```{r sanity-focal, results='asis'}
if (nrow(focal) == 0) {
  cat("> **No rows selected.** The focal visit/analyte/antigen labels did not match the data. ",
      "Re-check the discovery output above and edit the `focal-config` chunk. Common fixes: the ",
      "5-month visit may be spelled differently (e.g. 'InfMon5', 'M05', 'V05', '5m'), or total ",
      "IgG may be coded as 'IgG_total' rather than 'IgG'.\n\n")
} else {
  cat("Per-antigen row counts at the focal visit:\n\n")
  print(kable(as.data.frame(table(antigen = focal$.antigen)), caption = "n per antigen (focal visit, total IgG)"))
}
```

---

# Comparison engine

```{r gm-helpers}
gm    <- function(x) { x <- x[is.finite(x) & x > 0]; if (!length(x)) NA_real_ else exp(mean(log(x))) }
gm_ci <- function(x, conf = 0.95) {
  x <- x[is.finite(x) & x > 0]; n <- length(x)
  if (n < 2) return(c(gm = if (n) x else NA_real_, lo = NA_real_, hi = NA_real_))
  lx <- log(x); m <- mean(lx); se <- stats::sd(lx) / sqrt(n)
  tt <- stats::qt(1 - (1 - conf) / 2, df = n - 1)
  c(gm = exp(m), lo = exp(m - tt * se), hi = exp(m + tt * se))
}

# Summarise one infant subgroup vs the CHIM anchors for one antigen.
summarise_group <- function(vals, antigen_code, group_label) {
  vals <- vals[is.finite(vals) & vals > 0]
  a <- chim[chim$antigen == antigen_code, ]
  ci <- gm_ci(vals)
  gm_res <- a$gm_res; gm_col <- a$gm_col
  tibble::tibble(
    Antigen      = antigen_code,
    Group        = group_label,
    n            = length(vals),
    GM           = round(unname(ci["gm"]), 1),
    GM_CI        = if (is.na(ci["lo"])) NA_character_ else sprintf("%.1f\u2013%.1f", ci["lo"], ci["hi"]),
    Median       = round(stats::median(vals), 1),
    `xGM_vs_resistant` = if (is.na(gm_res) || is.na(ci["gm"])) NA_real_ else round(unname(ci["gm"]) / gm_res, 1),
    `%>_resistantGM`   = round(100 * mean(vals > gm_res), 0),
    `%>_colonisedGM`   = round(100 * mean(vals > gm_col), 0)
  )
}

# Assemble the infant subgroups that exist in the data.
build_subgroups <- function(fa, antigen_code) {
  s <- fa %>% filter(.antigen == antigen_code)
  out <- list()
  add <- function(v, lbl) if (length(v[is.finite(v) & v > 0]) >= 1)
    out[[length(out) + 1]] <<- summarise_group(v, antigen_code, lbl)

  # All infants
  add(s$.conc, "All GaPs infants (5 mo)")

  # Infant priming arm
  if (any(!is.na(s$.infarm))) {
    add(s$.conc[s$.infarm == "wP"], "Infant wP-primed (5 mo)")
    add(s$.conc[s$.infarm == "aP"], "Infant aP-primed (5 mo)")
  }

  # Maternal arm (the blunting axis)
  if (any(!is.na(s$.matarm))) {
    add(s$.conc[s$.matarm == "Tdap"], "Maternal-Tdap infants (blunted)")
    add(s$.conc[s$.matarm == "TT"],   "Maternal-TT infants (unblunted)")
    # Fully-crossed cells (the co-author's 'most blunted' = maternal-Tdap x wP)
    if (any(!is.na(s$.infarm))) {
      add(s$.conc[s$.matarm == "Tdap" & s$.infarm == "wP"], "Maternal-Tdap x wP (most blunted)")
      add(s$.conc[s$.matarm == "Tdap" & s$.infarm == "aP"], "Maternal-Tdap x aP")
      add(s$.conc[s$.matarm == "TT"   & s$.infarm == "wP"], "Maternal-TT x wP")
      add(s$.conc[s$.matarm == "TT"   & s$.infarm == "aP"], "Maternal-TT x aP")
    }
  }
  dplyr::bind_rows(out)
}
```

# Results table — PT IgG (the tentative correlate)

```{r pt-table, results='asis'}
if (nrow(focal) > 0 && "PT" %in% focal$.antigen) {
  a <- chim[chim$antigen == "PT", ]
  chim_rows <- tibble::tibble(
    Antigen = "PT",
    Group   = c("CHIM adult colonised (susceptible)", "CHIM adult resistant (protected)"),
    n       = c(a$n_col, a$n_res),
    GM      = c(a$gm_col, a$gm_res),
    GM_CI   = c(sprintf("%.1f\u2013%.1f", a$col_lo, a$col_hi),
                sprintf("%.1f\u2013%.1f", a$res_lo, a$res_hi)),
    Median  = c(NA_real_, NA_real_),
    `xGM_vs_resistant` = c(round(a$gm_col / a$gm_res, 1), 1),
    `%>_resistantGM`   = c(NA_real_, NA_real_),
    `%>_colonisedGM`   = c(NA_real_, NA_real_)
  )
  pt_tbl <- dplyr::bind_rows(chim_rows, build_subgroups(focal, "PT"))
  kable(pt_tbl, caption = "PT IgG (IU/mL). CHIM anchors from de Graaf Table 1; GaPs infant rows computed from d_set at the 5-month post-priming visit (total IgG). 'xGM_vs_resistant' = infant GM / 6.2. '%>_resistantGM' = share of infants above 6.2 IU/mL; '%>_colonisedGM' = share above 1.9 IU/mL.")
} else {
  cat("> PT rows unavailable (see selection diagnostics above).\n")
}
```

# Results tables — PRN, FHA, FIM IgG

```{r other-tables, results='asis'}
for (ag in c("PRN", "FHA", "FIM")) {
  if (nrow(focal) == 0 || !(ag %in% focal$.antigen)) {
    cat(sprintf("\n> %s rows unavailable.\n\n", ag)); next
  }
  a <- chim[chim$antigen == ag, ]
  unit <- a$unit
  chim_rows <- tibble::tibble(
    Antigen = ag,
    Group   = c("CHIM adult colonised (susceptible)", "CHIM adult resistant (protected)"),
    n       = c(a$n_col, a$n_res),
    GM      = c(a$gm_col, a$gm_res),
    GM_CI   = c(sprintf("%.1f\u2013%.1f", a$col_lo, a$col_hi),
                sprintf("%.1f\u2013%.1f", a$res_lo, a$res_hi)),
    Median  = c(NA_real_, NA_real_),
    `xGM_vs_resistant` = c(round(a$gm_col / a$gm_res, 1), 1),
    `%>_resistantGM`   = c(NA_real_, NA_real_),
    `%>_colonisedGM`   = c(NA_real_, NA_real_)
  )
  tbl <- dplyr::bind_rows(chim_rows, build_subgroups(focal, ag))
  print(kable(tbl, caption = sprintf(
    "%s IgG (%s). CHIM anchors from de Graaf Table 1. For %s the Table 1 colonised-vs-resistant difference was %s.",
    ag, unit, ag,
    if (ag == "FIM") "NOT significant (p=0.97) \u2014 so the resistant 'threshold' is not meaningful for FIM" else
      sprintf("significant (p=%s)", format(a$p_table1)))))
  cat("\n\n")
}
```

---

# Auto-generated summary sentences (for pasting into the LLM / manuscript)

```{r narrative, results='asis'}
if (nrow(focal) > 0 && "PT" %in% focal$.antigen) {
  pt <- build_subgroups(focal, "PT")
  res_gm <- chim$gm_res[chim$antigen == "PT"]   # 6.2
  col_gm <- chim$gm_col[chim$antigen == "PT"]   # 1.9
  say <- function(row) {
    sprintf(paste0("- **%s** (n=%d): GM = %.1f IU/mL%s, median = %.1f. That is **%.1f\u00d7** the ",
                   "resistant-adult PT IgG GM (6.2), with **%d%%** of infants above the resistant ",
                   "GM and **%d%%** above the colonised GM (1.9)."),
            row$Group, row$n, row$GM,
            ifelse(is.na(row$GM_CI), "", sprintf(" (95%% CI %s)", row$GM_CI)),
            row$Median, row$xGM_vs_resistant, row$`%>_resistantGM`, row$`%>_colonisedGM`)
  }
  cat("PT IgG, GaPs infants at 5 months vs the CHIM resistant-adult anchor (6.2 IU/mL):\n\n")
  for (i in seq_len(nrow(pt))) cat(say(pt[i, ]), "\n")
  cat("\n\n")

  # headline for the most-blunted cell if present
  mb <- pt[pt$Group == "Maternal-Tdap x wP (most blunted)", ]
  if (nrow(mb) == 1 && is.finite(mb$GM)) {
    verdict <- if (mb$GM > res_gm)
      sprintf("Even the **most blunted** GaPs cell (maternal-Tdap x wP) sits **above** the resistant-adult PT IgG GM (%.1f vs 6.2 IU/mL, %.1f\u00d7), with %d%% of those infants above 6.2 IU/mL.",
              mb$GM, mb$xGM_vs_resistant, mb$`%>_resistantGM`)
    else
      sprintf("The most blunted GaPs cell (maternal-Tdap x wP) has GM %.1f IU/mL, **below** the resistant-adult anchor (6.2); %d%% of those infants nonetheless exceed 6.2 IU/mL.",
              mb$GM, mb$`%>_resistantGM`)
    cat("**Headline:** ", verdict, "\n\n")
  }
} else {
  cat("> Narrative skipped: PT selection produced no rows (fix the focal-config chunk).\n")
}
```

# Caveats to keep in the narrative

- The CHIM correlate is **tentative and directional**: de Graaf show *higher* pre-inoculation
  PT IgG in resistant adults, not a validated protective threshold. 6.2 IU/mL is a group
  geometric mean, not a cut-off, and the confidence intervals for colonised (0.9\u20134.0) and
  resistant (3.0\u201312.8) adults **overlap**. Read "% above 6.2" as a descriptive framing device,
  not a protection probability.
- Whether the CHIM signal reflects **vaccine- or infection-induced** immunity is unknown; the
  adults were whole-cell-primed in infancy and were *not* selected on PT IgG, so some of the
  resistant-adult titre may reflect silent natural boosting.
- **Age / kinetics mismatch:** GaPs values are peak post-priming infant titres; the CHIM adults
  are decades removed from priming. High infant titres now do not imply durable protection at
  the ages when CHIM-type exposure occurs.
- **Units:** PT/PRN/FHA compared in RIVM IU/mL (directly comparable per the assay provenance you
  confirmed); FIM in AU/mL. The Table 1 FIM difference was not significant (p=0.97), so the FIM
  "resistant threshold" is not interpretable as a correlate.
- All GaPs numbers above are **computed live from `d_set`**; if the discovery output shows a
  different visit/analyte/group spelling, edit the single `focal-config` chunk and re-knit.

*Anchors: de Graaf et al., Lancet Microbe 2026, Table 1 (pre-inoculation serum IgG geometric means).*
