---
title: "Standards Curve Comparison: inf_baa vs mat_baa"
subtitle: "Standard-Curve Back-Interpolation for Cross-Set MFI Normalization"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    highlight: tango
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center"
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(knitr)
  library(kableExtra)
  library(DT)
  library(scales)
  library(htmltools)
})

load(file = here::here("./data/standards_sel.RData"))

# ── Palettes ──────────────────────────────────────────────────────────────────
plate_levels <- sort(unique(standards_sel$plate))
n_plates     <- length(plate_levels)

pal_base <- c(
  "#1F78B4","#33A02C","#E31A1C","#FF7F00","#6A3D9A",
  "#B15928","#A6CEE3","#B2DF8A","#FB9A99","#FDBF6F",
  "#CAB2D6","#FFFF99","#1B9E77","#D95F02","#7570B3"
)
plate_colors  <- setNames(pal_base[seq_along(plate_levels)], plate_levels)
source_shapes <- c(NIBSC06_140 = 16, SD = 17)
source_lty    <- c(NIBSC06_140 = "solid", SD = "dashed")
set_colors    <- c(inf_baa = "#2171B5", mat_baa = "#CB181D")

# ── Core normalization helpers ─────────────────────────────────────────────────

# Fit a single log-log linear standard curve; returns named list of stats
fit_log_lm <- function(log10_mfi, log10_conc) {
  tryCatch({
    m <- lm(log10_mfi ~ log10_conc)
    list(
      intercept = unname(coef(m)[1]),
      slope     = unname(coef(m)[2]),
      r_squared = summary(m)$r.squared,
      adj_r2    = summary(m)$adj.r.squared,
      rmse      = sqrt(mean(residuals(m)^2))
    )
  }, error = function(e) {
    list(intercept = NA_real_, slope = NA_real_, r_squared = NA_real_,
         adj_r2 = NA_real_, rmse = NA_real_)
  })
}

# Normalize raw MFI from one set onto a target set's scale.
# All arithmetic is in log10 space; the result is back-transformed.
#   a_src, b_src : source-set curve params  (log10 MFI = a + b*log10 conc)
#   a_tgt, b_tgt : target-set curve params
normalize_mfi <- function(raw_mfi, a_src, b_src, a_tgt, b_tgt) {
  log10_raw  <- log10(pmax(raw_mfi, 0.5))      # guard against 0
  log10_conc <- (log10_raw - a_src) / b_src    # invert source curve -> conc
  10^(a_tgt + b_tgt * log10_conc)              # forward-predict on target
}
```

---

## Approach

The two sets (`inf_baa`, `mat_baa`) run their standard curves at **different concentration
values**, so direct point-by-point ratio comparison is impossible. Instead this report uses
**standard-curve back-interpolation**:

1. Fit `log10(MFI) ~ log10(concentration)` **independently** for each set within every
   (feature × antigen × plate × source) group.
2. To normalise a measurement from set **A** onto set **B**'s scale:
   - *Invert* set A's curve to recover `log10(concentration)` from the observed MFI.
   - *Forward-predict* what set B's curve would give at that concentration.
3. In log10 space this is a linear transform:
   `log10(norm_MFI) = cross_intercept + cross_slope × log10(raw_MFI)`.
   A `cross_slope ≠ 1` means the correction is **concentration-dependent**.

The exported `normalization_params.csv` can be joined to any sample data frame and the
two-line formula applied to produce normalized MFI values comparable across sets.

---

## 1  Data Overview {.tabset}

### Feature × Antigen × Set

```{r overview-structure}

set_antigen_tbl <- standards_sel %>%
  group_by(set, feature, antigen) %>%
  dplyr::summarise(
    n_obs     = n(),
    n_plates  = n_distinct(plate),
    n_sources = n_distinct(source),
    n_conc    = n_distinct(concentration),
    .groups   = "drop"
  )

overlap_tbl <- set_antigen_tbl %>%
  pivot_wider(
    id_cols     = c(feature, antigen),
    names_from  = set,
    values_from = c(n_obs, n_plates, n_sources, n_conc)
  )

datatable(overlap_tbl,
          caption  = "Rows per feature x antigen by set (NA = absent)",
          filter   = "top", rownames = FALSE,
          options  = list(scrollX = TRUE, pageLength = 15))
```

### Antigen Name Mapping

*`mat_baa` antigens carry a batch-code suffix (e.g. `act_42`).  The base name
(column `antigen_base`) is used as the cross-set join key throughout this report.*

```{r overview-antigen-map}
antigen_map <- standards_sel %>%
  mutate(antigen_base = str_remove(antigen, "_\\d+$")) %>%
  distinct(set, antigen, antigen_base) %>%
  pivot_wider(names_from = set, values_from = antigen,
              values_fn = ~ paste(unique(.x), collapse = " / ")) %>%
  arrange(antigen_base)

datatable(antigen_map,
          caption  = "Antigen naming across sets (antigen_base = join key)",
          rownames = FALSE,
          options  = list(pageLength = 25, dom = "t"))
```

### Concentration Ranges by Set

```{r overview-concentrations}
conc_tbl <- standards_sel %>%
  group_by(set, feature, antigen) %>%
  dplyr::summarise(
    n_conc   = n_distinct(concentration),
    conc_min = min(concentration),
    conc_max = max(concentration),
    dil_min  = min(dilution),
    dil_max  = max(dilution),
    .groups  = "drop"
  ) %>%
  arrange(feature, antigen, set)

datatable(conc_tbl, caption = "Concentration range per feature x antigen x set",
          filter = "top", rownames = FALSE,
          options = list(scrollX = TRUE, pageLength = 15))
```

### Plate x Source

```{r overview-plate-source}
plate_src_tbl <- standards_sel %>%
  filter(!is.na(mfi)) %>%
  group_by(set, plate, source) %>%
  dplyr::summarise(
    n_obs      = n(),
    median_mfi = round(median(mfi), 1),
    n_features = n_distinct(feature),
    n_antigens = n_distinct(antigen),
    .groups    = "drop"
  )

datatable(plate_src_tbl, caption = "Observations per set x plate x source",
          filter = "top", rownames = FALSE,
          options = list(scrollX = TRUE, pageLength = 20))
```

---

## 2  Aggregate MFI

```{r aggregate}
# Median MFI over any replicate beads / wells within each biological combination
stds_agg <- standards_sel %>%
  filter(!is.na(mfi)) %>%
  group_by(set, feature, antigen, plate, source, concentration, dilution) %>%
  dplyr::summarise(
    median_mfi = median(mfi),
    mean_mfi   = mean(mfi),
    sd_mfi     = sd(mfi),
    cv_pct     = sd(mfi) / mean(mfi) * 100,
    n_reps     = n(),
    .groups    = "drop"
  ) %>%
  mutate(
    log10_conc   = log10(concentration),
    log10_mfi    = log10(pmax(median_mfi, 0.5)),
    # Strip trailing _digits suffix (e.g. "act_42" -> "act") so inf_baa and
    # mat_baa antigens can be matched even though their names differ by batch code
    antigen_base = str_remove(antigen, "_\\d+$")
  )

cat(sprintf("Aggregated rows : %d\n", nrow(stds_agg)))
cat(sprintf("Sets            : %s\n", paste(unique(stds_agg$set), collapse = ", ")))
cat(sprintf("Features        : %d\n", n_distinct(stds_agg$feature)))
cat(sprintf("Antigens        : %d\n", n_distinct(stds_agg$antigen)))
cat(sprintf("Plates          : %d\n", n_distinct(stds_agg$plate)))
```

---

## 3  Standard Curve Fits

A `log10(MFI) ~ log10(concentration)` model is fitted **separately** for each
(set x feature x antigen x plate x source) group. Requires at least 3 distinct
concentration points with non-zero variance.

```{r fit-curves}
curve_fits <- stds_agg %>%
  group_by(set, feature, antigen, plate, source) %>%
  filter(n() >= 3, sd(log10_conc) > 0) %>%
  nest() %>%
  mutate(
    n_pts      = map_int(data,  nrow),
    conc_min   = map_dbl(data,  ~ min(.x$concentration)),
    conc_max   = map_dbl(data,  ~ max(.x$concentration)),
    mfi_min    = map_dbl(data,  ~ min(.x$median_mfi)),
    mfi_max    = map_dbl(data,  ~ max(.x$median_mfi)),
    stats      = map(data, ~ fit_log_lm(.x$log10_mfi, .x$log10_conc)),
    intercept  = map_dbl(stats, "intercept"),
    slope      = map_dbl(stats, "slope"),
    r_squared  = map_dbl(stats, "r_squared"),
    adj_r2     = map_dbl(stats, "adj_r2"),
    rmse       = map_dbl(stats, "rmse")
  ) %>%
  select(-data, -stats) %>%
  ungroup() %>%
  # Derive antigen_base here too so it is available for the cross-set pivot
  mutate(antigen_base = str_remove(antigen, "_\\d+$"))

cat(sprintf("Curve fits total  : %d\n", nrow(curve_fits)))
cat(sprintf("  R2 >= 0.99      : %d  (%.0f%%)\n",
            sum(curve_fits$r_squared >= 0.99, na.rm = TRUE),
            100 * mean(curve_fits$r_squared >= 0.99, na.rm = TRUE)))
cat(sprintf("  R2 < 0.95       : %d  (%.0f%%)\n",
            sum(curve_fits$r_squared < 0.95, na.rm = TRUE),
            100 * mean(curve_fits$r_squared < 0.95, na.rm = TRUE)))
```

### 3.1  Fit parameters table

```{r fit-quality-table}
datatable(
  curve_fits %>%
    mutate(across(c(intercept, slope, r_squared, adj_r2, rmse), ~ round(.x, 4))) %>%
    arrange(set, feature, antigen, source, plate),
  caption  = "Standard curve fit parameters per set x feature x antigen x plate x source",
  filter   = "top", rownames = FALSE,
  options  = list(scrollX = TRUE, pageLength = 20)
) %>%
  formatStyle("r_squared",
    background         = styleColorBar(c(0, 1), "lightgreen"),
    backgroundSize     = "100% 70%",
    backgroundRepeat   = "no-repeat",
    backgroundPosition = "center"
  ) %>%
  formatStyle("rmse",
    backgroundColor = styleInterval(c(0.05, 0.15),
                                    c("white", "#FFF3CD", "#FFCCCC"))
  )
```

### 3.2  Curves: observed data + fitted lines {.tabset}

*Points = observed median MFI.  Lines = fitted log-log linear model.
Both sets overlaid per antigen.  Colour = plate, linetype = source.*

```{r std-curve-plot-fn}
plot_std_curves_fitted <- function(feat) {
  obs  <- stds_agg   %>% filter(feature == feat)
  fits <- curve_fits %>% filter(feature == feat, !is.na(intercept))

  if (nrow(fits) == 0) return(NULL)

  pred_lines <- fits %>%
    rowwise() %>%
    mutate(log10_conc_seq = list(seq(log10(conc_min), log10(conc_max),
                                     length.out = 60))) %>%
    ungroup() %>%
    unnest(log10_conc_seq) %>%
    mutate(log10_mfi_fit = intercept + slope * log10_conc_seq)

  ggplot() +
    geom_point(data = obs,
               aes(x = log10_conc, y = log10_mfi,
                   colour = plate, shape = set),
               size = 1.8, alpha = 0.6) +
    geom_line(data = pred_lines,
              aes(x = log10_conc_seq, y = log10_mfi_fit,
                  colour = plate, linetype = source,
                  group  = interaction(set, plate, source)),
              linewidth = 0.55, alpha = 0.85) +
    scale_colour_manual(values = plate_colors) +
    scale_linetype_manual(values = source_lty) +
    scale_shape_manual(values = c(inf_baa = 16, mat_baa = 17)) +
    facet_wrap(~ antigen_base, scales = "free_y", ncol = 4) +
    labs(title    = paste("Standard curves (both sets) --", feat),
         x        = "log10(Concentration)",
         y        = "log10(Median MFI)",
         colour   = "Plate", linetype = "Source", shape = "Set") +
    theme_bw(base_size = 9) +
    theme(legend.position = "bottom", legend.box = "vertical",
          strip.text = element_text(face = "bold", size = 8))
}
```

```{r std-curves-fitted, results='asis', fig.width=16, fig.height=11}
for (feat in sort(unique(stds_agg$feature))) {
  cat(sprintf("\n### %s\n\n", feat))
  p <- plot_std_curves_fitted(feat)
  if (!is.null(p)) print(p) else cat("*No data.*\n\n")
  cat("\n\n")
}
```

---

## 4  Cross-Set Normalization Parameters

For each (feature x antigen x plate x source) group, the fitted coefficients from both
sets are paired.  The **cross-curve linear parameters** fully define the normalization:

```
log10(norm_MFI) = cross_intercept + cross_slope x log10(raw_MFI)
```

- `cross_slope = b_target / b_source`  -- deviation from 1 = concentration-dependent correction
- `cross_intercept = a_target - a_source * cross_slope`  -- the overall log-scale offset

Both directions (inf->mat and mat->inf) are pre-computed for convenience.

```{r cross-params}
norm_params <- curve_fits %>%
  select(set, feature, antigen_base, plate, source,
         intercept, slope, r_squared, rmse,
         n_pts, conc_min, conc_max, mfi_min, mfi_max) %>%
  pivot_wider(
    id_cols     = c(feature, antigen_base, plate, source),
    names_from  = set,
    values_from = c(intercept, slope, r_squared, rmse,
                    n_pts, conc_min, conc_max, mfi_min, mfi_max)
  ) %>%
  filter(!is.na(intercept_inf_baa), !is.na(intercept_mat_baa)) %>%
  mutate(
    # inf_baa -> mat_baa
    cross_slope_inf2mat     = slope_mat_baa / slope_inf_baa,
    cross_intercept_inf2mat = intercept_mat_baa -
                              intercept_inf_baa * cross_slope_inf2mat,
    # mat_baa -> inf_baa
    cross_slope_mat2inf     = slope_inf_baa / slope_mat_baa,
    cross_intercept_mat2inf = intercept_inf_baa -
                              intercept_mat_baa * cross_slope_mat2inf,
    # How far from concentration-independent (slope=1)
    slope_dev_inf2mat = cross_slope_inf2mat - 1,
    slope_dev_mat2inf = cross_slope_mat2inf - 1
  )

cat(sprintf("Groups with both curves fitted: %d\n", nrow(norm_params)))
```

### 4.1  Parameter table

```{r norm-param-table}
datatable(
  norm_params %>%
    select(feature, antigen_base, plate, source,
           cross_slope_mat2inf, cross_intercept_mat2inf,
           slope_dev_inf2mat,
           r_squared_inf_baa, r_squared_mat_baa,
           rmse_inf_baa, rmse_mat_baa) %>%
    mutate(across(where(is.numeric), ~ round(.x, 4))),
  caption  = paste("Cross-curve normalization parameters.",
    "Apply: log10(norm_MFI) = cross_intercept + cross_slope x log10(raw_MFI).",
    "inf2mat = normalize inf_baa onto mat_baa scale; mat2inf = reverse."),
  filter   = "top", rownames = FALSE,
  options  = list(scrollX = TRUE, pageLength = 20)
) %>%
  formatStyle("slope_dev_inf2mat",
    backgroundColor = styleInterval(c(-0.1, 0.1),
                                    c("#FDBB84","white","#FDBB84"))
  )
```

### 4.2  Cross-slope distribution by feature

```{r cross-slope-plot, fig.width=14, fig.height=8}
norm_params_long <- norm_params %>%
  select(feature, antigen_base, plate, source,
         `inf -> mat` = cross_slope_inf2mat,
         `mat -> inf` = cross_slope_mat2inf) %>%
  pivot_longer(c(`inf -> mat`, `mat -> inf`),
               names_to = "direction", values_to = "cross_slope")

ggplot(norm_params_long,
       aes(x = cross_slope, y = antigen_base,
           colour = plate, shape = source)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey40") +
  geom_point(size = 2.5, alpha = 0.85,
             position = position_jitter(height = 0.18, width = 0)) +
  scale_colour_manual(values = plate_colors) +
  scale_shape_manual(values = source_shapes) +
  facet_grid(direction ~ feature, scales = "free_y") +
  labs(
    title    = "Cross-slope by feature, antigen, plate, source",
    subtitle = "1.0 = concentration-independent correction; != 1 = concentration-dependent",
    x        = "Cross-slope  (b_target / b_source)",
    y        = "Antigen",
    colour   = "Plate", shape = "Source"
  ) +
  theme_bw(base_size = 9) +
  theme(legend.position = "bottom", legend.box = "vertical",
        strip.text = element_text(face = "bold"))
```

---

## 5  Scatter Plots: Raw inf_baa vs Predicted mat_baa {.tabset}

For each observed **inf_baa** data point:

1. The inf_baa model is inverted to back-calculate `log10(concentration)` from `log10(observed MFI)`.
2. The mat_baa model is forward-evaluated at that concentration to predict `mat_baa MFI`.

**x = observed inf_baa MFI; y = predicted mat_baa MFI (both log10 scaled).**
Grey dashed 1:1 line = perfect agreement after normalization.
Paths connect points in order of increasing concentration.

```{r scatter-fn}
plot_scatter_feature <- function(feat) {
  params <- norm_params %>% filter(feature == feat)
  if (nrow(params) == 0) return(NULL)

  inf_obs <- stds_agg %>%
    filter(feature == feat, set == "inf_baa") %>%
    left_join(
      params %>% select(feature, antigen_base, plate, source,
                        a_src = intercept_inf_baa, b_src = slope_inf_baa,
                        a_tgt = intercept_mat_baa, b_tgt = slope_mat_baa),
      by = c("feature","antigen_base","plate","source")
    ) %>%
    filter(!is.na(a_src)) %>%
    mutate(
      mfi_x = median_mfi,
      mfi_y = normalize_mfi(median_mfi, a_src, b_src, a_tgt, b_tgt)
    ) %>%
    filter(mfi_x > 0, mfi_y > 0, is.finite(log(mfi_x)), is.finite(log(mfi_y))) %>%
    arrange(antigen_base, plate, source, concentration)

  if (nrow(inf_obs) == 0) return(NULL)

  ggplot(inf_obs,
         aes(x = mfi_x, y = mfi_y, colour = plate, shape = source)) +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", colour = "grey40", linewidth = 0.5) +
    geom_path(aes(group = interaction(plate, source)),
              alpha = 0.35, linewidth = 0.4) +
    geom_point(aes(size = log10_conc), alpha = 0.85) +
    scale_x_log10(labels = label_scientific()) +
    scale_y_log10(labels = label_scientific()) +
    annotation_logticks(sides = "bl", size = 0.2, alpha = 0.4) +
    scale_colour_manual(values = plate_colors) +
    scale_shape_manual(values = source_shapes) +
    scale_size_continuous(name = "log10(Conc)", range = c(1, 4)) +
    facet_wrap(~ antigen_base, scales = "free", ncol = 4) +
    labs(
      title    = paste("Raw inf_baa MFI vs Back-Interpolated mat_baa MFI --", feat),
      subtitle = "Dashed = 1:1 target. Path: low to high concentration.",
      x        = "Observed inf_baa MFI",
      y        = "Predicted mat_baa MFI  (back-interpolated via standard curves)",
      colour   = "Plate", shape = "Source"
    ) +
    theme_bw(base_size = 9) +
    theme(legend.position = "bottom", legend.box = "vertical",
          strip.text = element_text(face = "bold", size = 8))
}
```

```{r scatter-plots, results='asis', fig.width=16, fig.height=11}
for (feat in sort(unique(stds_agg$feature))) {
  cat(sprintf("\n### %s\n\n", feat))
  p <- plot_scatter_feature(feat)
  if (!is.null(p)) print(p) else cat("*No groups with both sets fitted.*\n\n")
  cat("\n\n")
}
```

---

## 6  Normalization Factor Curves {.tabset}

The **concentration-dependent correction factor** is `mfi_mat_predicted / mfi_inf_predicted`
evaluated across the shared concentration range.  Plotted as log2(factor) vs log10(conc):

- **Flat line at 0** = scalar correction (same multiplier at every concentration).
- **Sloped line** = concentration-dependent; large-MFI and small-MFI measurements get different corrections.

```{r norm-curve-fn}
build_norm_curves <- function(feat) {
  params <- norm_params %>%
    filter(feature == feat, !is.na(cross_slope_inf2mat)) %>%
    mutate(
      conc_lo = pmax(conc_min_inf_baa, conc_min_mat_baa),
      conc_hi = pmin(conc_max_inf_baa, conc_max_mat_baa)
    ) %>%
    filter(conc_lo <= conc_hi)

  if (nrow(params) == 0) return(NULL)

  params %>%
    rowwise() %>%
    mutate(log10_conc_seq = list(seq(log10(conc_lo), log10(conc_hi),
                                     length.out = 80))) %>%
    ungroup() %>%
    unnest(log10_conc_seq) %>%
    mutate(
      mfi_inf_pred  = 10^(intercept_inf_baa + slope_inf_baa * log10_conc_seq),
      mfi_mat_pred  = 10^(intercept_mat_baa + slope_mat_baa * log10_conc_seq),
      norm_factor   = mfi_mat_pred / mfi_inf_pred,
      log2_nf       = log2(norm_factor),
      concentration = 10^log10_conc_seq
    )
}

plot_norm_curves <- function(feat) {
  d <- build_norm_curves(feat)
  if (is.null(d)) return(NULL)

  ggplot(d, aes(x = log10_conc_seq, y = log2_nf,
                colour = plate, linetype = source,
                group  = interaction(plate, source))) +
    geom_hline(yintercept = 0, linetype = "dotted", colour = "grey40") +
    geom_line(linewidth = 0.7, alpha = 0.9) +
    scale_colour_manual(values = plate_colors) +
    scale_linetype_manual(values = source_lty) +
    facet_wrap(~ antigen_base, scales = "free_y", ncol = 4) +
    labs(
      title    = paste("Concentration-dependent normalization factor --", feat),
      subtitle = "log2(mat_baa model / inf_baa model) at each concentration",
      x        = "log10(Concentration)",
      y        = "log2(Normalization factor)",
      colour   = "Plate", linetype = "Source"
    ) +
    theme_bw(base_size = 9) +
    theme(legend.position = "bottom", legend.box = "vertical",
          strip.text = element_text(face = "bold", size = 8))
}
```

```{r norm-curves, results='asis', fig.width=16, fig.height=11}
for (feat in sort(unique(norm_params$feature))) {
  cat(sprintf("\n### %s\n\n", feat))
  p <- plot_norm_curves(feat)
  if (!is.null(p)) print(p) else cat("*Insufficient overlapping range.*\n\n")
  cat("\n\n")
}
```

---

## 7  Normalization Heatmaps

### 7.1  Cross-intercept (log-scale offset, inf -> mat)

```{r heatmap-intercept, fig.width=17, fig.height=9}
ggplot(norm_params,
       aes(x = plate, y = antigen_base,
           fill = cross_intercept_inf2mat)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  geom_text(aes(label = round(cross_intercept_inf2mat, 2)),
            size = 2, colour = "black") +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#D6604D",
                       midpoint = 0, name = "cross-\nintercept") +
  facet_grid(source ~ feature, scales = "free", space = "free") +
  labs(title = "Cross-intercept (inf->mat): overall log-scale offset",
       x = "Plate", y = "Antigen") +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        strip.text  = element_text(face = "bold", size = 7))
```

### 7.2  Cross-slope deviation from 1.0

*Values away from 0 = concentration-dependent correction needed.*

```{r heatmap-slope, fig.width=17, fig.height=9}
ggplot(norm_params,
       aes(x = plate, y = antigen_base,
           fill = slope_dev_inf2mat)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  geom_text(aes(label = round(slope_dev_inf2mat, 3)),
            size = 2, colour = "black") +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#D6604D",
                       midpoint = 0, name = "slope\ndev") +
  facet_grid(source ~ feature, scales = "free", space = "free") +
  labs(
    title    = "Cross-slope deviation from 1 (inf->mat)",
    subtitle = "0 = scalar correction;  != 0 = concentration-dependent correction",
    x = "Plate", y = "Antigen"
  ) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        strip.text  = element_text(face = "bold", size = 7))
```

---

## 8  Validation: Normalized mat_baa vs Observed inf_baa {.tabset}

Each observed **mat_baa** MFI is normalized to the inf_baa scale and overlaid with the
raw inf_baa observations.  Good normalization = orange triangles (normalized mat_baa) lying
on top of blue circles (observed inf_baa).

```{r validation}
validation <- stds_agg %>%
  filter(set == "mat_baa") %>%
  left_join(
    norm_params %>%
      select(feature, antigen_base, plate, source,
             a_src = intercept_mat_baa, b_src = slope_mat_baa,
             a_tgt = intercept_inf_baa, b_tgt = slope_inf_baa),
    by = c("feature","antigen_base","plate","source")
  ) %>%
  filter(!is.na(a_src)) %>%
  mutate(
    mfi_normalized  = normalize_mfi(median_mfi, a_src, b_src, a_tgt, b_tgt),
    log10_mfi_norm  = log10(pmax(mfi_normalized, 0.5))
  )

cat(sprintf("mat_baa observations with normalization applied: %d\n", nrow(validation)))
```

```{r validate-plot-fn}
plot_validate_feature <- function(feat) {
  inf_d <- stds_agg %>%
    filter(feature == feat, set == "inf_baa") %>%
    mutate(layer = "inf_baa observed", log10_mfi_plot = log10_mfi)

  mat_d <- validation %>%
    filter(feature == feat) %>%
    mutate(layer = "mat_baa normalised", log10_mfi_plot = log10_mfi_norm)

  combined <- bind_rows(
    inf_d %>% select(feature, antigen_base, plate, source,
                     log10_conc, log10_mfi_plot, layer),
    mat_d %>% select(feature, antigen_base, plate, source,
                     log10_conc, log10_mfi_plot, layer)
  ) %>% filter(!is.na(log10_mfi_plot))

  if (nrow(combined) == 0) return(NULL)

  ggplot(combined,
         aes(x = log10_conc, y = log10_mfi_plot,
             colour = plate, shape = layer,
             group  = interaction(plate, source, layer))) +
    geom_line(aes(linetype = source), linewidth = 0.45, alpha = 0.6) +
    geom_point(size = 2, alpha = 0.8) +
    scale_colour_manual(values = plate_colors) +
    scale_shape_manual(values = c("inf_baa observed" = 16,
                                  "mat_baa normalised" = 17)) +
    scale_linetype_manual(values = source_lty) +
    facet_wrap(~ antigen_base, scales = "free_y", ncol = 4) +
    labs(
      title    = paste("Validation --", feat),
      subtitle = "inf_baa observed (circles) vs mat_baa normalised to inf_baa scale (triangles)",
      x        = "log10(Concentration)",
      y        = "log10(MFI)  [inf_baa scale]",
      colour   = "Plate", shape = "Layer", linetype = "Source"
    ) +
    theme_bw(base_size = 9) +
    theme(legend.position = "bottom", legend.box = "vertical",
          strip.text = element_text(face = "bold", size = 8))
}
```

```{r validate-plots, results='asis', fig.width=16, fig.height=11}
for (feat in sort(unique(validation$feature))) {
  cat(sprintf("\n### %s\n\n", feat))
  p <- plot_validate_feature(feat)
  if (!is.null(p)) print(p) else cat("*No data.*\n\n")
  cat("\n\n")
}
```

### Residual bias after normalization

```{r residuals, fig.width=16, fig.height=10}
resid_tbl <- validation %>%
  mutate(
    mfi_inf_modelpred = 10^(a_tgt + b_tgt * log10_conc),
    residual_log2     = log2(mfi_normalized / mfi_inf_modelpred)
  ) %>%
  filter(is.finite(residual_log2))

ggplot(resid_tbl,
       aes(x = log10_conc, y = residual_log2,
           colour = plate, shape = source)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(size = 1.4, alpha = 0.7) +
  scale_colour_manual(values = plate_colors) +
  scale_shape_manual(values = source_shapes) +
  facet_grid(antigen_base ~ feature, scales = "free_y") +
  labs(
    title    = "Post-normalization residuals",
    subtitle = "log2(normalized mat_baa / inf_baa model prediction) -- target = 0",
    x        = "log10(Concentration)",
    y        = "log2(residual)",
    colour   = "Plate", shape = "Source"
  ) +
  theme_bw(base_size = 7) +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 6))
```

---

## 9  Export

```{r export}
# Standard curve fit parameters (one row per set x feature x antigen x plate x source)
write.csv(curve_fits, "std_curve_fits.csv", row.names = FALSE)

# Cross-curve normalization lookup table (one row per feature x antigen x plate x source)
write.csv(
  norm_params %>%
    select(feature, antigen_base, plate, source,
           intercept_inf_baa, slope_inf_baa,
           intercept_mat_baa, slope_mat_baa,
           cross_slope_inf2mat, cross_intercept_inf2mat,
           cross_slope_mat2inf, cross_intercept_mat2inf,
           slope_dev_inf2mat, slope_dev_mat2inf,
           r_squared_inf_baa, r_squared_mat_baa,
           rmse_inf_baa, rmse_mat_baa),
  "normalization_params.csv", row.names = FALSE
)

cat(sprintf("std_curve_fits.csv       : %d rows\n", nrow(curve_fits)))
cat(sprintf("normalization_params.csv : %d rows\n", nrow(norm_params)))

cat("
# ─────────────────────────────────────────────────────────────────────────────
# HOW TO APPLY THE NORMALIZATION TO SAMPLE DATA
# ─────────────────────────────────────────────────────────────────────────────
#
# sample_df must have columns: feature, antigen, plate, source, set, mfi
# IMPORTANT: add antigen_base by stripping the trailing _digits suffix:
#   sample_df <- sample_df %>%
#     mutate(antigen_base = str_remove(antigen, '_\\\\d+$'))
#
norm_lookup <- read.csv('normalization_params.csv')
#
# Normalize mat_baa samples onto the inf_baa scale:
samples_normalised <- sample_df %>%
  filter(set == 'mat_baa') %>%
  mutate(antigen_base = str_remove(antigen, '_\\\\d+$')) %>%
  left_join(norm_lookup, by = c('feature','antigen_base','plate','source')) %>%
  mutate(
    mfi_norm = 10^(cross_intercept_mat2inf +
                   cross_slope_mat2inf * log10(pmax(mfi, 0.5)))
  )
#
# Normalize inf_baa samples onto the mat_baa scale (reverse direction):
samples_normalised_rev <- sample_df %>%
  filter(set == 'inf_baa') %>%
  mutate(antigen_base = str_remove(antigen, '_\\\\d+$')) %>%
  left_join(norm_lookup, by = c('feature','antigen_base','plate','source')) %>%
  mutate(
    mfi_norm = 10^(cross_intercept_inf2mat +
                   cross_slope_inf2mat * log10(pmax(mfi, 0.5)))
  )
# ─────────────────────────────────────────────────────────────────────────────
")
```

---

## 10  Interpretation Guide

| Observation | Meaning | Action |
|---|---|---|
| `cross_slope ~= 1` | Curves are parallel; correction is the same at all concentrations | Simple multiplier: `10^cross_intercept` |
| `cross_slope != 1` | Curves diverge/converge; correction grows or shrinks with MFI | Must use the full linear transform |
| Flat lines in §6 | Concentration-independent bias | Same factor everywhere |
| Sloped lines in §6 | Concentration-dependent bias | Factor varies -- check R2 of fits |
| Triangles overlay circles in §8 | Normalization successful | Proceed |
| Triangles offset from circles | Systematic residual bias | Check R2; consider 4PL/5PL fits |
| High RMSE in fit table | Log-log linear model is a poor fit | Flag those groups; consider non-linear model |
| Plate-to-plate spread in §4.2 | Plate-specific normalization matters | Always apply plate-stratified `norm_params` |

**Core formula (apply to any sample observation):**

```
# mat_baa -> inf_baa:
log10(norm_MFI) = cross_intercept_mat2inf  +  cross_slope_mat2inf  x  log10(raw_MFI)

# inf_baa -> mat_baa:
log10(norm_MFI) = cross_intercept_inf2mat  +  cross_slope_inf2mat  x  log10(raw_MFI)
```

---

*Report generated with R `r R.version.string`.*
