---
title: "Maternal Vaccination Effect Across the Serology Panel"
subtitle: "TdaP vs TT (referent TT) by antigen, analyte, visit, and infant arm \u2014 untransformed and IgG-standardized"
author: "Scot Zens"
date: "`r Sys.Date()`"
output:
  html_document:        # -> hand this to the interpretation (secondary) engine
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:        # -> hand this to the analysis recipient
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
```

```{r bootstrap, include=FALSE}
library(here)
suppressPackageStartupMessages({
  library(stringr)
  if (requireNamespace("data.table", quietly = TRUE)) library(data.table)
})

# ---- self-locating resolver (same as the endpoint masters) ----
# ---- shared path resolver: single definition in R/find_proj_file.R ----
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R",
                "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}

# ---- shared configuration + helper library (sourced once) ----
source(find_proj_file("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"), local = FALSE, encoding = "UTF-8")

# ---- load + recode the antibody data ONCE (applies VISIT_RECODE) ----
data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))

# ---- one participant N-tracker ----
nt <- new_ntracker()
```

# Maternal Vaccination Effect Overview {-}

This is the fourth analysis part. It is **not endpoint-parameterized** (it surveys the
whole serology panel rather than a single outcome), but it reuses the same shared
library as the SBA / WT_IgG / PTNA pipeline: the global visit recode, the participant
N-tracker, the interpretation-mapping callouts, and the arm colours.

**Visit naming** is recoded throughout for the recipient:
`r paste(paste0(names(VISIT_RECODE), "\u2192", VISIT_RECODE), collapse = ", ")`.

**Comparison:** TdaP vs TT maternal vaccination (referent **TT**), shown overall and
stratified by infant priming (aP, wP; referent **wP**), on untransformed and
IgG-standardized scales.

**Downstream use:** the knitted **HTML** is for the secondary interpretation engine;
the **Word** document is for the recipient. Inline "🧭 Maps to interpretation" callouts
tie each section to the six questions the secondary maternal-effect summary will address.

---

```{r part-bubbles, child=find_proj_file('parts/04_response_bubbles.Rmd')}
```

---

# Global Participant Accounting {#global-naccounting}

```{r global-naccount, results='asis'}
nt_render(nt, caption = "Participant accounting across the bubble analysis.")
```

---

*Generated by the shared serology pipeline (Part 4). Knit both formats with*
`rmarkdown::render("response_bubbles_analysis.Rmd", output_format = "all")`.
*The IgG-standardized section requires the residual files named in
`config/endpoints.R` (`BUBBLE_RESIDUAL_AP/WP`) and the `data.table` package; if absent,
that section is skipped and the untransformed section still renders.*
