---
title: "Systems Serology — Maternal/Cord/Infant Analysis"
subtitle: "Endpoint-parameterized pipeline · primary stratification on maternal arm (TdaP vs TT)"
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
params:
  endpoint: "SBA"      # "SBA" | "WT_IgG" | "PTNA"  <-- the ONLY thing to change per endpoint
output:
  html_document:        # -> hand this to the summary engine
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:        # -> hand this to the analysis recipient
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
```

```{r bootstrap, include=FALSE}
library(here)

# ---- self-locating resolver -------------------------------------------------
# Finds config/, R/, parts/, data/ whether they sit beside this .Rmd, under an
# analysis/ subfolder, or at the project root — and works even when rmarkdown
# renders from a temp directory. Returns the first candidate that exists.
# ---- shared path resolver: single definition in R/find_proj_file.R ----
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R",
                "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}

# ---- shared configuration + helper library (sourced once) ----
source(find_proj_file("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"), local = FALSE, encoding = "UTF-8")

# ---- resolve the endpoint config from the single param ----
cfg <- get_endpoint_config(params$endpoint)

# ---- load + recode the antibody data ONCE; every part reuses it ----
data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))  # applies VISIT_RECODE

# ---- one participant N-tracker shared across all parts ----
nt <- new_ntracker()
```

# `r cfg$outcome_label`: Maternal/Cord/Infant Antibody Analysis {-}

This document is generated from a **single config-driven pipeline**. The endpoint is
set once (`params$endpoint = "`r params$endpoint`"`), and the outcome
(**`r cfg$outcome_long`**, `r cfg$outcome_feature`), the visit set, and all labels are
derived from `config/endpoints.R`. Re-running for `WT_IgG` or `PTNA` is a one-line
change to that parameter.

**Visit naming.** All visits are recoded for clarity throughout:
`r paste(paste0(names(VISIT_RECODE), "\u2192", VISIT_RECODE), collapse = ", ")`.

**Stratification.** Primary axis is the **maternal vaccination arm (TdaP vs TT)**;
within each, results are reported pooled across infant arms and separately for each
infant priming level (aP, wP).

**Downstream use.** The knitted **HTML** is intended for a summary engine; the
**Word** document is for the analysis recipient. Inline "🧭 Maps to summary"
callouts tie each analysis block to the matching section / synthesis question of the
written interpretation summary for this endpoint.

---

```{r part-maternal-chain, child=find_proj_file('parts/01_maternal_chain.Rmd')}
```

```{r part-clinical, echo=FALSE, child=find_proj_file('parts/02_clinical_covariates.Rmd', optional=TRUE)}
# Clinical-covariate part — refactored in the next increment.
# Auto-included once parts/02_clinical_covariates.Rmd exists.
```

```{r part-concurrent, echo=FALSE, child=find_proj_file('parts/03_concurrent_networks.Rmd', optional=TRUE)}
# Concurrent within-visit network / mixed-model part — next increment.
# Auto-included once parts/03_concurrent_networks.Rmd exists.
```

```{r part-antigen-subclass, echo=FALSE, child=find_proj_file('parts/04_antigen_subclass_predictors.Rmd', optional=TRUE)}
# Antigen × IgG subclass, FcG secondary binding, and (WT_IgG only) cord blood
# pre-INFMon2 predictor blocks (Blocks 7–10).
# Auto-included once parts/04_antigen_subclass_predictors.Rmd exists.
```

```{r part-effector-functions, echo=FALSE, child=find_proj_file('parts/05_effector_functions.Rmd', optional=TRUE)}
# ADCD / ADCP / ADNP effector function blocks (Blocks 11–14).
# Kept as a separate group to preserve analytical distinction between
# binding-level (Parts 1 & 4) and functional-level (Part 5) predictors.
# Auto-included once parts/05_effector_functions.Rmd exists.
```

---

# Synthesis Questions to Address {#synthesis}

*The results above (with their "Maps to summary" callouts) are intended to be
read against the written interpretation summary for `r cfg$outcome_label`. The
questions below are restated inline so the summary engine has the prompts
alongside the evidence.*

1. **Which block explains the most variance?** Is the infant outcome determined most
   by the maternal supply (Block 1), the quantity transferred (Block 3), or the
   cord→infant change (Block 5)?
2. **Dominant predictors within the leading block** — which antigen/subclass carries
   the signal (cross-reference antigen-total and full-12 relative-importance tables)?
3. **Are the ratio effects (Blocks 2/4/6) independent of quantity?** Is there a
   reproducible qualitative (subclass-balance) layer, or is the outcome purely
   quantitative?
4. **Do aP and wP infants follow different predictive pathways?** Use the infant-split
   results within each maternal arm and the relative-importance comparisons.
5. **Does the maternal vaccine (TdaP vs TT) change the chain?** Is the maternal-supply
   signal booster-dependent (strong in TdaP, attenuated/collapsed in TT)?
6. **Mechanistic model** — at which step is the signal most amplified or attenuated,
   and which subclass carries it through the chain? (Attach the Block-5 positive-median
   caveat to any cord→infant term.)
7. **Convergence with the concurrent `r paste(cfg$concurrent_visits, collapse="/")`
   architecture** — does the upstream maternal chain point to the same dominant
   antigen as the concurrent within-visit analysis?
8. **Antigen specificity at INFMon2 (Block 7)** — which antigen × IgG subclass
   combination best predicts `r cfg$outcome_label`? Is the signal concentrated in
   pertussis-specific antigens (PT / PRN / FHA) or carrier/toxoid antigens
   (DT / TT / FIM), and does this differ by maternal vaccination arm?
9. **FcG binding layer (Block 9)** — do Fc-receptor binding features add predictive
   power beyond IgG subclass quantities (cross-reference elastic-net in Blocks 7–9)?
   Which FcγR shows the strongest concurrent association, and is it consistent across
   antigens?
10. **Cord blood temporal predictor (Block 10)** —
    `r if (!is.null(cfg$cord_predictor)) paste0("After controlling for maternal arm, does cord blood ", cfg$cord_predictor_label, " independently predict INFMon2 ", cfg$outcome_label, "? Does the cord level mediate or attenuate the maternal chain signal from Block 1?") else paste0("[Not applicable for ", cfg$outcome_label, " — no pre-INFMon2 cord-blood measurement is defined for this endpoint. This question is active for the WT_IgG analysis.]")`
11. **Effector-to-whole-cell alignment (Blocks 11–14)** — do ADCD, ADCP, or ADNP
    associate with `r cfg$outcome_label` at INFMon2? Is there a consistent effector
    hierarchy across antigens (e.g., ADNP > ADCP > ADCD), and is the dominant
    antigen for effector alignment the same as the dominant antigen for subclass
    binding (cross-reference Blocks 7 and 14)?

---

# Global Participant Accounting {#global-naccounting}

A single attrition ledger across **all** parts of this analysis, so differing sample
sizes between sections are fully traceable.

```{r global-naccount, results='asis'}
nt_render(nt, caption = "Participant accounting across all analysis parts.")
```

---

*Generated by the config-driven serology pipeline. Set `params$endpoint` to
`SBA`, `WT_IgG`, or `PTNA`. Knit both formats with*
`rmarkdown::render("SBA_analysis.Rmd", output_format = "all")`.
