```{r bub-setup, include=FALSE}
# CHILD document. Assumes the master created data_raw, nt, shared_utils(),
# and sourced config/endpoints.R + R/C1_serology_helpers.R.
source(shared_utils("R/C0_connection_transform.R"), local = FALSE, encoding = "UTF-8")
include_feature    <- BUBBLE_INCLUDE_FEATURE     # globals used by standardized_set defaults
include_feature_wo <- BUBBLE_INCLUDE_FEATURE_WO
have_dt  <- requireNamespace("data.table", quietly = TRUE)
res_ap   <- shared_utils(BUBBLE_RESIDUAL_AP, optional = TRUE)
res_wp   <- shared_utils(BUBBLE_RESIDUAL_WP, optional = TRUE)
have_std <- have_dt && !is.null(res_ap) && !is.null(res_wp)
```

# Maternal Vaccination Effect Across Antigens and Analytes {#bubbles}

This part surveys the **effect of maternal vaccination (TdaP vs TT, referent TT)** on
the full serology panel — antigens `r paste(BUBBLE_ANTIGENS, collapse=", ")` × analytes
`r paste(BUBBLE_ANALYTES, collapse=", ")` — at the three infant visits
(`r paste(BUBBLE_VISITS, collapse=", ")`). Each antigen × analyte cell is a Wilcoxon
rank-sum test of TT vs TdaP; the bubble encodes effect size (median difference),
direction, and significance. Two scales are shown: **untransformed** (log10 MFI) and
**IgG-standardized** (residuals after removing the total-IgG signal).

Within each visit, effects are shown **overall (all infants pooled)** and **stratified
by infant priming (aP, wP — referent wP)**, so the maternal effect and its modification
by infant arm are both visible. The bubble figures are retained from the original
analysis; the per-grid **tables are new** (n, medians, effect, direction, raw and
adjusted p), as is a cross-visit effect summary intended to seed the secondary analysis.

```{r bub-mapnote-top, echo=FALSE, results='asis'}
mapnote("This part feeds a forthcoming **secondary summary analysis** of the maternal ",
        "vaccination effect. The output here is organized to answer:",
        feeds = c(
          "Q1 \u2014 Magnitude: which antigen\u00d7analyte effects are largest, overall and per visit? (bubble figures + per-grid tables)",
          "Q2 \u2014 Antigen specificity: are effects concentrated in vaccine antigens (PT/FHA/PRN) vs DT/TT? (antigen axis of every grid)",
          "Q3 \u2014 Function vs binding: do effects appear in functional analytes (ADCD/ADCP/ADNP, FcgR) as well as IgG/subclasses? (analyte axis)",
          "Q4 \u2014 Temporal persistence: do InfMon2 effects persist or wane by InfMon5/InfMon9? (cross-visit summary tables)",
          "Q5 \u2014 Infant-arm modification: do effects differ in aP vs wP infants? (overall vs aP vs wP panels)",
          "Q6 \u2014 IgG independence: which effects survive IgG standardization? (untransformed vs standardized sections)"))
if (!have_std)
  cat("\n> **Note:** IgG-standardized residual files not found (or data.table missing); ",
      "the standardized section is skipped. The untransformed section is unaffected.\n\n")
```

---

## Data Preparation

```{r bub-data, results='asis'}
unstandardized_data <- data_raw %>%
  dplyr::filter(feature %in% BUBBLE_INCLUDE_FEATURE,
                arm_name %in% MATERNAL_ARMS,
                visit_name %in% BUBBLE_VISITS) %>%
  dplyr::select(subject_accession, visit_name, arm_name, infant_arm,
                antigen, analyte, feature, value_reported, log_assay_value) %>%
  distinct(subject_accession, visit_name, feature, .keep_all = TRUE) %>%
  mutate(across(c(antigen, analyte, feature, arm_name, infant_arm), as.character))
unstandardized_data$standstat <- "untransformed"

if (have_std) {
  std_wp <- standardized_set("IgG", unstandardized_data[unstandardized_data$infant_arm == "wP", ], res_wp)
  std_ap <- standardized_set("IgG", unstandardized_data[unstandardized_data$infant_arm == "aP", ], res_ap)
  std_not   <- dplyr::bind_rows(std_wp, std_ap)
  total_igg <- unstandardized_data[unstandardized_data$feature %in% BUBBLE_IGGLIST, ]
  standardized_data <- dplyr::bind_rows(std_not, total_igg)
  standardized_data$standstat <- "standardized"
  data_all <- dplyr::bind_rows(unstandardized_data, standardized_data)
} else {
  data_all <- unstandardized_data
}

# Recompute antigen/analyte from feature so both scales are consistent.
data_all$antigen <- stringr::str_split_i(as.character(data_all$feature), "_", 1)
data_all$analyte <- stringr::str_split_i(as.character(data_all$feature), "_", 2)
data_all$arm_name <- factor(data_all$arm_name, levels = c("TT", "TdaP"))   # TT referent

ft <- as.data.frame(table(data_all$standstat, data_all$visit_name))
colnames(ft) <- c("Scale", "Visit", "n_obs")
ft <- ft %>% dplyr::filter(n_obs > 0) %>% pivot_wider(names_from = Visit, values_from = n_obs, values_fill = 0)
knitr::kable(ft, caption = "Table B1. Observation rows by scale and visit (after recode & filter).") %>% print()

# Per visit x infant-arm subject counts by maternal arm
sc <- data_all %>% dplyr::filter(standstat == "untransformed") %>%
  distinct(subject_accession, visit_name, infant_arm, arm_name) %>%
  count(visit_name, infant_arm, arm_name) %>%
  pivot_wider(names_from = arm_name, values_from = n, values_fill = 0)
knitr::kable(sc, caption = "Table B2. Unique subjects by visit, infant arm, and maternal arm.") %>% print()
```

```{r bub-run-setup, results='hide'}
strata_defs <- list(`Overall (all infants)` = NULL, `aP infants` = "aP", `wP infants` = "wP")
bubble_stats_acc <- list()   # accumulate every grid's stats for the cross-visit summary

run_bubble_cell <- function(ss, visit, strat_name) {
  iarm <- strata_defs[[strat_name]]
  sub  <- data_all %>% dplyr::filter(standstat == ss, visit_name == visit)
  if (!is.null(iarm)) sub <- sub %>% dplyr::filter(infant_arm == iarm)

  # participant accounting (unique subjects per maternal arm)
  subj <- sub %>% distinct(subject_accession, arm_name)
  n_TT <- sum(subj$arm_name == "TT"); n_TdaP <- sum(subj$arm_name == "TdaP")
  nt_log_manual(nt, "Bubbles", paste0(ss, " / ", visit), strat_name,
                "TdaP vs TT effect grid", n_entered = n_TT + n_TdaP, n_used = n_TT + n_TdaP,
                note = sprintf("TT = %d, TdaP = %d subjects; per-cell n in table", n_TT, n_TdaP))

  cat("####", strat_name, sprintf("(TT n = %d, TdaP n = %d)\n\n", n_TT, n_TdaP))
  if (n_TT < 2 || n_TdaP < 2) { cat("*Too few subjects in one arm; grid skipped.*\n\n"); return(invisible()) }

  res <- bubble_arm_effect(sub, standstat = ss,
                           title = sprintf("%s \u2014 %s \u2014 %s (TdaP vs TT)", ss, visit, strat_name))
  print(res$plot); cat("\n\n")
  print(bubble_stats_kable(res$stats,
        caption = sprintf("Effect grid \u2014 %s, %s, %s.", ss, visit, strat_name)))
  cat("\n")
  bubble_stats_acc[[paste(ss, visit, strat_name, sep = " | ")]] <<-
    res$stats %>% mutate(scale = ss, visit = visit, stratum = strat_name)
}
```

---

```{r bub-loops, results='asis', fig.width=8, fig.height=5}
scales <- if (have_std) c("untransformed", "standardized") else "untransformed"
for (ss in scales) {
  cat("##", ss, "scale\n\n")
  mapnote(ss, " effects across antigen\u00d7analyte, all visits and infant strata.",
          feeds = if (ss == "standardized")
            "Q6 (effects independent of total IgG) of the secondary summary" else
            c("Q1\u2013Q3 (magnitude, antigen specificity, function) of the secondary summary"))
  for (visit in BUBBLE_VISITS) {
    cat("###", visit, "\n\n")
    for (strat_name in names(strata_defs)) run_bubble_cell(ss, visit, strat_name)
  }
}
```

---

## Cross-Visit Effect Summary (for the secondary analysis)

```{r bub-cross-mapnote, echo=FALSE, results='asis'}
mapnote("These tables line up each antigen\u00d7analyte effect across the three visits, ",
        "so the secondary analysis can read temporal persistence and infant-arm ",
        "modification directly.",
        feeds = c("Q4 (temporal persistence) and Q5 (infant-arm modification) of the secondary summary"))
```

```{r bub-cross, results='asis'}
all_bubble_stats <- dplyr::bind_rows(bubble_stats_acc)
if (nrow(all_bubble_stats) > 0) {
  for (ss in scales) for (strat_name in names(strata_defs)) {
    sub <- all_bubble_stats %>%
      dplyr::filter(scale == ss, stratum == strat_name) %>%
      mutate(cell = sprintf("%+.2f %s", effect_size, pvalue_to_label(p_adj)))
    # keep antigen x analyte that are significant in at least one visit
    keep <- sub %>% group_by(antigen, analyte) %>%
      dplyr::summarize(any_sig = any(significant, na.rm = TRUE), .groups = "drop") %>%
      dplyr::filter(any_sig)
    sub <- sub %>% semi_join(keep, by = c("antigen", "analyte"))
    if (nrow(sub) == 0) {
      cat(sprintf("\n**%s \u2014 %s:** no antigen\u00d7analyte reached adjusted significance at any visit.\n\n",
                  ss, strat_name)); next
    }
    wide <- sub %>%
      dplyr::select(antigen, analyte, visit, cell) %>%
      pivot_wider(names_from = visit, values_from = cell) %>%
      arrange(antigen, analyte)
    cat(sprintf("\n### %s \u2014 %s\n\n", ss, strat_name))
    print(knitr::kable(wide,
      caption = sprintf("Signed effect (median TT\u2212TdaP) with significance, by visit \u2014 %s, %s. Significant in \u22651 visit only.", ss, strat_name),
      align = c("l","l", rep("r", ncol(wide) - 2))))
    cat("\n")
  }
} else cat("*No grids computed.*\n\n")
```

---

## Participant Accounting (Bubbles) {#bubbles-naccounting}

```{r bub-naccount, results='asis'}
nt_render(nt, part = "Bubbles",
          caption = "Table B3. Participant accounting per effect grid (per-cell n shown in each grid table).")
```
