```{r conc-setup, include=FALSE}
# =============================================================================
# CHILD document — CONCURRENT WITHIN-VISIT ARCHITECTURE (v2, expanded + staged)
# -----------------------------------------------------------------------------
# Assumes cfg, data_raw, nt, shared_utils() exist (created by the driver).
#
# WHAT CHANGED vs v1
#   1. Feature set expanded beyond PT/FHA/PRN subclasses:
#        - FIM antigen (total + subclasses)
#        - FcγR binding: FcgR2a, FcgR3b   (all 6 antigens)
#        - Effector functions: ADCD       (all 6 antigens)
#                              ADCP, ADNP  (PT/PRN/DT/TT only — FHA & FIM absent)
#      Vocabulary is built from the ACTUAL analyte casing in the data
#      (ADCD/ADCP/ADNP upper-case, FcgR2a/FcgR3b mixed-case) and then filtered
#      to features actually present, so a coverage/casing gap is reported, not
#      silently dropped by complete-case deletion.
#   2. DATA MODE: raw (log_assay_value) only for now. The two-mode machinery
#      (raw + total-IgG-standardised) is retained but igg_std is DEFERRED:
#      DATA_MODES defaults to "raw", so standardize_wide() is never called.
#      To re-enable later, set DATA_MODES <- c("raw","igg_std") and finalise
#      the standardize_wide() definition.
#   3. Visit restricted to the infant 8-week timepoint (InfMon2) only.
#   4. Staged execution with console/terminal logging: cheap layers
#      (regression / relimp / ratio / mixed model) run first; expensive layers
#      (GGM networks, then NCT permutations) run last and are individually
#      toggleable, so a fast pass is a one-line change.
# =============================================================================
set.seed(20240614)

# ---- control knobs (a driver may pre-define any of these to override) -------
if (!exists("CONC_VISIT"))    CONC_VISIT    <- "InfMon2"          # 8-week infant timepoint
if (!exists("DATA_MODES"))    DATA_MODES    <- c("raw")   # igg_std deferred — raw only for now
if (!exists("NCT_IT"))        NCT_IT        <- 1000L              # NCT permutations (lower for quick pass)
if (!exists("RUN_NETWORKS"))  RUN_NETWORKS  <- TRUE               # STAGE 4
if (!exists("RUN_NCT"))       RUN_NCT       <- TRUE               # STAGE 5 (most expensive)
if (!exists("RUN_EFFECTOR_NET")) RUN_EFFECTOR_NET <- TRUE         # optional 2nd (FcγR+ADCD) network

have_net  <- all(vapply(c("bootnet", "qgraph", "networktools", "NetworkComparisonTest"),
                        requireNamespace, logical(1), quietly = TRUE))
have_lmer <- requireNamespace("lme4",     quietly = TRUE)
have_corr <- requireNamespace("corrplot", quietly = TRUE)
RUN_NETWORKS <- RUN_NETWORKS && have_net
RUN_NCT      <- RUN_NCT      && have_net

# ---- console / terminal logging --------------------------------------------
# message() writes to stderr, so these lines appear live in the R console /
# terminal (and the render log) while the document knits — the "better
# feedback" requested. They are independent of the knitted HTML/Word output.
.RUNLOG <- new.env(parent = emptyenv())
.RUNLOG$t0   <- Sys.time()
.RUNLOG$rows <- list()
.hms <- function() format(Sys.time(), "%H:%M:%S")
.el  <- function() as.numeric(difftime(Sys.time(), .RUNLOG$t0, units = "secs"))
.stage <- function(k, K, msg) message(sprintf("\n[%s | +%5.1fs] === STAGE %d/%d — %s ===",
                                              .hms(), .el(), k, K, msg))
.info  <- function(...) message(sprintf("[%s | +%5.1fs]     %s", .hms(), .el(), sprintf(...)))
.mark  <- function(label) {
  .RUNLOG$rows[[length(.RUNLOG$rows) + 1]] <-
    data.frame(step = label, elapsed_s = round(.el(), 1), stringsAsFactors = FALSE)
  invisible()
}
N_STAGES <- 5L
message(sprintf("[%s] concurrent v2 START — endpoint=%s outcome=%s visit=%s modes=%s",
                .hms(), cfg$endpoint, cfg$outcome_feature, CONC_VISIT,
                paste(DATA_MODES, collapse = "+")))
.info("capabilities: networks=%s lmer=%s corrplot=%s | NCT_IT=%d RUN_NCT=%s",
      have_net, have_lmer, have_corr, NCT_IT, RUN_NCT)

# ---- expanded feature vocabulary (built in the data's own casing) -----------
ANTIGENS6      <- c("PT", "FHA", "PRN", "DT", "TT", "FIM")
ADCPNP_ANTIGENS <- c("PT", "PRN", "DT", "TT")     # FHA & FIM have no ADCP/ADNP
FCG_RECEPTORS  <- c("FcgR2a", "FcgR3b")           # mixed-case, as stored
EFF_ADCD       <- "ADCD"                           # upper-case, as stored

ANTIGEN_TOTALS6 <- paste0(ANTIGENS6, "_IgG")                                   # 6
FCG_ALL         <- as.vector(t(outer(ANTIGENS6, FCG_RECEPTORS, paste, sep = "_")))  # 12
ADCD_ALL        <- paste0(ANTIGENS6, "_ADCD")                                  # 6
ADCPNP_ALL      <- as.vector(t(outer(ADCPNP_ANTIGENS, c("ADCP", "ADNP"), paste, sep = "_")))  # 8
# PT/FHA/PRN/FIM subclass vectors already defined in config (PT_SUBCLASSES …).
SUBCLASS_ALL    <- c(PT_SUBCLASSES, FHA_SUBCLASSES, PRN_SUBCLASSES, FIM_SUBCLASSES)  # +FIM

# ---- verify presence against the loaded (post-recode) feature levels --------
# load_serology_data() rebuilds `feature = paste(antigen, analyte)` in mixed
# case; verify_predictor_casing() already guards the chain/subclass set, but it
# does NOT cover FcγR / effector features, so we check them here and report.
present_feats <- as.character(unique(data_raw$feature))
keep_present  <- function(v) v[v %in% present_feats]
report_gap    <- function(name, v) {
  miss <- setdiff(v, present_feats)
  if (length(miss)) .info("%-22s %d/%d present; ABSENT: %s",
                          name, length(v) - length(miss), length(v),
                          paste(miss, collapse = ", "))
  else .info("%-22s %d/%d present", name, length(v), length(v))
}
report_gap("antigen totals",  ANTIGEN_TOTALS6)
report_gap("subclasses(+FIM)", SUBCLASS_ALL)
report_gap("FcγR binding",     FCG_ALL)
report_gap("ADCD",             ADCD_ALL)
report_gap("ADCP/ADNP",        ADCPNP_ALL)

ANTIGEN_TOTALS6 <- keep_present(ANTIGEN_TOTALS6)
FCG_ALL         <- keep_present(FCG_ALL)
ADCD_ALL        <- keep_present(ADCD_ALL)
ADCPNP_ALL      <- keep_present(ADCPNP_ALL)

# All antigen-specific features that get standardised in igg_std mode (totals
# and the whole-cell outcome are NOT standardised — they carry quantity, and
# the outcome has no antigen total to divide by).
STD_TARGETS <- keep_present(c(SUBCLASS_ALL, FCG_ALL, ADCD_ALL, ADCPNP_ALL))

# Predictor blocks used by the regression layer, in run order.
BLOCKS <- list(
  antigen  = list(label = "Antigen totals (6)",                 feats = ANTIGEN_TOTALS6, sec = "Antigen",  relimp = TRUE),
  PT       = list(label = "PT subclasses",                      feats = keep_present(PT_SUBCLASSES),  sec = "Subclass", relimp = TRUE),
  FHA      = list(label = "FHA subclasses",                     feats = keep_present(FHA_SUBCLASSES), sec = "Subclass", relimp = TRUE),
  PRN      = list(label = "PRN subclasses",                     feats = keep_present(PRN_SUBCLASSES), sec = "Subclass", relimp = TRUE),
  FIM      = list(label = "FIM subclasses",                     feats = keep_present(FIM_SUBCLASSES), sec = "Subclass", relimp = TRUE),
  fcgr     = list(label = "FcγR binding (all antigens)",        feats = FCG_ALL,         sec = "FcgR",     relimp = TRUE),
  adcd     = list(label = "ADCD (all antigens)",                feats = ADCD_ALL,        sec = "Effector", relimp = TRUE),
  adcpnp   = list(label = "ADCP + ADNP (PT/PRN/DT/TT)",         feats = ADCPNP_ALL,      sec = "Effector", relimp = TRUE),
  full_sub = list(label = "Full subclass (PT/FHA/PRN/DT/TT ×4)", feats = keep_present(cfg$all_subclasses), sec = "Full", relimp = TRUE),
  ratio    = list(label = "Pro/tolerogenic ratios",             feats = c("PT_ratio", "FHA_ratio", "PRN_ratio"), sec = "Ratio", relimp = FALSE)
)

# Interaction antigens for the mixed model (must have full IgG1–4 subclasses).
INTERACTION_ANTIGENS <- c("PT", "FHA", "PRN")

# Network node sets.
NET_SUBCLASS_NODES <- keep_present(c(cfg$outcome_feature, NETWORK_SUBCLASSES))  # outcome + PT/FHA/PRN/FIM ×4
NET_EFFECTOR_NODES <- keep_present(c(cfg$outcome_feature, FCG_ALL, ADCD_ALL))   # outcome + FcγR + ADCD

# =============================================================================
# igg_std STANDARDISATION  —  PROVISIONAL DEFINITION, CONFIRM BEFORE USE
# -----------------------------------------------------------------------------
# Placeholder = per-antigen total-IgG normalisation on the log scale:
#     feature_std = log_value(A_analyte) − log_value(A_IgG)
# i.e. express each antigen-specific subclass / FcγR / effector RELATIVE to the
# same antigen's total IgG at the same draw. Totals and the whole-cell outcome
# are left untouched. Pro-/tolerogenic ratios are invariant to this transform
# (the total cancels), so the ratio block is identical across modes.
#
# If the project's "total-IgG-standardised" values are instead (a) stored in
# ebaa_extra2 under a column/flag, or (b) the precomputed igg_standard_residuals
# objects, replace the body of this function with a lookup/join — the rest of
# the pipeline is agnostic to how the standardised matrix is produced.
# =============================================================================
standardize_wide <- function(dw, targets = STD_TARGETS) {
  for (f in targets) {
    ag  <- sub("_.*$", "", f)
    tot <- paste0(ag, "_IgG")
    if (tot %in% names(dw) && f %in% names(dw))
      dw[[f]] <- dw[[f]] - dw[[tot]]
  }
  dw
}

# ---- build one mode's wide matrix at the single concurrent visit ------------
prep_mode <- function(mode) {
  feats <- keep_present(c(cfg$outcome_feature, ANTIGEN_TOTALS6, STD_TARGETS))
  dl <- data_raw %>%
    dplyr::filter(feature %in% feats,
                  visit_name == CONC_VISIT,
                  arm_name %in% cfg$maternal_arms) %>%
    dplyr::select(subject_accession, arm_name, infant_arm, feature, log_assay_value) %>%
    dplyr::distinct(subject_accession, feature, .keep_all = TRUE)
  dw <- dl %>%
    tidyr::pivot_wider(id_cols = c(subject_accession, arm_name, infant_arm),
                       names_from = feature, values_from = log_assay_value) %>%
    add_concurrent_ratios()                       # ratios from RAW subclasses (mode-invariant)
  if (identical(mode, "igg_std")) dw <- standardize_wide(dw)
  dw
}

# ---- expanded per-stratum block runner (uses generic run_lm → N-tracker) ----
run_blocks_x <- function(d, stratum, mode, tracker = nt) {
  oc <- cfg$outcome_feature
  out <- list()
  for (bn in names(BLOCKS)) {
    b <- BLOCKS[[bn]]
    preds <- intersect(b$feats, names(d))
    if (length(preds) < 1) { out[[bn]] <- NULL; next }
    out[[bn]] <- run_lm(oc, preds, d,
                        label = sprintf("%s [%s]", b$label, mode),
                        show_relimp_calc = b$relimp,
                        tracker = tracker, part = "Concurrent",
                        section = paste0(b$sec, " · ", mode), stratum = stratum)
  }
  out
}

show_stratum_blocks <- function(res, full = TRUE) {
  ids <- if (full) names(res) else c("antigen", "PT", "FHA", "PRN", "FIM",
                                     "fcgr", "adcd", "adcpnp", "full_sub", "ratio")
  for (id in ids) if (!is.null(res[[id]]))
    show_model(res[[id]], show_relimp = !identical(id, "ratio"))
}
```

# Concurrent Within-Visit Architecture (Expanded) {#concurrent}

This part models the infant's **`r cfg$outcome_label`** against antibody features
**measured at the same infant visit**, restricted to the **8-week timepoint
(`r CONC_VISIT`)**. Relative to the original version it adds the **FIM** antigen,
the **FcγR binding** features (FcgR2a, FcgR3b), and the **antigen-specific effector
functions** (ADCD for all six antigens; ADCP and ADNP for PT/PRN/DT/TT only), on the
**untransformed (raw) log assay values**. The total-IgG-standardised mode is deferred
and not run here.

**Stratification is maternal-arm primary** (TdaP / TT), with infant splits (aP, wP)
nested. Networks / NCT need ≥ `r NETWORK_MIN` complete cases; underpowered cells are
skipped and recorded in the participant accounting. Execution is staged
(cheap → expensive) with live console logging.

```{r conc-mapnote, echo=FALSE, results='asis'}
mapnote("Expanded within-visit layer at ", CONC_VISIT, " for **", cfg$outcome_label,
        "**, raw + total-IgG-standardised.",
        feeds = c(
          "Antigen / subclass / effector relative importance \u2192 \u00a71, \u00a73",
          "FHA negative association \u2192 \u00a72",
          "Ratio models \u2192 \u00a74",
          "Antigen \u00d7 subclass \u00d7 level interaction \u2192 \u00a75",
          "GGM networks + bridge centrality \u2192 \u00a76",
          "NCT (arm / infant-arm contrasts at 8 weeks) \u2192 \u00a77"))
if (!have_net)
  cat("\n> **Note:** network packages not installed; STAGES 4\u20135 skipped.\n\n")
```

---

## STAGE 1 — Data Preparation and Coverage

```{r conc-prep, results='asis'}
.stage(1, N_STAGES, "data preparation + coverage")
wide_by_mode  <- list()
strata_by_mode <- list()
for (md in DATA_MODES) {
  dw <- prep_mode(md)
  wide_by_mode[[md]]   <- dw
  strata_by_mode[[md]] <- build_maternal_strata(dw)
  .info("mode=%-7s wide = %d rows \u00d7 %d cols", md, nrow(dw), ncol(dw))
  cat(sprintf("**Mode `%s`** — %d subjects \u00d7 %d columns at %s.  \n",
              md, nrow(dw), ncol(dw), CONC_VISIT))
  cnt <- as.data.frame(table(dw$arm_name), stringsAsFactors = FALSE)
  colnames(cnt) <- c("Maternal arm", "n_subjects")
  knitr::kable(cnt, caption = sprintf("Subjects by maternal arm (mode = %s, %s).",
                                      md, CONC_VISIT)) %>% print()
  cat("\n")
}
.mark("stage1_prep")
```

---

## STAGE 2 — Regression Layer (relative importance + ratios) {#regression}

```{r conc-reg-mapnote, echo=FALSE, results='asis'}
mapnote("Relative-importance and ratio models across the expanded feature set ",
        "(subclass, FcγR, ADCD, ADCP/ADNP), raw + standardised.",
        feeds = c("\u00a71 (dominant antigen/analyte), \u00a73 (arm signatures)",
                  "\u00a72 (FHA negative association)", "\u00a74 (ratio model)"))
```

```{r conc-reg, results='asis'}
.stage(2, N_STAGES, "regression / relative importance / ratios")
conc_by_mode <- list()
for (md in DATA_MODES) {
  cat("\n## Mode: `", md, "`\n\n", sep = "")
  sx <- strata_by_mode[[md]]
  conc <- lapply(names(sx), function(k) {
    r <- run_blocks_x(sx[[k]]$data, stratum = sx[[k]]$label, mode = md)
    .info("mode=%-7s stratum=%-26s blocks fitted", md, sx[[k]]$label)
    r
  })
  names(conc) <- names(sx)
  conc_by_mode[[md]] <- conc
  for (m in cfg$maternal_arms) {
    cat("### ", m, " mothers\n\n", sep = "")
    show_stratum_blocks(conc[[m]], full = TRUE)
    for (i in cfg$infant_arms) {
      key <- paste0(m, "_", i)
      if (!is.null(conc[[key]])) {
        cat("#### ", sx[[key]]$label, "\n\n", sep = "")
        show_stratum_blocks(conc[[key]], full = FALSE)
      }
    }
  }
}
.mark("stage2_regression")
```

### Cross-stratum full-subclass relative importance (TdaP vs TT)

```{r conc-relimp-compare, results='asis'}
for (md in DATA_MODES) {
  cc <- conc_by_mode[[md]]
  rT <- cc[["TdaP"]]$full_sub$relimp; rN <- cc[["TT"]]$full_sub$relimp
  if (!is.null(rT) && !is.null(rN)) {
    cmp <- merge(data.frame(feature = names(rT@lmg), lmg_TdaP = round(rT@lmg * 100, 2)),
                 data.frame(feature = names(rN@lmg), lmg_TT   = round(rN@lmg * 100, 2)),
                 by = "feature") %>%
      dplyr::mutate(delta_lmg = round(lmg_TdaP - lmg_TT, 2)) %>%
      dplyr::arrange(dplyr::desc(abs(delta_lmg)))
    knitr::kable(cmp, caption = sprintf("Full-subclass relative importance (lmg %%): TdaP vs TT — mode %s.", md),
                 align = c("l", "r", "r", "r")) %>% print()
    cat("\n")
  }
}
```

---

## STAGE 3 — Antigen × Subclass × Level Interaction (mixed model) {#interaction}

```{r conc-lmer-mapnote, echo=FALSE, results='asis'}
mapnote("Three-way antigen \u00d7 subclass \u00d7 binding-level interaction. At a single ",
        "visit the (1|subject) random intercept is identified across the ",
        "antigen\u00d7subclass rows, not across visits.",
        feeds = "\u00a75 (three-way interaction)")
```

```{r conc-lmer, results='asis'}
.stage(3, N_STAGES, "mixed-model interaction (lmer)")
if (have_lmer) {
  patt <- paste0("(", paste(INTERACTION_ANTIGENS, collapse = "|"), ")_(IgG[1-4])")
  for (md in DATA_MODES) {
    cat("### Mode: `", md, "`\n\n", sep = "")
    for (m in cfg$maternal_arms) {
      d <- strata_by_mode[[md]][[m]]$data
      dl <- d %>%
        dplyr::select(subject_accession, dplyr::all_of(cfg$outcome_feature),
                      dplyr::matches("(PT|FHA|PRN)_IgG[1-4]$")) %>%
        tidyr::pivot_longer(dplyr::matches("IgG[1-4]$"),
                            names_to = c("antigen", "subclass"),
                            names_pattern = patt, values_to = "subclass_value") %>%
        tidyr::drop_na(tidyselect::all_of(c("subclass_value", cfg$outcome_feature)))
      fml <- stats::as.formula(paste0(cfg$outcome_feature,
                    " ~ antigen * subclass * subclass_value + (1 | subject_accession)"))
      mm <- tryCatch(lme4::lmer(fml, data = dl), error = function(e) NULL)
      if (is.null(mm)) { cat(sprintf("*%s: mixed model could not be fit.*\n\n", m)); next }
      av <- as.data.frame(anova(mm)); av <- cbind(Term = rownames(av), round(av, 3)); rownames(av) <- NULL
      knitr::kable(av, caption = sprintf("Type-III ANOVA of fixed effects — %s mothers (mode %s).", m, md),
                   row.names = FALSE) %>% print(); cat("\n")
      .info("mode=%-7s arm=%-4s lmer fitted (n_rows=%d)", md, m, nrow(dl))
    }
  }
} else cat("*lme4 not installed; mixed-model interaction skipped.*\n\n")
.mark("stage3_lmer")
```

---

## STAGE 4 — GGM Networks and Bridge Centrality {#networks}

```{r conc-net-mapnote, echo=FALSE, results='asis'}
mapnote("Partial-correlation (EBICglasso) networks and bridge centrality at ",
        CONC_VISIT, ". A subclass(+FIM) network, and — if enabled — a FcγR+ADCD ",
        "network.", feeds = "\u00a76 (network architecture, bridge nodes)")
```

```{r conc-networks, results='asis', fig.height=6, fig.width=7}
.stage(4, N_STAGES, "GGM networks + bridge centrality")
bridge_store <- list()   # bridge_store[[mode]][[which]][[arm]]
run_one_network <- function(dat, nodes, communities, groups_idx, group_colors,
                            title, tracker_label, mode, which) {
  d <- dat %>% dplyr::select(dplyr::all_of(nodes)) %>% tidyr::drop_na()
  nt_log(nt, "Concurrent", paste0(which, " network @ ", CONC_VISIT),
         title, tracker_label, nodes, dat)
  cat(sprintf("\n**%s** (n = %d)\n\n", title, nrow(d)))
  .info("mode=%-7s %s network: %s n=%d", mode, which, title, nrow(d))
  if (nrow(d) < NETWORK_MIN) { cat(sprintf("*n < %d; skipped.*\n\n", NETWORK_MIN)); return(NULL) }
  net <- tryCatch(bootnet::estimateNetwork(d, default = "EBICglasso", threshold = TRUE),
                  error = function(e) NULL)
  if (is.null(net)) { cat("*Network estimation failed.*\n\n"); return(NULL) }
  qgraph::qgraph(net$graph, groups = groups_idx, color = group_colors,
                 layout = "spring", labels = nodes, legend = TRUE,
                 legend.cex = 0.33, vsize = 7, title = title)
  br <- networktools::bridge(net$graph, communities = communities)
  bdf <- data.frame(Variable = names(br$`Bridge Strength`),
                    Bridge_Strength = round(br$`Bridge Strength`, 3),
                    Bridge_EI = round(br$`Bridge Expected Influence (1-step)`, 3),
                    row.names = NULL) %>% dplyr::arrange(dplyr::desc(Bridge_Strength))
  print(knitr::kable(bdf, caption = paste0("Bridge centrality — ", title))); cat("\n")
  bdf
}

# community assignment for the subclass(+FIM) network, robust to node order
sub_comm <- function(nodes) {
  comm <- integer(length(nodes))
  comm[nodes == cfg$outcome_feature] <- 1L
  comm[nodes %in% PT_SUBCLASSES]  <- 2L
  comm[nodes %in% FHA_SUBCLASSES] <- 3L
  comm[nodes %in% PRN_SUBCLASSES] <- 4L
  comm[nodes %in% FIM_SUBCLASSES] <- 5L
  comm
}
sub_groups <- function(nodes) list(
  OUT = which(nodes == cfg$outcome_feature),
  PT  = which(nodes %in% PT_SUBCLASSES),  FHA = which(nodes %in% FHA_SUBCLASSES),
  PRN = which(nodes %in% PRN_SUBCLASSES), FIM = which(nodes %in% FIM_SUBCLASSES))
sub_cols <- c("gold", "tomato", "steelblue", "seagreen", "mediumpurple")

if (RUN_NETWORKS) {
  for (md in DATA_MODES) {
    cat("### Mode: `", md, "`\n\n", sep = "")
    bridge_store[[md]] <- list(subclass = list(), effector = list())
    for (m in cfg$maternal_arms) {
      dat <- strata_by_mode[[md]][[m]]$data   # already single-visit (InfMon2); no visit_name column
      # subclass(+FIM) network
      nodes <- NET_SUBCLASS_NODES
      bdf <- run_one_network(dat, nodes, sub_comm(nodes), sub_groups(nodes), sub_cols,
                             sprintf("%s — %s subclass network @ %s", m, cfg$outcome_label, CONC_VISIT),
                             m, md, "subclass")
      if (!is.null(bdf)) bridge_store[[md]]$subclass[[m]] <- bdf
      # optional FcγR + ADCD "secondary/effector" network
      if (RUN_EFFECTOR_NET) {
        nod2 <- NET_EFFECTOR_NODES
        comm2 <- integer(length(nod2)); comm2[nod2 == cfg$outcome_feature] <- 1L
        comm2[nod2 %in% FCG_ALL] <- 2L; comm2[nod2 %in% ADCD_ALL] <- 3L
        grp2 <- list(OUT = which(nod2 == cfg$outcome_feature),
                     FcgR = which(nod2 %in% FCG_ALL), ADCD = which(nod2 %in% ADCD_ALL))
        bdf2 <- run_one_network(dat, nod2, comm2, grp2, c("gold", "darkorange", "purple"),
                                sprintf("%s — %s FcγR+ADCD network @ %s", m, cfg$outcome_label, CONC_VISIT),
                                m, md, "effector")
        if (!is.null(bdf2)) bridge_store[[md]]$effector[[m]] <- bdf2
      }
    }
  }
} else cat("*Networks disabled (RUN_NETWORKS = FALSE or packages absent).*\n\n")
.mark("stage4_networks")
```

---

## STAGE 5 — Network Comparison Tests (NCT) {#nct}

```{r conc-nct-mapnote, echo=FALSE, results='asis'}
mapnote("NCT at the single 8-week visit: maternal arm (TdaP vs TT) and infant arm ",
        "(aP vs wP) within maternal arm. Temporal contrasts are not applicable at a ",
        "single visit.", feeds = "\u00a77 (cross-arm network comparison)")
```

```{r conc-nct, results='asis'}
.stage(5, N_STAGES, sprintf("NCT permutations (it = %d)", NCT_IT))
nct_subset <- function(dat, nodes) dat %>%
  dplyr::select(dplyr::all_of(nodes)) %>% tidyr::drop_na()
run_nct <- function(d1, d2, lab) {
  cat(sprintf("\n**%s** — n1 = %d, n2 = %d\n\n", lab, nrow(d1), nrow(d2)))
  .info("NCT %-48s n1=%d n2=%d", lab, nrow(d1), nrow(d2))
  if (nrow(d1) < NETWORK_MIN || nrow(d2) < NETWORK_MIN) {
    cat(sprintf("*One or both groups n < %d; NCT skipped.*\n\n", NETWORK_MIN)); return(invisible()) }
  res <- tryCatch(NetworkComparisonTest::NCT(d1, d2, it = NCT_IT, test.edges = FALSE,
                  test.centrality = TRUE, centrality = c("strength", "expectedInfluence")),
                  error = function(e) NULL)
  if (is.null(res)) { cat("*NCT failed.*\n\n"); return(invisible()) }
  cat(sprintf("- Network-structure invariance (M) p = %.3f\n", res$nwinv.pval))
  cat(sprintf("- Global-strength invariance (S) p = %.3f\n\n", res$glstrinv.pval))
}
if (RUN_NCT) {
  nodes <- NET_SUBCLASS_NODES
  for (md in DATA_MODES) {
    cat("### Mode: `", md, "`\n\n", sep = "")
    sx <- strata_by_mode[[md]]
    run_nct(nct_subset(sx[["TdaP"]]$data, nodes), nct_subset(sx[["TT"]]$data, nodes),
            sprintf("Maternal arm: TdaP vs TT @ %s", CONC_VISIT))
    for (m in cfg$maternal_arms) {
      ka <- paste0(m, "_aP"); kw <- paste0(m, "_wP")
      if (!is.null(sx[[ka]]) && !is.null(sx[[kw]]))
        run_nct(nct_subset(sx[[ka]]$data, nodes), nct_subset(sx[[kw]]$data, nodes),
                sprintf("Infant arm within %s: aP vs wP @ %s", m, CONC_VISIT))
    }
  }
} else cat("*NCT disabled (RUN_NCT = FALSE or packages absent).*\n\n")
.mark("stage5_nct")
```

---

## Run Manifest and Participant Accounting {#concurrent-naccounting}

```{r conc-manifest, results='asis'}
message(sprintf("[%s | +%5.1fs] concurrent v2 DONE — endpoint=%s", .hms(), .el(), cfg$endpoint))
tm <- dplyr::bind_rows(.RUNLOG$rows)
if (nrow(tm)) {
  tm$stage_seconds <- round(c(tm$elapsed_s[1], diff(tm$elapsed_s)), 1)
  knitr::kable(tm, caption = "Stage timing (cumulative and per-stage seconds).",
               col.names = c("Step", "Cumulative (s)", "This stage (s)")) %>% print()
  cat("\n")
}
nt_render(nt, part = "Concurrent",
          caption = "Participant accounting for every concurrent model / network (both modes).")
```
