# Effector Function Associations with the Whole-Cell Outcome at INFMon2 {#effector-functions}

*🧭 Maps to summary: §11 (ADCD), §12 (ADCP), §13 (ADNP), §14 (combined profile)*

**Effector coverage note.** ADCD is measured for all six antigens (PT, PRN,
FHA, DT, TT, FIM). ADCP and ADNP are measured only for PT, PRN, DT, and TT;
FHA and FIM have ADCD only. Blocks 12 and 13 reflect this reduced antigen panel.

---

```{r p5-setup, include=FALSE}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2)
  library(purrr); library(tibble); library(forcats)
})

# ── Standalone bootstrap ─────────────────────────────────────────────────────
if (!exists("cfg")) {
  .find <- function(rel) {
    cand <- character(0)
    if (requireNamespace("here", quietly = TRUE))
      cand <- c(here::here(rel), here::here("analysis", rel))
    rmd_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                        error = function(e) NA_character_)
    if (!is.na(rmd_dir))
      cand <- c(cand, file.path(rmd_dir, rel), file.path(dirname(rmd_dir), rel))
    cand <- unique(c(cand, file.path(getwd(), rel), rel))
    hit  <- cand[file.exists(cand)]
    if (length(hit)) return(normalizePath(hit[1], winslash="/", mustWork=FALSE))
    stop("Cannot locate '", rel, "'. Run from the project root.", call.=FALSE)
  }
  source(.find("config/endpoints.R"),           local=FALSE, encoding="UTF-8")
  tryCatch(
    source(.find("config/endpoints_additions.R"), local=FALSE, encoding="UTF-8"),
    error = function(e) NULL
  )
  source(.find("R/C1_serology_helpers.R"),          local=FALSE, encoding="UTF-8")
  cfg  <- get_endpoint_config("SBA")  # <- change to "WT_IgG"/"PTNA" to test
  data <- load_serology_data(.find("data/d_set.RData"))
  nt   <- new_ntracker()
}
if (!exists("nt_add")) nt_add <- function(...) invisible(NULL)
if (!exists("nt"))     nt     <- list()
if (!exists("make_wide")) {
  make_wide <- function(data_long, visit_label, features = NULL,
                        visit_col="visit_name", feature_col="feature",
                        value_col="log_assay_value",
                        id_cols=c("subject_accession","maternal_arm","infant_arm")) {
    if (is.null(visit_label)||length(visit_label)==0L||!nzchar(visit_label))
      stop("visit_label is NULL/empty.", call.=FALSE)
    df <- dplyr::filter(data_long, .data[[visit_col]] == visit_label)
    if (!is.null(features))
      df <- dplyr::filter(df, .data[[feature_col]] %in% features)
    tidyr::pivot_wider(df, id_cols=dplyr::all_of(id_cols),
      names_from=dplyr::all_of(feature_col),
      values_from=dplyr::all_of(value_col), values_fn=mean)
  }
}

# ── Variable-name bridge ──────────────────────────────────────────────────────
# The parent .Rmd loads data as 'data_raw'; child parts use 'data'.
# When called via child= the bootstrap is skipped (cfg already exists), so
# 'data' may not have been created. Alias data_raw → data when needed.
if (!exists("data") || !is.data.frame(data)) {
  if (exists("data_raw") && is.data.frame(data_raw)) {
    data <- data_raw
  }
}

# ── Feature groups ────────────────────────────────────────────────────────────
if (!exists("ANTIGENS"))
  ANTIGENS <- c("PT", "PRN", "FHA", "DT", "TT", "FIM")

ANTIGENS_FULL_EFF  <- c("PT", "PRN", "DT", "TT")   # ADCD + ADCP + ADNP
ANTIGENS_ADCD_ONLY <- c("FHA", "FIM")               # ADCD only

ADCD_FEATS      <- paste0(ANTIGENS,           "_ADCD")
ADCP_FEATS      <- paste0(ANTIGENS_FULL_EFF,  "_ADCP")
ADNP_FEATS      <- paste0(ANTIGENS_FULL_EFF,  "_ADNP")
ALL_EFF_FEATS   <- c(ADCD_FEATS, ADCP_FEATS, ADNP_FEATS)

OUTCOME_COL    <- cfg$outcome_feature
.derive_visit <- function(cfg_val, raw_code, field_name) {
  # 1. Explicit config value wins
  if (!is.null(cfg_val) && length(cfg_val) > 0L && nzchar(cfg_val)) return(cfg_val)

  # 2. Inspect actual visit labels in the data.
  #    Guard with is.data.frame() — base R also has a data() *function* so
  #    exists("data") alone is not sufficient to detect the data frame.
  visits_in_data <- if (exists("data") && is.data.frame(data) &&
                        "visit_name" %in% names(data))
                      as.character(unique(data$visit_name))
                    else character(0)

  # 3. Prefer raw code if it is directly in the data (no VISIT_RECODE needed)
  if (raw_code %in% visits_in_data) {
    message("[Part 5] cfg$", field_name, " not set — using raw code '",
            raw_code, "' (found in data$visit_name).\n",
            "  Add ", field_name, " = \"", raw_code,
            "\" to config/endpoints.R to suppress this.")
    return(raw_code)
  }

  # 4. Try VISIT_RECODE only when the raw code is NOT directly in the data
  if (exists("VISIT_RECODE") && !is.na(VISIT_RECODE[raw_code])) {
    recoded <- unname(VISIT_RECODE[raw_code])
    if (length(visits_in_data) == 0L || recoded %in% visits_in_data) {
      message("[Part 5] cfg$", field_name, " not set — using VISIT_RECODE label '",
              recoded, "'.\n",
              "  Add ", field_name, " = \"", recoded,
              "\" to config/endpoints.R to suppress this.")
      return(recoded)
    }
    warning("[Part 5] VISIT_RECODE[\"", raw_code, "\"] = \"", recoded,
            "\" is not in data$visit_name either. Falling back to raw code.")
  }

  message("[Part 5] cfg$", field_name, " not set — using raw code '", raw_code,
          "' as last-resort fallback.")
  raw_code
}
INFMON2_VISIT  <- .derive_visit(cfg$infmon2_visit,       "P02", "infmon2_visit")
MATERNAL_VISIT <- .derive_visit(cfg$maternal_pred_visit, "M09", "maternal_pred_visit")
# ↑ ADCD/ADCP/ADNP are only measured at maternal visits; M09 is the default.
#   Set cfg$maternal_pred_visit in config/endpoints.R to override.

# ── Two wide frames ───────────────────────────────────────────────────────────
# P02 wide — needed only for the outcome column
if (!exists("data_infmon2") || !is.data.frame(data_infmon2) ||
    !("subject_accession" %in% names(data_infmon2))) {
  data_infmon2 <- make_wide(data, visit_label = INFMON2_VISIT)
}

# Maternal wide — effectors are here (not at P02)
if (!exists("data_maternal") || !is.data.frame(data_maternal) ||
    !("subject_accession" %in% names(data_maternal))) {
  data_maternal <- make_wide(data, visit_label = MATERNAL_VISIT)
}

# Joined frame: maternal effectors + P02 outcome
if (!exists("data_mat_p02") || !is.data.frame(data_mat_p02) ||
    nrow(data_mat_p02) == 0L) {
  if (isTRUE(OUTCOME_COL %in% names(data_infmon2))) {
    .p02_out <- dplyr::select(data_infmon2, subject_accession,
                               dplyr::all_of(OUTCOME_COL))
    data_mat_p02 <- dplyr::inner_join(data_maternal, .p02_out,
                                      by = "subject_accession")
  } else {
    warning("[Part 5] Outcome '", OUTCOME_COL, "' not found at ", INFMON2_VISIT)
    data_mat_p02 <- data.frame()
  }
}

# ── Availability guards ───────────────────────────────────────────────────────
# Effectors come from the maternal frame
avail_adcd <- intersect(ADCD_FEATS,    names(data_maternal))
avail_adcp <- intersect(ADCP_FEATS,    names(data_maternal))
avail_adnp <- intersect(ADNP_FEATS,    names(data_maternal))
avail_eff  <- c(avail_adcd, avail_adcp, avail_adnp)

.warn_miss <- function(req, avail, lbl)
  if (length(m <- setdiff(req, avail)))
    message("[Part 5 | ", lbl, "] Missing: ",
            paste(m[seq_len(min(6L, length(m)))], collapse = ", "),
            if (length(m) > 6L) " ...")

.warn_miss(ADCD_FEATS,  avail_adcd, "ADCD")
.warn_miss(ADCP_FEATS,  avail_adcp, "ADCP")
.warn_miss(ADNP_FEATS,  avail_adnp, "ADNP")
```

## Effector Feature Inventory {#p5-inventory}

```{r p5-inventory-tbl, results='asis'}
tibble::tibble(
  Effector  = c("ADCD", "ADCP", "ADNP"),
  `Antigens requested` = c(
    paste(ANTIGENS,          collapse = ", "),
    paste(ANTIGENS_FULL_EFF, collapse = ", "),
    paste(ANTIGENS_FULL_EFF, collapse = ", ")
  ),
  `Features found` = c(length(avail_adcd), length(avail_adcp), length(avail_adnp)),
  Note = c(
    "Antibody-Dependent Complement Deposition — all 6 antigens",
    "Antibody-Dependent Cellular Phagocytosis — PT/PRN/DT/TT only",
    "Antibody-Dependent Neutrophil Phagocytosis — PT/PRN/DT/TT only"
  )
) |>
  knitr::kable(caption = paste("Part 5 effector feature availability for",
                               cfg$outcome_label, "at INFMon2."))
```

---

## Block 11: ADCD Associations {#block11}

*🧭 Maps to summary §11 — complement deposition (all 6 antigens)*

```{r block11-adcd, fig.width=9, fig.height=4.5, fig.cap="Block 11 — ADCD Spearman ρ vs. outcome by antigen and maternal arm."}
if (length(avail_adcd) == 0L) {
  message("Block 11 skipped: no ADCD features found.")
} else {

  cor11 <- data_mat_p02 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_adcd),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      n = n(), .groups = "drop"
    ) |>
    tidyr::pivot_longer(dplyr::all_of(avail_adcd), names_to = "feature", values_to = "rho") |>
    dplyr::mutate(antigen = factor(sub("_ADCD$", "", feature), levels = ANTIGENS))

  ggplot(cor11,
         aes(x = fct_reorder(antigen, abs(rho)), y = rho, fill = maternal_arm)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
    scale_fill_manual(values = c("TdaP" = "#e08214", "TT" = "#542788"),
                      name = "Maternal arm") +
    coord_flip() +
    labs(title    = paste("Block 11 — ADCD vs.", cfg$outcome_label, "at INFMon2"),
         subtitle = "Spearman ρ (log_assay_value); by maternal arm — exploratory",
         x = "Antigen", y = expression(rho[s])) +
    theme_minimal(base_size = 11)
}
```

---

## Block 12: ADCP Associations {#block12}

*🧭 Maps to summary §12 — cellular phagocytosis (PT / PRN / DT / TT)*

```{r block12-adcp, fig.width=8, fig.height=4, fig.cap="Block 12 — ADCP Spearman ρ. Note: FHA and FIM not measured for ADCP."}
if (length(avail_adcp) == 0L) {
  message("Block 12 skipped: no ADCP features found.")
} else {

  cor12 <- data_mat_p02 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_adcp),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(dplyr::all_of(avail_adcp), names_to = "feature", values_to = "rho") |>
    dplyr::mutate(antigen = factor(sub("_ADCP$", "", feature), levels = ANTIGENS_FULL_EFF))

  ggplot(cor12,
         aes(x = fct_reorder(antigen, abs(rho)), y = rho, fill = maternal_arm)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
    scale_fill_manual(values = c("TdaP" = "#e08214", "TT" = "#542788"),
                      name = "Maternal arm") +
    coord_flip() +
    labs(title    = paste("Block 12 — ADCP vs.", cfg$outcome_label, "at INFMon2"),
         subtitle = "Spearman ρ; PT/PRN/DT/TT only (FHA and FIM not measured)",
         x = "Antigen", y = expression(rho[s])) +
    theme_minimal(base_size = 11)
}
```

---

## Block 13: ADNP Associations {#block13}

*🧭 Maps to summary §13 — neutrophil phagocytosis (PT / PRN / DT / TT)*

```{r block13-adnp, fig.width=8, fig.height=4, fig.cap="Block 13 — ADNP Spearman ρ. Note: FHA and FIM not measured for ADNP."}
if (length(avail_adnp) == 0L) {
  message("Block 13 skipped: no ADNP features found.")
} else {

  cor13 <- data_mat_p02 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_adnp),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(dplyr::all_of(avail_adnp), names_to = "feature", values_to = "rho") |>
    dplyr::mutate(antigen = factor(sub("_ADNP$", "", feature), levels = ANTIGENS_FULL_EFF))

  ggplot(cor13,
         aes(x = fct_reorder(antigen, abs(rho)), y = rho, fill = maternal_arm)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
    scale_fill_manual(values = c("TdaP" = "#e08214", "TT" = "#542788"),
                      name = "Maternal arm") +
    coord_flip() +
    labs(title    = paste("Block 13 — ADNP vs.", cfg$outcome_label, "at INFMon2"),
         subtitle = "Spearman ρ; PT/PRN/DT/TT only (FHA and FIM not measured)",
         x = "Antigen", y = expression(rho[s])) +
    theme_minimal(base_size = 11)
}
```

---

## Block 14: Combined Effector Profile {#block14}

### Block 14a: Effector × antigen heatmap {#block14a}

*🧭 Maps to summary §14 — effector hierarchy and whole-cell alignment*

Placing all available effector–antigen combinations in one view reveals whether the
outcome aligns with an **effector mechanism** (row pattern) or an **antigen target**
(column pattern). Grey tiles indicate feature not measured for that antigen.

```{r block14a-heatmap, fig.width=9, fig.height=5, fig.cap="Block 14a — All effector functions vs. outcome. Grey = not measured."}
if (length(avail_eff) == 0L) {
  message("Block 14 skipped: no effector features available.")
} else {

  # Build a complete antigen × effector grid; fill missing with NA
  all_combos <- expand.grid(
    antigen  = ANTIGENS,
    effector = c("ADCD", "ADCP", "ADNP"),
    stringsAsFactors = FALSE
  ) |>
    dplyr::mutate(feature = paste(antigen, effector, sep = "_"))

  # Correlations for available features only
  cor14_present <- data_mat_p02 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_eff),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(-maternal_arm, names_to = "feature", values_to = "rho")

  # Join onto complete grid so FHA/FIM ADCP/ADNP show as NA
  cor14 <- all_combos |>
    dplyr::left_join(cor14_present, by = "feature") |>
    dplyr::mutate(
      antigen  = factor(antigen,  levels = ANTIGENS),
      effector = factor(effector, levels = c("ADCD", "ADCP", "ADNP"))
    )

  ggplot(cor14, aes(x = antigen, y = fct_rev(effector), fill = rho)) +
    geom_tile(colour = "white", linewidth = 0.7) +
    geom_text(aes(label = ifelse(is.na(rho), "—", sprintf("%.2f", rho))),
              size = 3.2) +
    scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#d6604d",
                         midpoint = 0, limits = c(-1, 1), na.value = "grey88",
                         name = expression(rho[s])) +
    facet_wrap(~ maternal_arm, labeller = label_both) +
    labs(
      title    = paste("Block 14a — Effector functions vs.", cfg$outcome_label,
                       "at INFMon2"),
      subtitle = "Spearman ρ; grey tiles = not measured in this dataset",
      x = "Antigen", y = "Effector function"
    ) +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank())
}
```

### Block 14b: Elastic-net importance across all effector features {#block14b}

```{r block14b-elasticnet, fig.width=9, fig.height=5, fig.cap="Block 14b — Elastic-net coefficients; colour = effector type."}
if (length(avail_eff) >= 3L && requireNamespace("glmnet", quietly = TRUE)) {
  library(glmnet)

  df14 <- data_mat_p02 |>
    dplyr::select(dplyr::all_of(c(OUTCOME_COL, avail_eff))) |>
    tidyr::drop_na()

  nt_add(nt, "Block 14b — effector importance complete cases", nrow(df14))

  set.seed(20240202)
  cv14   <- glmnet::cv.glmnet(as.matrix(df14[, avail_eff]),
                               df14[[OUTCOME_COL]],
                               alpha = 0.5, nfolds = 5)
  coef14 <- as.matrix(coef(cv14, s = "lambda.1se")) |>
    as.data.frame() |>
    tibble::rownames_to_column("feature") |>
    setNames(c("feature", "coef")) |>
    dplyr::filter(feature != "(Intercept)", coef != 0) |>
    dplyr::arrange(desc(abs(coef))) |>
    dplyr::mutate(effector = factor(sub(".*_(ADCD|ADCP|ADNP)$", "\\1", feature),
                             levels = c("ADCD", "ADCP", "ADNP")))

  if (nrow(coef14) == 0L) {
    message("Block 14b: all effector features shrunk to zero at lambda.1se.")
  } else {
    ggplot(coef14,
           aes(x = fct_reorder(feature, abs(coef)), y = coef,
               fill = effector, alpha = coef > 0)) +
      geom_col() +
      coord_flip() +
      scale_fill_manual(
        values = c("ADCD" = "#4dac26", "ADCP" = "#d01c8b", "ADNP" = "#f1a340"),
        name = "Effector") +
      scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.6), guide = "none") +
      labs(
        title    = paste("Block 14b — Effector elastic-net vs.", cfg$outcome_label),
        subtitle = paste0("λ.1se; ", nrow(coef14), " non-zero features",
                          "; N = ", nrow(df14)),
        x = NULL, y = "Coefficient"
      ) +
      theme_minimal(base_size = 11)
  }
} else if (length(avail_eff) < 3L) {
  message("Block 14b: fewer than 3 effector features available — skipped.")
} else {
  message("Install package 'glmnet' for Block 14b.")
}
```

### Block 14c: Full BH-corrected ρ table {#block14c}

*All antigen × effector Spearman ρ values, pooled across arms, with
Benjamini–Hochberg FDR correction.*

```{r block14c-fdr, results='asis'}
if (length(avail_eff) > 0L) {

  df_all <- data_mat_p02 |> dplyr::filter(!is.na(.data[[OUTCOME_COL]]))

  fdr_tbl <- map_dfr(avail_eff, function(feat) {
    ct <- cor.test(df_all[[feat]], df_all[[OUTCOME_COL]],
                   method = "spearman", exact = FALSE)
    tibble(feature = feat, rho = ct$estimate, p_value = ct$p.value,
           n = sum(!is.na(df_all[[feat]]) & !is.na(df_all[[OUTCOME_COL]])))
  }) |>
    dplyr::mutate(
      antigen  = factor(sub("_(ADCD|ADCP|ADNP)$", "", feature), levels = ANTIGENS),
      effector = factor(sub(".*_(ADCD|ADCP|ADNP)$", "\\1", feature),
                        levels = c("ADCD", "ADCP", "ADNP")),
      q_BH     = p.adjust(p_value, method = "BH")
    ) |>
    dplyr::arrange(q_BH) |>
    dplyr::select(Antigen = antigen, Effector = effector,
           `ρ` = rho, `p` = p_value, `q (BH)` = q_BH, `n` = n)

  knitr::kable(fdr_tbl, digits = 3,
    caption = paste("Block 14c — Effector × antigen Spearman ρ vs.",
                    cfg$outcome_label,
                    "at INFMon2 (pooled arms; BH-corrected)."))
}
```

---

## ADCD pathway & ADCD↔whole-cell relationship — SCAFFOLD (inert) {#adcd-scaffold}

*🧭 Requirement #7. This section is intentionally **inert**: the ADCD transition
data are not yet in hand, so every chunk below is `eval=FALSE`. The structure
mirrors the three within-subject transitions used for the antigen-specific pool
(production / transfer / decay) and adds the concurrent ADCD↔whole-cell-assay
relationship at InfMon2. When the data arrive, flip `ADCD_READY` to `TRUE`
(or remove `eval=FALSE`) and confirm ADCD is measured at the transition visits.*

```{r adcd-scaffold-setup, eval=FALSE}
# Readiness gate — keep FALSE until ADCD is confirmed at the transition visits.
# ADCD is present in d_set.RData for all six antigens, but the longitudinal
# transition readout (ADCD at PregEarly/MatBirth/CordBlood as well as InfMon2)
# must be verified before these run; until then the section is design-only.
ADCD_READY <- FALSE

# ADCD feature vocabulary (antigen_analyte). ADCD is assayed for all six
# antigens; ADCP/ADNP for PT/PRN/DT/TT only (see the coverage note above).
ADCD_FEATS <- paste0(ANTIGEN_ORDER, "_ADCD")

# Same three transitions as the antibody pathway (results_B_to_F / Figure 3).
ADCD_TRANSITIONS <- list(
  c(from = "PregEarly", to = "MatBirth",  label = "ADCD production (PregEarly\u2192MatBirth)"),
  c(from = "MatBirth",  to = "CordBlood", label = "ADCD transfer (MatBirth\u2192CordBlood)"),
  c(from = "CordBlood", to = "InfMon2",   label = "ADCD decay (CordBlood\u2192InfMon2)"))
```

```{r adcd-transition-pathway, eval=FALSE}
# Per-antigen ADCD change across each transition, reusing the shared paired_change /
# summarise_paired engine. Mirrors build_all_tr() in results_B_to_F.Rmd so the ADCD
# pathway can be plotted with the same plot_pathway() renderer once ADCD_READY.
adcd_all_tr <- dplyr::bind_rows(lapply(ADCD_TRANSITIONS, function(tr) {
  paired <- paired_change(data, tr["from"], tr["to"])           # 'data' = recoded long frame
  paired <- paired[paired$feature %in% ADCD_FEATS, , drop = FALSE]
  summarise_paired(paired, min_n = 10L) %>%
    dplyr::left_join(dplyr::distinct(paired, feature, antigen, analyte), by = "feature") %>%
    dplyr::mutate(transition = tr["label"], from = tr["from"], to = tr["to"])
}))
knitr::kable(adcd_all_tr, caption = "ADCD within-subject change per antigen across the three transitions (SCAFFOLD).")
```

```{r adcd-wholecell-infmon2, eval=FALSE}
# Concurrent ADCD <-> whole-cell-assay relationship at InfMon2: does per-antigen
# ADCD track the whole-bacterium functional outcome (PTNA / SBA / WT_IgG) on the
# same InfMon2 draw? Spearman per antigen, by maternal arm. cfg$outcome_feature
# selects the active whole-cell outcome (e.g. WHOLE_SBA).
infmon2 <- cfg$infmon2_visit %||% "InfMon2"
wide_adcd <- make_wide(data, visit_label = infmon2,
                       features = c(ADCD_FEATS, cfg$outcome_feature))
adcd_vs_whole <- tidyr::expand_grid(antigen = ANTIGEN_ORDER) %>%
  dplyr::mutate(feat = paste0(antigen, "_ADCD")) %>%
  dplyr::filter(feat %in% names(wide_adcd)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    rho = suppressWarnings(stats::cor(wide_adcd[[feat]],
                                      wide_adcd[[cfg$outcome_feature]],
                                      method = "spearman",
                                      use = "pairwise.complete.obs"))) %>%
  dplyr::ungroup()
knitr::kable(adcd_vs_whole, caption = paste0("Concurrent InfMon2 ADCD vs ",
             cfg$outcome_label, " (Spearman; SCAFFOLD)."))
```
