# Antigen-Specific IgG Subclass & FcG Predictors at INFMon2 {#antigen-subclass}

*🧭 Maps to summary: §8 (IgG subclass profiles), §9 (FcG secondary binding),
§10 (cord blood pre-INFMon2 predictor — WT_IgG only)*

This part asks whether **antigen identity** (PT / PRN / FHA / DT / TT / FIM),
**IgG subclass balance** (IGG1–IGG4), and **Fc-receptor binding profile**
(FCGR2A / FCGR3B) carry information about the endpoint outcome beyond the bulk
quantities examined in Part 1. All features are measured concurrently at INFMon2
and are already on the log scale (`log_assay_value`). For **WT_IgG only**,
Block 10 adds the cord blood level as a genuine temporal predictor.

---

```{r p4-setup, include=FALSE}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2)
  library(purrr); library(tibble); library(forcats)
})

# ── Standalone bootstrap ──────────────────────────────────────────────────────
# When called via child= the parent already has cfg/data/nt/make_wide in scope
# and this entire block is skipped. When knitted standalone it sources the
# pipeline so the rest of the chunk can run.
if (!exists("cfg")) {
  .find <- function(rel) {
    cand <- character(0)
    if (requireNamespace("here", quietly = TRUE))
      cand <- c(here::here(rel), here::here("analysis", rel))
    rmd_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                        error = function(e) NA_character_)
    if (!is.na(rmd_dir))
      cand <- c(cand, file.path(rmd_dir, rel), file.path(dirname(rmd_dir), rel))
    cand <- unique(c(cand, file.path(getwd(), rel), rel))
    hit  <- cand[file.exists(cand)]
    if (length(hit)) return(normalizePath(hit[1], winslash="/", mustWork=FALSE))
    stop("Cannot locate '", rel, "'. Run from project root.", call.=FALSE)
  }
  source(.find("config/endpoints.R"),           local=FALSE, encoding="UTF-8")
  tryCatch(
    source(.find("config/endpoints_additions.R"), local=FALSE, encoding="UTF-8"),
    error = function(e) NULL
  )
  source(.find("R/serology_helpers.R"),          local=FALSE, encoding="UTF-8")
  cfg  <- get_endpoint_config("SBA")  # <- change to "WT_IgG" or "PTNA" to test
  data <- load_serology_data(.find("data/c_set.RData"))
  nt   <- new_ntracker()
}
if (!exists("nt_add"))   nt_add   <- function(...) invisible(NULL)
if (!exists("nt"))       nt       <- list()
if (!exists("make_wide")) {
  make_wide <- function(data_long, visit_label, features=NULL,
                        visit_col="visit_name", feature_col="feature",
                        value_col="log_assay_value",
                        id_cols=c("subject_accession","maternal_arm","infant_arm")) {
    if (is.null(visit_label)||length(visit_label)==0L||!nzchar(visit_label))
      stop("visit_label is NULL/empty — check cfg$infmon2_visit.", call.=FALSE)
    df <- dplyr::filter(data_long, .data[[visit_col]] == visit_label)
    if (!is.null(features))
      df <- dplyr::filter(df, .data[[feature_col]] %in% features)
    tidyr::pivot_wider(df,
      id_cols=dplyr::all_of(id_cols),
      names_from=dplyr::all_of(feature_col),
      values_from=dplyr::all_of(value_col),
      values_fn=mean)
  }
}

# ── Variable-name bridge ──────────────────────────────────────────────────────
# The parent .Rmd loads data as 'data_raw'; child parts use 'data'.
# When called via child= the bootstrap is skipped (cfg already exists), so
# 'data' may not have been created. Alias data_raw → data when needed.
if (!exists("data") || !is.data.frame(data)) {
  if (exists("data_raw") && is.data.frame(data_raw)) {
    data <- data_raw
  }
}

# ── Feature groups (from config/endpoints_additions.R) -----------------------
# Inline fallbacks used if config hasn't been updated yet.
if (!exists("ANTIGENS"))
  ANTIGENS <- c("PT", "PRN", "FHA", "DT", "TT", "FIM")
if (!exists("SUBCLASSES"))
  SUBCLASSES <- c("IGG1", "IGG2", "IGG3", "IGG4")
if (!exists("ANTIGEN_SUBCLASS_FEATS"))
  ANTIGEN_SUBCLASS_FEATS <- as.vector(outer(ANTIGENS, SUBCLASSES, paste, sep = "_"))
if (!exists("FCG_RECEPTORS"))
  FCG_RECEPTORS <- c("FCGR2A", "FCGR3B")
if (!exists("FCG_FEATS"))
  FCG_FEATS <- as.vector(outer(ANTIGENS, FCG_RECEPTORS, paste, sep = "_"))

# ── Endpoint config -----------------------------------------------------------
# Fields marked ** are new additions from endpoints_additions.R.
# Until config/endpoints.R is updated, the fallbacks below auto-derive the
# visit labels from VISIT_RECODE and emit an explanatory message.
OUTCOME_COL <- cfg$outcome_feature        # e.g. "WHOLE_SBA"
CORD_COL    <- cfg$cord_predictor         # "WHOLE_WT_IgG" or NULL  **
CORD_LABEL  <- cfg$cord_predictor_label   # display string or NULL   **

# ** INFMON2_VISIT: prefer cfg$infmon2_visit; fall back to VISIT_RECODE["P02"].
.derive_visit <- function(cfg_val, raw_code, field_name) {
  # 1. Explicit config value wins
  if (!is.null(cfg_val) && length(cfg_val) > 0L && nzchar(cfg_val)) return(cfg_val)

  # 2. Inspect actual visit labels in the data.
  #    Guard with is.data.frame() — base R also has a data() *function* so
  #    exists("data") alone is not sufficient to detect the data frame.
  visits_in_data <- if (exists("data") && is.data.frame(data) &&
                        "visit_name" %in% names(data))
                      as.character(unique(data$visit_name))
                    else character(0)

  # 3. Prefer raw code if it is directly in the data (no VISIT_RECODE needed)
  if (raw_code %in% visits_in_data) {
    message("[Part 4] cfg$", field_name, " not set — using raw code '",
            raw_code, "' (found in data$visit_name).\n",
            "  Add ", field_name, " = \"", raw_code,
            "\" to config/endpoints.R to suppress this.")
    return(raw_code)
  }

  # 4. Try VISIT_RECODE only when the raw code is NOT directly in the data
  if (exists("VISIT_RECODE") && !is.na(VISIT_RECODE[raw_code])) {
    recoded <- unname(VISIT_RECODE[raw_code])
    if (length(visits_in_data) == 0L || recoded %in% visits_in_data) {
      message("[Part 4] cfg$", field_name, " not set — using VISIT_RECODE label '",
              recoded, "'.\n",
              "  Add ", field_name, " = \"", recoded,
              "\" to config/endpoints.R to suppress this.")
      return(recoded)
    }
    warning("[Part 4] VISIT_RECODE[\"", raw_code, "\"] = \"", recoded,
            "\" is not in data$visit_name either. Falling back to raw code.")
  }

  message("[Part 4] cfg$", field_name, " not set — using raw code '", raw_code,
          "' as last-resort fallback.")
  raw_code
}
INFMON2_VISIT  <- .derive_visit(cfg$infmon2_visit,       "P02", "infmon2_visit")
CORD_VISIT     <- .derive_visit(cfg$cord_visit,          "M09", "cord_visit")
MATERNAL_VISIT <- .derive_visit(cfg$maternal_pred_visit, "M09", "maternal_pred_visit")
# ↑ FCGR2A/FCGR3B and all effectors are only measured at maternal visits;
#   M09 is a near-term ante-partum default.  Set cfg$maternal_pred_visit
#   (and cfg$cord_visit) to the correct visit code in config/endpoints.R.

# ── Two wide frames ───────────────────────────────────────────────────────────
# (a) P02 — IgG subclasses (IGG1-4) + outcome: concurrent analysis.
#     FIM_IGG1-4 are zero at P02 so will be auto-excluded by intersect().
data_infmon2 <- make_wide(data, visit_label = INFMON2_VISIT)

# (b) Maternal visit — FcG + effectors: longitudinal maternal→infant analysis.
#     FCGR2A/FCGR3B/ADCD/ADCP/ADNP are only measured at maternal visits.
data_maternal <- make_wide(data, visit_label = MATERNAL_VISIT)

# Join maternal predictors with the P02 outcome for Blocks 9, 11-14.
if (isTRUE(OUTCOME_COL %in% names(data_infmon2))) {
  .p02_out <- dplyr::select(data_infmon2, subject_accession,
                             dplyr::all_of(OUTCOME_COL))
  data_mat_p02 <- dplyr::inner_join(data_maternal, .p02_out,
                                    by = "subject_accession")
} else {
  warning("[Part 4] Outcome column '", OUTCOME_COL, "' not found at visit '",
          INFMON2_VISIT, "'. Blocks 9/10 and Part 5 will be skipped.\n",
          "  Check cfg$outcome_feature in config/endpoints.R.")
  data_mat_p02 <- data.frame()
}

# ── Availability guards ───────────────────────────────────────────────────────
# IgG subclasses from P02; FcG from maternal visit
avail_ant_sub <- intersect(ANTIGEN_SUBCLASS_FEATS, names(data_infmon2))
avail_fcg     <- intersect(FCG_FEATS,              names(data_maternal))
avail_cord    <- if (!is.null(CORD_COL)) CORD_COL else NULL

.warn_missing <- function(req, avail, lbl) {
  miss <- setdiff(req, avail)
  if (length(miss))
    message("[Part 4 | ", lbl, "] Missing from data: ",
            paste(miss[seq_len(min(8L, length(miss)))], collapse = ", "),
            if (length(miss) > 8L) " ...")
}
.warn_missing(ANTIGEN_SUBCLASS_FEATS, avail_ant_sub, "IgG subclass (P02)")
.warn_missing(FCG_FEATS,              avail_fcg,     "FcG binding (maternal)")

nt_add(nt, "Part 4 — P02 participants (wide)",       nrow(data_infmon2))
nt_add(nt, "Part 4 — maternal participants (wide)",  nrow(data_maternal))
nt_add(nt, "Part 4 — maternal+P02 joined",           nrow(data_mat_p02))
```

## Feature Inventory {#p4-inventory}

```{r p4-inventory-tbl, results='asis'}
tibble::tibble(
  `Predictor group`        = c("Antigen × IgG subclass (Blocks 7–8)",
                               "FcG binding (Block 9)",
                               "Cord blood pre-predictor (Block 10)"),
  `Requested`              = c(length(ANTIGEN_SUBCLASS_FEATS),
                                length(FCG_FEATS),
                                as.integer(!is.null(CORD_COL))),
  `Found in data`          = c(length(avail_ant_sub),
                                length(avail_fcg),
                                as.integer(!is.null(avail_cord))),
  `Endpoint applicability` = c("SBA / WT_IgG / PTNA",
                                "SBA / WT_IgG / PTNA",
                                ifelse(!is.null(CORD_COL),
                                       cfg$outcome_label,
                                       "WT_IgG only — N/A here"))
) |>
  knitr::kable(caption = paste("Part 4 predictor availability for",
                               cfg$outcome_label, "at INFMon2."))
```

---

## Block 7: Antigen-Specific IgG Subclass Associations {#block7}

*🧭 Maps to summary §8a — IgG subclass quantity by antigen*

Spearman ρ between each antigen × IgG subclass feature and `r cfg$outcome_label`
at INFMon2. Values are on the log scale from `log_assay_value`; no further
transformation applied. Stratified by maternal vaccination arm.

```{r block7-heatmap, fig.width=10, fig.height=6, fig.cap="Block 7 — IgG subclass Spearman ρ. Rows = antigens, columns = subclasses, panels = maternal arm."}
if (length(avail_ant_sub) == 0L) {
  message("Block 7 skipped: no antigen-subclass features in data.")
} else {

  cor7 <- data_infmon2 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_ant_sub),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      n = n(), .groups = "drop"
    ) |>
    tidyr::pivot_longer(dplyr::all_of(avail_ant_sub),
                 names_to = "feature", values_to = "rho") |>
    dplyr::mutate(
      antigen  = factor(sub("_IGG[0-9]$", "", feature), levels = ANTIGENS),
      subclass = factor(sub(".*_(IGG[0-9])$", "\\1", feature), levels = SUBCLASSES)
    )

  nt_add(nt, "Block 7 — n per maternal arm",
         paste(cor7 |> dplyr::distinct(maternal_arm, n) |> dplyr::pull(n), collapse = " / "))

  ggplot(cor7, aes(x = subclass, y = fct_rev(antigen), fill = rho)) +
    geom_tile(colour = "white", linewidth = 0.6) +
    geom_text(aes(label = sprintf("%.2f", rho)), size = 3.2) +
    scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#d6604d",
                         midpoint = 0, limits = c(-1, 1),
                         name = expression(rho[s])) +
    facet_wrap(~ maternal_arm, labeller = label_both) +
    labs(
      title    = paste("Block 7 —", cfg$outcome_label,
                       "vs. antigen-specific IgG subclasses at INFMon2"),
      subtitle = "Spearman ρ on log_assay_value; by maternal arm",
      x = "IgG subclass", y = "Antigen"
    ) +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank())
}
```

### Block 7b: Stratified by infant priming arm {#block7b}

```{r block7b-heatmap, fig.width=12, fig.height=7, fig.cap="Block 7b — IgG subclass ρ by maternal AND infant priming arm."}
if (length(avail_ant_sub) > 0L) {
  cor7b <- data_infmon2 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm, infant_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_ant_sub),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(dplyr::all_of(avail_ant_sub),
                 names_to = "feature", values_to = "rho") |>
    dplyr::mutate(
      antigen  = factor(sub("_IGG[0-9]$", "", feature), levels = ANTIGENS),
      subclass = factor(sub(".*_(IGG[0-9])$", "\\1", feature), levels = SUBCLASSES),
      arm_lbl  = paste0(maternal_arm, " / ", infant_arm)
    )

  ggplot(cor7b, aes(x = subclass, y = fct_rev(antigen), fill = rho)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = sprintf("%.2f", rho)), size = 2.7) +
    scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#d6604d",
                         midpoint = 0, limits = c(-1, 1),
                         name = expression(rho[s])) +
    facet_wrap(~ arm_lbl, nrow = 2) +
    labs(
      title    = paste("Block 7b —", cfg$outcome_label, "vs. IgG subclasses"),
      subtitle = "Stratified by maternal arm (TdaP/TT) × infant priming (aP/wP)",
      x = "IgG subclass", y = "Antigen"
    ) +
    theme_minimal(base_size = 10) +
    theme(panel.grid = element_blank())
}
```

---

## Block 8: IgG Subclass Ratio (Complement-Activating Class Bias) {#block8}

*🧭 Maps to summary §8b — qualitative subclass layer independent of quantity*

The ratio IGG1 / (IGG1 + IGG3) per antigen captures complement-activating subclass
bias. Because `log_assay_value` is on the log scale, we back-transform to linear
before computing the proportion; the resulting ratio (0–1) is not re-logged.

```{r block8-ratio, fig.width=9, fig.height=5, fig.cap="Block 8 — IGG1/(IGG1+IGG3) per antigen vs. outcome; Spearman ρ by maternal arm."}
antigens_with_13 <- ANTIGENS[
  paste0(ANTIGENS, "_IGG1") %in% avail_ant_sub &
  paste0(ANTIGENS, "_IGG3") %in% avail_ant_sub
]

if (length(antigens_with_13) == 0L) {
  message("Block 8 skipped: IGG1 and/or IGG3 absent for all antigens.")
} else {

  # Back-transform log values → linear, compute ratio, store as new columns
  for (ag in antigens_with_13) {
    v1 <- 10 ^ data_infmon2[[paste0(ag, "_IGG1")]]
    v3 <- 10 ^ data_infmon2[[paste0(ag, "_IGG3")]]
    denom <- v1 + v3
    data_infmon2[[paste0(ag, "_RATIO13")]] <-
      ifelse(is.finite(denom) & denom > 0, v1 / denom, NA_real_)
  }

  ratio_cols <- paste0(antigens_with_13, "_RATIO13")

  cor8 <- data_infmon2 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(ratio_cols),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(-maternal_arm, names_to = "feature", values_to = "rho") |>
    dplyr::mutate(antigen = factor(sub("_RATIO13$", "", feature),
                            levels = antigens_with_13))

  ggplot(cor8, aes(x = fct_reorder(antigen, abs(rho)), y = rho,
                   fill = maternal_arm)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
    scale_fill_manual(values = c("TdaP" = "#e08214", "TT" = "#542788"),
                      name = "Maternal arm") +
    coord_flip() +
    labs(
      title    = paste("Block 8 — IGG1/(IGG1+IGG3) ratio vs.", cfg$outcome_label),
      subtitle = "Spearman ρ at INFMon2 by maternal arm (exploratory, uncorrected)",
      x = "Antigen", y = expression(rho[s])
    ) +
    theme_minimal(base_size = 11)
}
```

---

## Block 9: FcG Receptor Binding Associations {#block9}

*🧭 Maps to summary §9 — Fc-receptor quality layer*

Two Fc-receptor binding features are measured in this dataset: **FCGR2A** and
**FCGR3B**. These capture Fc-mediated effector recruitment capacity beyond
IgG subclass quantity.

```{r block9-heatmap, fig.width=8, fig.height=5, fig.cap="Block 9 — FcG binding Spearman ρ. Rows = antigens, columns = receptors, panels = maternal arm."}
if (length(avail_fcg) == 0L) {
  message("Block 9 skipped: no FcG features in data_maternal.")
} else if (nrow(data_mat_p02) == 0L) {
  message("Block 9 skipped: data_mat_p02 is empty — outcome not found at P02.")
} else {

  cor9 <- data_mat_p02 |>
    dplyr::filter(!is.na(.data[[OUTCOME_COL]])) |>
    dplyr::group_by(maternal_arm) |>
    dplyr::summarise(
      dplyr::across(dplyr::all_of(avail_fcg),
             ~ cor(.x, .data[[OUTCOME_COL]],
                   method = "spearman", use = "pairwise.complete.obs"),
             .names = "{.col}"),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(-maternal_arm, names_to = "feature", values_to = "rho") |>
    dplyr::mutate(
      antigen  = factor(sub("_(FCGR2A|FCGR3B)$", "", feature), levels = ANTIGENS),
      receptor = sub(".*_(FCGR2A|FCGR3B)$", "\\1", feature)
    )

  ggplot(cor9, aes(x = receptor, y = fct_rev(antigen), fill = rho)) +
    geom_tile(colour = "white", linewidth = 0.7) +
    geom_text(aes(label = sprintf("%.2f", rho)), size = 3.4) +
    scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#d6604d",
                         midpoint = 0, limits = c(-1, 1),
                         name = expression(rho[s])) +
    facet_wrap(~ maternal_arm, labeller = label_both) +
    labs(
      title    = paste("Block 9 — FcG binding vs.", cfg$outcome_label, "at INFMon2"),
      subtitle = "Spearman ρ on log_assay_value; by maternal arm",
      x = "FcG receptor", y = "Antigen"
    ) +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank())
}
```

---

## Block 10: Cord Blood Pre-INFMon2 Predictor (WT_IgG only) {#block10}

```{r block10-header, results='asis', echo=FALSE}
if (is.null(avail_cord)) {
  cat("> **Block 10 not applicable for", cfg$outcome_label, ".**",
      "No cord-blood predictor is defined for this endpoint.",
      "Block 10 is active only for **WT_IgG**, where cord blood",
      "provides a genuine temporal predictor of the INFMon2 outcome.\n\n")
}
```

```{r block10-scatter, eval=!is.null(avail_cord) && !is.null(CORD_VISIT), fig.width=8.5, fig.height=5, fig.cap="Block 10 — Cord blood WHOLE_WT_IgG → INFMon2 WT_IgG; linear smoother per maternal arm."}
# Pivot cord blood visit to wide; extract the single cord predictor column.
cord_wide <- make_wide(data, visit_label = CORD_VISIT,
                       features = avail_cord)

cord_df <- cord_wide |>
  dplyr::select(subject_accession, maternal_arm, infant_arm,
         cord_val = dplyr::all_of(avail_cord)) |>
  dplyr::inner_join(
    data_infmon2 |>
      dplyr::select(subject_accession, outcome_val = dplyr::all_of(OUTCOME_COL)),
    by = "subject_accession"
  ) |>
  dplyr::filter(!is.na(cord_val), !is.na(outcome_val))

nt_add(nt, "Block 10 — cord × INFMon2 paired N", nrow(cord_df))

ggplot(cord_df,
       aes(x = cord_val, y = outcome_val, colour = maternal_arm)) +
  geom_point(alpha = 0.55, size = 1.9) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 0.9) +
  scale_colour_manual(values = c("TdaP" = "#e08214", "TT" = "#542788"),
                      name = "Maternal arm") +
  facet_wrap(~ infant_arm, labeller = label_both) +
  labs(
    title    = paste0("Block 10 — Cord blood ", CORD_LABEL,
                      " → INFMon2 ", cfg$outcome_label),
    subtitle = "log_assay_value scale; one point per participant; smoother ± SE",
    x        = paste0(CORD_LABEL, " [cord blood, log scale]"),
    y        = paste0(cfg$outcome_label, " [INFMon2, log scale]")
  ) +
  theme_minimal(base_size = 11)
```

```{r block10-partial-cor, eval=!is.null(avail_cord) && !is.null(CORD_VISIT), results='asis'}
if (requireNamespace("ppcor", quietly = TRUE)) {
  pcor_df <- cord_df |>
    dplyr::transmute(cord_val, outcome_val,
              arm_num = as.numeric(factor(maternal_arm))) |>
    tidyr::drop_na()

  pc <- ppcor::pcor.test(pcor_df$cord_val, pcor_df$outcome_val,
                         pcor_df$arm_num, method = "spearman")
  cat(sprintf(
    "\n**Partial Spearman ρ** (cord vs INFMon2, adjusting for maternal arm): ",
    "ρ = %.3f, *p* = %.4f, *n* = %d\n\n",
    pc$estimate, pc$p.value, nrow(pcor_df)
  ))
} else {
  cat("> Install package **ppcor** for partial correlation in Block 10.\n\n")
}
```

---

## Blocks 7–9 Cross-Feature Importance {#p4-importance}

*🧭 Maps to summary §8c — which antigen / subclass / FcG feature dominates?*

Elastic-net regression (α = 0.5, 5-fold CV at λ.1se) across all available
antigen-subclass features (concurrent, P02) and FcG features (longitudinal,
from the maternal visit). Since they come from different data frames, only
subjects present in both `data_infmon2` and `data_mat_p02` are included.

```{r p4-elasticnet, fig.width=10, fig.height=6, fig.cap="Blocks 7-9 elastic-net — top 20 non-zero coefficients vs. outcome."}
all_p4 <- c(avail_ant_sub, avail_fcg)

if (length(all_p4) < 4L) {
  message("Blocks 7-9 importance: fewer than 4 features available — skipped.")
} else if (!requireNamespace("glmnet", quietly = TRUE)) {
  message("Install package 'glmnet' for elastic-net importance.")
} else {
  library(glmnet)

  # Combine concurrent (P02) and longitudinal (maternal) features
  p02_cols <- intersect(all_p4, names(data_infmon2))
  mat_cols <- intersect(all_p4, names(data_mat_p02))
  mod_df <- dplyr::inner_join(
    dplyr::select(data_infmon2,  subject_accession,
                  dplyr::all_of(c(OUTCOME_COL, p02_cols))),
    dplyr::select(data_mat_p02, subject_accession,
                  dplyr::any_of(mat_cols)),
    by = "subject_accession"
  ) |>
    dplyr::select(-subject_accession) |>
    tidyr::drop_na()

  nt_add(nt, "Blocks 7-9 importance — complete cases", nrow(mod_df))

  X      <- as.matrix(mod_df[, all_p4])
  y      <- mod_df[[OUTCOME_COL]]

  set.seed(20240101)
  cv_fit  <- glmnet::cv.glmnet(X, y, alpha = 0.5, nfolds = 5)
  coef_df <- as.matrix(coef(cv_fit, s = "lambda.1se")) |>
    as.data.frame() |>
    tibble::rownames_to_column("feature") |>
    setNames(c("feature", "coef")) |>
    dplyr::filter(feature != "(Intercept)", coef != 0) |>
    dplyr::arrange(desc(abs(coef))) |>
    head(20)

  if (nrow(coef_df) == 0L) {
    message("Elastic net: all features shrunk to zero at lambda.1se.")
  } else {
    ggplot(coef_df,
           aes(x = fct_reorder(feature, abs(coef)), y = coef,
               fill = coef > 0)) +
      geom_col() +
      coord_flip() +
      scale_fill_manual(values = c("TRUE" = "#d6604d", "FALSE" = "#2166ac"),
                        labels = c("TRUE" = "Positive", "FALSE" = "Negative"),
                        name = "Direction") +
      labs(
        title    = paste("Blocks 7-9 elastic-net vs.", cfg$outcome_label),
        subtitle = paste0("λ.1se; top ", nrow(coef_df), " non-zero features",
                          "; N = ", nrow(mod_df)),
        x = NULL, y = "Coefficient"
      ) +
      theme_minimal(base_size = 11)
  }
}
```
