---
title: "Cord Blood → 8 Weeks: Subclass, ADCD & WT-IgG Decay (Supplementary Tables S.4.2 / S.4.3)"
subtitle: "Extension of the total-IgG half-life analysis (S.4.1) to IgG subclasses, ADCD, and whole-cell IgG — absolute (S.4.2) and total-IgG-standardized (S.4.3). Built from d_set (ebaa_extra2)."
author: "Scot Zens"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:
    toc: true
    toc_depth: 3
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r bootstrap, include=FALSE}
library(here)
# ---- shared path resolver: single definition in R/find_proj_file.R ----
# (identical bootstrap to cord_to_infmon2_analysis_refactored.Rmd and pathway_analysis.Rmd,
#  so this file runs from the same project tree with no extra setup)
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R",
                "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}
source(find_proj_file("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"), local = FALSE, encoding = "UTF-8")

library(dplyr)
library(tidyr)

# ---- IMPORTANT: this analysis uses d_set (object ebaa_extra2), NOT c_set ----
data_raw <- load_serology_data(find_proj_file("data/d_set.RData"),
                               object_name = "ebaa_extra2")

FROM_VISIT <- "CordBlood"   # raw code M00 (cord, sample_type C), recoded by load_serology_data()
TO_VISIT   <- "InfMon2"     # raw code M02 (infant 8-week / 2-month visit)
MIN_PAIRED <- 10L

# Oguti et al. (2022) log2 conversion: one unit on log2 = one halving.
LOG10_TO_LOG2 <- 1 / log10(2)          # ~3.321928 ; log2 delta = log10 delta * this
N_BONF        <- 6L                     # Bonferroni base (six DTP antigens, as in S.4.1)

# analyte sets for the two tables
S42_ANALYTES <- c("IgG1", "IgG2", "IgG3", "IgG4", "ADCD", "WT_IgG")   # absolute (raw)
S43_ANALYTES <- c("IgG1", "IgG2", "IgG3", "IgG4", "ADCD")             # standardized by total IgG

# presentation ordering
ANTIGEN_ORDER_DISP <- c("PT", "PRN", "FHA", "FIM", "TT", "DT",
                        "IPV1", "IPV2", "IPV3", "PENTAMER", "WHOLE")
ANALYTE_ORDER_DISP <- c("IgG1", "IgG2", "IgG3", "IgG4", "ADCD", "WT_IgG")
```

# Purpose and relationship to Table S.4.1

Supplementary Table **S.4.1** (unchanged) reports the half-life of transplacentally
transferred **total IgG** for the six DTP antigens using the method of Oguti et al. (2022),
alongside the published pooled estimates. This document **extends that decay analysis** to the
other analytes that were measured in cord blood in `d_set`:

- **Table S.4.2** — the same decay columns as S.4.1 (the Oguti-method half-life; the Oguti
  comparison columns are dropped, since Oguti et al. measured only total IgG), **plus** the
  absolute (raw) amount of decay across **all** paired subjects, the absolute decay restricted
  to subjects whose level actually **declined**, and the **proportion of samples declining**
  from cord to 8 weeks. Rows are the **IgG subclasses (IgG1–IgG4)** and **ADCD** for every
  available antigen, and **WT_IgG** (whole-bacterium IgG).
- **Table S.4.3** — the identical column set, but computed on values **standardized by total
  IgG** (each antigen-specific feature residualized on that antigen's total IgG, using the
  standardization from `pathway_analysis.Rmd`). WT_IgG is excluded here because it has no
  antigen-specific total-IgG denominator.

**Convention (as in the parent analysis):** Δ = value at **InfMon2** − value at **CordBlood**
(log₁₀). A **negative** Δ is the expected **decay**; a positive Δ is an increase. The
Oguti-method half-life uses the per-infant log₂ decay rate and, following the source paper,
excludes infants whose level increased over the interval (rate ≥ 0, taken to indicate likely
infection).

```{r visit-check, results='asis'}
vv <- as.character(unique(data_raw$visit_name))
cat(sprintf("Loaded `d_set` (`ebaa_extra2`): **%d** rows. Visit labels present include: %s.\n\n",
            nrow(data_raw), paste(sort(vv), collapse = ", ")))
if (!all(c(FROM_VISIT, TO_VISIT) %in% vv))
  cat(sprintf("> **Warning:** expected visit labels '%s' and '%s' were not both found after ",
              "`load_serology_data()`. Check the visit recode.\n\n", FROM_VISIT, TO_VISIT))
```

```{r pairing-helpers, include=FALSE}
## ---------------------------------------------------------------------------
## Manual within-subject pairing (used for BOTH tables so the pairing logic is
## identical for raw and standardized values). Takes a long frame with columns
## subject_accession / visit_name / feature / value and returns one row per
## subject x feature that has a finite value at both visits, with:
##   v_from (CordBlood), v_to (InfMon2), delta = v_to - v_from.
## ---------------------------------------------------------------------------
make_paired <- function(long_df) {
  w <- long_df %>%
    dplyr::filter(visit_name %in% c(FROM_VISIT, TO_VISIT), is.finite(value)) %>%
    dplyr::distinct(subject_accession, visit_name, feature, .keep_all = TRUE) %>%
    dplyr::select(subject_accession, visit_name, feature, value) %>%
    tidyr::pivot_wider(names_from = visit_name, values_from = value)
  ## If either visit is absent (e.g. a residual source that covers only one visit),
  ## there is nothing to pair -- return an empty frame rather than erroring, so the
  ## caller can fall back cleanly.
  if (!all(c(FROM_VISIT, TO_VISIT) %in% names(w)))
    return(data.frame(subject_accession = character(0), feature = character(0),
                      v_from = numeric(0), v_to = numeric(0), delta = numeric(0),
                      stringsAsFactors = FALSE))
  w %>%
    dplyr::filter(is.finite(.data[[FROM_VISIT]]), is.finite(.data[[TO_VISIT]])) %>%
    dplyr::mutate(v_from = .data[[FROM_VISIT]],
                  v_to   = .data[[TO_VISIT]],
                  delta  = v_to - v_from) %>%
    dplyr::select(subject_accession, feature, v_from, v_to, delta)
}

## native feature -> analyte map (clean casing from the data), antigen from prefix
feat_map <- data_raw %>%
  dplyr::distinct(feature, analyte) %>%
  dplyr::mutate(feature = as.character(feature),
                analyte = as.character(analyte),
                antigen = sub("_.*", "", feature))

attach_meta <- function(pd) {
  pd %>%
    dplyr::mutate(feature = as.character(feature)) %>%
    dplyr::left_join(feat_map, by = "feature")
}
```

```{r interval-days, include=FALSE}
## ---------------------------------------------------------------------------
## Per-infant CordBlood -> 8-week interval in days (same source as S.4.1).
## Loaded from data/INTERVAL_DAYS_BY_SUBJECT.RData if present; otherwise every
## infant falls back to the cohort/nominal interval. Because each infant's rate
## is divided by that infant's own interval, half-lives are on a true per-day
## footing; the fallback affects only absolute half-life scaling, never the
## relative comparison between analytes.
## ---------------------------------------------------------------------------
NOMINAL_INTERVAL_DAYS <- 8.7 * 7        # 60.9 d (Oguti Table 1 pooled median; fallback only)

if (!exists("INTERVAL_DAYS_BY_SUBJECT")) {
  .iv_path <- tryCatch(find_proj_file("data/INTERVAL_DAYS_BY_SUBJECT.RData", optional = TRUE),
                       error = function(e) NULL)
  if (!is.null(.iv_path) && file.exists(.iv_path)) load(.iv_path)
}

if (exists("INTERVAL_DAYS_BY_SUBJECT") &&
    is.data.frame(INTERVAL_DAYS_BY_SUBJECT) &&
    all(c("subject_accession", "interval_days") %in% names(INTERVAL_DAYS_BY_SUBJECT))) {
  .iv_tbl <- INTERVAL_DAYS_BY_SUBJECT[, c("subject_accession", "interval_days")]
  .iv_tbl$subject_accession <- as.character(.iv_tbl$subject_accession)
  .iv_tbl$interval_days     <- as.numeric(.iv_tbl$interval_days)
  .iv_med <- stats::median(.iv_tbl$interval_days, na.rm = TRUE)
  INTERVAL_IS_NOMINAL <- FALSE
  INTERVAL_SOURCE <- sprintf(
    "exact per-infant cord->8-week interval from sample-receipt dates (n=%d infants, median %.1f d)",
    sum(is.finite(.iv_tbl$interval_days)), .iv_med)
} else {
  .iv_tbl <- NULL
  .iv_med <- NOMINAL_INTERVAL_DAYS
  INTERVAL_IS_NOMINAL <- TRUE
  INTERVAL_SOURCE <- sprintf(paste0("nominal constant of %.1f days (INTERVAL_DAYS_BY_SUBJECT not ",
                                    "found in the session or under data/)"), NOMINAL_INTERVAL_DAYS)
}

attach_interval <- function(pd) {
  pd$subject_accession <- as.character(pd$subject_accession)
  if (INTERVAL_IS_NOMINAL) {
    pd$interval_days <- NOMINAL_INTERVAL_DAYS
  } else {
    .fb <- if (is.finite(.iv_med)) .iv_med else NOMINAL_INTERVAL_DAYS
    pd <- pd %>% dplyr::left_join(.iv_tbl, by = "subject_accession") %>%
      dplyr::mutate(interval_days = ifelse(is.finite(interval_days), interval_days, .fb))
  }
  pd
}
```

```{r decay-builder, include=FALSE}
## ---------------------------------------------------------------------------
## Oguti-method intercept-only fit -> half-life + 95% CI, from a vector of
## per-infant log2/day decay rates. Half-life = 1 / |mean rate|.
## ---------------------------------------------------------------------------
.hl_fit <- function(rate) {
  r <- rate[is.finite(rate)]
  n <- length(r)
  if (n < 3L || stats::sd(r) == 0) return(NULL)
  fit <- stats::lm(r ~ 1)
  est <- unname(stats::coef(fit)[1])
  ci  <- unname(stats::confint(fit)[1, ])
  p   <- stats::coef(summary(fit))[1, 4]
  hl_ci <- sort(1 / abs(ci))
  data.frame(n_used = n, rate = est, half_life = 1 / abs(est),
             hl_lo = hl_ci[1], hl_hi = hl_ci[2], p_raw = p)
}

## ---------------------------------------------------------------------------
## One decay summary row per (antigen, analyte). `pd` must carry:
##   subject_accession, antigen, analyte, delta (log10), interval_days.
## Columns produced (S.4.2 / S.4.3 layout):
##   n_pair        - subjects paired at both visits (all subjects)
##   n_used        - subjects contributing to the half-life (decliners; matches
##                   S.4.1 "This study N", which excludes increases per Oguti)
##   half_life,hl_lo,hl_hi,p_raw  - Oguti-method half-life + 95% CI
##   med_all       - median absolute decay Δlog10 across ALL paired subjects
##   n_dec, med_dec- decliners count and median Δlog10 among decliners only
##   pct_dec       - % of paired samples declining (Δ < 0)
## ---------------------------------------------------------------------------
build_decay_table <- function(pd, exclude_increases = TRUE) {
  pd <- pd %>%
    dplyr::filter(is.finite(delta), is.finite(interval_days), interval_days > 0)
  if (!nrow(pd)) return(data.frame())
  keys <- pd %>% dplyr::distinct(antigen, analyte)
  rows <- lapply(seq_len(nrow(keys)), function(i) {
    ant <- keys$antigen[i]; an <- keys$analyte[i]
    d <- pd %>% dplyr::filter(antigen == ant, analyte == an)
    n_pair <- nrow(d)
    if (n_pair < MIN_PAIRED) return(NULL)
    dec     <- d %>% dplyr::filter(delta < 0)
    n_dec   <- nrow(dec)
    med_all <- stats::median(d$delta)
    med_dec <- if (n_dec) stats::median(dec$delta) else NA_real_
    pct_dec <- 100 * n_dec / n_pair
    d$rate  <- (d$delta * LOG10_TO_LOG2) / d$interval_days
    use     <- if (exclude_increases) d$rate[d$rate < 0] else d$rate
    hl      <- .hl_fit(use)
    data.frame(
      antigen = ant, analyte = an, n_pair = n_pair,
      n_used    = if (is.null(hl)) NA_integer_ else hl$n_used,
      half_life = if (is.null(hl)) NA_real_    else hl$half_life,
      hl_lo     = if (is.null(hl)) NA_real_    else hl$hl_lo,
      hl_hi     = if (is.null(hl)) NA_real_    else hl$hl_hi,
      p_raw     = if (is.null(hl)) NA_real_    else hl$p_raw,
      med_all = med_all, n_dec = n_dec, med_dec = med_dec, pct_dec = pct_dec,
      stringsAsFactors = FALSE)
  })
  out <- dplyr::bind_rows(rows)
  if (!nrow(out)) return(out)
  out$p_bonf <- pmin(1, out$p_raw * N_BONF)
  out %>%
    dplyr::mutate(
      antigen = factor(antigen, levels = union(ANTIGEN_ORDER_DISP, sort(unique(antigen)))),
      analyte = factor(analyte, levels = union(ANALYTE_ORDER_DISP, sort(unique(analyte))))) %>%
    dplyr::arrange(analyte, antigen)
}

## Pretty-printer shared by both tables (columns mirror S.4.1 minus Oguti, + 3 new)
render_decay_table <- function(tbl, caption, half_life_dagger = FALSE) {
  hl_hdr <- if (half_life_dagger) "Half-life (days) (95% CI) \u2020" else "Half-life (days) (95% CI)"
  ## Build with plain, syntactic column names, then attach the display headers via
  ## col.names. (Unicode \u escapes are legal in normal double-quoted strings but
  ## NOT inside backtick-quoted names, so the fancy headers live only in col.names.)
  disp <- tbl %>% dplyr::transmute(
    antigen   = as.character(antigen),
    analyte   = as.character(analyte),
    n_pair    = n_pair,
    n_used    = ifelse(is.na(n_used), "\u2014", as.character(n_used)),
    half_life = ifelse(is.na(half_life), "\u2014",
                       sprintf("%.1f (%.1f \u2013 %.1f)", half_life, hl_lo, hl_hi)),
    med_all   = sprintf("%+.3f", med_all),
    med_dec   = ifelse(is.na(med_dec), "\u2014", sprintf("%+.3f", med_dec)),
    pct_dec   = sprintf("%.1f%%", pct_dec))
  knitr::kable(
    disp, caption = caption,
    col.names = c("Antigen", "Analyte", "N paired", "N used (decay)", hl_hdr,
                  "Median \u0394log\u2081\u2080, all subjects",
                  "Median \u0394log\u2081\u2080, decliners",
                  "% declining")) %>% print()
  invisible(disp)
}
```

# Coverage of the extended analytes in `d_set`

```{r coverage, results='asis'}
cov <- data_raw %>%
  dplyr::mutate(visit_name = as.character(visit_name),
                antigen = sub("_.*", "", as.character(feature)),
                analyte = as.character(analyte)) %>%
  dplyr::filter(analyte %in% S42_ANALYTES, visit_name %in% c(FROM_VISIT, TO_VISIT)) %>%
  dplyr::distinct(subject_accession, visit_name, feature, .keep_all = TRUE) %>%
  dplyr::group_by(antigen, analyte) %>%
  dplyr::summarize(
    n_cord   = dplyr::n_distinct(subject_accession[visit_name == FROM_VISIT]),
    n_inf    = dplyr::n_distinct(subject_accession[visit_name == TO_VISIT]),
    .groups  = "drop")

## paired counts
raw_long0 <- data_raw %>%
  dplyr::transmute(subject_accession, visit_name = as.character(visit_name),
                   feature = as.character(feature), value = log_assay_value)
paired_all <- make_paired(raw_long0) %>% attach_meta()
pair_cnt <- paired_all %>%
  dplyr::filter(analyte %in% S42_ANALYTES) %>%
  dplyr::group_by(antigen, analyte) %>%
  dplyr::summarize(n_paired = dplyr::n(), .groups = "drop")

cov <- cov %>% dplyr::left_join(pair_cnt, by = c("antigen", "analyte")) %>%
  dplyr::mutate(n_paired = ifelse(is.na(n_paired), 0L, n_paired),
                antigen = factor(antigen, levels = union(ANTIGEN_ORDER_DISP, sort(unique(antigen)))),
                analyte = factor(analyte, levels = union(ANALYTE_ORDER_DISP, sort(unique(analyte))))) %>%
  dplyr::arrange(analyte, antigen)

knitr::kable(
  cov %>% dplyr::transmute(Antigen = as.character(antigen), Analyte = as.character(analyte),
                           `n CordBlood` = n_cord, `n InfMon2` = n_inf, `n paired` = n_paired),
  caption = paste0("Coverage of the extended analytes in d_set: subjects with CordBlood, with ",
                   "InfMon2, and with BOTH (the n used). Only analyte x antigen cells with ",
                   sprintf("n paired \u2265 %d ", MIN_PAIRED),
                   "enter Tables S.4.2 / S.4.3.")) %>% print()

cat(sprintf("\nInterval source for all half-lives: **%s**.\n\n", INTERVAL_SOURCE))
```

# Table S.4.2 — Absolute (raw) decay: IgG subclasses, ADCD, and WT-IgG

```{r s42, results='asis'}
pd_raw <- paired_all %>%
  dplyr::filter(analyte %in% S42_ANALYTES) %>%
  attach_interval()

s42 <- build_decay_table(pd_raw, exclude_increases = TRUE)

render_decay_table(
  s42,
  caption = paste0(
    "Supplementary Table S.4.2. Decay of transplacental antibody from cord blood to the ",
    "8-week infant visit for the IgG subclasses (IgG1\u2013IgG4) and ADCD across every available ",
    "antigen, and for whole-bacterium IgG (WT_IgG). Columns mirror S.4.1 (Oguti-method half-life ",
    "on the per-infant log2 decay rate; the Oguti reference columns are omitted because that ",
    "study measured only total IgG), with three additions: the absolute (raw) median decay ",
    "\u0394log10 (InfMon2 \u2212 CordBlood; negative = decay) across all paired subjects, the same ",
    "restricted to subjects whose level declined, and the percentage of paired samples that ",
    "declined. 'N used (decay)' is the half-life sample after excluding infants whose level ",
    "increased over the interval (rate \u2265 0), matching Oguti et al. and the 'This study N' of S.4.1."))
```

```{r s42-raw-values, results='asis'}
## Full-precision values (for transcription into the supplement document)
knitr::kable(
  s42 %>% dplyr::transmute(
    antigen = as.character(antigen), analyte = as.character(analyte),
    n_pair, n_used,
    half_life = round(half_life, 2), hl_lo = round(hl_lo, 2), hl_hi = round(hl_hi, 2),
    p_raw = signif(p_raw, 3), p_bonf = signif(p_bonf, 3),
    med_all = round(med_all, 4), n_dec, med_dec = round(med_dec, 4),
    pct_dec = round(pct_dec, 1)),
  caption = "Table S.4.2 (full-precision underlying values).") %>% print()
```

# Standardization by total IgG (for Table S.4.3)

```{r standardize-setup, include=FALSE}
## ===========================================================================
## Total-IgG standardization.
## PRIMARY (preferred): the project's own bubble-plot transform
##   (R/connection_transform.R :: standardized_set(), per infant arm), exactly as in
##   the standardization section of pathway_analysis.Rmd, using the COMBINED residual
##   files that standardize M00 (CordBlood), M02 (InfMon2) and M05 together under a
##   single parameterization:
##     igg_standard_residuals_{ap,wp}_..._k_trn.RData
##   Both visits of the decay contrast are therefore standardized the SAME way, so the
##   per-subject cord->8-week change is not confounded by differing model fits.
## FALLBACK 1 (period-specific residuals, two separate model fits): CordBlood from the
##   maternal/cord ("_mat") residuals and InfMon2 from the infant/8-week residuals.
##   Used only if the combined "_trn" files are unavailable; kept internally consistent
##   (either BOTH visits from _trn, or BOTH from the period-specific pair).
## FALLBACK 2 (only if the transform/files/helpers are unavailable): residualize each
##   antigen-specific feature (log10) on that antigen's total IgG (log10) in-document,
##   fit per feature x infant-arm across both visits. The path used is reported in the
##   HTML diagnostics.
## In all cases the standardized value is a residual on total IgG; features with no
## antigen-specific total IgG (IPV*, PENTAMER, WHOLE/WT_IgG) drop out.
## ===========================================================================
STD_METHOD  <- "none"
std_paired  <- NULL
STD_DIAG    <- character(0)
.dg <- function(...) STD_DIAG <<- c(STD_DIAG, sprintf(...))

## Known filenames of the project bubble residuals (per infant arm), in data/.
## PREFERRED: the combined "_trn" files standardize CordBlood (M00), InfMon2 (M02) and
## M05 together under one parameterization, so the cord->8-week contrast uses a single
## consistent model. The period-specific "_mat" (maternal+cord) and infant/8-week files
## are retained as an internally-consistent fallback pair. The project constant is tried
## first (the symbol pathway_analysis.Rmd uses), then these literal names.
RESID_TRN_AP <- "igg_standard_residuals_ap_matpm_prevacvacc_k_trn.RData"   # combined M00+M02+M05
RESID_TRN_WP <- "igg_standard_residuals_wp_matpm_prevacvacc_k_trn.RData"
RESID_INF_AP <- "igg_standard_residuals_ap_matpm_prevacvacc_k.RData"       # 8-week (InfMon2)
RESID_INF_WP <- "igg_standard_residuals_wp_matpm_prevacvacc_k.RData"
RESID_MAT_AP <- "igg_standard_residuals_ap_matpm_prevacvacc_k_mat.RData"   # maternal + cord
RESID_MAT_WP <- "igg_standard_residuals_wp_matpm_prevacvacc_k_mat.RData"

.resolve_resid <- function(const_name, literal) {
  ## 1) project constant, resolved through find_proj_file (as in pathway_analysis.Rmd)
  if (exists(const_name, inherits = TRUE)) {
    cval <- get(const_name, inherits = TRUE)
    p <- tryCatch(find_proj_file(cval, optional = TRUE), error = function(e) NULL)
    if (!is.null(p) && file.exists(p)) { .dg("%s: resolved via constant %s -> %s", literal, const_name, p); return(p) }
  }
  ## 2) known filename through find_proj_file (bare and under data/)
  for (cand in c(literal, file.path("data", literal))) {
    p <- tryCatch(find_proj_file(cand, optional = TRUE), error = function(e) NULL)
    if (!is.null(p) && file.exists(p)) { .dg("%s: resolved via find_proj_file('%s') -> %s", literal, cand, p); return(p) }
  }
  ## 3) plain relative existence
  for (cand in c(file.path("data", literal), literal, file.path("..", "data", literal))) {
    if (file.exists(cand)) { .dg("%s: resolved via relative path -> %s", literal, cand); return(normalizePath(cand)) }
  }
  .dg("%s: NOT found (constant %s %s)", literal, const_name,
      if (exists(const_name, inherits = TRUE)) paste0("= '", get(const_name, inherits = TRUE), "'") else "undefined")
  NULL
}

## ---- attempt the project transform ----------------------------------------
.ct_err <- tryCatch({
  source(find_proj_file("R/connection_transform.R"), local = FALSE, encoding = "UTF-8"); NA_character_
}, error = function(e) conditionMessage(e))
if (!is.na(.ct_err)) .dg("source(R/connection_transform.R) ERROR: %s", .ct_err)

have_ct  <- exists("standardized_set", mode = "function")
have_dt  <- requireNamespace("data.table", quietly = TRUE)
have_inc <- exists("BUBBLE_INCLUDE_FEATURE")
.dg("standardized_set()=%s | BUBBLE_INCLUDE_FEATURE=%s | data.table=%s", have_ct, have_inc, have_dt)

.inc_feat <- if (have_inc) BUBBLE_INCLUDE_FEATURE else character(0)

## Resolve all residual sources (project constant first, then literal filename).
.res_trn_ap <- .resolve_resid("BUBBLE_RESIDUAL_TRN_AP", RESID_TRN_AP)
.res_trn_wp <- .resolve_resid("BUBBLE_RESIDUAL_TRN_WP", RESID_TRN_WP)
.res_inf_ap <- .resolve_resid("BUBBLE_RESIDUAL_AP",     RESID_INF_AP)
.res_inf_wp <- .resolve_resid("BUBBLE_RESIDUAL_WP",     RESID_INF_WP)
.res_mat_ap <- .resolve_resid("BUBBLE_RESIDUAL_MAT_AP", RESID_MAT_AP)
.res_mat_wp <- .resolve_resid("BUBBLE_RESIDUAL_MAT_WP", RESID_MAT_WP)

## Standardize ONE visit's slice with a specified residual pair, mirroring how
## pathway_analysis.Rmd calls standardized_set() on a single visit at a time. Returns
## a long frame (subject_accession, visit_name = v, feature, value) or NULL.
std_one_visit <- function(v, res_ap, res_wp, src_label) {
  if (is.null(res_ap) || is.null(res_wp)) return(NULL)
  cv <- data_raw %>%
    dplyr::filter(feature %in% .inc_feat, visit_name == v) %>%
    dplyr::select(subject_accession, visit_name, arm_name, infant_arm,
                  antigen, analyte, feature, value_reported, log_assay_value) %>%
    dplyr::distinct(subject_accession, feature, .keep_all = TRUE) %>%
    dplyr::mutate(dplyr::across(c(visit_name, antigen, analyte, feature, arm_name, infant_arm),
                                as.character))
  if (!nrow(cv)) return(NULL)
  s <- tryCatch(
    dplyr::bind_rows(
      standardized_set("IgG", cv[cv$infant_arm == "wP", ], res_wp),
      standardized_set("IgG", cv[cv$infant_arm == "aP", ], res_ap)),
    error = function(e) { .dg("visit '%s' via %s: standardized_set ERROR: %s", v, src_label, conditionMessage(e)); NULL })
  if (is.null(s)) return(NULL)
  if (!all(c("subject_accession", "feature", "log_assay_value") %in% names(s))) {
    .dg("visit '%s' via %s: returned cols {%s} (required subject_accession/feature/log_assay_value missing)",
        v, src_label, paste(names(s), collapse = ", "))
    return(NULL)
  }
  s <- s %>% dplyr::filter(is.finite(log_assay_value))
  .dg("visit '%s' via %s: %d usable rows (%d subjects, %d features)",
      v, src_label, nrow(s), dplyr::n_distinct(s$subject_accession), dplyr::n_distinct(s$feature))
  if (!nrow(s)) return(NULL)
  s %>% dplyr::transmute(subject_accession = as.character(subject_accession),
                         visit_name = v, feature = as.character(feature),
                         value = log_assay_value)
}

## Build a paired standardized set from a FROM-visit source and a TO-visit source.
## Returns a paired frame (with a "method" attribute) or NULL if either visit is empty.
build_std <- function(from_src, to_src, method_label) {
  if (is.null(from_src$ap) || is.null(to_src$ap)) return(NULL)
  f <- std_one_visit(FROM_VISIT, from_src$ap, from_src$wp, from_src$label)
  t <- std_one_visit(TO_VISIT,   to_src$ap,   to_src$wp,   to_src$label)
  sl <- dplyr::bind_rows(f, t)
  if (is.null(sl) || !nrow(sl) ||
      !any(sl$visit_name == FROM_VISIT) || !any(sl$visit_name == TO_VISIT)) return(NULL)
  cand <- make_paired(sl)
  .dg("  -> paired standardized rows (finite at BOTH visits): %d", nrow(cand))
  if (!nrow(cand)) return(NULL)
  attr(cand, "method") <- method_label
  cand
}

if (have_ct && have_dt && have_inc) {
  .src_trn <- list(ap = .res_trn_ap, wp = .res_trn_wp, label = "combined M00+M02+M05 residuals (_trn)")
  .src_mat <- list(ap = .res_mat_ap, wp = .res_mat_wp, label = "maternal/cord residuals (_mat)")
  .src_inf <- list(ap = .res_inf_ap, wp = .res_inf_wp, label = "infant/8-week residuals")

  cand <- NULL
  ## Preferred: BOTH visits from the combined (_trn) residuals -- one parameterization.
  if (!is.null(.src_trn$ap)) {
    .dg("Trying combined (_trn) residuals for both %s and %s ...", FROM_VISIT, TO_VISIT)
    cand <- build_std(.src_trn, .src_trn,
      "project transform (standardized_set on the combined '_trn' residuals): CordBlood and InfMon2 standardized under a single M00+M02+M05 parameterization, differenced per subject")
  }
  ## Fallback: period-specific residuals -- CordBlood from _mat, InfMon2 from infant.
  if (is.null(cand) && !is.null(.src_mat$ap) && !is.null(.src_inf$ap)) {
    .dg("Combined (_trn) path unavailable/empty; trying period-specific residuals (cord<-_mat, 8wk<-infant).")
    cand <- build_std(.src_mat, .src_inf,
      "project transform (standardized_set): CordBlood from maternal/cord (_mat) residuals and InfMon2 from infant residuals, differenced per subject (separate model fits)")
  }
  if (!is.null(cand)) {
    std_paired <- cand
    STD_METHOD  <- attr(cand, "method")
  } else {
    .dg("No project-transform path produced a paired standardized set; using fallback.")
  }
} else {
  .dg("Primary preconditions not all met (see flags above) -- using fallback.")
}

## ---- fallback: in-document residual-on-total-IgG ---------------------------
if (is.null(std_paired)) {
  .dg("FALLBACK engaged: in-document residual-on-total-IgG (per feature x infant_arm, both visits pooled).")
  base <- data_raw %>%
    dplyr::mutate(visit_name = as.character(visit_name),
                  feature = as.character(feature),
                  analyte = as.character(analyte),
                  infant_arm = as.character(infant_arm),
                  antigen = sub("_.*", "", feature)) %>%
    dplyr::filter(visit_name %in% c(FROM_VISIT, TO_VISIT)) %>%
    dplyr::distinct(subject_accession, visit_name, feature, .keep_all = TRUE)

  tot <- base %>% dplyr::filter(analyte == "IgG") %>%
    dplyr::transmute(subject_accession, visit_name, antigen, tot_igg = log_assay_value)

  targ <- base %>%
    dplyr::filter(analyte %in% S43_ANALYTES) %>%
    dplyr::left_join(tot, by = c("subject_accession", "visit_name", "antigen")) %>%
    dplyr::filter(is.finite(log_assay_value), is.finite(tot_igg))

  targ <- targ %>%
    dplyr::group_by(feature, infant_arm) %>%
    dplyr::mutate(std_val = tryCatch(
      as.numeric(stats::residuals(stats::lm(log_assay_value ~ tot_igg))),
      error = function(e) rep(NA_real_, dplyr::n()))) %>%
    dplyr::ungroup()

  .stdlong <- targ %>%
    dplyr::transmute(subject_accession, visit_name, feature, value = std_val)
  std_paired <- make_paired(.stdlong)
  STD_METHOD <- "in-document residual-on-total-IgG (project transform unavailable or not covering both visits -- see diagnostics)"
}

std_paired <- std_paired %>% attach_meta()
```

```{r standardize-note, results='asis'}
cat(sprintf(paste0(
  "Standardization method used at knit time: **%s**. Each antigen-specific feature is expressed ",
  "as its residual on that antigen's **total IgG**, so the reported decay reflects change in ",
  "antibody *quality/composition* net of the bulk total-IgG quantity change. Only antigens with ",
  "an antigen-specific total IgG at both visits are standardizable (in `d_set` that is ",
  "DT, FHA, PRN, PT, TT), so IPV, PENTAMER and WT_IgG do not appear in S.4.3.\n\n"),
  STD_METHOD))

## Full standardization diagnostics (so the path taken is auditable in the HTML).
cat("**Standardization diagnostics (knit-time):**\n\n")
for (.m in STD_DIAG) cat(sprintf("- %s\n", .m))
cat("\n")
```

# Table S.4.3 — Total-IgG-standardized decay: IgG subclasses and ADCD

```{r s43, results='asis'}
pd_std <- std_paired %>%
  dplyr::filter(analyte %in% S43_ANALYTES) %>%
  attach_interval()

s43 <- build_decay_table(pd_std, exclude_increases = TRUE)

render_decay_table(
  s43,
  half_life_dagger = TRUE,
  caption = paste0(
    "Supplementary Table S.4.3. Decay from cord blood to the 8-week infant visit for the IgG ",
    "subclasses (IgG1\u2013IgG4) and ADCD on all available antigens, computed on values ",
    "standardized by total IgG (each feature residualized on its antigen's total IgG, per the ",
    "standardization section of pathway_analysis.Rmd). Column layout is identical to S.4.2; here ",
    "\u0394 is the change in the total-IgG-standardized value (InfMon2 \u2212 CordBlood), so a ",
    "negative value indicates decay in the quantity-adjusted (quality) measure and '% declining' ",
    "is the share of subjects whose standardized value fell. \u2020 Because the standardized value ",
    "is a residual (not a raw titre), the half-life is a formal extension of the Oguti calculation ",
    "and should be read as a decay-rate summary of the standardized quantity rather than a ",
    "titre-halving time."))
```

```{r s43-raw-values, results='asis'}
knitr::kable(
  s43 %>% dplyr::transmute(
    antigen = as.character(antigen), analyte = as.character(analyte),
    n_pair, n_used,
    half_life = round(half_life, 2), hl_lo = round(hl_lo, 2), hl_hi = round(hl_hi, 2),
    p_raw = signif(p_raw, 3), p_bonf = signif(p_bonf, 3),
    med_all = round(med_all, 4), n_dec, med_dec = round(med_dec, 4),
    pct_dec = round(pct_dec, 1)),
  caption = "Table S.4.3 (full-precision underlying values).") %>% print()
```

# Notes

- **Data source.** All quantities are computed from `d_set` (object `ebaa_extra2`) via the shared
  pipeline (`config/endpoints.R`, `R/serology_helpers.R`, the global visit recode). This is the
  same object used by `pathway_analysis.Rmd`. The cord (`CordBlood`, raw code `M00`,
  sample_type C) and 8-week (`InfMon2`, raw code `M02`) visits are paired within subject.
- **Half-life method.** Identical to S.4.1 / Oguti et al. (2022): per-infant decay rate =
  Δlog₂ ÷ (days between draws); per (antigen × analyte) intercept-only fit; half-life = 1/|mean
  rate| with reciprocal-of-CI limits; infants whose level increased (rate ≥ 0) are excluded
  ('N used'). p-values are Bonferroni-scaled by 6 (as in S.4.1) in the full-precision tables.
- **Why the two tables differ only in the value.** S.4.2 uses the raw log₁₀ assay value
  ("absolute"); S.4.3 uses the total-IgG-standardized residual. Every column definition is the
  same, so the two tables are directly comparable: a subclass/ADCD that decays in S.4.2 but not
  in S.4.3 was decaying only because total IgG was decaying with it.
- **Reproduce with:** `rmarkdown::render("cord_to_infmon2_decay_subclass_adcd.Rmd")`.
