---
title: "Clinical Predictors of Infant M02 WT_IgG"
subtitle: |
  Univariate screening, continuous vs tertile form selection, and multivariate
  model building — stratified by infant arm (aP vs wP) and maternal arm (TdaP vs TT)
author: "Scot Zens"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 5
params:
  outcome_feature: "WHOLE_WT_IgG"
  outcome_label:   "WT_IgG"
  outcome_y_label: "Log10 WT_IgG"
  alpha:           0.05        # significance threshold for univariate screening
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo    = TRUE
)
```

---

# Analytical Framework

This document asks whether **clinical characteristics of the mother or infant**
predict infant **`r params$outcome_label` at M02** (pertussis toxin
neutralization, measured immediately before first vaccination), after
stratifying by infant vaccination arm (aP vs wP) and by maternal vaccination
arm (TdaP vs TT).

The analysis runs in four steps within each stratum:

1. **Univariate screen.** Each clinical covariate is tested alone against the
   outcome at the stratum level. Categorical variables are kept at two
   levels; continuous variables are tested in both continuous and
   tertile-categorical forms.
2. **Form selection.** For any continuous variable significant in both forms,
   the form with the higher adjusted R² is retained. For variables significant
   in only one form, that form is retained.
3. **Multivariate model.** If more than one predictor passes the univariate
   screen (p~model~ < `r params$alpha`), all selected predictors are entered
   jointly into a single linear model. If only one passes, the univariate
   result stands. If none pass, that is reported.
4. **Cross-stratum summary.** Significant predictors across all strata are
   tabulated for comparison.

The six antibody-chain predictor blocks in `predict_M02_PTNA_from_maternal_v2.Rmd`
are not included here; this file focuses exclusively on clinical covariates.

---

# Clinical Covariates

The ten covariates come from the `clin_assess` data object:

```{r covariate-list, echo=FALSE, results='asis'}
cov_meta <- data.frame(
  Label       = c("Birth Weight",
                  "Infant Sex",
                  "Delivery Mode",
                  "Parity",
                  "Gestational Age at Vaccine",
                  "Interval: Vaccination to Delivery (days)",
                  "Gestational Age at Delivery",
                  "Maternal Age at Vaccination",
                  "Maternal BMI at Vaccination",
                  "Maternal Haemoglobin"),
  Variable    = c("birth_weight", "infant_sex", "delivery_mode", "parity",
                  "gestational_age_vaccination", "vaccine_birth_interval_days",
                  "gestational_age_birth", "maternal_age",
                  "maternal_bmi", "maternal_Hb"),
  Type        = c("Continuous", "Categorical (2 levels)", "Categorical (2 levels)",
                  "Categorical (2 levels)", "Continuous", "Continuous",
                  "Continuous", "Continuous", "Continuous", "Continuous"),
  `Forms tested` = c("continuous + tertile", "factor", "factor (vaginal vs cesarean)",
                     "factor (nulliparous vs parous)", "continuous + tertile",
                     "continuous + tertile", "continuous + tertile",
                     "continuous + tertile", "continuous + tertile",
                     "continuous + tertile"),
  stringsAsFactors = FALSE
)
knitr::kable(cov_meta,
  col.names = c("Label", "Variable", "Type", "Forms tested"),
  caption   = "Table 1. Clinical covariates and testing approach."
)
```

---

# Setup

```{r setup}
library(data.table)
library(tidyverse)
library(here)
library(broom)
library(knitr)

# ---- Outcome definitions ----------------------------------------
outcome_feature <- params$outcome_feature
outcome_label   <- params$outcome_label
outcome_y_label <- params$outcome_y_label
outcome_col     <- paste0(outcome_feature, "_M02")
alpha           <- params$alpha

# ---- Stratification arms ----------------------------------------
arms        <- c("TdaP", "TT")
infant_arms <- c("aP", "wP")
iarm_colors <- c("aP" = "dodgerblue2", "wP" = "forestgreen")
marm_colors <- c("TdaP" = "skyblue",   "TT" = "salmon")

# ---- Named covariate list (label → variable) --------------------
covariates_named_list <- list(
  "Birth Weight"                              = "birth_weight",
  "Infant Sex"                                = "infant_sex",
  "Delivery Mode"                             = "delivery_mode",
  "Parity"                                    = "parity",
  "Gestational Age at Vaccine"                = "gestational_age_vaccination",
  "Interval: Vaccination to Delivery (days)"  = "vaccine_birth_interval_days",
  "Gestational Age at Delivery"               = "gestational_age_birth",
  "Maternal Age at Vaccination"               = "maternal_age",
  "Maternal BMI at Vaccination"               = "maternal_bmi",
  "Maternal Haemoglobin"                      = "maternal_Hb"
)

# Variables to treat as categorical (2 levels)
cat_vars <- c("infant_sex", "delivery_mode", "parity")

# Continuous variables (will be tested as continuous + tertile)
cont_vars <- setdiff(unname(unlist(covariates_named_list)), cat_vars)

# ---- Helper: p-value label -------------------------------------
pvalue_to_label <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
}

# ---- Helper: tertile split -------------------------------------
# Returns an ordered factor T1 (low) / T2 (mid) / T3 (high).
make_tertile <- function(x) {
  qs <- quantile(x, probs = c(1/3, 2/3), na.rm = TRUE)
  cut(x,
      breaks         = c(-Inf, qs[1], qs[2], Inf),
      labels         = c("T1 (low)", "T2 (mid)", "T3 (high)"),
      include.lowest = TRUE,
      ordered_result = FALSE)
}

# ---- Helper: run one lm and return a structured list -----------
run_lm_clin <- function(data, outcome_col, predictor_col, label = "") {
  f <- reformulate(predictor_col, response = outcome_col)
  m <- tryCatch(
    lm(f, data = data, na.action = na.omit),
    error = function(e) NULL
  )
  if (is.null(m)) {
    return(list(label = label, predictor = predictor_col,
                R2 = NA, adj_R2 = NA, n_obs = NA, p_model = NA,
                tidy_coef = NULL, model = NULL))
  }
  s <- summary(m)
  p_mod <- tryCatch(
    pf(s$fstatistic[1], s$fstatistic[2], s$fstatistic[3], lower.tail = FALSE),
    error = function(e) NA
  )
  tidy_coef <- tidy(m) %>%
    mutate(
      sig     = pvalue_to_label(p.value),
      across(c(estimate, std.error, statistic), ~round(., 4)),
      p.value = signif(p.value, 3)
    ) %>%
    rename(Predictor = term, `β` = estimate, SE = std.error,
           t = statistic, p = p.value, ` ` = sig)
  list(
    label     = label,
    predictor = predictor_col,
    R2        = round(s$r.squared,     4),
    adj_R2    = round(s$adj.r.squared, 4),
    n_obs     = nobs(m),
    p_model   = p_mod,
    tidy_coef = tidy_coef,
    model     = m
  )
}

# ---- Helper: display a stored model result ---------------------
show_lm_clin <- function(res) {
  if (is.null(res$tidy_coef)) {
    cat("*Model could not be estimated.*\n\n")
    return(invisible(NULL))
  }
  cat(sprintf(
    "\n**n = %d &nbsp;|&nbsp; R² = %.3f &nbsp;|&nbsp; adj-R² = %.3f &nbsp;|&nbsp; p~model~ = %.3g**\n\n",
    res$n_obs, res$R2, res$adj_R2, res$p_model
  ))
  print(knitr::kable(
    res$tidy_coef,
    caption = paste0("Coefficients — ", res$label),
    align   = c("l", rep("r", ncol(res$tidy_coef) - 1))
  ))
  cat("\n")
}

# ---- Helper: run multivariate model ----------------------------
run_multi <- function(data, outcome_col, predictors, label = "Multivariate") {
  if (length(predictors) < 2) return(NULL)
  f <- reformulate(predictors, response = outcome_col)
  m <- tryCatch(
    lm(f, data = data, na.action = na.omit),
    error = function(e) NULL
  )
  if (is.null(m)) return(NULL)
  s <- summary(m)
  p_mod <- tryCatch(
    pf(s$fstatistic[1], s$fstatistic[2], s$fstatistic[3], lower.tail = FALSE),
    error = function(e) NA
  )
  tidy_coef <- tidy(m) %>%
    mutate(
      sig     = pvalue_to_label(p.value),
      across(c(estimate, std.error, statistic), ~round(., 4)),
      p.value = signif(p.value, 3)
    ) %>%
    rename(Predictor = term, `β` = estimate, SE = std.error,
           t = statistic, p = p.value, ` ` = sig)
  list(
    label = label, R2 = round(s$r.squared, 4),
    adj_R2 = round(s$adj.r.squared, 4),
    n_obs = nobs(m), p_model = p_mod,
    tidy_coef = tidy_coef, model = m
  )
}
```

---

# Data Preparation

## Load Antibody and Clinical Data

```{r data-load}
# ---- Antibody data (same source as v2) --------------------------
load(file = here::here("./data/c_set.RData"))
data_raw          <- ebaa_extra
data_raw$arm_name <- data_raw$maternal_arm
data_raw$arm_name <- fct_recode(data_raw$arm_name, "TT" = "TT", "TdaP" = "TdaP")
data_raw$feature  <- factor(paste(data_raw$antigen, data_raw$analyte, sep = "_"))

# Pull the outcome at M02 and arm columns (no IgG predictors needed)
data_outcome <- data_raw %>%
  dplyr::filter(feature    == outcome_feature,
                visit_name == "M02",
                arm_name   %in% arms) %>%
  dplyr::select(subject_accession, arm_name, infant_arm, log_assay_value) %>%
  distinct(subject_accession, .keep_all = TRUE) %>%
  rename(!!outcome_col := log_assay_value)

cat(sprintf("Outcome rows: **%d**  |  `%s` non-missing: **%d**\n",
            nrow(data_outcome),
            outcome_col,
            sum(!is.na(data_outcome[[outcome_col]]))))

# ---- Clinical assessments ---------------------------------------
load(file = here::here("./data/clin_assess.RData"))

# Merge: left join so every antibody subject is kept
data <- merge(data_outcome, clin_assess, by = "subject_accession", all.x = TRUE)

cat(sprintf("After merge: **%d subjects × %d columns**\n",
            nrow(data), ncol(data)))
```

## Prepare Clinical Covariates

```{r cov-prep}
# ---- Infant sex: ensure 2-level factor --------------------------
data$infant_sex <- factor(data$infant_sex)

# ---- Delivery mode: collapse to vaginal vs cesarean -------------
# Inspect first to confirm actual levels in the data
cat("delivery_mode levels: ", paste(levels(factor(data$delivery_mode)), collapse = ", "), "\n")
data$delivery_mode <- fct_collapse(
  factor(data$delivery_mode),
  vaginal   = grep("(?i)vag|normal|spontan|SVD",
                   levels(factor(data$delivery_mode)), value = TRUE, perl = TRUE),
  cesarean  = grep("(?i)ces|csec|section|C/S",
                   levels(factor(data$delivery_mode)), value = TRUE, perl = TRUE),
  other     = grep("(?i)vag|normal|spontan|SVD|ces|csec|section|C/S",
                   levels(factor(data$delivery_mode)), value = TRUE, perl = TRUE,
                   invert = TRUE)
) %>%
  # Drop "other" if very few, keeping binary structure
  fct_drop()
data$delivery_mode <- relevel(data$delivery_mode, ref = "vaginal")
cat("delivery_mode after collapse: ", paste(levels(data$delivery_mode), collapse = ", "), "\n")

# ---- Parity: 0 (nulliparous) vs 1+ (parous) --------------------
data$parity <- factor(
  ifelse(data$parity == 0, "nulliparous", "parous"),
  levels = c("nulliparous", "parous")
)

# ---- Tertile columns for every continuous variable -------------
for (v in cont_vars) {
  tname <- paste0(v, "_T")
  if (v %in% names(data)) {
    data[[tname]] <- make_tertile(data[[v]])
  }
}

# Column names for tertile variables
cont_vars_T <- paste0(cont_vars, "_T")

cat("Tertile columns added:", paste(cont_vars_T, collapse = ", "), "\n")
```

## Covariate Descriptives by Stratum

```{r cov-descriptives, results='asis'}
# Continuous variables: median (IQR) by infant_arm and arm_name
for (strat_col in c("infant_arm", "arm_name")) {
  cat(sprintf("\n### Descriptives by %s\n\n", strat_col))
  
  rows <- lapply(names(covariates_named_list), function(lbl) {
    v <- covariates_named_list[[lbl]]
    if (v %in% cat_vars) {
      # Count table
      tab <- table(data[[v]], data[[strat_col]], useNA = "no")
      paste_fn <- function(col) {
        ns <- tab[, col, drop = FALSE]
        paste(rownames(tab), ns, sep = "=", collapse = "; ")
      }
      lvls <- colnames(tab)
      row <- c(lbl, "categorical",
               sapply(lvls, paste_fn),
               sum(!is.na(data[[v]])))
    } else {
      # Median [Q1–Q3]
      lvls <- sort(unique(data[[strat_col]]))
      miq <- function(x) {
        q <- quantile(x, c(.25, .5, .75), na.rm = TRUE)
        sprintf("%.1f [%.1f–%.1f]", q[2], q[1], q[3])
      }
      row <- c(lbl, "continuous",
               sapply(lvls, function(lv) miq(data[[v]][data[[strat_col]] == lv])),
               sum(!is.na(data[[v]])))
    }
    row
  })
  
  lvl_cols <- sort(unique(data[[strat_col]]))
  desc_df  <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
  colnames(desc_df) <- c("Covariate", "Type", lvl_cols, "n (non-missing)")
  print(knitr::kable(desc_df,
    caption = sprintf("Covariate descriptives by %s. Continuous: median [Q1–Q3].", strat_col)))
  cat("\n")
}
```

---

# Univariate Screening and Model Selection

The core analysis function runs for a single stratum. It:

1. Loops over all covariates and fits univariate models (categorical = 1 model;
   continuous = 2 models: continuous form and tertile form).
2. Selects significant predictors at p~model~ < `r params$alpha`.
3. For continuous variables significant in both forms, retains the form with
   the higher adjusted R².
4. Runs a multivariate model if ≥ 2 predictors are selected.

```{r engine-function}
# ================================================================
# run_clinical_screen()
#   d          : data frame for one stratum
#   outcome_col: string, the outcome column name
#   cov_list   : named list (label → variable name)
#   alpha      : significance threshold
#
# Returns a list with:
#   univ_table   : data.frame summarising all univariate results
#   selected     : named vector (label → selected column name)
#   form_notes   : explanation of form selection for continuous vars
#   multi_result : run_multi() output (or NULL)
# ================================================================
run_clinical_screen <- function(d, outcome_col, cov_list,
                                alpha = 0.05) {

  cat_v  <- c("infant_sex", "delivery_mode", "parity")
  cont_v <- setdiff(unname(unlist(cov_list)), cat_v)

  # ---- 1. Univariate models ------------------------------------
  univ_rows <- list()
  univ_res  <- list()   # full result objects for form-selection

  for (lbl in names(cov_list)) {
    v <- cov_list[[lbl]]

    if (v %in% cat_v) {
      # Single model: factor as-is
      r <- run_lm_clin(d, outcome_col, v, label = lbl)
      univ_rows[[paste0(v, "_cat")]] <- data.frame(
        Label  = lbl, Variable = v, Form = "categorical",
        n      = ifelse(is.na(r$n_obs), NA_integer_, r$n_obs),
        R2     = r$R2, adj_R2 = r$adj_R2,
        p_model = r$p_model,
        sig    = ifelse(is.na(r$p_model), "—",
                        pvalue_to_label(r$p_model)),
        stringsAsFactors = FALSE
      )
      univ_res[[paste0(v, "_cat")]] <- r

    } else {
      # Two models: continuous and tertile
      vT <- paste0(v, "_T")
      r_cont <- run_lm_clin(d, outcome_col, v,  label = paste0(lbl, " (continuous)"))
      r_tert <- run_lm_clin(d, outcome_col, vT, label = paste0(lbl, " (tertile)"))

      for (tag in c("cont", "tert")) {
        r   <- if (tag == "cont") r_cont else r_tert
        frm <- if (tag == "cont") "continuous" else "tertile"
        var <- if (tag == "cont") v else vT
        univ_rows[[paste0(v, "_", tag)]] <- data.frame(
          Label  = lbl, Variable = var, Form = frm,
          n      = ifelse(is.na(r$n_obs), NA_integer_, r$n_obs),
          R2     = r$R2, adj_R2 = r$adj_R2,
          p_model = r$p_model,
          sig    = ifelse(is.na(r$p_model), "—",
                          pvalue_to_label(r$p_model)),
          stringsAsFactors = FALSE
        )
      }
      univ_res[[paste0(v, "_cont")]] <- r_cont
      univ_res[[paste0(v, "_tert")]] <- r_tert
    }
  }

  univ_table <- bind_rows(univ_rows)

  # ---- 2. Form selection for continuous vars -------------------
  selected   <- list()   # label → chosen column name
  form_notes <- list()   # label → explanation string

  for (lbl in names(cov_list)) {
    v <- cov_list[[lbl]]

    if (v %in% cat_v) {
      r <- univ_res[[paste0(v, "_cat")]]
      if (!is.na(r$p_model) && r$p_model < alpha) {
        selected[[lbl]]   <- v
        form_notes[[lbl]] <- "categorical — significant"
      }
    } else {
      rc <- univ_res[[paste0(v, "_cont")]]
      rt <- univ_res[[paste0(v, "_tert")]]
      sc <- !is.na(rc$p_model) && rc$p_model < alpha
      st <- !is.na(rt$p_model) && rt$p_model < alpha

      if (sc && st) {
        # Both significant: pick higher adj-R2
        if (!is.na(rc$adj_R2) && !is.na(rt$adj_R2) &&
            rc$adj_R2 >= rt$adj_R2) {
          selected[[lbl]]   <- v
          form_notes[[lbl]] <- sprintf(
            "continuous (adj-R²=%.3f) preferred over tertile (adj-R²=%.3f) — both sig",
            rc$adj_R2, rt$adj_R2)
        } else {
          selected[[lbl]]   <- paste0(v, "_T")
          form_notes[[lbl]] <- sprintf(
            "tertile (adj-R²=%.3f) preferred over continuous (adj-R²=%.3f) — both sig",
            rt$adj_R2, rc$adj_R2)
        }
      } else if (sc) {
        selected[[lbl]]   <- v
        form_notes[[lbl]] <- "continuous — only form significant"
      } else if (st) {
        selected[[lbl]]   <- paste0(v, "_T")
        form_notes[[lbl]] <- "tertile — only form significant"
      }
      # else neither significant → not added to selected
    }
  }

  # ---- 3. Multivariate model -----------------------------------
  sel_predictors <- unname(unlist(selected))
  multi_result   <- NULL
  if (length(sel_predictors) > 1) {
    multi_result <- run_multi(d, outcome_col, sel_predictors,
                              label = "Multivariate (all significant predictors)")
  }

  list(
    univ_table   = univ_table,
    selected     = selected,
    form_notes   = form_notes,
    multi_result = multi_result,
    univ_res     = univ_res
  )
}
```

---

# Run Analysis — Infant Arm (aP vs wP)

```{r run-infant-arm, results='hide'}
screens_iarm <- list()
for (iarm in infant_arms) {
  d_sub <- data %>%
    dplyr::filter(infant_arm == iarm, !is.na(.data[[outcome_col]]))
  screens_iarm[[iarm]] <- run_clinical_screen(
    d_sub, outcome_col, covariates_named_list, alpha = alpha
  )
}
```

---

# Run Analysis — Maternal Arm (TdaP vs TT)

```{r run-maternal-arm, results='hide'}
screens_marm <- list()
for (marm in arms) {
  d_sub <- data %>%
    dplyr::filter(arm_name == marm, !is.na(.data[[outcome_col]]))
  screens_marm[[marm]] <- run_clinical_screen(
    d_sub, outcome_col, covariates_named_list, alpha = alpha
  )
}
```

---

# Results: Infant Arm

## aP Infants

```{r display-ap, results='asis'}
s <- screens_iarm[["aP"]]
n_ap <- nrow(data %>% filter(infant_arm == "aP", !is.na(.data[[outcome_col]])))
cat(sprintf("**Stratum: aP infants** | n with outcome = %d\n\n", n_ap))

# Univariate summary table
tbl <- s$univ_table %>%
  mutate(
    R2      = ifelse(is.na(R2),      "—", sprintf("%.3f", R2)),
    adj_R2  = ifelse(is.na(adj_R2),  "—", sprintf("%.3f", adj_R2)),
    p_model = ifelse(is.na(p_model), "—", sprintf("%.3g", p_model))
  )
knitr::kable(
  tbl %>% select(Label, Form, n, R2, adj_R2, p_model, sig),
  col.names = c("Covariate", "Form", "n", "R²", "adj-R²", "p (model)", ""),
  caption   = "Table: Univariate screening — aP infants.",
  align     = c("l","l","r","r","r","r","c")
) %>% print()
cat("\n")

# Form selection notes
if (length(s$selected) == 0) {
  cat("**No covariates reached p <", alpha, "in the univariate screen.**\n\n")
} else {
  cat(sprintf("**%d covariate(s) significant at p < %g:**\n\n",
              length(s$selected), alpha))
  
  # Form-selection detail
  form_df <- data.frame(
    Covariate     = names(s$selected),
    `Selected form` = unname(unlist(s$selected)),
    Note          = unname(unlist(s$form_notes)),
    stringsAsFactors = FALSE
  )
  knitr::kable(form_df,
    col.names = c("Covariate", "Selected predictor column", "Form selection rationale"),
    caption   = "Form selection — aP infants."
  ) %>% print()
  cat("\n")

  # Show each selected univariate model
  cat("### Significant univariate models (aP)\n\n")
  for (lbl in names(s$selected)) {
    v   <- cov_list_raw <- covariates_named_list[[lbl]]
    sel <- s$selected[[lbl]]
    key <- if (sel == v) paste0(v, "_cat") else {
             if (endsWith(sel, "_T")) paste0(v, "_tert") else paste0(v, "_cont")
           }
    # Correct key for categorical
    if (v %in% cat_vars) key <- paste0(v, "_cat")
    res <- s$univ_res[[key]]
    if (!is.null(res)) show_lm_clin(res)
  }

  # Multivariate model (if applicable)
  if (length(s$selected) > 1 && !is.null(s$multi_result)) {
    cat("### Multivariate model (aP)\n\n")
    cat(sprintf(
      "All %d selected predictors entered jointly: %s\n\n",
      length(s$selected),
      paste(unlist(s$selected), collapse = ", ")
    ))
    show_lm_clin(s$multi_result)
  } else if (length(s$selected) == 1) {
    cat("*Only one predictor was significant; no multivariate model fitted.*\n\n")
  }
}
```

## wP Infants

```{r display-wp, results='asis'}
s <- screens_iarm[["wP"]]
n_wp <- nrow(data %>% filter(infant_arm == "wP", !is.na(.data[[outcome_col]])))
cat(sprintf("**Stratum: wP infants** | n with outcome = %d\n\n", n_wp))

tbl <- s$univ_table %>%
  mutate(
    R2      = ifelse(is.na(R2),      "—", sprintf("%.3f", R2)),
    adj_R2  = ifelse(is.na(adj_R2),  "—", sprintf("%.3f", adj_R2)),
    p_model = ifelse(is.na(p_model), "—", sprintf("%.3g", p_model))
  )
knitr::kable(
  tbl %>% select(Label, Form, n, R2, adj_R2, p_model, sig),
  col.names = c("Covariate", "Form", "n", "R²", "adj-R²", "p (model)", ""),
  caption   = "Table: Univariate screening — wP infants.",
  align     = c("l","l","r","r","r","r","c")
) %>% print()
cat("\n")

if (length(s$selected) == 0) {
  cat("**No covariates reached p <", alpha, "in the univariate screen.**\n\n")
} else {
  cat(sprintf("**%d covariate(s) significant at p < %g:**\n\n",
              length(s$selected), alpha))

  form_df <- data.frame(
    Covariate     = names(s$selected),
    `Selected form` = unname(unlist(s$selected)),
    Note          = unname(unlist(s$form_notes)),
    stringsAsFactors = FALSE
  )
  knitr::kable(form_df,
    col.names = c("Covariate", "Selected predictor column", "Form selection rationale"),
    caption   = "Form selection — wP infants."
  ) %>% print()
  cat("\n")

  cat("### Significant univariate models (wP)\n\n")
  for (lbl in names(s$selected)) {
    v   <- covariates_named_list[[lbl]]
    sel <- s$selected[[lbl]]
    key <- if (v %in% cat_vars) paste0(v, "_cat") else {
             if (endsWith(sel, "_T")) paste0(v, "_tert") else paste0(v, "_cont")
           }
    res <- s$univ_res[[key]]
    if (!is.null(res)) show_lm_clin(res)
  }

  if (length(s$selected) > 1 && !is.null(s$multi_result)) {
    cat("### Multivariate model (wP)\n\n")
    cat(sprintf(
      "All %d selected predictors entered jointly: %s\n\n",
      length(s$selected),
      paste(unlist(s$selected), collapse = ", ")
    ))
    show_lm_clin(s$multi_result)
  } else if (length(s$selected) == 1) {
    cat("*Only one predictor was significant; no multivariate model fitted.*\n\n")
  }
}
```

---

# Results: Maternal Arm

## TdaP Mothers

```{r display-tdap, results='asis'}
s <- screens_marm[["TdaP"]]
n_tdap <- nrow(data %>% filter(arm_name == "TdaP", !is.na(.data[[outcome_col]])))
cat(sprintf("**Stratum: TdaP mothers** | n with outcome = %d\n\n", n_tdap))

tbl <- s$univ_table %>%
  mutate(
    R2      = ifelse(is.na(R2),      "—", sprintf("%.3f", R2)),
    adj_R2  = ifelse(is.na(adj_R2),  "—", sprintf("%.3f", adj_R2)),
    p_model = ifelse(is.na(p_model), "—", sprintf("%.3g", p_model))
  )
knitr::kable(
  tbl %>% select(Label, Form, n, R2, adj_R2, p_model, sig),
  col.names = c("Covariate", "Form", "n", "R²", "adj-R²", "p (model)", ""),
  caption   = "Table: Univariate screening — TdaP mothers.",
  align     = c("l","l","r","r","r","r","c")
) %>% print()
cat("\n")

if (length(s$selected) == 0) {
  cat("**No covariates reached p <", alpha, "in the univariate screen.**\n\n")
} else {
  cat(sprintf("**%d covariate(s) significant at p < %g:**\n\n",
              length(s$selected), alpha))

  form_df <- data.frame(
    Covariate     = names(s$selected),
    `Selected form` = unname(unlist(s$selected)),
    Note          = unname(unlist(s$form_notes)),
    stringsAsFactors = FALSE
  )
  knitr::kable(form_df,
    col.names = c("Covariate", "Selected predictor column", "Form selection rationale"),
    caption   = "Form selection — TdaP mothers."
  ) %>% print()
  cat("\n")

  cat("### Significant univariate models (TdaP)\n\n")
  for (lbl in names(s$selected)) {
    v   <- covariates_named_list[[lbl]]
    sel <- s$selected[[lbl]]
    key <- if (v %in% cat_vars) paste0(v, "_cat") else {
             if (endsWith(sel, "_T")) paste0(v, "_tert") else paste0(v, "_cont")
           }
    res <- s$univ_res[[key]]
    if (!is.null(res)) show_lm_clin(res)
  }

  if (length(s$selected) > 1 && !is.null(s$multi_result)) {
    cat("### Multivariate model (TdaP)\n\n")
    cat(sprintf(
      "All %d selected predictors entered jointly: %s\n\n",
      length(s$selected),
      paste(unlist(s$selected), collapse = ", ")
    ))
    show_lm_clin(s$multi_result)
  } else if (length(s$selected) == 1) {
    cat("*Only one predictor was significant; no multivariate model fitted.*\n\n")
  }
}
```

## TT Mothers

```{r display-tt, results='asis'}
s <- screens_marm[["TT"]]
n_tt <- nrow(data %>% filter(arm_name == "TT", !is.na(.data[[outcome_col]])))
cat(sprintf("**Stratum: TT mothers** | n with outcome = %d\n\n", n_tt))

tbl <- s$univ_table %>%
  mutate(
    R2      = ifelse(is.na(R2),      "—", sprintf("%.3f", R2)),
    adj_R2  = ifelse(is.na(adj_R2),  "—", sprintf("%.3f", adj_R2)),
    p_model = ifelse(is.na(p_model), "—", sprintf("%.3g", p_model))
  )
knitr::kable(
  tbl %>% select(Label, Form, n, R2, adj_R2, p_model, sig),
  col.names = c("Covariate", "Form", "n", "R²", "adj-R²", "p (model)", ""),
  caption   = "Table: Univariate screening — TT mothers.",
  align     = c("l","l","r","r","r","r","c")
) %>% print()
cat("\n")

if (length(s$selected) == 0) {
  cat("**No covariates reached p <", alpha, "in the univariate screen.**\n\n")
} else {
  cat(sprintf("**%d covariate(s) significant at p < %g:**\n\n",
              length(s$selected), alpha))

  form_df <- data.frame(
    Covariate     = names(s$selected),
    `Selected form` = unname(unlist(s$selected)),
    Note          = unname(unlist(s$form_notes)),
    stringsAsFactors = FALSE
  )
  knitr::kable(form_df,
    col.names = c("Covariate", "Selected predictor column", "Form selection rationale"),
    caption   = "Form selection — TT mothers."
  ) %>% print()
  cat("\n")

  cat("### Significant univariate models (TT)\n\n")
  for (lbl in names(s$selected)) {
    v   <- covariates_named_list[[lbl]]
    sel <- s$selected[[lbl]]
    key <- if (v %in% cat_vars) paste0(v, "_cat") else {
             if (endsWith(sel, "_T")) paste0(v, "_tert") else paste0(v, "_cont")
           }
    res <- s$univ_res[[key]]
    if (!is.null(res)) show_lm_clin(res)
  }

  if (length(s$selected) > 1 && !is.null(s$multi_result)) {
    cat("### Multivariate model (TT)\n\n")
    cat(sprintf(
      "All %d selected predictors entered jointly: %s\n\n",
      length(s$selected),
      paste(unlist(s$selected), collapse = ", ")
    ))
    show_lm_clin(s$multi_result)
  } else if (length(s$selected) == 1) {
    cat("*Only one predictor was significant; no multivariate model fitted.*\n\n")
  }
}
```

---

# Cross-Stratum Summary

```{r cross-stratum-summary, results='asis'}
# Build a combined table: for each covariate, show which strata found it significant
all_strata <- list(
  "aP (infant)"    = screens_iarm[["aP"]],
  "wP (infant)"    = screens_iarm[["wP"]],
  "TdaP (maternal)"= screens_marm[["TdaP"]],
  "TT (maternal)"  = screens_marm[["TT"]]
)
strata_names <- names(all_strata)

# Collect all (covariate label × form) combinations
all_rows <- bind_rows(lapply(all_strata, function(x) x[["univ_table"]]), .id = "Stratum") %>%
  mutate(
    p_fmt   = ifelse(is.na(p_model), "—", sprintf("%.3g", p_model)),
    sig_chr = ifelse(is.na(p_model), "—", pvalue_to_label(p_model))
  )

# Wide pivot: one row per (covariate, form), one column per stratum showing sig + p
summary_wide <- all_rows %>%
  select(Stratum, Label, Form, R2, adj_R2, p_fmt, sig_chr) %>%
  pivot_wider(
    id_cols     = c(Label, Form),
    names_from  = Stratum,
    values_from = c(R2, adj_R2, p_fmt, sig_chr),
    names_sep   = "__"
  )

# For display, show p_fmt and sig_chr merged per stratum
cat("## R² and Significance Across All Strata\n\n")
display_tbl <- all_rows %>%
  mutate(cell = sprintf("%.3f (%s)", R2, sig_chr)) %>%
  select(Stratum, Label, Form, cell) %>%
  pivot_wider(
    id_cols     = c(Label, Form),
    names_from  = Stratum,
    values_from = cell
  )
knitr::kable(
  display_tbl,
  col.names = c("Covariate", "Form", strata_names),
  caption   = "Table: R² (significance) across all four strata. * p<0.05  ** p<0.01  *** p<0.001",
  align     = c("l", "l", rep("r", length(strata_names)))
) %>% print()
cat("\n")

# Table of selected predictors per stratum
cat("## Selected Predictors per Stratum\n\n")
sel_rows <- lapply(names(all_strata), function(sn) {
  sc <- all_strata[[sn]]
  if (length(sc$selected) == 0) {
    data.frame(Stratum = sn, Covariate = "none", Selected_form = "—",
               Rationale = "—", stringsAsFactors = FALSE)
  } else {
    data.frame(
      Stratum       = sn,
      Covariate     = names(sc$selected),
      Selected_form = unname(unlist(sc$selected)),
      Rationale     = unname(unlist(sc$form_notes)),
      stringsAsFactors = FALSE
    )
  }
})
sel_df <- bind_rows(sel_rows)
knitr::kable(
  sel_df,
  col.names = c("Stratum", "Covariate", "Selected column", "Form selection rationale"),
  caption   = "Table: Predictors selected for multivariate models in each stratum.",
  align     = c("l","l","l","l")
) %>% print()
cat("\n")

# Multivariate R² summary
cat("## Multivariate Model Summary\n\n")
multi_tbl <- lapply(names(all_strata), function(sn) {
  sc  <- all_strata[[sn]]
  mr  <- sc$multi_result
  n_s <- length(sc$selected)
  data.frame(
    Stratum          = sn,
    N_sig_predictors = n_s,
    Multivariate_R2  = if (!is.null(mr)) sprintf("%.3f", mr$R2)  else
                       ifelse(n_s > 1, "model failed", "—"),
    Multivariate_adjR2 = if (!is.null(mr)) sprintf("%.3f", mr$adj_R2) else "—",
    Multivariate_p   = if (!is.null(mr)) sprintf("%.3g", mr$p_model) else "—",
    stringsAsFactors = FALSE
  )
})
multi_df <- bind_rows(multi_tbl)
knitr::kable(
  multi_df,
  col.names = c("Stratum", "N significant", "Multi R²", "Multi adj-R²", "Multi p"),
  caption   = "Table: Multivariate model fit per stratum.",
  align     = c("l","r","r","r","r")
) %>% print()
```

---

# Visualisations

## Forest-Style Plot: Univariate R² by Stratum

```{r forest-plot, fig.cap="Figure 1. Univariate R² for each covariate form across all four strata. Filled = p < 0.05; open = p ≥ 0.05. Both forms shown for continuous variables.", fig.height=7}
plot_df <- all_rows %>%
  mutate(
    sig_fill = !is.na(p_model) & p_model < alpha,
    strat_type = ifelse(grepl("infant", Stratum, ignore.case = TRUE),
                        "Infant arm", "Maternal arm"),
    Covariate_form = paste0(Label, "\n(", Form, ")")
  )

ggplot(plot_df,
       aes(x = R2, y = reorder(Covariate_form, R2),
           color = Stratum, shape = sig_fill)) +
  geom_point(size = 3, position = position_dodge(0.5)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 1),
                     labels = c(`TRUE` = paste0("p < ", alpha),
                                `FALSE` = paste0("p ≥ ", alpha))) +
  scale_color_manual(values = c(
    "aP (infant)"    = iarm_colors["aP"],
    "wP (infant)"    = iarm_colors["wP"],
    "TdaP (maternal)"= marm_colors["TdaP"],
    "TT (maternal)"  = marm_colors["TT"]
  )) +
  labs(x = "R² (univariate)", y = NULL,
       color = "Stratum", shape = "Significance",
       title = paste0("Univariate R² — ", outcome_label, " at M02")) +
  theme_bw(base_size = 11) +
  theme(axis.text.y = element_text(size = 8))
```

## Significant Univariate Predictors: Coefficient Plots

```{r coef-plots, fig.cap="Figure 2. Coefficients for significant univariate predictors by stratum. Error bars = 95% CI. Only significant models shown.", fig.height=5}
# Collect coefficients from all significant univariate models
sig_coef_rows <- lapply(names(all_strata), function(sn) {
  sc <- all_strata[[sn]]
  if (length(sc$selected) == 0) return(NULL)
  lapply(names(sc$selected), function(lbl) {
    v   <- covariates_named_list[[lbl]]
    sel <- sc$selected[[lbl]]
    key <- if (v %in% cat_vars) paste0(v, "_cat") else {
             if (endsWith(sel, "_T")) paste0(v, "_tert") else paste0(v, "_cont")
           }
    res <- sc$univ_res[[key]]
    if (is.null(res) || is.null(res$tidy_coef)) return(NULL)
    res$tidy_coef %>%
      filter(Predictor != "(Intercept)") %>%
      mutate(Stratum = sn, Covariate = lbl,
             lower95 = `β` - 1.96 * SE,
             upper95 = `β` + 1.96 * SE)
  }) %>% bind_rows()
}) %>% bind_rows()

if (!is.null(sig_coef_rows) && nrow(sig_coef_rows) > 0) {
  ggplot(sig_coef_rows,
         aes(x = `β`, y = Predictor, color = Stratum,
             xmin = lower95, xmax = upper95)) +
    geom_point(size = 2.5, position = position_dodge(0.5)) +
    geom_errorbarh(height = 0.2, position = position_dodge(0.5)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
    facet_wrap(~Covariate, scales = "free", ncol = 2) +
    scale_color_manual(values = c(
      "aP (infant)"    = iarm_colors["aP"],
      "wP (infant)"    = iarm_colors["wP"],
      "TdaP (maternal)"= marm_colors["TdaP"],
      "TT (maternal)"  = marm_colors["TT"]
    )) +
    labs(x = paste0("β (→ ", outcome_label, " M02)"),
         y = NULL, color = "Stratum",
         title = paste0("Significant predictor coefficients — ", outcome_label)) +
    theme_bw(base_size = 10) +
    theme(strip.text = element_text(size = 8))
} else {
  cat("*No significant univariate predictors found in any stratum.*\n")
}
```

---

# Synthesis and Interpretation

*Complete after reviewing the results above.*

## 1. Which Clinical Factors Predict PTNA, and in Which Strata?

> **[To complete]** Summarise which covariates passed the p < `r params$alpha`
> threshold. Are any significant across all four strata? Are any stratum-specific
> (e.g., maternal-arm-only, or infant-arm-only)? Note that the maternal-arm
> stratification (TdaP vs TT) splits by the maternal vaccine and thus by
> pertussis-antigen exposure, so predictors that differ between TdaP and TT
> may reflect confounding with the vaccine effect rather than independent biology.

## 2. Continuous vs Tertile: Which Form Best Captures the Relationship?

> **[To complete]** For each significant continuous variable, report which form
> was selected and by how much adj-R² differed between forms. A large difference
> in favour of tertiles suggests a non-linear or threshold relationship; a small
> difference suggests linearity is adequate.

## 3. Multivariate Models: Do the Clinical Factors Explain Independent Variance?

> **[To complete]** For strata where ≥ 2 predictors were significant, report the
> multivariate adj-R² and compare it to the univariate adj-R² values. Substantial
> drop in individual predictors' coefficients in the joint model relative to their
> univariate estimates suggests collinearity. A large joint R² compared to the
> individual R² values suggests additive/independent contributions.

## 4. Clinical vs Antibody Predictors: Relative Contribution?

> **[To complete]** The antibody-chain analysis
> (`predict_M02_PTNA_from_maternal_v2.Rmd`) showed that maternal PT-IgG at P09
> explains ~43–48% of infant PTNA variance in the infant arms. Compare the best
> clinical-predictor R² from this analysis to that benchmark. Are the clinical
> covariates weaker, comparable, or (unexpectedly) stronger predictors?

## 5. Implications for Confounder Adjustment?

> **[To complete]** If clinical covariates are significant predictors of PTNA,
> they are potential confounders (or effect modifiers) of the maternal-antibody
> → infant-PTNA relationship. Identify which, if any, should be considered as
> adjustment variables in the antibody models. Strong candidates are covariates
> that: (a) are significant here, and (b) are plausibly related to maternal IgG
> levels (e.g., gestational age at vaccine, maternal BMI, parity).

---

*Document generated by `predict_M02_PTNA_clinical_covariates_v1.Rmd`.*  
*Adapt `params$outcome_feature` to run for `WHOLE_SBA` or `WHOLE_WT_IgG`.*  
*Adapt `params$alpha` to change the screening threshold (default 0.05).*
