---
title: "Transition-response subgroup heterogeneity (per antigen × transition)"
output: html_document
params:
  analytes: !r c("IgG", "IgG1", "IgG2", "IgG3", "IgG4")
  min_complete: 20
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({ library(tidyverse) })

# ---- shared path resolver: single definition in R/find_proj_file.R ----
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R",
                "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R \u2014 place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}

source(find_proj_file("config/endpoints.R"),       local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"),      local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/within_visit_helpers.R"),  local = FALSE, encoding = "UTF-8")

data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))

TRANSITIONS <- list(
  c(from = "PregEarly", to = "MatBirth",  label = "Production (PregEarly\u2192MatBirth)"),
  c(from = "MatBirth",  to = "CordBlood", label = "Transfer (MatBirth\u2192CordBlood)"),
  c(from = "CordBlood", to = "InfMon2",   label = "Decay (CordBlood\u2192InfMon2)"))
tr_lab <- vapply(TRANSITIONS, `[`, "", "label")
```

# Transition-response subgroup heterogeneity

For each vaccine antigen (`r paste(ANTIGEN_ORDER, collapse = ", ")`) and each of the
three within-subject transitions, subjects are clustered on their per-analyte change
(Δlog₁₀ = later − earlier visit) across total IgG and IgG1–IgG4. The question is
whether the transition response is **heterogeneous** — i.e. splits into distinct
subgroups (for example strong vs weak placental transfer) — rather than a single
population. Clustering uses PAM with a silhouette-chosen *k* (2–4) when the `cluster`
package is available and ≥ `r params$min_complete` subjects have paired data;
otherwise a transparent tertile split on the total-IgG change is used. The maternal-arm
association column tests whether the detected subgroups line up with the randomised
maternal arm (Fisher exact, simulated *p*). This is an exploratory secondary output.

```{r scan}
scan <- transition_subgroup_scan(
  data_raw, ANTIGEN_ORDER, TRANSITIONS,
  analytes     = params$analytes,
  min_complete = as.integer(params$min_complete))

summ <- scan$summary %>%
  mutate(antigen    = factor(antigen,    levels = ANTIGEN_ORDER),
         transition = factor(transition, levels = tr_lab)) %>%
  arrange(transition, antigen)

knitr::kable(
  summ,
  col.names = c("Transition", "Antigen", "Method", "n", "Silhouette",
                "Subgroups (n, median Δ)", "p (arm assoc.)"),
  caption = paste0("Subgroup heterogeneity of the within-subject change, per antigen ",
                   "and per transition. Higher silhouette = cleaner separation; the ",
                   "subgroup column gives each subgroup's size and median total-IgG ",
                   "change. p (arm assoc.) tests subgroup \u00d7 maternal-arm independence."))
```

## Distribution of the total-IgG change by detected subgroup

```{r fig-dist, fig.width=11, fig.height=7}
asn <- scan$assignments %>%
  mutate(antigen    = factor(antigen,    levels = ANTIGEN_ORDER),
         transition = factor(transition, levels = tr_lab))

ggplot(asn, aes(x = antigen, y = total_delta, fill = subgroup)) +
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.4) +
  geom_boxplot(outlier.size = 0.5, linewidth = 0.35, position = position_dodge2(preserve = "single")) +
  facet_wrap(~ transition, ncol = 1, scales = "free_y") +
  scale_fill_brewer(palette = "RdBu", direction = -1, name = "Subgroup") +
  labs(x = "Antigen", y = expression("total-IgG " * Delta * "log"[10] * " (later \u2212 earlier)"),
       title = "Transition-response subgroups separate high- vs low-change subjects per antigen") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(),
        strip.text = element_text(face = "bold"))
```

*Each box is one detected subgroup within an antigen × transition cell; clean vertical
separation indicates the response is heterogeneous (distinct high/low-change subjects)
rather than a single population. Antigens whose subclasses are unavailable before
InfMon2 (FIM) are clustered on total IgG alone for the production and transfer panels.*
