---
title: "C0 · Data harmonisation & IgG standardisation"
subtitle: "Cross-batch MFI normalisation (standard-curve regression) and total-IgG standardisation of functional assays"
author: "Anja Saso and Michael Scot Zens"
output:
  workflowr::wflow_html:
    toc: true
    number_sections: false
    theme: cosmo
    highlight: textmate
    fig_path_ext: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(cgam)
library(MASS)
library(DT)
library(kableExtra)
library(scales)

# Force dplyr to win
select  <- dplyr::select
filter  <- dplyr::filter
mutate  <- dplyr::mutate
arrange <- dplyr::arrange

source(here::here("config", "endpoints.R"),            local = TRUE)
source(here::here("config", "endpoints_additions.R"),  local = TRUE)
source(here::here("R", "C0_connection_transform.R"), local = TRUE)
source(here::here("R", "colors.R"),               local = TRUE)

## Gate the slow standardisation refit (cgam nsim=100 x every feature).
## FALSE = document only; set TRUE to regenerate residuals into data/.
run_standardisation <- FALSE
```

## Overview

C0 prepares every downstream input for C1–C8/CK from the canonical
d_set.RData (ebaa_extra2). Two stages:

Harmonisation — cross-batch MFI normalisation. The maternal / cord-blood
assays (`mat_baa`) were run ~1 year after the infant assays (`inf_baa`), by
different staff, on different plate layouts, using different biosamples. The
**only** shared reference is the standard: the same composition, run over the
same dilution series in both batches. We therefore normalise `mat_baa` onto the
`inf_baa` scale by regressing the two batches' standard responses against each
other, pooled across plates, and apply the fitted relationship to the
maternal/cord test samples.
IgG standardisation — regressing each functional / subclass assay on
its antigen-specific total IgG; whole-bacterium assays are standardised
against a single chosen antigen (SBA & WT-IgG → PRN_IgG; PTNA → PT_IgG).
All intermediate outputs are written to data/ and are git-ignored.

## Stage 1 · Harmonisation
```{r c0sc-setup, include=FALSE}
## Child fragment of analysis/C0_data_harmonisation.Rmd
## Packages (tidyverse, knitr, kableExtra, DT, scales, htmltools) are loaded
## by the C0 parent setup chunk. Only fragment-specific data + helpers here.

load(file = here::here("data", "standards_sel.RData"))       # observed standards, both sets
load(file = here::here("data", "standard_grid_sel.RData"))   # 200-pt sigmoid fits, both sets
load(file = here::here("data", "ebaa_extra2_plated.RData"))  # plated test samples

# ── Normalisation configuration ───────────────────────────────────────────────
grid_analytes <- c("ADCD", "IgG1", "IgG2", "IgG3", "IgG4")            # grid `feature`
grid_antigens <- c("DT", "FHA", "IPV1", "IPV2", "IPV3", "PRN", "PT", "TT")
src_pref      <- c("NIBSC06_140", "SD")   # standard source: primary, then fallback
R2_MIN        <- 0.50                     # >= : concentration-dependent regression
                                          #  < : constant mean-shift fallback
norm_types    <- c("M", "C")              # maternal + cord normalised; B = reference
N_GRID        <- 200L                     # common dilution grid resolution

# ── Palettes ──────────────────────────────────────────────────────────────────
plate_levels <- sort(unique(standards_sel$plate))
n_plates     <- length(plate_levels)

pal_base <- c(
  "#1F78B4","#33A02C","#E31A1C","#FF7F00","#6A3D9A",
  "#B15928","#A6CEE3","#B2DF8A","#FB9A99","#FDBF6F",
  "#CAB2D6","#FFFF99","#1B9E77","#D95F02","#7570B3"
)
plate_colors  <- setNames(pal_base[seq_along(plate_levels)], plate_levels)
source_shapes <- c(NIBSC06_140 = 16, SD = 17)
source_lty    <- c(NIBSC06_140 = "solid", SD = "dashed")
set_colors    <- c(inf_baa = "#2171B5", mat_baa = "#CB181D")
method_colors <- c(regression = "#2171B5", mean_shift = "#D95F02")

# ── Observed standards: harmonised antigen base + shared log10(dilution) axis ──
# The maternal antigens carry a batch-code suffix (e.g. `pt_42`); antigen_base is
# the cross-set join key. standards_sel$concentration is dilution-derived (it
# falls as MFI rises); log10(dilution) is the axis both batches share exactly.
std <- standards_sel %>%
  dplyr::filter(!is.na(dilution), !is.na(mfi), dilution > 0) %>%
  dplyr::mutate(antigen_base = str_to_upper(str_remove(antigen, "_\\d+$")),
                ld           = log10(dilution))

# ── Core helper: build one mat->inf calibration for feature x antigen x source ─
# Pairs, at each shared dilution, observed INFANT MFI with the MATERNAL MFI read
# off the maternal predicted grid (pooled across plates), then fits
#   log10(inf_MFI) ~ log10(mat_MFI).
# Also returns the slope-1 mean offset (the mean-shift fallback) and the maternal
# MFI support used for out-of-range flagging.  Returns a 1-row tibble or NULL.
build_calibration <- function(ft, agn, src) {
  gm <- standard_grid_sel %>%
    dplyr::filter(set == "mat_baa", feature == ft, antigen == agn, source == src)
  mo <- std %>% dplyr::filter(set == "mat_baa", feature == ft, antigen_base == agn, source == src)
  io <- std %>% dplyr::filter(set == "inf_baa", feature == ft, antigen_base == agn, source == src)
  if (nrow(gm) == 0 || nrow(mo) == 0 || nrow(io) == 0) return(NULL)

  # map each maternal grid curve onto log10(dilution) via its own observed range.
  # Grid x rises with true concentration = FALLS with dilution, so orientation is
  # inverse: grid_x_min -> most dilute (ld_max);  grid_x_max -> least dilute (ld_min).
  rng <- mo %>% group_by(plate) %>%
    dplyr::summarise(dlo = min(ld), dhi = max(ld), nu = n_distinct(ld), .groups = "drop") %>%
    dplyr::filter(nu >= 3)
  if (nrow(rng) == 0) return(NULL)

  gm2 <- gm %>% inner_join(rng, by = "plate") %>%
    group_by(plate) %>%
    dplyr::mutate(ld = dhi + (concentration - min(concentration)) /
                    (max(concentration) - min(concentration)) * (dlo - dhi)) %>%
    ungroup()

  # pooled maternal curve: median grid MFI across plates on a common ld grid
  common <- seq(quantile(gm2$ld, 0.02), quantile(gm2$ld, 0.98), length.out = N_GRID)
  mat_by_plate <- gm2 %>% group_by(plate) %>%
    group_map(~ approx(arrange(.x, ld)$ld, arrange(.x, ld)$mfi, xout = common, rule = 1)$y)
  mat_pooled <- apply(do.call(rbind, mat_by_plate), 2, median, na.rm = TRUE)

  # infant observed anchors -> pooled maternal MFI at the same dilution
  io <- io %>%
    dplyr::mutate(log10_inf = log10(pmax(mfi, 0.5)),
                  log10_mat = approx(common, mat_pooled, xout = ld, rule = 1)$y) %>%
    dplyr::filter(is.finite(log10_inf), is.finite(log10_mat))
  if (nrow(io) < 10) return(NULL)

  fit <- lm(log10_inf ~ log10_mat, data = io)
  tibble(
    feature = ft, antigen = agn, source = src, n = nrow(io),
    intercept = unname(coef(fit)[1]), slope = unname(coef(fit)[2]),
    r_squared = summary(fit)$r.squared,
    shift     = mean(io$log10_inf - io$log10_mat),   # slope-1 (mean-shift) offset
    mat_lo    = min(io$log10_mat), mat_hi = max(io$log10_mat)
  )
}
```

## Approach

`mat_baa` is placed on the `inf_baa` scale one **feature × antigen** at a time,
pooled across plates (plate identity does not carry between the two runs):

1. Both batches ran the standard over the **same dilution series**, so
   `log10(dilution)` is the shared x-axis. At each dilution we pair the
   **observed infant MFI** with the **maternal MFI** read off the maternal
   predicted standard-curve grid (linearly interpolated; each grid curve is first
   mapped back onto the dilution axis using its own observed range).
2. Fit `log10(inf_MFI) ~ log10(mat_MFI)`. If `R² ≥ R2_MIN` the fitted line is the
   **concentration-dependent** normalisation. If `R² < R2_MIN` (a near-flat,
   low-signal standard where the slope is ill-determined) we fall back to a
   **constant mean-shift** — `log10(norm) = log10(raw) + mean(log10_inf − log10_mat)` —
   i.e. a slope-1, batch-offset correction for that feature × antigen.
3. Apply to the maternal/cord (`sample_type` M, C) `mbaa` test samples. Infant
   (`B`) samples are the reference scale and are left unchanged. Everything stays
   in MFI space — nothing is expressed in concentration units.

The exported `maternal_to_infant_calibration.csv` (written to `data/`) carries the
per feature × antigen coefficients and the method used, and can be joined to any
sample frame to reproduce the normalisation.

---

## 1  Data Overview {.tabset}

### Feature × Antigen × Set

```{r c0sc-overview-structure}
set_antigen_tbl <- std %>%
  group_by(set, feature, antigen_base) %>%
  dplyr::summarise(
    n_obs     = n(),
    n_plates  = n_distinct(plate),
    n_sources = n_distinct(source),
    n_dil     = n_distinct(dilution),
    .groups   = "drop"
  )

overlap_tbl <- set_antigen_tbl %>%
  pivot_wider(
    id_cols     = c(feature, antigen_base),
    names_from  = set,
    values_from = c(n_obs, n_plates, n_sources, n_dil)
  )

datatable(overlap_tbl,
          caption  = "Observed standard rows per feature x antigen by set (NA = absent)",
          filter   = "top", rownames = FALSE,
          options  = list(scrollX = TRUE, pageLength = 15))
```

### Dilution Ranges by Set

```{r c0sc-overview-dilutions}
dil_tbl <- std %>%
  group_by(set, feature, antigen_base) %>%
  dplyr::summarise(
    n_dil   = n_distinct(dilution),
    dil_min = min(dilution),
    dil_max = max(dilution),
    .groups = "drop"
  ) %>%
  dplyr::arrange(feature, antigen_base, set)

datatable(dil_tbl, caption = "Dilution range per feature x antigen x set",
          filter = "top", rownames = FALSE,
          options = list(scrollX = TRUE, pageLength = 15))
```

### Antigen Name Mapping

*`mat_baa` antigens carry a batch-code suffix (e.g. `pt_42`). The base name
(`antigen_base`) is the cross-set join key throughout.*

```{r c0sc-overview-antigen-map}
antigen_map <- standards_sel %>%
  dplyr::mutate(antigen_base = str_to_upper(str_remove(antigen, "_\\d+$"))) %>%
  distinct(set, antigen, antigen_base) %>%
  pivot_wider(names_from = set, values_from = antigen,
              values_fn = ~ paste(unique(.x), collapse = " / ")) %>%
  dplyr::arrange(antigen_base)

datatable(antigen_map,
          caption  = "Antigen naming across sets (antigen_base = join key)",
          rownames = FALSE,
          options  = list(pageLength = 25, dom = "t"))
```

### Grid Coverage

*The sigmoid standard-curve grid exists for a subset of features and antigens;
normalisation is only defined where the grid and both observed sets overlap.*

```{r c0sc-overview-grid}
grid_cov <- standard_grid_sel %>%
  group_by(feature, antigen) %>%
  dplyr::summarise(sets    = paste(sort(unique(set)), collapse = "/"),
                   plates  = n_distinct(plate),
                   sources = paste(sort(unique(source)), collapse = "/"),
                   pts     = n(), .groups = "drop")

datatable(grid_cov, caption = "standard_grid_sel coverage (feature x antigen)",
          filter = "top", rownames = FALSE,
          options = list(scrollX = TRUE, pageLength = 20))
```

---

## 2  Paired Standard Reference

At every shared dilution the observed infant MFI is paired with the pooled
maternal grid MFI, producing the regression training set for each
feature × antigen.

```{r c0sc-paired-reference}
build_pairs <- function(ft, agn, src) {
  gm <- standard_grid_sel %>%
    dplyr::filter(set == "mat_baa", feature == ft, antigen == agn, source == src)
  mo <- std %>% dplyr::filter(set == "mat_baa", feature == ft, antigen_base == agn, source == src)
  io <- std %>% dplyr::filter(set == "inf_baa", feature == ft, antigen_base == agn, source == src)
  if (nrow(gm) == 0 || nrow(mo) == 0 || nrow(io) == 0) return(NULL)
  rng <- mo %>% group_by(plate) %>%
    dplyr::summarise(dlo = min(ld), dhi = max(ld), nu = n_distinct(ld), .groups = "drop") %>%
    dplyr::filter(nu >= 3)
  if (nrow(rng) == 0) return(NULL)
  gm2 <- gm %>% inner_join(rng, by = "plate") %>%
    group_by(plate) %>%
    dplyr::mutate(ld = dhi + (concentration - min(concentration)) /
                    (max(concentration) - min(concentration)) * (dlo - dhi)) %>%
    ungroup()
  common <- seq(quantile(gm2$ld, 0.02), quantile(gm2$ld, 0.98), length.out = N_GRID)
  mat_pooled <- apply(
    do.call(rbind, gm2 %>% group_by(plate) %>%
              group_map(~ approx(arrange(.x, ld)$ld, arrange(.x, ld)$mfi, xout = common, rule = 1)$y)),
    2, median, na.rm = TRUE)
  io %>%
    dplyr::transmute(feature = ft, antigen = agn, source = src, ld,
                     log10_inf = log10(pmax(mfi, 0.5)),
                     log10_mat = approx(common, mat_pooled, xout = ld, rule = 1)$y) %>%
    dplyr::filter(is.finite(log10_inf), is.finite(log10_mat))
}

# resolve source per (feature, antigen): NIBSC first, else SD
src_choice <- map_dfr(grid_analytes, function(ft) {
  map_dfr(grid_antigens, function(agn) {
    for (src in src_pref) {
      p <- build_pairs(ft, agn, src)
      if (!is.null(p) && nrow(p) >= 10) return(tibble(feature = ft, antigen = agn, source = src))
    }
    tibble()
  })
})

pairs_tbl <- src_choice %>%
  pmap_dfr(function(feature, antigen, source) build_pairs(feature, antigen, source))

cat(sprintf("Feature x antigen with a paired reference : %d\n", nrow(src_choice)))
cat(sprintf("Paired standard points (total)            : %d\n", nrow(pairs_tbl)))
```

---

## 3  Calibration Fits

`log10(inf_MFI) ~ f(log10(mat_MFI))` is fitted per feature × antigen (pooled
across plates). With `CALIB_METHOD = "spline"` the mapping is a **monotone-
increasing constrained smoother** (`cgam::s.incr`), which is shape-free but
order-preserving. Low-signal or point-poor standards fall back to a constant
**mean-shift**. `CALIB_METHOD = "linear"` restores the legacy regression /
mean-shift behaviour (`R² ≥ `r R2_MIN``).

```{r c0sc-fit-calibrations}
# ── Config reminder (define these in the c0sc-setup chunk) ───────────────────
#   CALIB_METHOD <- "spline"   # "linear" (legacy) | "spline" (constrained GAM)
#   N_CAL_GRID   <- 200L       # resolution of the exported calibration lookup
#   MIN_PTS_SPL  <- 20L        # min paired points to attempt a spline
#   MIN_SPAN_LD  <- 0.5        # min log10(mat) span (decades) to attempt a spline

if (!exists("CALIB_METHOD")) CALIB_METHOD <- "spline"
if (!exists("N_CAL_GRID"))   N_CAL_GRID   <- 200L
if (!exists("MIN_PTS_SPL"))  MIN_PTS_SPL  <- 20L
if (!exists("MIN_SPAN_LD"))  MIN_SPAN_LD  <- 0.5

# ── Core fit: returns a 1-row tibble carrying a nested lookup grid, or NULL ────
# Reuses build_pairs() for the (log10_mat, log10_inf) training set, then fits
# either a monotone-increasing spline or the legacy linear/mean-shift model.
# In every case the fitted mapping is materialised as a dense, monotone lookup
# grid over [mat_lo, mat_hi] so it can be applied by interpolation and exported
# as a flat CSV (no model objects persisted).
build_calibration <- function(ft, agn, src) {
  pr <- build_pairs(ft, agn, src)
  if (is.null(pr) || nrow(pr) < 3) return(NULL)

  mat_lo <- min(pr$log10_mat, na.rm = TRUE)
  mat_hi <- max(pr$log10_mat, na.rm = TRUE)
  if (!is.finite(mat_lo) || !is.finite(mat_hi) || mat_hi <= mat_lo) return(NULL)

  xg   <- seq(mat_lo, mat_hi, length.out = N_CAL_GRID)
  span <- mat_hi - mat_lo

  # Diagnostics computed for every feature x antigen (feed §6 heatmaps as before)
  fit_lm <- lm(log10_inf ~ log10_mat, data = pr)
  r2     <- summary(fit_lm)$r.squared
  slope  <- unname(coef(fit_lm)[2])
  interc <- unname(coef(fit_lm)[1])
  shift  <- mean(pr$log10_inf - pr$log10_mat)

  # Decide method
  want_spline <- CALIB_METHOD == "spline" &&
                 nrow(pr) >= MIN_PTS_SPL &&
                 span     >= MIN_SPAN_LD

  method <- NA_character_; yg <- NULL

  if (want_spline) {
    fit_sp <- tryCatch(
      cgam::cgam(log10_inf ~ s.incr(log10_mat), data = pr),
      error = function(e) NULL
    )
    if (!is.null(fit_sp)) {
      yg <- tryCatch(
        as.numeric(predict(fit_sp,
                    newData = data.frame(log10_mat = xg))$fit),
        error = function(e) NULL
      )
    }
    if (!is.null(yg) && all(is.finite(yg))) {
      yg     <- cummax(yg)          # defensive numerical monotonicity
      method <- "spline_incr"
    }
  }

  # Fallbacks (also the whole path when CALIB_METHOD == "linear")
  if (is.na(method)) {
    if (CALIB_METHOD == "linear" && r2 >= R2_MIN) {
      yg <- interc + slope * xg
      method <- "regression"
    } else {
      yg <- xg + shift
      method <- "mean_shift"
    }
  }

  tibble(
    feature = ft, antigen = agn, source = src, n = nrow(pr),
    method  = method, r_squared = r2,
    intercept = interc, slope = slope, shift = shift,
    mat_lo = mat_lo, mat_hi = mat_hi,
    grid = list(tibble(log10_mat_grid = xg, log10_inf_pred = yg))
  )
}

calibration <- src_choice %>%
  pmap_dfr(function(feature, antigen, source)
             build_calibration(feature, antigen, source))

cat(sprintf("Calibration method            : %s\n", CALIB_METHOD))
cat(sprintf("Calibrations built            : %d\n", nrow(calibration)))
cat(sprintf("  monotone spline             : %d\n",
            sum(calibration$method == "spline_incr")))
cat(sprintf("  linear regression           : %d\n",
            sum(calibration$method == "regression")))
cat(sprintf("  mean-shift fallback         : %d\n",
            sum(calibration$method == "mean_shift")))
```

---

## 4  Apply Normalisation to Test Samples

The calibration is applied to maternal/cord `mbaa` rows by **monotone
interpolation** against the fitted lookup grid; infant (`B`) rows are the
reference scale and pass through unchanged. `rule = 2` clamps predictions at the
fitted support ends, and `out_of_range` still flags any raw MFI beyond
`[mat_lo, mat_hi]` for downstream censoring.

```{r c0sc-apply}
apply_lookup <- calibration %>%
  dplyr::select(analyte = feature, antigen, norm_source = source,
                norm_method = method, norm_slope = slope,
                norm_intercept = intercept, norm_shift = shift,
                norm_r2 = r_squared, mat_lo, mat_hi)

# Long-format lookup grid, keyed the same way as apply_lookup
apply_grid <- calibration %>%
  dplyr::select(analyte = feature, antigen, grid) %>%
  tidyr::unnest(grid)

ebaa <- ebaa_extra2_plated %>% dplyr::mutate(.row = row_number())

# Per feature x antigen interpolation helper (monotone grid -> predicted infant)
interp_norm <- function(x_raw, ana, agn) {
  g <- apply_grid %>% dplyr::filter(analyte == ana, antigen == agn)
  if (nrow(g) < 2) return(rep(NA_real_, length(x_raw)))
  approx(g$log10_mat_grid, g$log10_inf_pred, xout = x_raw, rule = 2)$y
}

applied <- ebaa %>%
  dplyr::filter(assay == "mbaa", sample_type %in% norm_types,
                !is.na(value_reported), value_reported > 0) %>%
  inner_join(apply_lookup, by = c("analyte", "antigen")) %>%
  dplyr::mutate(log10_mfi_raw = log10(value_reported)) %>%
  group_by(analyte, antigen) %>%
  dplyr::mutate(
    log10_mfi_norm = interp_norm(log10_mfi_raw,
                                 dplyr::first(analyte),
                                 dplyr::first(antigen)),
    mfi_norm       = 10^log10_mfi_norm,
    out_of_range   = log10_mfi_raw < mat_lo | log10_mfi_raw > mat_hi,
    norm_applied   = TRUE
  ) %>%
  ungroup() %>%
  dplyr::select(.row, norm_source, norm_method, norm_slope, norm_intercept,
                norm_shift, norm_r2, log10_mfi_norm, mfi_norm,
                out_of_range, norm_applied)

ebaa_extra2_normalised <- ebaa %>%
  left_join(applied, by = ".row") %>%
  dplyr::mutate(
    norm_applied = coalesce(norm_applied, FALSE),
    norm_status  = dplyr::case_when(
      norm_applied                                              ~ paste0("normalized_", norm_method),
      assay != "mbaa"                                           ~ "not_mbaa",
      sample_type == "B"                                        ~ "reference_infant",
      !sample_type %in% norm_types                              ~ "other_sampletype",
      !analyte %in% grid_analytes | !antigen %in% grid_antigens ~ "no_curve_feature",
      is.na(value_reported) | value_reported <= 0              ~ "nonpositive_mfi",
      TRUE                                                      ~ "no_curve_feature"
    ),
    # analysis-ready infant-scale MFI: normalised for M/C, raw for infant B
    log10_mfi_infantscale = dplyr::case_when(
      norm_applied                                           ~ log10_mfi_norm,
      norm_status == "reference_infant" & value_reported > 0 ~ log10(value_reported),
      TRUE                                                   ~ NA_real_
    ),
    mfi_infantscale = 10^log10_mfi_infantscale
  ) %>%
  dplyr::select(-.row)

cat(sprintf("Rows normalised (maternal/cord mbaa): %d\n",
            sum(ebaa_extra2_normalised$norm_applied)))
```

```{r c0sc-apply-summary}
status_tbl <- ebaa_extra2_normalised %>%
  dplyr::filter(assay == "mbaa", sample_type %in% norm_types) %>%
  dplyr::count(norm_status, name = "n") %>%
  dplyr::arrange(desc(n))

datatable(status_tbl, caption = "Normalisation status (maternal/cord mbaa scope)",
          rownames = FALSE, options = list(dom = "t"))

cat(sprintf("\nOut-of-range (flagged) among normalised: %d (%.1f%%)\n",
            sum(ebaa_extra2_normalised$out_of_range, na.rm = TRUE),
            100 * mean(ebaa_extra2_normalised$out_of_range[ebaa_extra2_normalised$norm_applied])))
```

---

## 5  Validation: Normalised standards vs observed infant {.tabset}

Applying each calibration back to its own maternal standard points should recover
the infant standard response (residual ≈ 0 in log10 space).

```{r c0sc-validation}
validation <- pairs_tbl %>%
  left_join(calibration, by = c("feature", "antigen", "source")) %>%
  dplyr::mutate(
    log10_norm    = if_else(method == "regression",
                            intercept + slope * log10_mat, log10_mat + shift),
    residual_log10 = log10_norm - log10_inf
  ) %>%
  dplyr::filter(is.finite(residual_log10))

resid_summary <- validation %>%
  group_by(feature, antigen, method) %>%
  dplyr::summarise(bias = round(mean(residual_log10), 3),
                   rmse = round(sqrt(mean(residual_log10^2)), 3),
                   .groups = "drop")

datatable(resid_summary,
          caption = "Post-normalisation residuals on the standards (target bias = 0)",
          filter = "top", rownames = FALSE,
          options = list(scrollX = TRUE, pageLength = 20))
```

```{r c0sc-validate-plot, fig.width=16, fig.height=9}
ggplot(validation,
       aes(x = log10_mat, y = residual_log10, colour = method)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(size = 0.9, alpha = 0.4) +
  scale_colour_manual(values = method_colors) +
  facet_grid(feature ~ antigen, scales = "free_x") +
  labs(title = "Post-normalisation residuals on the standard curves",
       subtitle = "log10(normalised maternal) - log10(observed infant); target = 0",
       x = "log10(maternal MFI)", y = "residual (log10)", colour = "Method") +
  theme_bw(base_size = 7) +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 6))
```

---

## 6  Calibration Summary Heatmaps

### 6.1  Cross-batch slope (concentration dependence)

```{r c0sc-heatmap-slope, fig.width=13, fig.height=6}
ggplot(calibration, aes(x = antigen, y = feature, fill = slope)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  geom_text(aes(label = ifelse(method == "regression", sprintf("%.2f", slope), "shift")),
            size = 2.6, colour = "black") +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#D6604D",
                       midpoint = 1, name = "slope") +
  labs(title = "Maternal->infant slope (1.0 = pure scale match)",
       subtitle = "cells labelled 'shift' used the mean-shift fallback (R2 < threshold)",
       x = "Antigen", y = "Feature") +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 6.2  Batch offset (mean log10 shift)

```{r c0sc-heatmap-shift, fig.width=13, fig.height=6}
ggplot(calibration, aes(x = antigen, y = feature, fill = shift)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  geom_text(aes(label = sprintf("%.2f", shift)), size = 2.6, colour = "black") +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#D6604D",
                       midpoint = 0, name = "log10\nshift") +
  labs(title = "Mean batch offset (log10 maternal -> infant)",
       x = "Antigen", y = "Feature") +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

## 7  Export

Two calibration artefacts are written to `data/` (git-ignored): a human-readable
per feature × antigen **summary** (method + linear diagnostics + support), and a
long-format **lookup grid** that reproduces the (possibly nonlinear) mapping by
monotone interpolation — the same mechanism used in §4.

```{r c0sc-export}
# 1) Summary table (drop the nested grid list-column) -> data/
write.csv(
  calibration %>%
    dplyr::select(feature, antigen, source, method, n, slope, intercept,
                  shift, r_squared, mat_lo, mat_hi),
  c0_out("maternal_to_infant_calibration.csv"), row.names = FALSE
)

# 2) Long-format lookup grid (the applicable calibration) -> data/
calibration_grid <- calibration %>%
  dplyr::select(feature, antigen, source, method, grid) %>%
  tidyr::unnest(grid) %>%
  dplyr::select(feature, antigen, source, method,
                log10_mat_grid, log10_inf_pred)

write.csv(
  calibration_grid,
  c0_out("maternal_to_infant_calibration_grid.csv"), row.names = FALSE
)

# 3) Normalised, plated test-sample frame -> data/
save(ebaa_extra2_normalised, file = here::here("data", "ebaa_extra2_normalised.RData"))

cat(sprintf("maternal_to_infant_calibration.csv      : %d rows -> data/\n",
            nrow(calibration)))
cat(sprintf("maternal_to_infant_calibration_grid.csv : %d rows -> data/\n",
            nrow(calibration_grid)))
cat(sprintf("ebaa_extra2_normalised.RData            : %d rows -> data/\n",
            nrow(ebaa_extra2_normalised)))
```

---

## 8  Interpretation Guide

| Observation | Meaning | Action |
|---|---|---|
| `method = regression`, slope ≈ 1 | Batches differ by a near-constant scale | Use as-is |
| `method = regression`, slope ≠ 1 | Correction depends on MFI level (concentration-dependent) | Use as-is |
| `method = mean_shift` | Standard too flat/low-signal to fit a slope; constant offset applied | Treat those analytes cautiously |
| Residuals in §5 centred on 0 | Calibration recovers the infant scale | Proceed |
| `out_of_range = TRUE` | Sample MFI beyond the standard range (extrapolated) | Flag/censor downstream as needed |

**Core formula (apply to any maternal/cord sample MFI):**

```
# regression:  log10(norm_MFI) = intercept + slope * log10(raw_MFI)
# mean_shift:  log10(norm_MFI) = log10(raw_MFI) + shift
# join maternal_to_infant_calibration.csv by (feature/analyte, antigen)
```

Use `log10_mfi_infantscale` (or `mfi_infantscale`) as the analysis-ready,
batch-harmonised value: normalised for maternal/cord, observed for infant.

## Stage 2 · IgG standardisation {.tabset}

```{r standardise, results="asis"}
if (run_standardisation) {
  for (i in seq_len(nrow(std_config))) {
    cfg_row <- std_config[i, ]
    res <- knitr::knit_child(
  here::here("analysis", "parts", "C0_standardise.Rmd"),
  quiet = TRUE
)
    cat(res, sep = "\n")
  }
} else {
  cat("\n*Standardisation refit gated off ",
      "(`run_standardisation <- FALSE`). ",
      "Residual files in `data/` are used as-is by C1/C4/C5/C7.*\n")
}
```
