---
title: "C0 · Data harmonisation & IgG standardisation"
subtitle: "Cross-batch MFI normalisation and total-IgG standardisation of functional assays"
author: "Anja Saso and Michael Scot Zens"
output:
  workflowr::wflow_html:
    toc: true
    number_sections: false
    theme: cosmo
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(cgam)
library(MASS)
library(DT)
library(kableExtra)
library(scales)

# Force dplyr to win
select  <- dplyr::select
filter  <- dplyr::filter
mutate  <- dplyr::mutate
arrange <- dplyr::arrange

source(here::here("config", "endpoints.R"),            local = TRUE)
source(here::here("config", "endpoints_additions.R"),  local = TRUE)
source(here::here("R", "C0_connection_transform.R"), local = TRUE)
source(here::here("R", "colors.R"),               local = TRUE)

## Gate the slow standardisation refit (cgam nsim=100 x every feature).
## FALSE = document only; set TRUE to regenerate residuals into data/.
run_standardisation <- FALSE
```

## Overview

C0 prepares every downstream input for C1–C8/CK from the canonical
d_set.RData (ebaa_extra2). Two stages:

Harmonisation — cross-batch MFI normalisation via standard-curve
back-interpolation (inf_baa ↔ mat_baa scales).
IgG standardisation — regressing each functional / subclass assay on
its antigen-specific total IgG; whole-bacterium assays are standardised
against a single chosen antigen (SBA & WT-IgG → PRN_IgG; PTNA → PT_IgG).
All intermediate outputs are written to data/ and are git-ignored.

## Stage 1 · Harmonisation
``` {r harmonise, child = here::here("analysis","parts","C0_standards_comparison.Rmd")}


```

## Stage 2 · IgG standardisation {.tabset}

```{r standardise, results="asis"}
if (run_standardisation) {
  for (i in seq_len(nrow(std_config))) {
    cfg_row <- std_config[i, ]
    res <- knitr::knit_child(
      here::here("analysis","parts","C0_standardise.Rmd"),
      envir = environment(), quiet = TRUE)
    cat(res, sep = "\n")
  }
} else {
  cat("\n*Standardisation refit gated off ",
      "(`run_standardisation <- FALSE`). ",
      "Residual files in `data/` are used as-is by C1/C4/C5/C7.*\n")
}



```

