---
title: "Maternal Pre-Vaccination → At Birth: Within-Mother Antibody Change"
subtitle: "Paired within-subject change (maternal pre-vacc → maternal at birth) — missing data structure, direction, and per-antigen / per-analyte distributions"
author: "Scot Zens"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r bootstrap, include=FALSE}
library(here)
# ---- shared path resolver: single definition in R/shared_utils.R ----
if (!exists("shared_utils", mode = "function")) {
  .ff_cand <- c("R/shared_utils.R", "../R/shared_utils.R",
                "analysis/R/shared_utils.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/shared_utils.R"),
                  here::here("analysis/R/shared_utils.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/shared_utils.R"),
                  file.path(dirname(.ff_rmd), "R/shared_utils.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/shared_utils.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}
source(shared_utils("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C1_serology_helpers.R"), local = FALSE, encoding = "UTF-8")
data_raw <- load_serology_data(shared_utils("data/c_set.RData"))
nt <- new_ntracker()

# Actual visit_name levels in the dataset:
#   CordBlood, InfMon2, InfMon4, InfMon5, InfMon9, MatBirth, PregEarly
# Maternal pre-vaccination = "PregEarly"; maternal at birth = "MatBirth".
FROM_VISIT <- "PregEarly"
TO_VISIT   <- "MatBirth"
MIN_PAIRED <- 10L
```

# The Question

Maternal antibody is measured at an **early-pregnancy, pre-vaccination visit** (`PregEarly`)
and again **at birth/delivery** (`MatBirth`). Between these two visits the mother receives
the study vaccine (a pertussis-containing TdaP or the TT control), which is expected to
**boost** vaccine-targeted antibody. Whether a given feature is higher or lower by delivery
depends on the **interval between vaccination and birth**: a short interval may catch
antibody still rising, while a longer interval may capture post-peak waning. Direction is
therefore **not assumed** and is expected to vary by antigen and by analyte.

**Convention:** delta = log10 value at **maternal birth** − log10 value at
**maternal pre-vacc**. Positive delta = an *increase* by delivery; negative delta = a
*decrease* by delivery.

# Data and Coverage

```{r coverage, results='asis'}
cov_raw <- paired_coverage(data_raw, FROM_VISIT, TO_VISIT)
# Standardise column names regardless of what paired_coverage() returns.
# Columns are always: feature, n_at_from_visit, n_at_to_visit, n_paired.
names(cov_raw) <- c("feature", "n_cord", "n_inf", "n_paired")
cov <- cov_raw %>% arrange(desc(n_paired))

for (i in seq_len(nrow(cov)))
  nt_log_manual(nt, "PreVacc→Birth", "Coverage", cov$feature[i], "paired availability",
                n_entered = max(cov$n_cord[i], cov$n_inf[i]), n_used = cov$n_paired[i],
                note = sprintf("PregEarly n=%d, MatBirth n=%d", cov$n_cord[i], cov$n_inf[i]))
cat(sprintf("Features with at least one paired subject: **%d**; with ≥ %d paired: **%d**.\n\n",
            nrow(cov), MIN_PAIRED, sum(cov$n_paired >= MIN_PAIRED)))
knitr::kable(cov, col.names = c("Feature", "n PregEarly", "n MatBirth", "n paired (both)"),
             caption = "Table 1. Per-feature coverage: subjects with PregEarly, with MatBirth, and with BOTH (the n used). Features with n paired = 0 were not measured at one visit and are excluded from all change analyses.") %>% print()
```

```{r compute, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

paired <- paired_change(data_raw, FROM_VISIT, TO_VISIT) %>%
  mutate(
    feature_class = case_when(
      grepl("_IgG$",    feature)      ~ "Total IgG",
      grepl("_IgG[1-4]$", feature)   ~ "IgG subclass",
      grepl("_FcgR",    feature)      ~ "FcR binding",
      grepl("_AD",      feature)      ~ "Effector function",
      grepl("_Bmem$",   feature)      ~ "B memory",
      TRUE                            ~ "Other"),
    antigen = sub("_.*", "", feature),
    analyte = sub("^[^_]+_", "", feature)
  )

feat_summary <- summarise_paired(paired, min_n = MIN_PAIRED) %>%
  left_join(distinct(paired, feature, antigen, analyte, feature_class), by = "feature") %>%
  arrange(desc(pct_inc))

# headline diagnostics
n_feat         <- nrow(feat_summary)
n_mostly_inc   <- sum(feat_summary$direction == "mostly increase")
n_mostly_dec   <- sum(feat_summary$direction == "mostly decrease")
n_mixed        <- sum(feat_summary$direction == "mixed")
pooled_med     <- round(median(paired$delta), 3)
pooled_pct_inc <- round(100 * mean(paired$delta > 0), 1)

# total-IgG features
tot_igg       <- feat_summary %>% filter(feature_class == "Total IgG")
n_tot_igg_inc <- sum(tot_igg$direction == "mostly increase")

# per-subject mean shift
subj_off <- paired %>% group_by(subject_accession) %>%
  summarize(mean_delta = mean(delta), n_feat = n(), .groups = "drop")
subj_pct_pos <- round(100 * mean(subj_off$mean_delta > 0), 1)
subj_med_off <- round(median(subj_off$mean_delta), 3)

# per-antigen summaries
antigen_sum <- feat_summary %>%
  group_by(antigen) %>%
  summarize(
    n_features = n(),
    n_inc      = sum(direction == "mostly increase"),
    n_dec      = sum(direction == "mostly decrease"),
    n_mix      = sum(direction == "mixed"),
    med_delta  = round(median(median_delta), 3),
    dominant   = case_when(
      n_inc / n_features >= 0.6 ~ "predominantly increasing",
      n_dec / n_features >= 0.6 ~ "predominantly decreasing",
      TRUE                      ~ "mixed"),
    .groups = "drop"
  ) %>%
  arrange(med_delta)

# per-analyte summaries (for Figure 5 narrative)
analyte_sum <- feat_summary %>%
  group_by(analyte) %>%
  summarize(
    n_features  = n(),
    n_inc       = sum(direction == "mostly increase"),
    n_dec       = sum(direction == "mostly decrease"),
    n_mix       = sum(direction == "mixed"),
    med_delta   = round(median(median_delta), 3),
    pct_subj_inc = round(100 * mean(pct_inc / 100), 1),
    dominant    = case_when(
      n_inc / n_features >= 0.6 ~ "predominantly increasing",
      n_dec / n_features >= 0.6 ~ "predominantly decreasing",
      TRUE                      ~ "mixed"),
    .groups = "drop"
  ) %>%
  arrange(med_delta)

# helper: build per-analyte histogram plot
make_analyte_hist <- function(analyte_label, fill_col = "steelblue") {
  dat <- paired %>% filter(feature %in% feat_summary$feature, analyte == analyte_label)
  if (nrow(dat) == 0) return(NULL)
  feat_med_df <- dat %>%
    group_by(feature, antigen) %>%
    summarize(feat_med  = median(delta, na.rm = TRUE),
              n_subj    = n_distinct(subject_accession),
              .groups   = "drop") %>%
    mutate(feature_fct = reorder(feature, feat_med))
  dat <- dat %>%
    mutate(feature_fct = factor(feature, levels = levels(feat_med_df$feature_fct)))
  ggplot(dat, aes(x = delta)) +
    geom_histogram(bins = 30, fill = fill_col, color = "white", alpha = 0.85) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    geom_vline(data = feat_med_df, aes(xintercept = feat_med),
               color = "firebrick", linewidth = 0.8) +
    geom_text(data = feat_med_df,
              aes(x = feat_med, y = Inf,
                  label = sprintf("med=%+.2f\nn=%d", feat_med, n_subj)),
              vjust = 1.3,
              hjust = ifelse(feat_med_df$feat_med >= 0, -0.08, 1.08),
              size = 2.5, color = "firebrick") +
    facet_wrap(~ feature_fct, scales = "free_y") +
    labs(
      x     = "delta (log10): later − earlier visit",
      y     = "Subjects",
      title = sprintf("%s features by antigen (later vs earlier visit)", analyte_label),
      subtitle = "One panel per feature (antigen), ordered by median delta. Dashed = zero; red = feature median."
    ) +
    theme_bw(base_size = 10) +
    theme(strip.background = element_rect(fill = "grey92"),
          strip.text       = element_text(face = "bold", size = 8.5),
          panel.spacing    = unit(0.5, "lines"))
}

# Colour palette per analyte
analyte_colours <- c(
  "IgG"    = "#2166ac",
  "IgG1"   = "#4dac26",
  "IgG2"   = "#d01c8b",
  "IgG3"   = "#f1a340",
  "IgG4"   = "#7b3294",
  "WT_IgG" = "#636363"
)
```


# Missing Data: Sources, Structure, and Implications

Not all subjects contribute equally to every feature's paired analysis, and the reasons
differ in important ways. Understanding the missing-data structure is essential before
interpreting the change estimates: subjects who are missing at one or both visits may
differ systematically from those with complete data, and the pattern of missingness itself
reveals aspects of the assay design that affect what questions can be answered.

## Why paired counts differ across features

There are four distinct sources of missingness in this dataset, each with a different
cause and a different implication for bias:

**Type A — Analyte never assayed at one visit (structural panel gap).**
Some analytes may not have been run at one of the two visits at all, giving n = 0 there.
This is a design decision, not subject dropout, and is **not correctable** from the
existing data: no paired change can be computed for those analytes. (See Table 1 for
which features have n paired = 0.)

**Type B — One visit's panel run on a smaller sub-cohort than the other.**
For features measured at both visits, the number of observations at the two visits can
differ if a panel was rolled out to different subject batches. The per-analyte step-down
below shows the structure:

```{r missing-type-b-table, results='asis'}
cov_paired <- cov %>% filter(n_paired > 0)

stepdown <- cov_paired %>%
  mutate(
    analyte_grp = case_when(
      grepl("_IgG$",    feature) & !grepl("WT", feature) ~ "Total IgG (antigen-specific)",
      grepl("_IgG1$|_IgG4$", feature)                   ~ "IgG1 and IgG4",
      grepl("_IgG3$",   feature)                         ~ "IgG3",
      grepl("_IgG2$",   feature) & grepl("^(DT|FHA|PRN|PT|TT)_", feature) ~ "IgG2 — core antigens",
      grepl("_IgG2$",   feature) & grepl("^IPV",        feature)               ~ "IgG2 — IPV antigens",
      grepl("WT_IgG",   feature)                         ~ "WHOLE_WT_IgG",
      TRUE                                               ~ "Other"
    )
  ) %>%
  group_by(analyte_grp) %>%
  summarize(
    n_features       = n(),
    n_cord_typical   = median(n_cord),
    n_inf_typical    = median(n_inf),
    n_paired_typical = median(n_paired),
    cord_vs_inf_gap  = median(n_inf) - median(n_cord),
    .groups = "drop"
  ) %>%
  arrange(desc(n_cord_typical))

knitr::kable(stepdown %>%
  transmute(`Analyte group`             = analyte_grp,
            `N features`               = n_features,
            `Typical n PregEarly`      = n_cord_typical,
            `Typical n MatBirth`        = n_inf_typical,
            `Typical n paired`         = n_paired_typical,
            `MatBirth minus PregEarly`  = cord_vs_inf_gap),
  caption = paste0("Table 2. Per-analyte-group sample sizes at the earlier (PregEarly) and later (MatBirth) visit. ",
                   "Each row is a set of features sharing the same coverage. ",
                   "The gap (MatBirth minus PregEarly) quantifies subjects present at the later visit ",
                   "but not the earlier one for that analyte group — subjects who ",
                   "cannot contribute to any paired change analysis for those features.")) %>% print()
```

Any stepped structure here reflects which analyte panels were run on which subject
batches, rather than subject dropout. Where one analyte is available on fewer subjects than
another, the smaller subset may differ systematically (e.g. earlier-enrolled or a specific
plate batch); confirm before pooling those features with the better-covered ones. *Specifics
to be read from the rendered Table 2.*

**Type C — Dropout (subjects with a value at one visit but not the other).**
Subjects present at one visit but missing the other contribute to that visit's
cross-sectional n but not to the paired change. The size of this group is the gap between
each visit's n and the paired n in Table 1.

**Type D — Feature-level NA within otherwise assayed subjects.**
A subject may attend both visits yet still be dropped from a feature's paired analysis if
the value failed QC at one visit (insufficient volume, assay failure). This is the
difference between the per-visit n and the paired n not explained by Type B/C.

## Visualising missingness structure

```{r fig-miss-heatmap, fig.cap="Figure M1. Coverage heatmap: number of subjects with a valid measurement at the earlier and later visit for every well-covered feature (n >= 10 paired). Features are grouped by analyte type and ordered by earlier-visit n within group. Any step-down in coverage by analyte group (Type B missingness) is visible here.", fig.width=11, fig.height=8}
library(ggplot2)
library(dplyr)
library(tidyr)

# Build long-form coverage data for heatmap (cov already has n_cord, n_inf, n_paired)
cov_heat <- cov %>%
  filter(n_paired >= MIN_PAIRED) %>%
  mutate(
    analyte_grp = case_when(
      grepl("_IgG$",  feature) & !grepl("WT", feature) ~ "Total IgG",
      grepl("_IgG1$", feature)                          ~ "IgG1",
      grepl("_IgG2$", feature)                          ~ "IgG2",
      grepl("_IgG3$", feature)                          ~ "IgG3",
      grepl("_IgG4$", feature)                          ~ "IgG4",
      TRUE                                              ~ "Other"
    )
  ) %>%
  arrange(analyte_grp, desc(n_cord)) %>%
  mutate(feature_ord = factor(feature, levels = rev(unique(feature))))

# Reshape to long form without rename-in-select
cov_long <- bind_rows(
  cov_heat %>% mutate(visit = FROM_VISIT, n = n_cord),
  cov_heat %>% mutate(visit = TO_VISIT,   n = n_inf)
) %>%
  mutate(visit = factor(visit, levels = c(FROM_VISIT, TO_VISIT)))

ggplot(cov_long, aes(x = visit, y = feature_ord, fill = n)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(aes(label = n), size = 2.4, color = "white", fontface = "bold") +
  scale_fill_gradient(low = "#d9534f", high = "#2166ac",
                      name = "n subjects", limits = c(0, NA)) +
  facet_grid(analyte_grp ~ ., scales = "free_y", space = "free_y") +
  labs(x = NULL, y = NULL,
       title = "Subject coverage by feature and visit",
       subtitle = "Each cell = n subjects with a valid value at that visit. Features with n >= 10 paired only.") +
  theme_bw(base_size = 9) +
  theme(strip.background = element_rect(fill = "grey88"),
        strip.text.y     = element_text(angle = 0, face = "bold", size = 8),
        axis.text.y      = element_text(size = 6.5),
        legend.position  = "right",
        panel.spacing    = unit(0.3, "lines"))
```

```{r fig-miss-waterfall, fig.cap="Figure M2. Paired analysis waterfall: for each well-covered feature, overlapping bars show n subjects at the earlier visit (dark blue), at the later visit (medium blue), and paired at both (light blue). The gap between the later-visit bar and the paired bar is Type B+C+D missingness. Features ordered by n paired.", fig.width=10, fig.height=7}
cov_water <- cov %>%
  filter(n_paired >= MIN_PAIRED) %>%
  mutate(feature_ord = reorder(feature, n_paired))

# Reshape to long form using bind_rows to avoid pivot_longer column-name issues
cov_water_long <- bind_rows(
  cov_water %>% mutate(type = FROM_VISIT,      n = n_cord),
  cov_water %>% mutate(type = TO_VISIT,        n = n_inf),
  cov_water %>% mutate(type = "Paired (both)", n = n_paired)
) %>%
  mutate(type = factor(type, levels = c(FROM_VISIT, TO_VISIT, "Paired (both)")))

water_cols <- setNames(c("#2166ac", "#74add1", "#d1e5f0"),
                       c(FROM_VISIT, TO_VISIT, "Paired (both)"))

ggplot(cov_water_long, aes(x = n, y = feature_ord, fill = type)) +
  geom_col(position = "identity", alpha = 0.85, width = 0.7) +
  scale_fill_manual(values = water_cols, name = NULL) +
  geom_vline(xintercept = 0) +
  labs(x = "n subjects", y = NULL,
       title = "Paired analysis waterfall: earlier-to-later visit coverage",
       subtitle = "Each feature shows n at the earlier visit, the later visit, and n contributing to the paired change analysis.") +
  theme_bw(base_size = 9) +
  theme(axis.text.y     = element_text(size = 6.5),
        legend.position = "top")
```

## Subject-level missingness patterns

The per-feature view above (Figures M1–M2) is computed from the coverage summary. To see
missingness *at the subject level* we use the paired dataset directly: each subject
contributes a known set of features to the paired change analysis, and the **number of
features a subject contributes** reveals which assay batches they were part of. A subject
who received the full panel contributes up to `r MIN_PAIRED`+ features; a subject in an
early, total-IgG-only batch contributes far fewer.

One limitation to state plainly: the paired dataset contains only subjects who achieved
**at least one** valid earlier↔later visit pairing. Subjects who attended only one visit, or
whose every measurement failed pairing, do not appear here at all — quantifying that group
requires the per-visit raw records and is not attempted in this document. The per-feature
coverage gaps in Table 1 (e.g. later-visit n exceeding earlier-visit n) are the aggregate footprint
of those subjects.

```{r miss-subject-compute, include=FALSE}
# Subject-level coverage derived ONLY from the paired dataset (known-good schema:
# subject_accession, feature, antigen, analyte, feature_class). We do not touch the
# raw data frame, whose column names are not assumed here.

paired_well <- paired %>% filter(feature %in% feat_summary$feature)

# Per-subject count of paired features contributed
subj_cov <- paired_well %>%
  group_by(subject_accession) %>%
  summarize(n_paired_features = n_distinct(feature), .groups = "drop")

n_subjects_paired <- nrow(subj_cov)
max_features      <- n_distinct(paired_well$feature)

# Per-subject coverage broken down by analyte group (reveals batch structure)
subj_by_analyte <- paired_well %>%
  mutate(analyte_grp = case_when(
    grepl("_IgG$", feature) & !grepl("WT", feature) ~ "Total IgG",
    grepl("WT_IgG", feature)                         ~ "Total IgG",
    grepl("_IgG1$", feature)                         ~ "IgG1",
    grepl("_IgG2$", feature)                         ~ "IgG2",
    grepl("_IgG3$", feature)                         ~ "IgG3",
    grepl("_IgG4$", feature)                         ~ "IgG4",
    TRUE                                             ~ "Other")) %>%
  group_by(subject_accession, analyte_grp) %>%
  summarize(n_feat = n_distinct(feature), .groups = "drop")

# How many subjects reached each analyte group at all
analyte_reach <- subj_by_analyte %>%
  group_by(analyte_grp) %>%
  summarize(n_subjects = n_distinct(subject_accession), .groups = "drop") %>%
  arrange(desc(n_subjects))

# Distribution of per-subject total paired-feature counts, binned
subj_cov_binned <- subj_cov %>%
  mutate(coverage_band = cut(
    n_paired_features,
    breaks = c(0, 6, 20, 30, 40, Inf),
    labels = c("1\u20136 (total IgG only)", "7\u201320", "21\u201330", "31\u201340", "41+"),
    right = TRUE)) %>%
  count(coverage_band, name = "n_subjects")
```

```{r miss-subject-table, results='asis'}
knitr::kable(analyte_reach,
  col.names = c("Analyte group", "N subjects contributing \u2265 1 paired feature"),
  caption = paste0("Table 3. Number of subjects contributing at least one paired feature ",
                   "to each analyte group. Derived from the paired dataset only. The drop ",
                   "from Total IgG down through the subclasses mirrors any earlier-visit assay ",
                   "batch step-down in Table 2: fewer subjects had cord-blood subclass ",
                   "panels run, so fewer can contribute paired subclass features.")) %>% print()
```

```{r fig-miss-subject, fig.cap="Figure M3. Per-subject paired-feature coverage. Each bar is one subject; height = number of distinct features that subject contributes to the paired change analysis, sorted descending. The flat plateaus reflect the assay batch structure: subjects either received a given panel or did not, producing discrete coverage levels rather than a smooth gradient.", fig.width=10, fig.height=4.5}
subj_sorted <- subj_cov %>%
  arrange(desc(n_paired_features)) %>%
  mutate(rank = row_number())

ggplot(subj_sorted, aes(x = rank, y = n_paired_features)) +
  geom_col(fill = "#2166ac", width = 1) +
  labs(x = "Subjects (sorted by n paired features)",
       y = "N paired features contributed",
       title = "Per-subject coverage in the paired change analysis",
       subtitle = sprintf("%d subjects contribute at least one paired feature (max possible = %d).",
                          n_subjects_paired, max_features)) +
  theme_bw(base_size = 11)
```

```{r fig-miss-analyte-reach, fig.cap="Figure M4. Number of subjects contributing at least one paired feature, by analyte group. The step-down from Total IgG through the IgG subclasses is the subject-level signature of any assay batch rollout (Type B missingness): if a panel was run on fewer samples at one visit, progressively fewer subjects can contribute that paired feature.", fig.width=8, fig.height=4.5}
ggplot(analyte_reach,
       aes(x = reorder(analyte_grp, n_subjects), y = n_subjects, fill = analyte_grp)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = n_subjects), hjust = -0.15, size = 3.4) +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
  expand_limits(y = max(analyte_reach$n_subjects) * 1.08) +
  labs(x = NULL, y = "N subjects with \u2265 1 paired feature",
       title = "Subjects reaching each analyte group in the paired analysis",
       subtitle = "A step-down mirrors any assay batch structure (Type B missingness).") +
  theme_bw(base_size = 11)
```

## Implications of the missingness structure for change estimates

*Brief notes — to be expanded after reviewing the rendered tables and figures.*

- **Batch / sub-cohort structure (Table 2, Figures M1–M4).** If different analytes were run
  on different subject batches, the subjects contributing to one analyte's paired analysis
  may not be the same population as another's; check before pooling across analytes.
- **Paired loss with both visits present (Type D).** Subjects with both visit records but a
  missing paired value warrant a look: do they cluster by site, collection date, or
  maternal arm, and do their available single-visit values differ from the paired group?
- **Per-visit-only subjects (Type C).** These inflate the cross-sectional n at one visit
  relative to what is available for the paired change analysis.

# Direction of Change Across All Features

```{r headline, results='asis'}
cat(sprintf(paste0(
  "- Across the **%d** well-covered features, **%d** are *mostly increasing*, **%d** are *mostly decreasing*, ",
  "and **%d** are *mixed* (no clear net direction).\n",
  "- Pooling all subject×feature observations, **%.1f%%** show an increase (median delta = **%+.3f** log10).\n",
  "- Of the **%d** total-antigen IgG features, **%d** are *mostly increasing*.\n",
  "- Per subject (mean delta across all their features), **%.1f%%** have a net positive shift ",
  "(median per-subject offset = **%+.3f** log10).\n",
  "- Analytes with **zero observations at one or both visits** cannot contribute to this change analysis (see Table 1).\n\n"),
  n_feat, n_mostly_inc, n_mostly_dec, n_mixed,
  pooled_pct_inc, pooled_med,
  nrow(tot_igg), n_tot_igg_inc,
  subj_pct_pos, subj_med_off))

pct_dec_prop <- n_mostly_dec / max(n_feat, 1)
pct_inc_prop <- n_mostly_inc / max(n_feat, 1)

# computed (not hardcoded) descriptors for the concentrated-change branch
inc_analytes_h <- analyte_sum$analyte[grepl("increasing", analyte_sum$dominant)]
top_inc_ant_h  <- feat_summary %>%
  dplyr::filter(direction == "mostly increase") %>%
  dplyr::arrange(dplyr::desc(median_delta)) %>%
  dplyr::distinct(antigen) %>% utils::head(4) %>% dplyr::pull(antigen)

verdict <- if (pct_inc_prop >= 0.6) {
  sprintf(paste0("Most features **increase** from %s to %s (%.0f%% of features mostly increasing), ",
                 "and %.1f%% of subjects show a net positive shift. See the per-analyte (Table 5) ",
                 "and per-antigen (Table 6) breakdowns for where the change concentrates."),
          FROM_VISIT, TO_VISIT, 100*pct_inc_prop, subj_pct_pos)
} else if (pct_dec_prop >= 0.6) {
  sprintf(paste0("Most features **decrease** from %s to %s (%.0f%% of features mostly decreasing), ",
                 "and %.1f%% of subjects show a net positive shift. See the per-analyte (Table 5) ",
                 "and per-antigen (Table 6) breakdowns for where the change concentrates."),
          FROM_VISIT, TO_VISIT, 100*pct_dec_prop, subj_pct_pos)
} else {
  sprintf(paste0("The change is **concentrated, not uniform**: %d features increase, %d decrease, ",
                 "and %d are mixed, but %.1f%% of mothers show a net positive shift. The increases ",
                 "concentrate in %s and in the antigens %s (see Tables 5–6); the per-arm split, and ",
                 "the tetanus/diphtheria toxoid responses specifically, are in Tables 7–9 and ",
                 "Figure A1 and discussed in the Interpretation section below."),
          n_mostly_inc, n_mostly_dec, n_mixed, subj_pct_pos,
          if (length(inc_analytes_h)) paste(inc_analytes_h, collapse = ", ") else "no single analyte",
          if (length(top_inc_ant_h))  paste(top_inc_ant_h,  collapse = ", ") else "none")
}
cat("> **Verdict:** ", verdict, "\n\n")
```

```{r summary-table, results='asis'}
disp <- feat_summary %>%
  transmute(feature, class = feature_class, antigen, analyte, n_paired,
            `% increase` = pct_inc, `% decrease` = pct_dec,
            `median delta` = median_delta, `mean delta` = mean_delta,
            `p signed-rank` = p_signrank, `p sign-test` = p_sign, direction)
knitr::kable(disp,
  caption = "Table 4. Per-feature paired change (later vs earlier visit), sorted by % increasing. Positive delta = increase.") %>% print()
```

```{r analyte-summary-table, results='asis'}
knitr::kable(analyte_sum %>% arrange(med_delta) %>%
  transmute(Analyte = analyte, `N features` = n_features,
            `N mostly increasing` = n_inc, `N mostly decreasing` = n_dec, `N mixed` = n_mix,
            `Median delta (log10)` = med_delta,
            `Mean % subjects increasing` = pct_subj_inc,
            `Overall direction` = dominant),
  caption = "Table 5. Per-analyte summary across all antigens. Each row pools all antigens for that analyte type. Ordered by median delta (most negative first).") %>% print()
```

# Visualising Direction and Magnitude of Change

## Figure 1 — Feature-level lollipop

```{r fig-lollipop, fig.cap="Figure 1. Percent of subjects increasing (later visit > earlier visit), by feature, coloured by feature class. Dashed line = 50% (no net direction). Features to the right increase in most subjects; features to the left decrease in most subjects.", fig.height=9}
ggplot(feat_summary, aes(x = pct_inc, y = reorder(feature, pct_inc), color = feature_class)) +
  geom_segment(aes(x = 50, xend = pct_inc, yend = reorder(feature, pct_inc)), color = "grey75") +
  geom_point(size = 2.6) +
  geom_vline(xintercept = 50, linetype = "dashed") +
  scale_color_brewer(palette = "Dark2", name = "Feature class") +
  labs(x = "% of subjects increasing", y = NULL,
       title = "Direction of change per feature (later vs earlier visit)",
       subtitle = "Decrease = left of 50%; increase = right of 50%") +
  theme_bw(base_size = 9) + theme(axis.text.y = element_text(size = 6.5))
```

## Figure 2 — Delta distributions by feature class

```{r fig-delta-class, fig.cap="Figure 2. Distribution of per-subject delta (later − earlier visit) by feature class. Zero = no change.", fig.height=5}
ggplot(paired %>% filter(feature %in% feat_summary$feature),
       aes(x = feature_class, y = delta, fill = feature_class)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(outlier.size = 0.5, alpha = 0.7) +
  scale_fill_brewer(palette = "Dark2", guide = "none") +
  labs(x = NULL, y = "delta (log10): later − earlier visit",
       title = "Per-subject change by feature class") +
  theme_bw(base_size = 11) + theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

## Figure 3 — Per-subject average shift

```{r fig-subject-offset, fig.cap="Figure 3. Per-subject mean delta across all features. A distribution clearly displaced from zero would indicate a subject-level (sample/plate) offset; a distribution straddling zero indicates feature-structured rather than uniform change.", fig.height=4}
ggplot(subj_off, aes(x = mean_delta)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = subj_med_off, linetype = "solid", color = "firebrick", linewidth = 0.8) +
  annotate("text", x = subj_med_off + 0.02, y = Inf, vjust = 1.5, hjust = 0,
           label = sprintf("median = %+.3f", subj_med_off), color = "firebrick", size = 3.2) +
  labs(x = "Per-subject mean delta (log10), all features", y = "Subjects",
       title = "Subject-level shift (later vs earlier visit)") +
  theme_bw(base_size = 11)
```

## Figure 4 — Per-antigen histograms (analytes pooled)

Each panel pools all analytes measured for that antigen. The dashed line marks zero;
the red line marks the antigen-level median delta. Antigens are ordered from the most
negative (strongest decay) to the most positive (strongest increase) median.

```{r fig-per-antigen, fig.cap="Figure 4. Per-subject delta distributions by antigen (all analytes pooled within each antigen). Dashed = zero; red = antigen median. Ordered by median delta (most negative first).", fig.width=12, fig.height=14}
plot_data <- paired %>%
  filter(feature %in% feat_summary$feature) %>%
  group_by(antigen) %>%
  mutate(antigen_med = median(delta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(antigen_fct = reorder(antigen, antigen_med))

antigen_med_df <- plot_data %>%
  group_by(antigen_fct, antigen) %>%
  summarize(antigen_med = first(antigen_med),
            n_subj = n_distinct(subject_accession),
            n_feat = n_distinct(feature),
            .groups = "drop")

ggplot(plot_data, aes(x = delta)) +
  geom_histogram(bins = 35, fill = "steelblue", color = "white", alpha = 0.85) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_vline(data = antigen_med_df, aes(xintercept = antigen_med),
             color = "firebrick", linewidth = 0.8) +
  geom_text(data = antigen_med_df,
            aes(x = antigen_med, y = Inf,
                label = sprintf("med=%+.2f\nn=%d subj\n%d feat",
                                antigen_med, n_subj, n_feat)),
            vjust = 1.3,
            hjust = ifelse(antigen_med_df$antigen_med >= 0, -0.08, 1.08),
            size = 2.6, color = "firebrick") +
  facet_wrap(~ antigen_fct, ncol = 3, scales = "free_y") +
  labs(x = "delta (log10): later − earlier visit",
       y = "Observations (subject × feature)",
       title = "Delta distributions by antigen (later vs earlier visit)",
       subtitle = "All analytes pooled per antigen. Dashed = zero; red = antigen median. Ordered most negative first.") +
  theme_bw(base_size = 10) +
  theme(strip.background = element_rect(fill = "grey92"),
        strip.text       = element_text(face = "bold", size = 9),
        panel.spacing    = unit(0.6, "lines"))
```

```{r antigen-summary-table, results='asis'}
knitr::kable(antigen_sum %>%
  transmute(Antigen = antigen, `N features` = n_features,
            `N mostly increasing` = n_inc, `N mostly decreasing` = n_dec, `N mixed` = n_mix,
            `Median delta (log10)` = med_delta, `Overall direction` = dominant),
  caption = "Table 6. Per-antigen summary of change (later vs earlier visit), ordered by median delta (most negative first). Each row pools all analytes available for that antigen.") %>% print()
```

# Per-Analyte Histograms: Figures 5a–5f

Each figure in this section fixes **one analyte** (IgG, IgG1, IgG2, IgG3, IgG4, or
WHOLE_WT_IgG) and shows a separate histogram panel for each antigen measured with that
analyte. This layout isolates whether an analyte behaves consistently across antigens, or
whether the direction of change is antigen-specific. Features are ordered within each
figure by their median delta (most decreasing left/top, most increasing right/bottom).

**Note:** analytes with zero observations at one or both visits (e.g. effector-function,
FcR-binding, or B-memory assays, where applicable) cannot contribute to this paired change
analysis and do not appear below.

## Figure 5a — Total IgG (antigen-specific)

Total antigen-specific IgG, by antigen.

```{r fig-5a, fig.cap="Figure 5a. Per-subject delta distributions for total antigen-specific IgG features (one panel per antigen, ordered by median delta).", fig.width=10, fig.height=5}
make_analyte_hist("IgG", fill_col = analyte_colours["IgG"])
```

## Figure 5b — IgG1

IgG1 by antigen. IgG1 is the dominant IgG subclass in serum and is efficiently transferred across the placenta.

```{r fig-5b, fig.cap="Figure 5b. Per-subject delta distributions for IgG1 features by antigen (ordered by median delta).", fig.width=12, fig.height=6}
make_analyte_hist("IgG1", fill_col = analyte_colours["IgG1"])
```

## Figure 5c — IgG2

IgG2 by antigen. IgG2 is enriched in responses to polysaccharide antigens and is transferred across the placenta less efficiently than IgG1.

```{r fig-5c, fig.cap="Figure 5c. Per-subject delta distributions for IgG2 features by antigen (ordered by median delta).", fig.width=12, fig.height=6}
make_analyte_hist("IgG2", fill_col = analyte_colours["IgG2"])
```

## Figure 5d — IgG3

IgG3 by antigen. IgG3 has the shortest serum half-life of the IgG subclasses (~7 days vs ~21 days for IgG1).

```{r fig-5d, fig.cap="Figure 5d. Per-subject delta distributions for IgG3 features by antigen (ordered by median delta).", fig.width=12, fig.height=6}
make_analyte_hist("IgG3", fill_col = analyte_colours["IgG3"])
```

## Figure 5e — IgG4

IgG4 by antigen. IgG4 is typically the lowest-abundance IgG subclass and is elevated by chronic/repeated antigen exposure.

```{r fig-5e, fig.cap="Figure 5e. Per-subject delta distributions for IgG4 features by antigen (ordered by median delta).", fig.width=12, fig.height=6}
make_analyte_hist("IgG4", fill_col = analyte_colours["IgG4"])
```

## Figure 5f — WHOLE_WT_IgG

WHOLE_WT_IgG: a single whole-cell lysate total-IgG measurement integrating broadly across antigens.

```{r fig-5f, fig.cap="Figure 5f. Per-subject delta distribution for WHOLE_WT_IgG (single feature).", fig.width=5, fig.height=4}
make_analyte_hist("WT_IgG", fill_col = analyte_colours["WT_IgG"])
```

# Stratified by maternal arm (TdaP vs TT)

The combined (pooled) results above are retained unchanged. This section ADDS the
per-arm breakdown of the *same* PregEarly → MatBirth production change, computed
live from the loaded data. It exists to answer one specific question: **do the
tetanus (TT) and diphtheria (DT) toxoid antigens — which are vaccine components —
show a within-arm rise in IgG and IgG1 from early pregnancy to delivery?**

```{r by-arm-setup, include=FALSE}
paired_m <- paired %>% dplyr::filter(arm_name %in% MATERNAL_ARMS)

# Per (feature) production summary for an arbitrary subject subset, including the
# PregEarly/MatBirth level medians so a flat delta can be read against its baseline.
prod_by_arm <- function(pd) {
  pd %>% dplyr::group_by(feature, antigen, analyte) %>%
    dplyr::summarize(
      n_pair    = dplyr::n(),
      med_from  = round(stats::median(v_from), 3),
      med_to    = round(stats::median(v_to), 3),
      med_delta = round(stats::median(delta), 3),
      pct_inc   = round(100 * mean(delta > 0), 1),
      p_sr      = tryCatch(signif(stats::wilcox.test(delta, mu = 0, exact = FALSE)$p.value, 3),
                           error = function(e) NA_real_),
      .groups = "drop") %>%
    dplyr::filter(n_pair >= MIN_PAIRED)
}

prod_all <- dplyr::bind_rows(
  prod_by_arm(paired_m)                                   %>% dplyr::mutate(arm = "Pooled"),
  prod_by_arm(dplyr::filter(paired_m, arm_name == "TdaP")) %>% dplyr::mutate(arm = "TdaP"),
  prod_by_arm(dplyr::filter(paired_m, arm_name == "TT"))   %>% dplyr::mutate(arm = "TT")) %>%
  dplyr::mutate(arm = factor(arm, levels = c("Pooled", "TdaP", "TT")))

getv <- function(ant, an, a, col) {
  r <- prod_all[prod_all$antigen == ant & prod_all$analyte == an & prod_all$arm == a, ]
  if (nrow(r) == 0) NA_real_ else r[[col]][1]
}
fmtd <- function(x) ifelse(is.na(x), "n/a", sprintf("%+.3f", x))
```

## Direction summary by arm

```{r by-arm-direction, results='asis'}
arm_sum <- summarise_paired(paired_m, by = "arm_name", min_n = MIN_PAIRED) %>%
  dplyr::group_by(arm_name) %>%
  dplyr::summarize(features = dplyr::n(),
            `% feats mostly increasing` = round(100*mean(direction == "mostly increase"), 1),
            `median of feature median-deltas` = round(stats::median(median_delta), 3),
            .groups = "drop")
knitr::kable(arm_sum,
  caption = "Table 7. Direction summary by maternal arm. An arm-specific difference points toward vaccine-driven (biological) heterogeneity; an arm-shared pattern points to a shared technical or developmental effect.") %>% print()
```

## The toxoid check: TT and DT, by arm

```{r by-arm-toxoid, results='asis'}
tox <- prod_all %>%
  dplyr::filter(antigen %in% c("TT", "DT"), analyte %in% c("IgG", "IgG1")) %>%
  dplyr::arrange(antigen, analyte, arm) %>%
  dplyr::transmute(Antigen = antigen, Analyte = analyte, Arm = arm,
                   `n pairs` = n_pair,
                   `median PregEarly` = med_from, `median MatBirth` = med_to,
                   `median delta` = med_delta, `% increasing` = pct_inc,
                   `p signed-rank` = p_sr)
knitr::kable(tox,
  caption = paste0("Table 8. Tetanus (TT) and diphtheria (DT) antigen-specific total IgG and IgG1 ",
                   "production change (PregEarly -> MatBirth), pooled and by maternal arm. ",
                   "A positive median delta with a small p is a genuine within-arm rise; a near-zero ",
                   "delta sitting on an already-high PregEarly median is a ceiling effect, not an ",
                   "absent response.")) %>% print()
```

```{r by-arm-verdict, results='asis'}
cat("> **Toxoid verdict (computed from the loaded data):**\n\n")
for (ant in c("TT", "DT")) {
  nm <- if (ant == "TT") "Tetanus (TT)" else "Diphtheria (DT)"
  cat(sprintf(paste0("- **%s** \u2014 PregEarly baseline median (pooled, total IgG) %.2f log10. ",
      "Total-IgG production median delta: TdaP %s, TT %s. ",
      "IgG1 production median delta: TdaP %s, TT %s.\n"),
    nm, getv(ant, "IgG", "Pooled", "med_from"),
    fmtd(getv(ant, "IgG", "TdaP", "med_delta")), fmtd(getv(ant, "IgG", "TT", "med_delta")),
    fmtd(getv(ant, "IgG1", "TdaP", "med_delta")), fmtd(getv(ant, "IgG1", "TT", "med_delta"))))
}
cat(paste0("\nInterpret each delta against its PregEarly baseline. A near-zero delta on top of an ",
  "already-high baseline is a **ceiling effect** \u2014 mothers were already tetanus/diphtheria-immune ",
  "from prior vaccination, so a booster moves the log10 value little. A near-zero delta on a **low** ",
  "baseline would instead be a true non-response, worth checking (assay dynamic range, antigen ",
  "coating, or which toxoids each arm's vaccine actually contains: tetanus is in both TdaP and TT, ",
  "whereas diphtheria is expected only where the comparator is Td rather than tetanus-only).\n\n"))
```

## Full per-antigen production change (IgG and IgG1), by arm

```{r by-arm-full, results='asis'}
full_ig <- prod_all %>%
  dplyr::filter(antigen %in% ANTIGEN_ORDER, analyte %in% c("IgG", "IgG1")) %>%
  dplyr::mutate(antigen = factor(antigen, levels = ANTIGEN_ORDER)) %>%
  dplyr::arrange(analyte, antigen, arm) %>%
  dplyr::transmute(Analyte = analyte, Antigen = antigen, Arm = arm, n = n_pair,
                   `median delta` = med_delta, `% inc` = pct_inc, `p signed-rank` = p_sr)
knitr::kable(full_ig,
  caption = "Table 9. Per-antigen production change for total IgG and IgG1, by arm (six vaccine antigens in canonical order PT, FHA, PRN, FIM, TT, DT).") %>% print()
```

```{r by-arm-fig, fig.width=10, fig.height=6, fig.cap="Figure A1. Production change (median delta log10, MatBirth - PregEarly) for the six vaccine antigens, total IgG and IgG1, by maternal arm. Point size = percent of mothers increasing. Positive = rise from early pregnancy to delivery. The TT and DT rows show whether the toxoid antigens move within each arm."}
figdf <- prod_all %>%
  dplyr::filter(antigen %in% ANTIGEN_ORDER, analyte %in% c("IgG", "IgG1"), arm != "Pooled") %>%
  dplyr::mutate(antigen = factor(antigen, levels = rev(ANTIGEN_ORDER)),
                arm = forcats::fct_drop(arm))
ggplot2::ggplot(figdf, ggplot2::aes(x = med_delta, y = antigen, colour = arm)) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey40", linewidth = 0.4) +
  ggplot2::geom_segment(ggplot2::aes(x = 0, xend = med_delta, yend = antigen),
                        position = ggplot2::position_dodge(width = 0.6),
                        linewidth = 0.6, alpha = 0.4) +
  ggplot2::geom_point(ggplot2::aes(size = pct_inc),
                      position = ggplot2::position_dodge(width = 0.6)) +
  ggplot2::facet_wrap(~ analyte) +
  ggplot2::scale_colour_manual(values = c(TdaP = "#8A1538", TT = "#2C7A7B")) +
  ggplot2::scale_size_continuous(name = "% increasing", range = c(2, 6)) +
  ggplot2::labs(x = expression("median " * Delta * "log"[10] * " (MatBirth - PregEarly)"),
                y = "Antigen", colour = "Maternal arm",
                title = "Production change by maternal arm: do the toxoids (TT, DT) rise?") +
  ggplot2::theme_bw()
```

# Interpretation

*All quantities in this section are computed from the current data at knit time
(from `feat_summary`, `analyte_sum`, `prod_all`, the per-arm summaries and the
headline diagnostics). Nothing here is hardcoded, so the narrative always matches
the tables and figures above.*

From early pregnancy (`PregEarly`) to delivery (`MatBirth`) the mother receives the
study vaccine, so the expected signal is a vaccine-induced rise in antibody to the
vaccine antigens, concentrated in the isotype a protein-antigen booster recruits.

```{r interp-calc, include=FALSE}
# Helper builders: every narrative figure is derived here from computed objects.
top_feats <- function(an, k = 4) {
  d <- feat_summary %>%
    dplyr::filter(analyte == an, direction == "mostly increase") %>%
    dplyr::arrange(dplyr::desc(median_delta)) %>% utils::head(k)
  if (!nrow(d)) return("no features predominantly increasing")
  paste(sprintf("%s (median %+.2f, %.0f%% up)", d$antigen, d$median_delta, d$pct_inc),
        collapse = "; ")
}
inc_an   <- analyte_sum$analyte[grepl("increasing", analyte_sum$dominant)]
other_an <- analyte_sum$analyte[!grepl("increasing", analyte_sum$dominant)]

arm_dir <- summarise_paired(paired_m, by = "arm_name", min_n = MIN_PAIRED) %>%
  dplyr::group_by(arm_name) %>%
  dplyr::summarize(pct = round(100 * mean(direction == "mostly increase"), 1),
                   .groups = "drop")
arm_pct <- function(a) { v <- arm_dir$pct[arm_dir$arm_name == a]
  if (length(v)) format(v) else "n/a" }

ant_phrase <- function(ant) {
  g <- function(an, col) { r <- feat_summary[feat_summary$antigen == ant &
      feat_summary$analyte == an, ]; if (nrow(r)) r[[col]][1] else NA_real_ }
  parts <- c()
  if (!is.na(g("IgG",  "median_delta")))
    parts <- c(parts, sprintf("total IgG %+.2f (%.0f%% up)", g("IgG","median_delta"),  g("IgG","pct_inc")))
  if (!is.na(g("IgG1", "median_delta")))
    parts <- c(parts, sprintf("IgG1 %+.2f (%.0f%% up)",       g("IgG1","median_delta"), g("IgG1","pct_inc")))
  if (!length(parts)) return(NA_character_)
  sprintf("**%s** — %s", ant, paste(parts, collapse = "; "))
}

tox_phrase <- function(ant) {
  gp <- function(an, a, col) { r <- prod_all[prod_all$antigen == ant &
      prod_all$analyte == an & prod_all$arm == a, ]; if (nrow(r)) r[[col]][1] else NA_real_ }
  base <- gp("IgG", "Pooled", "med_from")
  sprintf(paste0("**%s** — total-IgG \u0394: TdaP %s, TT %s (PregEarly baseline median %s); ",
                 "IgG1 \u0394: TdaP %s, TT %s"),
    ant, fmtd(gp("IgG","TdaP","med_delta")), fmtd(gp("IgG","TT","med_delta")),
    ifelse(is.na(base), "n/a", sprintf("%.2f", base)),
    fmtd(gp("IgG1","TdaP","med_delta")), fmtd(gp("IgG1","TT","med_delta")))
}
```

## Overview

```{r interp-overview, results='asis'}
cat(sprintf(paste0(
  "Across the **%d** well-covered features, **%d** are mostly increasing, **%d** mostly ",
  "decreasing, and **%d** mixed. The pooled median change is **%+.3f** log10, while **%.1f%%** ",
  "of mothers show a net positive mean shift across their features (median per-subject offset ",
  "**%+.3f** log10). A near-zero pooled median combined with most subjects tilting positive is ",
  "the signature of a response concentrated in a subset of features (the vaccine targets) and ",
  "flat elsewhere \u2014 averaging the two washes out at the feature level but still tilts most ",
  "individual mothers positive.\n\n"),
  n_feat, n_mostly_inc, n_mostly_dec, n_mixed, pooled_med, subj_pct_pos, subj_med_off))
```

## What the data show, by analyte

```{r interp-analyte, results='asis'}
cat(sprintf("Analytes **predominantly increasing**: %s. Flat or mixed overall: %s.\n\n",
  if (length(inc_an))   paste(inc_an,   collapse = ", ") else "none",
  if (length(other_an)) paste(other_an, collapse = ", ") else "none"))
cat(sprintf("- **Total IgG** \u2014 top rises: %s.\n",  top_feats("IgG")))
cat(sprintf("- **IgG1** \u2014 top rises: %s.\n\n",       top_feats("IgG1")))
cat(paste0("A protein-antigen booster preferentially recruits IgG1, so a signal concentrated ",
  "in total IgG and IgG1 with little movement in IgG2/IgG4 is the immunologically expected ",
  "pattern rather than a missing signal.\n\n"))
```

## Focus: *Bordetella pertussis* antigens

```{r interp-pertussis, results='asis'}
cat("Mapping the change onto the acellular-pertussis components present in the data:\n\n")
for (ant in intersect(c("PT", "FHA", "PRN", "FIM"), unique(feat_summary$antigen))) {
  p <- ant_phrase(ant); if (!is.na(p)) cat("- ", p, "\n")
}
cat(paste0("\nIgG1 against the pertussis antigens is the subclass most efficiently transferred ",
  "across the placenta (see the companion `MatBirth -> CordBlood` analysis), so a vaccine that ",
  "loads this compartment is loading the one the placenta transfers best \u2014 the mechanistic ",
  "basis for maternal vaccination protecting the newborn.\n\n"))
```

## By maternal arm

```{r interp-arm, results='asis'}
cat(sprintf(paste0("Features mostly increasing by arm: **TdaP %s%%** vs **TT %s%%** (Table 7). ",
  "An arm-specific increase points to a vaccine-driven (biological) effect rather than a shared ",
  "technical or developmental drift.\n\n"), arm_pct("TdaP"), arm_pct("TT")))
cat("Toxoid antigens, by arm (does tetanus/diphtheria IgG rise from PregEarly to MatBirth?):\n\n")
for (ant in intersect(c("TT", "DT"), unique(prod_all$antigen))) cat("- ", tox_phrase(ant), "\n")
cat(paste0("\nRead each toxoid \u0394 against its PregEarly baseline: a small \u0394 on an already-high ",
  "baseline is a ceiling effect (pre-existing tetanus/diphtheria immunity), not an absent ",
  "response; a small \u0394 on a low baseline would be a genuine non-response worth investigating.\n\n"))
```

## Scale and comparability

Both `PregEarly` and `MatBirth` subclass values are normalized bead-array values (maternal
samples placed on a common assay scale); total IgG comes from a separate ELISA path and is not
affected by that normalization. Where the total-IgG and IgG1 signals agree for the same
antigens (compare the per-antigen lines above), the change is more likely biological than a
scaling artifact of either platform.

## What would confirm or refine this interpretation

- **Relate the per-mother boost to the vaccine-to-delivery interval** — where the mean exceeds
  the median for an antigen (a right-skewed response), mothers vaccinated longer before delivery
  or with lower baseline titres are expected to show the largest rises.
- **Read Figure A1 and Tables 8–9** to separate the TdaP boost from the TT baseline antigen by
  antigen and analyte, rather than relying on the pooled medians.
- **Check the vaccine formulation actually used** against any antigen whose response is muted,
  to confirm it reflects vaccine composition rather than assay coverage.


# Participant Accounting

```{r naccount, results='asis'}
nt_render(nt, part = "PreVacc→Birth",
          caption = "Table 10. Per-feature paired availability (n PregEarly / n MatBirth / n used).")
```

---

*Reuses the shared pipeline (`config/endpoints.R`, `R/C1_serology_helpers.R`, the global visit
recode, and the N-tracker). Knit both formats with*
`rmarkdown::render("cord_to_infmon2_analysis_refactored.Rmd", output_format = "all")`.
