---
title: "Phase 08 (Part A) — Decomposition and Predictors of the Maternal Low vs High Responder Categories"
subtitle: |
  Do clinical covariates or a maternal serological signature predict which
  mothers are LOW vs HIGH responders? Stratified on the maternal arm (TdaP vs TT)
  only — no infant-arm analysis.
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
params:
  responder_features: !r c("PT_IgG", "FHA_IgG", "PRN_IgG")
  maternal_visits:    !r c("PregEarly", "MatBirth")
  baseline_visit:     "PregEarly"     # pre/early-pregnancy draw = the clean predictive signal
  breadth_thresh:     0.6
  alpha:              0.05            # univariate screening threshold
  en_alpha:           0.5            # elastic-net mixing (0=ridge, 1=lasso)
  en_folds:           10
  min_coverage:       0.6            # keep serology features measured in >= 60% of mothers
  seed:               1
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
```

```{r bootstrap, include=FALSE}
library(here); library(dplyr); library(tidyr); library(ggplot2)

# ---- self-locating resolver (same contract as the rest of the pipeline) ----
# ---- shared path resolver: single definition in R/shared_utils.R ----
if (!exists("shared_utils", mode = "function")) {
  .ff_cand <- c("R/shared_utils.R", "../R/shared_utils.R",
                "analysis/R/shared_utils.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/shared_utils.R"),
                  here::here("analysis/R/shared_utils.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/shared_utils.R"),
                  file.path(dirname(.ff_rmd), "R/shared_utils.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/shared_utils.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}

# ---- shared config + helper libraries ----
# NOTE: the Phase 07 engine (C4_C4_within_visit_helpers.R) and the shared serology
# helpers are sourced READ-ONLY and are NOT modified by this branch. The Phase 08
# logic lives entirely in the NEW file R/C3_responder_helpers.R.
source(shared_utils("config/endpoints.R"),               local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C1_C1_serology_helpers.R"),             local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C4_C4_within_visit_helpers.R"),         local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C3_responder_helpers.R"),   local = FALSE, encoding = "UTF-8")  # <- Phase 08 engine

# ---- antibody data (recoded visits) + clinical assessments ----
data_raw <- load_serology_data(shared_utils("data/c_set.RData"))   # applies VISIT_RECODE
clin <- tryCatch(load_clinical_data(shared_utils("data/clin_assess.RData")),
                 error = function(e) NULL)
if (is.null(clin)) {                                   # fallback to the raw object name
  .e <- new.env(); load(shared_utils("data/clin_assess.RData"), envir = .e)
  clin <- get("clin_assess", envir = .e)
}

has_cluster <- requireNamespace("cluster", quietly = TRUE)
has_glmnet  <- requireNamespace("glmnet",  quietly = TRUE)

# ---- participant tracker (shared, optional) ----
nt <- if (exists("new_ntracker", mode = "function")) new_ntracker() else NULL
```

---

# Analytical Framework

Phase 07 Stage A split mothers into **Low** and **High** responders from their
binding antibody across the two maternal draws
(`r paste(params$maternal_visits, collapse = " + ")`;
features `r paste(params$responder_features, collapse = ", ")`). This branch asks
a single question about that split:

> **Is there a consistent clinical covariate or maternal serological signature
> that predicts which mothers are Low vs High responders?**

Two facts shape the whole analysis:

1. **The responder class is almost the vaccine arm.** In Phase 07 the maternal
   Low/High split tracked the arm closely (TdaP mostly High, TT mostly Low). So
   the genuinely informative question is not "does anything predict the class"
   (the arm does) but **"does anything predict it beyond the arm, and within an
   arm"** — e.g. among vaccinated (TdaP) mothers, who still responds poorly?
2. **Infant priming (aP/wP) is not relevant here.** This is a question about
   mothers; there is **no infant-arm stratification anywhere** in this document.
   The only stratifier is the **maternal arm (TdaP vs TT)**.

The analysis runs in three parts:

- **Part A — the arm baseline.** How much of the Low/High split is just the arm?
  This sets the bar every other predictor must clear.
- **Part B — clinical covariates.** The ten `clin_assess` covariates, screened
  univariately (continuous + tertile forms) against the responder **score**
  (linear) and the Low/High **class** (logistic), overall and within each arm,
  then a maternal-arm-adjusted multivariable model.
- **Part C — serological signature.** A maternal-serology feature matrix
  (all antigen × analyte features at the maternal visits, **excluding the `r length(params$responder_features)`
  features that define the score**), screened by univariate AUC and modelled with
  an arm-adjusted elastic net (cross-validated AUC), with a clean pre-vaccination
  (`r params$baseline_visit`) baseline model.

```{r build-class}
RESP_FEATS <- params$responder_features
MAT_VIS    <- params$maternal_visits

# Rebuild the Phase 07 Stage A Low/High class via the NEW helper (which reuses
# the C4_C4_within_visit_helpers engine under the hood for an identical construct).
mat_cls <- build_maternal_responder_class(
  data_raw, MAT_VIS, RESP_FEATS,
  breadth_thresh = params$breadth_thresh, has_cluster = has_cluster)

arm_tbl <- subject_maternal_arm(data_raw)
mat <- merge(mat_cls, arm_tbl, by = "subject_accession")
mat <- mat[mat$arm_name %in% c("TdaP", "TT"), , drop = FALSE]
mat$arm_name <- factor(mat$arm_name, levels = c("TT", "TdaP"))   # TT = reference

# Optional participant log (best-effort; this branch does not depend on the
# tracker API, so any signature mismatch is caught and ignored).
if (!is.null(nt) && exists("nt_log_manual", mode = "function")) {
  nt <- tryCatch(
    nt_log_manual(nt, "Mothers with Low/High class + arm", nrow(mat), stratum = "all"),
    error = function(e1) tryCatch(
      nt_log_manual(nt, "Mothers with Low/High class + arm", nrow(mat)),
      error = function(e2) nt))
}
```

---

# Decomposition of the responder score categories

The Low/High split is the mean Blom-standardised maternal response across the **three
defining features** (`r paste(RESP_FEATS, collapse = ", ")`) at the maternal visits
(`r paste(MAT_VIS, collapse = ", ")`). All three of PT, FHA and PRN enter the score; no
ACT or PENTAMER feature does (neither is measured in this dataset, and both are explicitly
excluded from every serology step below). This section decomposes the categories into the
contribution of each defining antigen, so the High/Low classes can be read in terms of the
antibodies that build them.

```{r score-decomp, results='asis'}
.blom_d <- function(x) { r <- rank(x, na.last = "keep"); n <- sum(!is.na(x)); stats::qnorm((r - 3/8) / (n + 1/4)) }
dec <- data_raw %>%
  dplyr::filter(visit_name %in% MAT_VIS, as.character(feature) %in% RESP_FEATS) %>%
  dplyr::group_by(subject_accession, feature, visit_name) %>%
  dplyr::summarise(v = mean(log_assay_value, na.rm = TRUE), .groups = "drop") %>%
  dplyr::mutate(key = paste(feature, visit_name, sep = "@")) %>%
  dplyr::group_by(key) %>% dplyr::mutate(z = .blom_d(v)) %>% dplyr::ungroup() %>%
  dplyr::group_by(subject_accession, feature) %>%
  dplyr::summarise(antigen_z = mean(z, na.rm = TRUE), .groups = "drop")
decm <- dec %>% dplyr::inner_join(mat_cls[, c("subject_accession", "responder", "score")],
                                  by = "subject_accession")
per_ant <- decm %>% dplyr::group_by(feature) %>%
  dplyr::summarise(
    n         = sum(is.finite(antigen_z)),
    mean_Low  = round(mean(antigen_z[responder == "Low"],  na.rm = TRUE), 3),
    mean_High = round(mean(antigen_z[responder == "High"], na.rm = TRUE), 3),
    diff      = round(mean(antigen_z[responder == "High"], na.rm = TRUE) -
                      mean(antigen_z[responder == "Low"],  na.rm = TRUE), 3),
    AUC       = round(auc_binary(antigen_z, responder), 3),
    loading   = round(stats::cor(antigen_z, score, use = "complete.obs"), 3),
    p         = tryCatch(signif(stats::wilcox.test(antigen_z ~ responder)$p.value, 3),
                         error = function(e) NA_real_),
    .groups   = "drop") %>%
  dplyr::arrange(dplyr::desc(diff))
print(knitr::kable(per_ant,
  caption = paste0("Table D1. Per-antigen decomposition of the responder categories. Each defining ",
    "antigen's maternal Blom-z (averaged over the maternal visits) by class, the High-minus-Low ",
    "difference, the antigen's discrimination of the class (AUC), and its loading (correlation) on the ",
    "overall score. All three of PT, FHA and PRN contribute positively; the ranking shows which antigen ",
    "separates the categories most.")))
top <- per_ant$feature[which.max(per_ant$diff)]
cat(sprintf(paste0("\nThe categories separate on **all three** defining antigens. The largest ",
                   "High-vs-Low gap is **%s** (\u0394z = %.3f, AUC = %.3f); every antigen loads ",
                   "positively on the score (loadings: %s).\n\n"),
            top, per_ant$diff[per_ant$feature == top], per_ant$AUC[per_ant$feature == top],
            paste(sprintf("%s %.2f", per_ant$feature, per_ant$loading), collapse = ", ")))
```

```{r score-decomp-fig, fig.height=3.2, fig.cap="Figure D1. Per-antigen Blom-standardised maternal response (averaged over the maternal visits) by responder class, for each of the three score-defining antigens."}
ggplot(decm, aes(responder, antigen_z, fill = responder)) +
  geom_violin(alpha = .5, color = NA) +
  geom_boxplot(width = .16, outlier.size = .5) +
  facet_wrap(~ feature, nrow = 1) +
  scale_fill_manual(values = c(Low = "#c0297a", High = "#1b5aa0"), guide = "none") +
  labs(x = NULL, y = "antigen Blom-z (maternal)") +
  theme_bw(base_size = 10)
```

---

# Clinical Covariates

The ten covariates from `clin_assess` (identical to the clinical-predictor
analyses). `infant_sex` is retained as a dyad covariate of the pregnancy, not as
an infant-arm axis.

```{r covariate-list, echo=FALSE}
covariates_named_list <- list(
  "Birth Weight"                             = "birth_weight",
  "Infant Sex"                               = "infant_sex",
  "Delivery Mode"                            = "delivery_mode",
  "Parity"                                   = "parity",
  "Gestational Age at Vaccine"               = "gestational_age_vaccination",
  "Interval: Vaccination to Delivery (days)" = "vaccine_birth_interval_days",
  "Gestational Age at Delivery"              = "gestational_age_birth",
  "Maternal Age at Vaccination"              = "maternal_age",
  "Maternal BMI at Vaccination"              = "maternal_bmi",
  "Maternal Haemoglobin"                     = "maternal_Hb")
cat_vars  <- c("infant_sex", "delivery_mode", "parity")
cont_vars <- setdiff(unname(unlist(covariates_named_list)), cat_vars)
knitr::kable(data.frame(
  Label = names(covariates_named_list),
  Variable = unname(unlist(covariates_named_list)),
  Type = ifelse(unlist(covariates_named_list) %in% cat_vars,
                "categorical (2 level)", "continuous (cont + tertile)")),
  caption = "Clinical covariates screened.")
```

```{r merge-clin}
data <- merge(mat, clin, by = "subject_accession", all.x = TRUE)

# ---- prepare covariates (mirrors the clinical-predictor pipeline) ----
if ("infant_sex" %in% names(data)) data$infant_sex <- factor(data$infant_sex)
if ("delivery_mode" %in% names(data) && requireNamespace("forcats", quietly = TRUE)) {
  dm <- factor(data$delivery_mode)
  data$delivery_mode <- forcats::fct_collapse(dm,
    vaginal  = grep("(?i)vag|normal|spontan|SVD", levels(dm), value = TRUE, perl = TRUE),
    cesarean = grep("(?i)ces|csec|section|C/S",   levels(dm), value = TRUE, perl = TRUE))
  data$delivery_mode <- tryCatch(stats::relevel(forcats::fct_drop(data$delivery_mode),
                                                ref = "vaginal"),
                                 error = function(e) data$delivery_mode)
}
if ("parity" %in% names(data))
  data$parity <- factor(ifelse(data$parity == 0, "nulliparous", "parous"),
                        levels = c("nulliparous", "parous"))
for (v in cont_vars) if (v %in% names(data)) data[[paste0(v, "_T")]] <- make_tertile(data[[v]])

cat(sprintf("Merged analysis frame: **%d mothers × %d columns**; Low = %d, High = %d.\n\n",
            nrow(data), ncol(data), sum(data$responder == "Low"),
            sum(data$responder == "High")))

ids_resp <- mat$subject_accession
ids_clin <- clin$subject_accession
cat("responder mothers:", length(ids_resp),
    "| in clin_assess:", sum(ids_resp %in% ids_clin),
    "| missing:", sum(!ids_resp %in% ids_clin), "\n")
# are the unmatched a format issue? peek at both ID styles:
head(setdiff(ids_resp, ids_clin)); str(ids_resp); str(ids_clin)
# is missingness related to responder status (not just arm)?
data$has_clin <- !is.na(data$maternal_age)
print(table(data$responder, data$has_clin))
```

---

# Part A — The Arm Baseline

🧭 *Maps to summary: "how much of the responder split is the vaccine itself?"*

```{r part-a, results='asis'}
tab <- table(arm = data$arm_name, responder = data$responder)
print(knitr::kable(as.data.frame.matrix(tab),
      caption = "Table A1. Maternal arm × responder class."))

w <- tryCatch(stats::wilcox.test(score ~ arm_name, data = data), error = function(e) NULL)
arm_auc <- auc_binary(as.numeric(data$arm_name == "TdaP"), data$responder)
ft <- tryCatch(stats::fisher.test(tab), error = function(e) NULL)
cat(sprintf(paste0("\nResponder **score** by arm: Wilcoxon %s. ",
                   "Arm alone separates the **class** at AUC = **%.3f** (Fisher p = %s). ",
                   "This is the bar Parts B and C must beat.\n\n"),
            if (!is.null(w)) sprintf("W = %.0f, p = %.3g", w$statistic, w$p.value) else "n/a",
            arm_auc, if (!is.null(ft)) signif(ft$p.value, 3) else "n/a"))
```

```{r part-a-plot, fig.height=3.4, fig.cap="Figure A1. Maternal responder score by arm (the score that defines Low/High)."}
ggplot(data, aes(arm_name, score, fill = arm_name)) +
  geom_violin(alpha = .5, color = NA) +
  geom_boxplot(width = .18, outlier.size = .6) +
  scale_fill_manual(values = c(TT = "#c0297a", TdaP = "#1b5aa0"), guide = "none") +
  labs(x = "maternal arm", y = "maternal responder score (z)") +
  theme_bw(base_size = 11)
```

---

# Part B — Clinical Covariates

Each covariate is screened against the responder **score** (linear model, mirrors
the clinical-predictor pipeline) and the Low/High **class** (logistic), in three
strata: **Overall**, **TdaP mothers**, **TT mothers**.

```{r run-screens, results='hide'}
strata <- list(
  Overall = data,
  TdaP    = dplyr::filter(data, arm_name == "TdaP"),
  TT      = dplyr::filter(data, arm_name == "TT"))

screen_score <- lapply(strata, function(d)
  run_clinical_screen(d, "score", covariates_named_list, cat_vars,
                      mode = "lm", alpha = params$alpha))
screen_class <- lapply(strata, function(d)
  run_clinical_screen(d, "responder", covariates_named_list, cat_vars,
                      mode = "logit", alpha = params$alpha))
```

```{r show-screen-helper, echo=FALSE}
show_univ <- function(s, title) {
  cat(sprintf("\n\n#### %s\n\n", title))
  tb <- s$univ_table
  tb$R2  <- ifelse(is.na(tb$R2), "\u2014", sprintf("%.3f", tb$R2))
  tb$auc <- ifelse(is.na(tb$auc), "\u2014", sprintf("%.3f", tb$auc))
  tb$p_model <- ifelse(is.na(tb$p_model), "\u2014", sprintf("%.3g", tb$p_model))
  print(knitr::kable(tb[, c("Label","Form","n","R2","auc","p_model","sig")],
        col.names = c("Covariate","Form","n","R2 / pseudo-R2","AUC","p (model)",""),
        caption = paste("Univariate screen —", title)))
  if (length(s$selected) == 0) {
    cat("\n*No covariate reached p <", params$alpha, "in this stratum.*\n\n")
  } else {
    cat(sprintf("\n**Selected (p < %g):** %s\n\n", params$alpha,
                paste(names(s$selected), collapse = "; ")))
    if (!is.null(s$multi_result))
      print(knitr::kable(s$multi_result$tidy,
            caption = sprintf("Multivariable model (%s) — %s", s$mode, title)))
  }
}
```

## Predicting the responder SCORE (linear)

```{r show-score, results='asis'}
for (st in names(screen_score)) show_univ(screen_score[[st]], paste0(st, " — score"))
```

## Predicting the Low/High CLASS (logistic)

```{r show-class, results='asis'}
for (st in names(screen_class)) show_univ(screen_class[[st]], paste0(st, " — class"))
```

## Maternal-arm-adjusted model (overall)

🧭 *Maps to summary: "does any clinical covariate predict the class beyond the arm?"*

```{r arm-adjusted, results='asis'}
# Enter arm + any covariate selected in ANY stratum (score or class), jointly.
sel_any <- unique(unlist(c(
  lapply(screen_score, function(s) unname(unlist(s$selected))),
  lapply(screen_class, function(s) unname(unlist(s$selected))))))
sel_any <- sel_any[!is.na(sel_any)]

if (length(sel_any)) {
  fla <- stats::reformulate(c("arm_name", sel_any), response = "responder")
  m_adj <- tryCatch(stats::glm(fla, data = data, family = stats::binomial(),
                               na.action = stats::na.omit), error = function(e) NULL)
  if (!is.null(m_adj)) {
    print(knitr::kable(.tidy_fit(m_adj, logit = TRUE),
          caption = "Table B-adj. Arm-adjusted logistic model for High responder (OR per term)."))
    au <- auc_binary(stats::predict(m_adj, type = "response"),
                     stats::model.frame(m_adj)[["responder"]])
    cat(sprintf("\nIn-sample AUC = **%.3f** (vs arm-only AUC = %.3f). ", au, arm_auc))
    cat("Only terms that stay significant *after* arm adjustment add information beyond the vaccine.\n\n")
  }
} else {
  cat("*No clinical covariate was selected in any stratum, so no arm-adjusted model is fitted — ",
      "clinical covariates do not predict responder status beyond the arm.*\n\n")
}
```

## Cross-stratum summary & figures

```{r cross-stratum, results='asis'}
all_rows <- dplyr::bind_rows(lapply(names(screen_class), function(sn)
  dplyr::mutate(screen_class[[sn]]$univ_table, Stratum = sn)))
disp <- all_rows %>%
  dplyr::mutate(cell = ifelse(is.na(auc), "\u2014",
                              sprintf("%.2f%s", auc, ifelse(sig %in% c("*","**","***"), sig, "")))) %>%
  dplyr::select(Label, Form, Stratum, cell) %>%
  tidyr::pivot_wider(names_from = Stratum, values_from = cell)
print(knitr::kable(disp,
      caption = "Table B-x. Univariate class AUC (with significance stars) across strata."))
```

```{r forest, fig.height=6.5, fig.cap="Figure B1. Univariate predictive AUC of each clinical covariate for the Low/High class, by stratum. Filled = p < alpha."}
pdat <- all_rows %>%
  dplyr::filter(!is.na(auc)) %>%
  dplyr::mutate(sig_fill = !is.na(p_model) & p_model < params$alpha,
                cov_form = paste0(Label, " (", Form, ")"))
ggplot(pdat, aes(auc, reorder(cov_form, auc), color = Stratum, shape = sig_fill)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60") +
  geom_point(size = 3, position = position_dodge(.5)) +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 1),
                     labels = c(`TRUE` = paste0("p < ", params$alpha),
                                `FALSE` = paste0("p \u2265 ", params$alpha))) +
  scale_color_manual(values = c(Overall = "#444444", TdaP = "#1b5aa0", TT = "#c0297a")) +
  labs(x = "univariate AUC (class = High)", y = NULL, color = "stratum", shape = NULL,
       title = "Clinical covariate predictive value by stratum") +
  theme_bw(base_size = 10)
```

---

# Part C — Maternal Serological Signature

🧭 *Maps to summary: "is there an antibody-side signature of the responder class
that the `r length(params$responder_features)` defining features don't already encode?"*

We assemble a maternal-serology matrix from **all** antigen × analyte features at
the maternal visits, **excluding the `r length(params$responder_features)` responder-defining features**
(`r paste(RESP_FEATS, collapse = ", ")`) so the signature is not circular with the
score. Features are Blom-standardised within each feature@visit.

```{r build-serology}
# Explicit guard: never let ACT or PENTAMER features enter the signature (they carry
# no measured rows in this dataset, but the exclusion is made explicit per protocol).
.excl_anti <- { .av <- unique(as.character(data_raw$feature)); .av[sub("_.*$", "", toupper(.av)) %in% c("ACT", "PENTAMER")] }
.excl <- union(RESP_FEATS, .excl_anti)
W_all  <- build_maternal_serology_matrix(data_raw, MAT_VIS, exclude_feats = .excl,
                                         min_coverage = params$min_coverage)
W_base <- build_maternal_serology_matrix(data_raw, params$baseline_visit,
                                         exclude_feats = .excl,
                                         min_coverage = params$min_coverage)
cat(sprintf("Serology predictor matrices: all maternal visits = **%d features**; ",
            ncol(W_all) - 1))
cat(sprintf("pre-vaccination (%s) baseline = **%d features**.\n\n",
            params$baseline_visit, ncol(W_base) - 1))
```

## Univariate AUC of each serology feature

```{r serology-univ, results='asis'}
su <- serology_univariate_auc(W_all, mat_cls)
su$base_feature <- sub("@.*$", "", su$feature)
cls_map <- classify_feature(unique(su$base_feature))
su_disp <- su %>%
  dplyr::left_join(cls_map, by = c("base_feature" = "feature")) %>%
  dplyr::mutate(auc = round(auc, 3), p = signif(p, 3), p_BH = signif(p_BH, 3)) %>%
  dplyr::select(feature, antigen, analyte_class, n, auc, p, p_BH)
print(knitr::kable(utils::head(su_disp, 20),
      caption = "Table C1. Top maternal serology features by univariate AUC for the Low/High class."))
```

```{r serology-univ-plot, fig.height=5, fig.cap="Figure C1. Top-15 maternal serology features by univariate AUC for the responder class."}
top <- utils::head(su, 15)
top$feature <- factor(top$feature, levels = rev(top$feature))
ggplot(top, aes(auc, feature, fill = auc >= 0.5)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60") +
  geom_col() +
  scale_fill_manual(values = c(`TRUE` = "#1b5aa0", `FALSE` = "#c0297a"),
                    labels = c(`TRUE` = "higher \u2192 High", `FALSE` = "higher \u2192 Low"),
                    name = NULL) +
  labs(x = "univariate AUC (class = High)", y = NULL,
       title = "Maternal serology features ranked by responder-class AUC") +
  theme_bw(base_size = 10)
```

## Arm-adjusted elastic-net signature (cross-validated)

The elastic net forces the **maternal arm in unpenalised** and lets the serology
features compete for the remaining signal; CV-AUC is the honest metric (in-sample
AUC is optimistic). We compare the pre-vaccination baseline matrix with the full
maternal matrix.

```{r elastic-net, results='asis'}
if (!has_glmnet) {
  cat("> **glmnet is not installed**, so the penalised model is skipped. ",
      "Install `glmnet` to run the cross-validated signature; the univariate AUC ",
      "ranking above still indicates which features carry signal.\n\n")
} else {
  en_base <- elastic_net_signature(W_base, mat_cls, arm_tbl = arm_tbl,
                                   alpha = params$en_alpha, nfolds = params$en_folds,
                                   seed = params$seed)
  en_all  <- elastic_net_signature(W_all,  mat_cls, arm_tbl = arm_tbl,
                                   alpha = params$en_alpha, nfolds = params$en_folds,
                                   seed = params$seed)
  cmp <- data.frame(
    model = c("arm only (reference)",
              sprintf("arm + baseline serology (%s)", params$baseline_visit),
              "arm + all maternal serology"),
    cv_auc = c(round(arm_auc, 3),
               if (isTRUE(en_base$ok)) round(en_base$cv_auc, 3) else NA,
               if (isTRUE(en_all$ok))  round(en_all$cv_auc, 3)  else NA),
    n_features_selected = c(0L,
               if (isTRUE(en_base$ok)) nrow(en_base$coefs) - 1L else NA_integer_,
               if (isTRUE(en_all$ok))  nrow(en_all$coefs)  - 1L else NA_integer_),
    stringsAsFactors = FALSE)
  print(knitr::kable(cmp,
        caption = "Table C2. Cross-validated AUC: does serology add to the arm? (arm-only is the Part-A rank statistic; serology rows are CV-AUC.)"))

  if (isTRUE(en_all$ok) && nrow(en_all$coefs) > 1) {
    co <- en_all$coefs[en_all$coefs$term != "(Intercept)", , drop = FALSE]
    co <- co[order(-abs(co$coef)), , drop = FALSE]
    co$coef <- round(co$coef, 3)
    print(knitr::kable(co, row.names = FALSE,
          caption = "Table C3. Non-zero elastic-net coefficients (arm forced in; standardised features)."))
  }
}
```

```{r en-roc, fig.height=4, fig.cap="Figure C2. In-sample ROC for the arm + all-maternal-serology elastic net (optimistic; compare CV-AUC in Table C2).", eval=has_glmnet}
if (has_glmnet && exists("en_all") && isTRUE(en_all$ok)) {
  pr <- as.numeric(predict(en_all$cv, newx = en_all$X, s = "lambda.min", type = "response"))
  rp <- roc_points(pr, en_all$y)
  ggplot(rp, aes(fpr, tpr)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey70") +
    geom_path(color = "#1b5aa0", linewidth = 1) +
    labs(x = "false-positive rate", y = "true-positive rate",
         title = "Arm + maternal serology — in-sample ROC") +
    coord_equal() + theme_bw(base_size = 10)
}
```

---

# Synthesis Questions to Address {#synthesis}

1. **Is the responder split mostly the vaccine?** Compare the arm-only AUC
   (Part A) to everything else. If clinical/serology AUCs barely exceed it, the
   Low/High class is essentially a vaccine-exposure label.
2. **Any clinical covariate beyond the arm?** In the arm-adjusted logistic
   (Table B-adj) and the within-arm screens, does any covariate
   (gestational age at vaccine, vaccine→birth interval, BMI, parity, Hb, age)
   stay significant? Within **TdaP** mothers especially — what predicts a
   vaccinated mother nonetheless responding poorly?
3. **Consistent across arms or arm-specific?** Does a predictor that matters in
   TdaP also matter in TT (a general "good responder" trait), or is it
   arm-specific (interacts with the vaccine)?
4. **Continuous vs threshold.** For any significant continuous covariate, did the
   continuous or tertile form win, and by how much (linear vs threshold effect)?
5. **Serological signature beyond the `r length(params$responder_features)` defining features.** Does the
   elastic-net CV-AUC (Table C2) exceed arm-only? Which features
   (Table C1/C3) — other pertussis antigens (FHA/FIM), carrier antigens
   (DT/TT), IgG subclasses, Fc-receptor binding, or effector functions — carry
   it, and is the signal **baseline (pre-vaccination)** or only concurrent at
   birth?
6. **Baseline predictivity (the clean test).** Does the pre-vaccination
   (`r params$baseline_visit`) baseline model predict who becomes a High
   responder? A positive baseline signal would mean responder status is partly a
   pre-existing maternal trait, not only a vaccine response.
7. **Mechanistic reading.** Pull Parts A–C together: is "Low responder" best
   described as (a) a TT/exposure label, (b) a clinical phenotype, (c) a
   pre-existing serological trait, or (d) largely unpredictable from the measured
   covariates?

---

# Global Participant Accounting {#global-naccounting}

```{r global-naccount, results='asis'}
manual_acct <- function()
  cat(sprintf("Mothers analysed: **%d** (Low = %d, High = %d); TdaP = %d, TT = %d.\n\n",
              nrow(data), sum(data$responder == "Low"), sum(data$responder == "High"),
              sum(data$arm_name == "TdaP"), sum(data$arm_name == "TT")))

if (!is.null(nt) && exists("nt_render", mode = "function")) {
  ok <- tryCatch({ nt_render(nt, caption = "Participant accounting for Phase 08."); TRUE },
                 error = function(e) FALSE)
  if (!ok) manual_acct()
} else {
  manual_acct()
}
```

---

*Phase 08 branch. The maternal Low/High class is the Phase 07 Stage A construct,
rebuilt via the new `R/C3_responder_helpers.R` (which calls the unmodified
`R/C4_C4_within_visit_helpers.R` engine). Outcome modelled as both the continuous
responder score (linear) and the binary class (logistic). Stratifier: maternal arm
only — no infant-arm analysis. Knit both formats with*
`rmarkdown::render("08_maternal_responder_predictors.Rmd", output_format = "all")`.
