---
title: "GaPs: maternal-group × antibody interaction tests and Deming (errors-in-variables) antibody–function regression"
subtitle: "Do the antibody–function slopes differ by maternal vaccine background? (PT↔PTNA, PRN↔SBA, PRN↔whole-cell IgG)"
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
    code_folding: show
  word_document:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 4.6
params:
  boot_B: 2000        # bootstrap replicates for Deming slope CIs / slope-difference test
  lambda: 1           # Deming error-variance ratio (sigma^2_y / sigma^2_x); 1 = orthogonal regression
  seed: 20260628
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(knitr)
})
has_car     <- requireNamespace("car",     quietly = TRUE)
has_deming  <- requireNamespace("deming",  quietly = TRUE)   # optional cross-check only
```

```{r bootstrap, include=FALSE}
## ---- self-locating resolver (same pattern as the other analysis files) ----
library(here)
shared_utils <- function(rel, optional = FALSE) {
  cand <- character(0)
  if (requireNamespace("here", quietly = TRUE))
    cand <- c(cand, here::here(rel), here::here("analysis", rel))
  rmd_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(rmd_dir)) cand <- c(cand, file.path(rmd_dir, rel),
                                 file.path(dirname(rmd_dir), rel))
  cand <- unique(c(cand, file.path(getwd(), rel), rel))
  hit  <- cand[file.exists(cand)]
  if (length(hit)) return(normalizePath(hit[1], winslash = "/", mustWork = FALSE))
  if (optional) return(NULL)
  stop("Could not locate '", rel, "'.", call. = FALSE)
}
source(shared_utils("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C1_serology_helpers.R"), local = FALSE, encoding = "UTF-8")

data_raw <- load_serology_data(shared_utils("data/c_set.RData"))   # applies VISIT_RECODE
clin     <- load_clinical_data(shared_utils("data/clin_assess.RData"))
clin_age <- clin %>% dplyr::select(subject_accession, maternal_age) %>% distinct()

ACCENT <- "#8A1538"; MARM_COL <- c(TdaP = "#3a8bd6", TT = "#e06a5a")
```

# Design

These two analyses formalise the section-D claim. Section D shows the **expanded
maternal model fits better after Tdap-IPV** (e.g. SBA R² 68.7% vs 17.9%), but a
difference in variance explained can arise from a different *slope*, a different
*predictor range* (TT mothers have little PRN to transfer, so a restricted
x-range deflates R² even at an identical slope), or different residual scatter.
To separate these we fit:

1. **Interaction models** — `function ~ maternal group * matched antibody + maternal age`,
   standardised, so the `group × antibody` term tests directly whether the
   antibody–function **slope** differs by maternal background. We run this for the
   matched **IgG1** predictor at two visits: maternal **delivery** (MatBirth —
   the section-D predictor) and **concurrent** infant (InfMon2 — the section-E
   predictor).
2. **Deming / errors-in-variables regression** — both antibody and function are
   measured with error, so OLS slopes are attenuated. Deming regression
   (λ = `r params$lambda`; λ = 1 is orthogonal regression) estimates the
   error-corrected slope. We fit it **within each maternal arm** and test the
   slope difference by stratified bootstrap. The concurrent (same-sample, InfMon2)
   pairing is the appropriate EIV setting; the maternal-delivery pairing is also
   reported to reconcile with table 2.

**Matched antibody–function pairs** (the dominant correlates identified earlier):
PT IgG1 ↔ PTNA, PRN IgG1 ↔ SBA, PRN IgG1 ↔ whole-cell IgG binding.

```{r spec}
PAIRS <- list(
  list(key="PTNA",   outcome="WHOLE_PTNA",   antigen="PT",  fn="PT-neutralising activity (PTNA)"),
  list(key="SBA",    outcome="WHOLE_SBA",    antigen="PRN", fn="serum bactericidal activity (SBA)"),
  list(key="WT_IgG", outcome="WHOLE_WT_IgG", antigen="PRN", fn="whole-cell IgG binding")
)
PRED_ANALYTE <- "IgG1"   # matched-subclass predictor (change to "IgG" to reconcile with table-2 totals)

## case-insensitive feature resolver (guards the IgG/IGG casing issue)
.avail <- as.character(unique(data_raw$feature))
resolve_feat <- function(w) { hit <- .avail[toupper(.avail) == toupper(w)]; if (length(hit)) hit[[1]] else NA_character_ }

## Build a one-row-per-subject frame: x = predictor antibody at pred_visit,
## y = whole-cell function at outcome_visit, + maternal arm + maternal age.
build_pair <- function(outcome_feat, pred_feat, pred_visit, outcome_visit = "InfMon2") {
  of <- resolve_feat(outcome_feat); pf <- resolve_feat(pred_feat)
  if (is.na(of) || is.na(pf)) return(NULL)
  d <- data_raw %>%
    dplyr::filter(feature %in% c(of, pf),
                  visit_name %in% c(pred_visit, outcome_visit),
                  as.character(arm_name) %in% MATERNAL_ARMS) %>%
    dplyr::mutate(key = paste(feature, visit_name, sep = "@")) %>%
    dplyr::distinct(subject_accession, arm_name, key, .keep_all = TRUE) %>%
    dplyr::select(subject_accession, arm_name, key, log_assay_value) %>%
    tidyr::pivot_wider(names_from = key, values_from = log_assay_value,
                       values_fn = function(z) mean(z, na.rm = TRUE))
  xcol <- paste(pf, pred_visit, sep = "@"); ycol <- paste(of, outcome_visit, sep = "@")
  if (!all(c(xcol, ycol) %in% names(d))) return(NULL)
  tibble::tibble(subject_accession = d$subject_accession,
                 arm = factor(as.character(d$arm_name), levels = MATERNAL_ARMS),
                 x = d[[xcol]], y = d[[ycol]]) %>%
    dplyr::left_join(clin_age, by = "subject_accession") %>%
    dplyr::filter(is.finite(x), is.finite(y))
}
```

# 1. Maternal vaccine group × antibody interaction tests

```{r interaction-engine}
## Standardised interaction model: function ~ arm * antibody + maternal age.
## Returns the group x antibody coefficient (difference in standardised slope,
## TdaP minus TT), each arm's simple slope, and model fit.
fit_interaction <- function(df) {
  d <- df %>% dplyr::filter(is.finite(x), is.finite(y), is.finite(maternal_age))
  if (nrow(d) < 30 || dplyr::n_distinct(d$arm) < 2) return(NULL)
  d$zx <- as.numeric(scale(d$x)); d$zy <- as.numeric(scale(d$y))
  d$za <- as.numeric(scale(d$maternal_age))
  d$arm <- relevel(factor(d$arm), ref = "TT")            # TT = comparator
  m  <- lm(zy ~ arm * zx + za, data = d)
  co <- summary(m)$coefficients
  it <- grep("(armTdaP:zx|zx:armTdaP)", rownames(co), value = TRUE)[1]
  ci <- suppressMessages(confint(m))
  slope_TT   <- co["zx", "Estimate"]
  slope_TdaP <- slope_TT + co[it, "Estimate"]
  # Type III test for the interaction (single df -> equals the t-test); use car if available
  p_int <- co[it, "Pr(>|t|)"]
  if (has_car) {
    a3 <- tryCatch(car::Anova(m, type = 3), error = function(e) NULL)
    if (!is.null(a3) && "arm:zx" %in% rownames(a3)) p_int <- a3["arm:zx", "Pr(>F)"]
  }
  data.frame(n = nrow(d),
             slope_TT = round(slope_TT, 3), slope_TdaP = round(slope_TdaP, 3),
             int_diff = round(co[it, "Estimate"], 3),
             ci_lo = round(ci[it, 1], 3), ci_hi = round(ci[it, 2], 3),
             p_interaction = signif(p_int, 3),
             adjR2 = round(summary(m)$adj.r.squared, 3))
}

run_interaction_set <- function(pred_visit) {
  rows <- lapply(PAIRS, function(p) {
    df <- build_pair(p$outcome, paste0(p$antigen, "_", PRED_ANALYTE), pred_visit)
    if (is.null(df)) return(NULL)
    f <- fit_interaction(df)
    if (is.null(f)) return(NULL)
    cbind(Outcome = p$fn, Predictor = paste0(p$antigen, " ", PRED_ANALYTE, " @ ", pred_visit), f)
  })
  do.call(rbind, rows)
}
```

## 1a. Maternal-delivery antibody predictor — addresses section D

```{r int-matbirth}
int_mat <- run_interaction_set("MatBirth")
kable(int_mat, row.names = FALSE,
      col.names = c("Infant outcome", "Predictor", "n",
                    "slope (TT)", "slope (TdaP)", "Δslope (TdaP−TT)",
                    "95% CI low", "95% CI high", "p (interaction)", "adj R²"),
      caption = "Table 1a. Standardised maternal-group × maternal-delivery IgG1 interaction. Slopes are standardised (per SD of antibody, in SD of function). A significant Δslope means the maternal antibody–infant function slope differs by maternal vaccine group; a non-significant Δslope means the section-D R² gap is better explained by predictor range / residual variance than by a slope difference.")
```

## 1b. Concurrent infant antibody predictor — addresses section E

```{r int-infmon2}
int_inf <- run_interaction_set("InfMon2")
kable(int_inf, row.names = FALSE,
      col.names = c("Infant outcome", "Predictor", "n",
                    "slope (TT)", "slope (TdaP)", "Δslope (TdaP−TT)",
                    "95% CI low", "95% CI high", "p (interaction)", "adj R²"),
      caption = "Table 1b. Standardised maternal-group × concurrent (InfMon2) infant IgG1 interaction — whether the same-sample antibody→function slope in the infant differs by maternal background.")
```

```{r int-plot, fig.width=9, fig.height=3.4}
## visualise the InfMon2 (concurrent) simple slopes
plot_df <- do.call(rbind, lapply(PAIRS, function(p) {
  df <- build_pair(p$outcome, paste0(p$antigen, "_", PRED_ANALYTE), "InfMon2")
  if (is.null(df)) return(NULL)
  d <- df %>% filter(is.finite(x), is.finite(y))
  d$Outcome <- p$fn; d
}))
ggplot(plot_df, aes(x, y, colour = arm)) +
  geom_point(alpha = .35, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  facet_wrap(~ Outcome, scales = "free", nrow = 1) +
  scale_colour_manual(values = MARM_COL, name = "Maternal arm") +
  labs(x = paste0("log10 matched ", PRED_ANALYTE, " (InfMon2)"), y = "log10 function (InfMon2)",
       title = "Concurrent antibody–function relationship by maternal arm (OLS lines)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold", colour = ACCENT))
```

# 2. Deming (errors-in-variables) regression by maternal arm

```{r deming-engine}
## Closed-form Deming slope/intercept for a given error-variance ratio lambda.
deming_fit <- function(x, y, lambda = 1) {
  ok <- is.finite(x) & is.finite(y); x <- x[ok]; y <- y[ok]; n <- length(x)
  if (n < 5) return(c(intercept = NA, slope = NA, n = n))
  mx <- mean(x); my <- mean(y)
  sxx <- sum((x - mx)^2) / (n - 1); syy <- sum((y - my)^2) / (n - 1)
  sxy <- sum((x - mx) * (y - my)) / (n - 1)
  b <- ((syy - lambda * sxx) + sqrt((syy - lambda * sxx)^2 + 4 * lambda * sxy^2)) / (2 * sxy)
  c(intercept = my - b * mx, slope = b, n = n)
}

## Per-arm Deming with stratified bootstrap CIs and a slope-difference test.
deming_by_arm <- function(df, lambda = 1, B = 2000, seed = 1) {
  set.seed(seed)
  arms <- MATERNAL_ARMS
  d <- df %>% filter(is.finite(x), is.finite(y))
  obs <- lapply(arms, function(a) deming_fit(d$x[d$arm == a], d$y[d$arm == a], lambda))
  names(obs) <- arms
  diff_obs <- unname(obs[[1]]["slope"] - obs[[2]]["slope"])
  bs <- matrix(NA_real_, B, length(arms), dimnames = list(NULL, arms))
  for (b in seq_len(B)) for (a in arms) {
    da <- d[d$arm == a, ]; idx <- sample(nrow(da), replace = TRUE)
    bs[b, a] <- deming_fit(da$x[idx], da$y[idx], lambda)["slope"]
  }
  diffs <- bs[, arms[1]] - bs[, arms[2]]
  q <- function(v) round(stats::quantile(v, c(.025, .975), na.rm = TRUE), 3)
  list(obs = obs, diff_obs = diff_obs,
       ci = lapply(arms, function(a) q(bs[, a])),
       ci_diff = q(diffs),
       p_diff = round(2 * min(mean(diffs <= 0, na.rm = TRUE), mean(diffs >= 0, na.rm = TRUE)), 4))
}

run_deming_set <- function(pred_visit) {
  res <- list(); rows <- list()
  for (p in PAIRS) {
    df <- build_pair(p$outcome, paste0(p$antigen, "_", PRED_ANALYTE), pred_visit)
    if (is.null(df)) next
    r <- deming_by_arm(df, lambda = params$lambda, B = params$boot_B, seed = params$seed)
    res[[p$key]] <- list(df = df, r = r, p = p)
    rows[[p$key]] <- data.frame(
      Outcome = p$fn, Predictor = paste0(p$antigen, " ", PRED_ANALYTE, " @ ", pred_visit),
      n_TdaP = unname(r$obs$TdaP["n"]), n_TT = unname(r$obs$TT["n"]),
      slope_TdaP = round(unname(r$obs$TdaP["slope"]), 3),
      ci_TdaP = paste0("[", r$ci[[1]][1], ", ", r$ci[[1]][2], "]"),
      slope_TT = round(unname(r$obs$TT["slope"]), 3),
      ci_TT = paste0("[", r$ci[[2]][1], ", ", r$ci[[2]][2], "]"),
      slope_diff = round(r$diff_obs, 3),
      ci_diff = paste0("[", r$ci_diff[1], ", ", r$ci_diff[2], "]"),
      p_diff = r$p_diff)
  }
  list(table = do.call(rbind, rows), res = res)
}

deming_plot <- function(set, pred_visit) {
  for (k in names(set$res)) {
    o <- set$res[[k]]; r <- o$r; d <- o$df %>% filter(is.finite(x), is.finite(y))
    g <- ggplot(d, aes(x, y, colour = arm)) +
      geom_point(alpha = .35, size = 1) +
      geom_abline(intercept = r$obs$TdaP["intercept"], slope = r$obs$TdaP["slope"],
                  colour = MARM_COL["TdaP"], linewidth = 1) +
      geom_abline(intercept = r$obs$TT["intercept"], slope = r$obs$TT["slope"],
                  colour = MARM_COL["TT"], linewidth = 1) +
      geom_smooth(aes(group = arm), method = "lm", se = FALSE, linetype = "dashed",
                  linewidth = .6, alpha = .7) +
      scale_colour_manual(values = MARM_COL, name = "Maternal arm") +
      labs(title = paste0(o$p$fn, " — Deming by arm (", pred_visit, " predictor)"),
           subtitle = "solid = Deming (errors-in-variables); dashed = OLS",
           x = paste0("log10 ", o$p$antigen, " ", PRED_ANALYTE, " (", pred_visit, ")"),
           y = paste0("log10 ", o$p$key)) +
      theme_minimal(base_size = 10) +
      theme(legend.position = "bottom", plot.title = element_text(face = "bold", colour = ACCENT, size = 11))
    print(g)
  }
}
```

## 2a. Concurrent (same-sample, InfMon2) — the appropriate EIV setting

```{r deming-infmon2, fig.width=6.2, fig.height=4}
dem_inf <- run_deming_set("InfMon2")
kable(dem_inf$table, row.names = FALSE,
      col.names = c("Outcome", "Predictor", "n (TdaP)", "n (TT)",
                    "slope TdaP", "95% CI", "slope TT", "95% CI",
                    "Δslope", "95% CI", "p (diff)"),
      caption = sprintf("Table 2a. Deming regression (λ = %s) of concurrent infant function on matched infant IgG1, fitted within each maternal arm. Δslope and its bootstrap CI/p test whether the error-corrected antibody→function slope differs by maternal background.", params$lambda))
deming_plot(dem_inf, "InfMon2")
```

## 2b. Maternal-delivery predictor — reconciles with table 2

```{r deming-matbirth, fig.width=6.2, fig.height=4}
dem_mat <- run_deming_set("MatBirth")
kable(dem_mat$table, row.names = FALSE,
      col.names = c("Outcome", "Predictor", "n (TdaP)", "n (TT)",
                    "slope TdaP", "95% CI", "slope TT", "95% CI",
                    "Δslope", "95% CI", "p (diff)"),
      caption = sprintf("Table 2b. Deming regression (λ = %s) of infant function (InfMon2) on maternal-delivery IgG1 (MatBirth), within each maternal arm. Use this to reconcile the slope with the maternal-model coefficients in table 2.", params$lambda))
deming_plot(dem_mat, "MatBirth")
```

# 3. Combined summary and how to read it

```{r summary}
get_row <- function(tab, fn, col) {
  if (is.null(tab) || !"Outcome" %in% names(tab)) return(NA)
  v <- tab[tab$Outcome == fn, col]; if (length(v)) v[1] else NA
}
out <- do.call(rbind, lapply(PAIRS, function(p) {
  data.frame(
    Outcome = p$fn,
    `Interaction p (MatBirth predictor)`  = get_row(int_mat, p$fn, "p_interaction"),
    `Interaction p (InfMon2 predictor)`   = get_row(int_inf, p$fn, "p_interaction"),
    `Deming slope diff (InfMon2)`         = get_row(dem_inf$table, p$fn, "slope_diff"),
    `Deming p (InfMon2)`                  = get_row(dem_inf$table, p$fn, "p_diff"),
    check.names = FALSE)
}))
kable(out, row.names = FALSE, caption = "Table 3. Combined slope-difference evidence. If interaction p and Deming p are non-significant, section D should remain phrased as a difference in variance explained, not a difference in slope. If significant (especially for the surface outcomes after Tdap-IPV), the antibody-to-function slope itself differs by maternal background and the wording can be strengthened.")
```

**Interpretation.**

* **Interaction tests (section 1).** The `group × antibody` term is the formal test
  the analyst note asks for. A non-significant interaction (with a clearly higher
  per-arm R² after Tdap-IPV in section D) indicates the section-D gap is driven by
  the *range and abundance* of transferable PRN antibody — there is more PRN to
  work with after Tdap-IPV — rather than by a steeper antibody→function conversion.
  In that case keep section D as "variance explained," as currently written. A
  significant interaction licenses a slope-difference statement.
* **Deming (section 2).** Error-corrected slopes are systematically steeper than the
  OLS (dashed) lines; the per-arm Deming slopes and their bootstrap difference give
  the errors-in-variables version of the same question and connect to the earlier
  GaPs Deming result. Table 2b (maternal-delivery predictor) is the panel to cite
  alongside results table 2.
* **λ caveat.** Deming assumes a known error-variance ratio λ (set to
  `r params$lambda`). With replicate-based assay CVs you can set λ to the true
  ratio; the slopes are mildly sensitive to λ, the *slope-difference* test much less so.

```{r optional-deming-pkg, eval=has_deming, echo=FALSE}
## Optional cross-check against the 'deming' package, if installed.
cat("Cross-check: the 'deming' package is available; closed-form slopes above can be ",
    "verified with deming::deming(y ~ x, data, ...). Not run by default.\n")
```

```{r sessioninfo}
sessionInfo()
```
