---
title: "Systems Serology — Concurrent Within-Visit Associations"
subtitle: "Focused endpoint-parameterized driver · SBA · WT_IgG · PTNA · maternal-arm primary (TdaP vs TT)"
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
params:
  endpoint: "SBA"      # "SBA" | "WT_IgG" | "PTNA"  <-- the ONLY thing to change per endpoint
  data_file: "data/d_set.RData"   # <-- switched from data/d_set.RData
  data_object: "ebaa_extra2"      # <-- switched from ebaa_extra2 (more ADCD at maternal/cord)
  nct_it: 1000                    # NCT permutations; set e.g. 100 for a fast pass
  run_nct: true                   # STAGE 5 (most expensive) on/off
output:
  html_document:        # -> hand this to the summary engine
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:        # -> hand this to the analysis recipient
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
```

```{r bootstrap, include=FALSE}
library(here)

# ---- self-locating resolver -------------------------------------------------
# Finds config/, R/, parts/, data/ whether they sit beside this .Rmd, under an
# analysis/ subfolder, or at the project root — and works even when rmarkdown
# renders from a temp directory. Returns the first candidate that exists.
# ---- shared path resolver: single definition in R/shared_utils.R ----
if (!exists("shared_utils", mode = "function")) {
  .ff_cand <- c("R/shared_utils.R", "../R/shared_utils.R",
                "analysis/R/shared_utils.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/shared_utils.R"),
                  here::here("analysis/R/shared_utils.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/shared_utils.R"),
                  file.path(dirname(.ff_rmd), "R/shared_utils.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/shared_utils.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}

# ---- shared configuration + helper library (sourced once) ----
# NOTE: identical bootstrap to the full *_analysis.Rmd drivers. We deliberately
# do NOT source config/endpoints_additions.R here: the concurrent part needs
# only ANTIGENS = c("PT","FHA","PRN") (endpoints.R) for add_concurrent_ratios(),
# and the FcG/effector vocabularies in the additions file belong to Parts 4/5.
source(shared_utils("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C1_serology_helpers.R"), local = FALSE, encoding = "UTF-8")

# ---- resolve the endpoint config from the single param ----
cfg <- get_endpoint_config(params$endpoint)

# ---- load + recode the antibody data ONCE ----
# Switched to d_set / ebaa_extra2 (same schema as d_set / ebaa_extra2, with more
# ADCD at maternal + cord blood). load_serology_data() rebuilds `feature` from
# antigen+analyte (mixed case), calls verify_predictor_casing() internally, and
# a benign "unmapped visit codes" warning (P05/P08) is expected. If the d_set
# file is absent this errors clearly via shared_utils(); drop it in data/.
.data_path <- shared_utils(params$data_file)
data_raw <- load_serology_data(.data_path, object_name = params$data_object)

# ---- staging controls consumed by parts/03 (see that file's conc-setup) ----
CONC_VISIT <- "InfMon2"                 # infant 8-week timepoint only
DATA_MODES <- c("raw")                  # untransformed only (igg_std deferred)
NCT_IT     <- as.integer(params$nct_it)
RUN_NCT    <- isTRUE(params$run_nct)

# ---- participant N-tracker (concurrent models + networks only) ----
nt <- new_ntracker()
```

# `r cfg$outcome_label`: Concurrent Within-Visit Antibody Analysis {-}

This is a **focused, expanded extract** of the full config-driven serology pipeline.
It runs **only the concurrent (within-visit) association layer** for the infant's
**`r cfg$outcome_long`** (`r cfg$outcome_feature`), **restricted to the 8-week infant
timepoint (`r CONC_VISIT`)**, and it now covers the expanded feature set — FIM plus
the FcγR binding and antigen-specific effector functions — on the untransformed (raw)
log assay values. The total-IgG-standardised mode is deferred. The maternal-chain,
clinical, and other pipeline parts are excluded.

The endpoint is set once (`params$endpoint = "`r params$endpoint`"`); switching to
`WT_IgG` or `PTNA` is a one-line change to that parameter (or use the
`render_concurrent_all()` helper in the appendix to knit all three at once).

**Visit naming.** All visits are recoded for clarity throughout:
`r paste(paste0(names(VISIT_RECODE), "\u2192", VISIT_RECODE), collapse = ", ")`.
This analysis is restricted to the **8-week infant timepoint (`r CONC_VISIT`)**.

**Stratification.** Primary axis is the **maternal vaccination arm (TdaP vs TT)**;
within each, results are reported pooled across infant arms and separately for each
infant priming level (aP, wP).

**Downstream use.** The knitted **HTML** is intended for the summary engine; the
**Word** document is for the analysis recipient. Inline "🧭 Maps to interpretation"
callouts tie each block to the matching section of the written concurrent-analysis
summary for this endpoint.

---

```{r part-concurrent, child=shared_utils('parts/C7_step00_networks.Rmd')}
# Expanded + staged concurrent within-visit part (FIM + FcγR + ADCD/ADCP/ADNP,
# untransformed/raw, 8-week visit only). Reuses cfg, data_raw, nt,
# shared_utils() and the staging controls set in the bootstrap chunk above.
```

---

# Synthesis Questions to Address (Concurrent) {#synthesis}

*These are the concurrent-analysis subset of the pipeline's synthesis prompts,
restated inline so the summary engine has the questions alongside the evidence
above. Each maps to a "Maps to interpretation" callout in the results.*

1. **Antigen / subclass relative importance** — within-visit, which antigen and
   which IgG subclass carries the signal for `r cfg$outcome_label`? Is PRN
   dominance reproduced (feeds §1), and is the FHA association negative (feeds §2)?
2. **Arm-specific signatures** — does the leading concurrent predictor differ
   between TdaP and TT mothers (full-12 relative-importance comparison; feeds §3)?
3. **Pro-/tolerogenic ratio effect** — do the PT/FHA/PRN IgG1+IgG3 vs IgG2+IgG4
   ratios add a qualitative (subclass-balance) layer to the within-visit outcome,
   overall and by visit, and does it differ by maternal arm (feeds §4)?
4. **Antigen × subclass × level interaction** — does *which* subclass matters
   depend on antigen and on binding level (the three-way mixed-model term; feeds §5)?
5. **Network architecture and bridge centrality** — at
   `r if ("InfMon5" %in% cfg$concurrent_visits) "InfMon5" else tail(cfg$concurrent_visits, 1)`,
   which nodes bridge the outcome to the antigen communities, and does the bridge
   structure differ between TdaP and TT (feeds §6)?
6. **Temporal and cross-arm network comparison (NCT)** — is the within-visit
   network invariant across the first vs last concurrent visit (temporal, feeds §7),
   and across maternal arm / infant arm at the reference visit (feeds §8)?
7. **Convergence with the maternal chain** — does the dominant concurrent antigen
   agree with the dominant antigen identified by the upstream maternal-chain part
   (cross-reference the full `r cfg$outcome_label` analysis)?

---

# Participant Accounting (Concurrent) {#naccounting}

A single attrition ledger for every concurrent model and network fitted in this
document, so differing sample sizes between sections are fully traceable. (This
mirrors the concurrent slice of the full pipeline's global ledger.)

```{r naccount, results='asis'}
nt_render(nt, caption = "Participant accounting across all concurrent models / networks.")
```

---

# Appendix — Reproducibility {#appendix}

**Environment.** Network and NCT sections require `bootnet`, `qgraph`,
`networktools`, and `NetworkComparisonTest`; the mixed model requires `lme4`;
partial-correlation heatmaps require `corrplot`. If any are absent the
corresponding sections are skipped with an inline note and the regression +
ratio layers still run. `NCT_IT` (permutations) is set inside the concurrent
part; lower it for a quick pass.

**Knit one endpoint.**

```r
rmarkdown::render("concurrent_analysis.Rmd",
                  params = list(endpoint = "SBA"),      # or "WT_IgG" / "PTNA"
                  output_file = "concurrent_SBA.html",
                  output_format = "html_document")
```

**Knit all three** (one HTML per endpoint, named by endpoint):

```r
render_concurrent_all <- function(rmd = "concurrent_analysis.Rmd",
                                  endpoints = c("SBA", "WT_IgG", "PTNA"),
                                  fmt = "html_document") {
  for (ep in endpoints) {
    ext <- if (grepl("word", fmt)) "docx" else "html"
    rmarkdown::render(
      rmd,
      params        = list(endpoint = ep),
      output_file   = sprintf("concurrent_%s.%s", ep, ext),
      output_format = fmt,
      envir         = new.env())   # fresh environment per endpoint
  }
}
# render_concurrent_all()
```

*Generated by the config-driven serology pipeline (concurrent-only driver). Place
this file beside the existing `*_analysis.Rmd` drivers so it can locate `config/`,
`R/`, `parts/`, and `data/`.*
