---
title: "GaPs (Part B): predicting infant InfMon2 whole-bacterial function from maternal responder status, in excess of the maternal arm"
subtitle: |
  An equal-chance test. Brings the Phase-07/08 maternal responder score into the
  infant InfMon2 functional models on the same footing as the matched antigen-specific
  antibody, using order-independent (Shapley) variance attribution, an orthogonalised
  breadth predictor for the genuinely non-redundant signal, honest out-of-sample
  (cross-validated) increments, and a within-arm de-confounding check.
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
    code_folding: show
params:
  responder_features: !r c("PT_IgG", "FHA_IgG", "PRN_IgG")
  maternal_visits:    !r c("PregEarly", "MatBirth")
  predictor_visit:    "MatBirth"     # forward predictor (maternal-at-delivery); not concurrent
  outcome_visit:      "InfMon2"
  breadth_thresh:     0.6
  min_coverage:       0.6
  min_breadth_feats:  5              # require >= this many non-matched maternal features for the breadth index
  cv_folds:           10
  boot_B:             1000           # bootstrap reps for Shapley-share CIs
  seed:               20260628
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
suppressPackageStartupMessages({ library(dplyr); library(tidyr); library(ggplot2); library(knitr) })
has_relaimpo <- requireNamespace("relaimpo", quietly = TRUE)   # optional cross-check for LMG
has_car      <- requireNamespace("car",      quietly = TRUE)   # optional VIF
```

```{r bootstrap, include=FALSE}
library(here)
# ---- shared path resolver: single definition in R/shared_utils.R ----
if (!exists("shared_utils", mode = "function")) {
  .ff_cand <- c("R/shared_utils.R", "../R/shared_utils.R",
                "analysis/R/shared_utils.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/shared_utils.R"),
                  here::here("analysis/R/shared_utils.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/shared_utils.R"),
                  file.path(dirname(.ff_rmd), "R/shared_utils.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/shared_utils.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}
source(shared_utils("config/endpoints.R"),             local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C1_serology_helpers.R"),           local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C4_within_visit_helpers.R"),       local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C3_responder_helpers.R"), local = FALSE, encoding = "UTF-8")

data_raw <- load_serology_data(shared_utils("data/d_set.RData"))     # applies VISIT_RECODE
clin     <- tryCatch(load_clinical_data(shared_utils("data/clin_assess.RData")), error = function(e) NULL)
if (is.null(clin)) { .e <- new.env(); load(shared_utils("data/clin_assess.RData"), envir = .e); clin <- get("clin_assess", envir = .e) }
clin_age <- clin %>% dplyr::select(subject_accession, maternal_age) %>% distinct()
has_cluster <- requireNamespace("cluster", quietly = TRUE)
ACCENT <- "#8A1538"; MARM_COL <- c(TdaP = "#1b5aa0", TT = "#c0297a")
```

```{r dplyr-guards, include=FALSE}
# Force dplyr to win over any masking packages loaded above (e.g. MASS, data.table).
select  <- dplyr::select
filter  <- dplyr::filter
mutate  <- dplyr::mutate
arrange <- dplyr::arrange
```

# What this tests

The responder score is a function of maternal PT, FHA and PRN IgG — the same antibodies that
already dominate the Section C–E infant-function models. A fair ("equal-chance") test must therefore
(1) put the score in the **same** InfMon2 outcome frame and CV protocol as those models, (2) credit
shared variance by an **order-independent Shapley decomposition** rather than by entry order, (3) isolate
the genuinely non-redundant part of the Phase-08 signal as a **breadth predictor orthogonalised to the
matched antibody**, and (4) keep the test **forward** (maternal-at-delivery) and **within-arm** so the
score cannot win by proxying the matched antigen or the vaccine label.

Matched antibody per outcome: **PT IgG1 → PTNA**, **PRN IgG1 → SBA and whole-cell IgG**.

```{r build-score}
RESP_FEATS <- params$responder_features
MAT_VIS    <- params$maternal_visits

mat_cls <- build_maternal_responder_class(data_raw, MAT_VIS, RESP_FEATS,
                                          breadth_thresh = params$breadth_thresh, has_cluster = has_cluster)
# expected columns: subject_accession, responder (Low/High), score (continuous)
stopifnot(all(c("subject_accession", "responder", "score") %in% names(mat_cls)))
arm_tbl <- subject_maternal_arm(data_raw)

# Maternal serology matrix (Blom-standardised; score-defining features excluded) for the breadth index
.excl_anti <- { .av <- unique(as.character(data_raw$feature)); .av[sub("_.*$", "", toupper(.av)) %in% c("ACT", "PENTAMER")] }
W_all <- build_maternal_serology_matrix(data_raw, MAT_VIS, exclude_feats = union(RESP_FEATS, .excl_anti),
                                        min_coverage = params$min_coverage)
feat_cols <- setdiff(names(W_all), "subject_accession")
```

```{r helpers}
## case-insensitive feature resolver
.avail <- as.character(unique(data_raw$feature))
resolve_feat <- function(w) { hit <- .avail[toupper(.avail) == toupper(w)]; if (length(hit)) hit[[1]] else NA_character_ }

## per-subject log value of one feature at one visit
subj_value <- function(feat, visit, newname) {
  f <- resolve_feat(feat); if (is.na(f)) return(NULL)
  data_raw %>%
    dplyr::filter(feature == f, visit_name == visit) %>%
    dplyr::group_by(subject_accession) %>%
    dplyr::summarise(!!newname := mean(log_assay_value, na.rm = TRUE), .groups = "drop")
}

## breadth index for an outcome = rowMeans of Blom-standardised maternal features,
## EXCLUDING every feature of the matched antigen (so it is non-redundant with the matched antibody)
breadth_index <- function(matched_antigen) {
  drop <- feat_cols[startsWith(sub("@.*$", "", feat_cols), paste0(matched_antigen, "_"))]
  keep <- setdiff(feat_cols, drop)
  M <- as.matrix(W_all[, keep, drop = FALSE])
  nok <- rowSums(is.finite(M))
  br <- ifelse(nok >= params$min_breadth_feats, rowMeans(M, na.rm = TRUE), NA_real_)
  tibble::tibble(subject_accession = W_all$subject_accession, breadth_raw = br)
}

## k-fold CV R^2, folds stratified by maternal arm so every fold has both levels
cv_r2 <- function(formula, data, k = params$cv_folds, seed = params$seed) {
  set.seed(seed)
  resp <- all.vars(formula)[1]
  fold <- integer(nrow(data))
  for (a in unique(data$arm)) { idx <- which(data$arm == a); fold[idx] <- sample(rep(seq_len(k), length.out = length(idx))) }
  pred <- numeric(nrow(data))
  for (i in seq_len(k)) {
    tr <- data[fold != i, , drop = FALSE]; te <- data[fold == i, , drop = FALSE]
    m  <- lm(formula, tr)
    pred[fold == i] <- tryCatch(predict(m, te), error = function(e) rep(mean(tr[[resp]]), nrow(te)))
  }
  y <- data[[resp]]
  1 - sum((y - pred)^2) / sum((y - mean(y))^2)
}
r2_of <- function(terms, base, resp, data) summary(lm(reformulate(c(base, terms), resp), data))$r.squared

## Shapley shares of candidate predictors over the R^2 increment beyond a fixed base set
shapley <- function(resp, base, cands, data) {
  k <- length(cands)
  subsets_of <- function(v) unlist(lapply(0:length(v), function(m) combn(v, m, simplify = FALSE)), recursive = FALSE)
  phi <- setNames(numeric(k), cands)
  for (j in cands) {
    others <- setdiff(cands, j)
    for (S in subsets_of(others)) {
      w <- factorial(length(S)) * factorial(k - length(S) - 1) / factorial(k)
      phi[j] <- phi[j] + w * (r2_of(c(S, j), base, resp, data) - r2_of(S, base, resp, data))
    }
  }
  phi  # sums to R2(full) - R2(base)
}
```

```{r outcome-spec}
OUTCOMES <- list(
  PTNA = list(out = "WHOLE_PTNA",   antigen = "PT",  matched = "PT_IgG1",  label = "PT-neutralising activity"),
  SBA  = list(out = "WHOLE_SBA",    antigen = "PRN", matched = "PRN_IgG1", label = "serum bactericidal activity"),
  WT   = list(out = "WHOLE_WT_IgG", antigen = "PRN", matched = "PRN_IgG1", label = "whole-cell IgG binding")
)

## Build the per-outcome modelling frame and z-score predictors within the analysis sample.
build_df <- function(spec) {
  y  <- subj_value(spec$out,     params$outcome_visit,   "y")
  ab <- subj_value(spec$matched, params$predictor_visit, "ab")
  if (is.null(y) || is.null(ab)) return(NULL)
  br <- breadth_index(spec$antigen)
  d <- y %>%
    dplyr::inner_join(ab, by = "subject_accession") %>%
    dplyr::inner_join(mat_cls[, c("subject_accession", "score", "responder")], by = "subject_accession") %>%
    dplyr::inner_join(arm_tbl, by = "subject_accession") %>%
    dplyr::left_join(clin_age, by = "subject_accession") %>%
    dplyr::left_join(br, by = "subject_accession") %>%
    dplyr::filter(as.character(arm_name) %in% MATERNAL_ARMS, is.finite(y), is.finite(ab),
                  is.finite(score), is.finite(maternal_age))
  d$arm <- relevel(factor(as.character(d$arm_name), levels = MATERNAL_ARMS), ref = "TT")
  z <- function(v) as.numeric(scale(v))
  d$ab_z <- z(d$ab); d$score_z <- z(d$score); d$age_z <- z(d$maternal_age); d$y <- as.numeric(scale(d$y))
  # orthogonalise breadth to the matched antibody (the non-redundant part), on rows where breadth exists
  d$breadth_resid_z <- NA_real_
  ok <- is.finite(d$breadth_raw)
  if (sum(ok) > 20) {
    bz <- z(d$breadth_raw[ok])
    d$breadth_resid_z[ok] <- as.numeric(scale(residuals(lm(bz ~ d$ab_z[ok]))))
  }
  d
}
DFS <- lapply(OUTCOMES, build_df)
```

# Responder status in excess of the maternal arm

The primary question of Part B: do the maternal responder predictors — the **High vs Low class** and
the continuous **responder score** — predict each infant InfMon2 whole-bacterial outcome (SBA, whole-cell
IgG, PTNA) **over and above the maternal arm (Tdap-IPV vs TT)**? The maternal arm is the base block, and
we report the honest cross-validated gain from adding responder status on top of it. Because the arm
largely *creates* the High/Low split (Part A), most of the responder signal is expected to be shared with
the arm; what matters is the part that is *unique* to responder status. All predictors are
maternal-at-delivery (forward).

```{r excess-of-arm, results='asis'}
excess_tbl <- function(d) {
  base <- c("arm", "age_z")
  models <- list(
    "Arm only"          = c("arm"),
    "Arm + age (base)"  = base,
    "+ responder class" = c(base, "responder"),
    "+ responder score" = c(base, "score_z"))
  rows <- lapply(names(models), function(nm) {
    tr <- models[[nm]]
    dd <- d[stats::complete.cases(d[, intersect(tr, names(d)), drop = FALSE]), , drop = FALSE]
    f  <- reformulate(tr, "y"); m <- lm(f, dd)
    data.frame(Model = nm, n = nrow(dd),
               R2    = round(summary(m)$r.squared, 3),
               adjR2 = round(summary(m)$adj.r.squared, 3),
               CV_R2 = round(cv_r2(f, dd), 3))
  })
  tab <- do.call(rbind, rows)
  base_cv <- tab$CV_R2[tab$Model == "Arm + age (base)"]
  tab$dCV_vs_arm <- round(tab$CV_R2 - base_cv, 3)
  tab$dCV_vs_arm[tab$Model %in% c("Arm only", "Arm + age (base)")] <- NA
  tab
}
for (k in names(OUTCOMES)) {
  d <- DFS[[k]]; if (is.null(d)) next
  cat(sprintf("\n\n### %s (%s); n = %d\n\n", k, OUTCOMES[[k]]$label, nrow(d)))
  print(kable(excess_tbl(d), row.names = FALSE,
        caption = sprintf("Incremental prediction of infant %s by maternal responder status beyond the maternal arm. dCV_vs_arm is the cross-validated R\u00b2 gain over the arm + age base; a positive, stable value is responder information the vaccine label does not already carry.", k)))
}
```

## How much of the {arm + responder} signal is unique to the responder?

```{r excess-shapley, results='asis'}
ex <- lapply(names(OUTCOMES), function(k) {
  d <- DFS[[k]]; if (is.null(d)) return(NULL)
  dd <- d[stats::complete.cases(d[, c("arm", "age_z", "score_z")]), ]
  sh <- shapley("y", "age_z", c("arm", "score_z"), dd)          # order-independent split, beyond age
  r_arm  <- r2_of("arm", "age_z", "y", dd)               - r2_of(character(0), "age_z", "y", dd)
  r_scr  <- r2_of("score_z", "age_z", "y", dd)           - r2_of(character(0), "age_z", "y", dd)
  r_both <- r2_of(c("arm", "score_z"), "age_z", "y", dd) - r2_of(character(0), "age_z", "y", dd)
  uniq_arm <- r_both - r_scr; uniq_scr <- r_both - r_arm; common <- r_both - uniq_arm - uniq_scr
  data.frame(Outcome = k,
             `Shapley arm`   = round(unname(sh["arm"]), 3),
             `Shapley score` = round(unname(sh["score_z"]), 3),
             `Unique arm`    = round(uniq_arm, 3),
             `Unique score`  = round(uniq_scr, 3),
             `Shared`        = round(common, 3),
             check.names = FALSE)
})
kable(do.call(rbind, ex), row.names = FALSE,
      caption = "Order-independent (Shapley) and commonality split of the joint {arm, responder-score} explained variance, beyond maternal age. 'Unique score' is the responder signal in excess of the arm; 'Shared' is the variance the arm and the score both carry (expected large, since the arm largely creates the High/Low split).")
```

# 1. Nested models — does the score add out-of-sample beyond the matched antibody?

All predictors are maternal-at-delivery (forward). Base = maternal arm + maternal age. We report in-sample
R^2/adjusted R^2 and honest 10-fold cross-validated R^2 (CV-R^2). The score earns its place only if it
improves **CV-R^2** beyond the matched antibody — in-sample gains alone do not count.

```{r nested, results='asis'}
nested_tbl <- function(spec, d) {
  base <- c("arm", "age_z")
  models <- list(
    "Base (arm + age)"                 = base,
    "+ matched IgG1"                   = c(base, "ab_z"),
    "+ responder score"               = c(base, "score_z"),
    "+ matched IgG1 + score"          = c(base, "ab_z", "score_z"),
    "+ matched IgG1 + breadth(resid)" = c(base, "ab_z", "breadth_resid_z")
  )
  rows <- lapply(names(models), function(nm) {
    tr <- models[[nm]]
    dd <- d[stats::complete.cases(d[, intersect(tr, names(d)), drop = FALSE]), , drop = FALSE]
    f  <- reformulate(tr, "y")
    m  <- lm(f, dd)
    data.frame(Model = nm, n = nrow(dd),
               R2 = round(summary(m)$r.squared, 3),
               adjR2 = round(summary(m)$adj.r.squared, 3),
               CV_R2 = round(cv_r2(f, dd), 3))
  })
  do.call(rbind, rows)
}
for (k in names(OUTCOMES)) {
  d <- DFS[[k]]; if (is.null(d)) next
  cat(sprintf("\n\n### %s (%s); n = %d\n\n", k, OUTCOMES[[k]]$label, nrow(d)))
  print(kable(nested_tbl(OUTCOMES[[k]], d), row.names = FALSE,
        caption = sprintf("Nested forward models for infant %s. The decisive rows are '+ matched IgG1' vs '+ matched IgG1 + score' (does the score add beyond the antibody already used?) and '+ matched IgG1 + breadth(resid)' (does the orthogonal breadth add?). Judge on CV-R\u00b2.", k)))
}
```

# 2. Equal-chance variance attribution (Shapley)

Entry order decides nothing here: each predictor's share of the explained variance (beyond arm + age) is
averaged over all entry orders. We do this for the matched antibody vs the score (how much is *shared* vs
*unique to the score*), and for the matched antibody vs the orthogonal breadth.

```{r shapley, results='asis'}
boot_shapley <- function(resp, base, cands, data, B = params$boot_B, seed = params$seed) {
  obs <- shapley(resp, base, cands, data)
  set.seed(seed)
  M <- matrix(NA_real_, B, length(cands), dimnames = list(NULL, cands))
  for (b in seq_len(B)) {
    db <- data[sample(nrow(data), replace = TRUE), , drop = FALSE]
    M[b, ] <- tryCatch(shapley(resp, base, cands, db), error = function(e) rep(NA, length(cands)))
  }
  ci <- apply(M, 2, function(v) round(quantile(v, c(.025, .975), na.rm = TRUE), 3))
  data.frame(predictor = cands, shapley_R2 = round(obs, 3),
             ci_low = ci[1, ], ci_high = ci[2, ], row.names = NULL)
}
for (k in names(OUTCOMES)) {
  d <- DFS[[k]]; if (is.null(d)) next
  base <- c("arm", "age_z")
  cat(sprintf("\n\n### %s\n\n", k))
  s1 <- boot_shapley("y", base, c("ab_z", "score_z"), d[stats::complete.cases(d[, c("ab_z","score_z")]), ])
  print(kable(s1, row.names = FALSE,
        caption = sprintf("Shapley shares (matched antibody vs responder score) for %s. A score share near zero with the antibody share carrying the increment means the score is redundant; a non-trivial score share with a CI excluding ~0 means it carries information the matched antibody does not.", k)))
  db <- d[stats::complete.cases(d[, c("ab_z","breadth_resid_z")]), ]
  if (nrow(db) > 30) {
    s2 <- boot_shapley("y", base, c("ab_z", "breadth_resid_z"), db)
    print(kable(s2, row.names = FALSE,
          caption = sprintf("Shapley shares (matched antibody vs orthogonal maternal breadth) for %s. The breadth term is the part of the coordinated maternal panel that is NOT the matched antigen; a non-trivial, CI-positive share here is the result that would license including the breadth construct in the model.", k)))
  }
}
```

# 3. Collinearity — how much do the score and the matched antibody overlap?

```{r collinearity, results='asis'}
co <- lapply(names(OUTCOMES), function(k) {
  d <- DFS[[k]]; if (is.null(d)) return(NULL)
  r <- cor(d$ab_z, d$score_z, use = "complete.obs")
  vif <- 1 / (1 - r^2)   # 2-predictor VIF for {matched antibody, score}
  rb <- suppressWarnings(cor(d$ab_z, d$breadth_resid_z, use = "complete.obs"))
  data.frame(Outcome = k, `cor(antibody, score)` = round(r, 3), `VIF` = round(vif, 2),
             `cor(antibody, breadth_resid)` = round(rb, 3), check.names = FALSE)
})
kable(do.call(rbind, co), row.names = FALSE,
      caption = "Overlap diagnostics. A high antibody-score correlation (and VIF) is exactly why entry order would mislead and why the Shapley split is needed. The orthogonal breadth term should correlate ~0 with the matched antibody by construction.")
```

# 4. Within-arm de-confounding (Tdap-IPV mothers only)

If the score predicts infant function only because it labels the vaccine arm, it should add nothing
*within* an arm. Refit the decisive increments among Tdap-IPV mothers (where responder status is not the
arm), dropping the arm term.

```{r within-arm, results='asis'}
within_tbl <- function(spec, d) {
  d <- d[d$arm == "TdaP", , drop = FALSE]
  if (nrow(d) < 40) return(data.frame(note = sprintf("Too few Tdap-IPV mothers (n=%d) for a stable within-arm fit.", nrow(d))))
  base <- c("age_z")
  out <- function(tr, nm) { dd <- d[stats::complete.cases(d[, intersect(tr, names(d))]), ]; f <- reformulate(tr, "y")
    data.frame(Model = nm, n = nrow(dd), CV_R2 = round(cv_r2(f, dd), 3)) }
  rbind(out(c(base, "ab_z"), "+ matched IgG1"),
        out(c(base, "ab_z", "score_z"), "+ matched IgG1 + score"),
        out(c(base, "ab_z", "breadth_resid_z"), "+ matched IgG1 + breadth(resid)"))
}
for (k in names(OUTCOMES)) {
  d <- DFS[[k]]; if (is.null(d)) next
  cat(sprintf("\n\n### %s — within Tdap-IPV\n\n", k))
  print(kable(within_tbl(OUTCOMES[[k]], d), row.names = FALSE,
        caption = sprintf("Within-arm CV-R\u00b2 for %s. A score/breadth CV gain that survives here is not the vaccine label.", k)))
}
```

# 5. Decision summary and how to read it

```{r decision, results='asis'}
summ <- lapply(names(OUTCOMES), function(k) {
  d <- DFS[[k]]; if (is.null(d)) return(NULL)
  base <- c("arm", "age_z")
  f_ab   <- reformulate(c(base, "ab_z"), "y")
  f_both <- reformulate(c(base, "ab_z", "score_z"), "y")
  f_brd  <- reformulate(c(base, "ab_z", "breadth_resid_z"), "y")
  d2  <- d[stats::complete.cases(d[, c("ab_z","score_z")]), ]
  db  <- d[stats::complete.cases(d[, c("ab_z","breadth_resid_z")]), ]
  sh  <- shapley("y", base, c("ab_z", "score_z"), d2)
  shb <- if (nrow(db) > 30) shapley("y", base, c("ab_z", "breadth_resid_z"), db) else c(breadth_resid_z = NA)
  data.frame(
    Outcome = k,
    `CV gain: score beyond antibody`  = round(cv_r2(f_both, d2) - cv_r2(f_ab, d2), 3),
    `CV gain: breadth beyond antibody`= if (nrow(db) > 30) round(cv_r2(f_brd, db) - cv_r2(f_ab, db), 3) else NA,
    `Shapley share: score`            = round(unname(sh["score_z"]), 3),
    `Shapley share: breadth`          = round(unname(shb["breadth_resid_z"]), 3),
    check.names = FALSE)
})
kable(do.call(rbind, summ), row.names = FALSE,
      caption = "Decision table. Positive CV gain + a Shapley share whose bootstrap CI (section 2) excludes ~0 + survival within-arm (section 4) = the construct earns a place in the model. A null here empirically confirms the current caveat (the score is a re-packaging of antibodies already modelled), which is itself a stronger statement than 'not tested'.")
```

**Reading the result.**

* *Score, beyond the matched antibody.* Expected to be small/redundant — the score contains the matched
  antibody. The honest report is the Shapley split: how much of the explained variance is shared vs unique
  to the score.
* *Orthogonal breadth.* This is the term that can change the conclusion. If it adds CV-R^2 beyond the
  matched antibody, holds a CI-positive Shapley share, and survives within-arm, then the coordinated
  maternal breadth phenotype has **incremental forward predictive value for infant function**, and the
  responder construct moves from a side-analysis to a modelled predictor.
* *Causality remains bounded.* Even a positive breadth result is associational and partly mediated by the
  antibody the mother transfers; the within-arm + forward + orthogonalised design makes it a credible
  *independent maternal contribution*, not a causal claim. The "Phase 08 does not test infant function"
  caveat is retired either way, because this analysis tests exactly that.

```{r session}
sessionInfo()
```
