---
title: "Subclass polarisation and infant SBA: a compositional (balance) re-analysis with a complement-mediation check"
subtitle: "Rationale, methods, and staged results — GaPS cohort, SBA at the 8-week baseline"
date: "`r format(Sys.Date())`"
params:
  endpoint: "SBA"
  analysis_antigens: "PT,FHA,PRN,TT,DT"   # PT/FHA/PRN = pertussis; TT/DT = vaccine-control antigens
  med_antigen: "PRN"          # antigen whose cord-blood balance is the mediation exposure
  med_x_visit: "CordBlood"    # upstream stage used as the mediation exposure
output:
  html_document: { toc: true, toc_depth: 3, toc_float: true, number_sections: true, df_print: kable }
  word_document: { toc: true, toc_depth: 3 }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({ library(dplyr); library(tidyr); library(ggplot2) })

`%||%` <- function(a, b) if (is.null(a) || length(a) == 0L) b else a

# ---- locate + source the pipeline (standalone bootstrap) -------------------
.stub <- function(rel, optional = FALSE) {
  cand <- character(0)
  if (requireNamespace("here", quietly = TRUE))
    cand <- c(cand, here::here(rel), here::here("analysis", rel))
  rmd_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) NA_character_)
  if (!is.na(rmd_dir)) cand <- c(cand, file.path(rmd_dir, rel), file.path(dirname(rmd_dir), rel))
  cand <- unique(c(cand, file.path(getwd(), rel), rel))
  hit  <- cand[file.exists(cand)]
  if (length(hit)) return(normalizePath(hit[1], winslash = "/", mustWork = FALSE))
  if (optional) return(NULL)
  stop("Cannot locate '", rel, "'. Run from the project root.", call. = FALSE)
}
if (!exists("shared_utils")) source(.stub("R/shared_utils.R"))
source(shared_utils("config/endpoints.R"),            encoding = "UTF-8")
source(shared_utils("R/C1_serology_helpers.R"),          encoding = "UTF-8")
source(shared_utils("R/C8_subclass_balance_helpers.R"),  encoding = "UTF-8")

cfg      <- get_endpoint_config(params$endpoint)
data_raw <- load_serology_data(here::here("./data/d_set.RData"), object_name = "ebaa_extra2")
verify_predictor_casing(data_raw)
nt       <- new_ntracker()
roles    <- chain_visit_roles(cfg)

# ---- antigen set for this analysis (pertussis core + TT/DT controls) -------
ants <- trimws(strsplit(params$analysis_antigens, ",")[[1]])
pert_ants <- intersect(ants, cfg$antigens)                    # PT, FHA, PRN
ctrl_ants <- setdiff(ants, cfg$antigens)                      # TT, DT (vaccine-control)

# ---- discover the complement (ADCD) measure available in this dataset ------
adcd_info      <- derive_adcd_tag(data_raw, ants, pattern = "ADCD")
adcd_available <- !is.null(adcd_info)
adcd_tag       <- if (adcd_available) adcd_info$tag else NA_character_
adcd_feats     <- if (adcd_available) adcd_info$feats else character(0)

# ---- wide chain frame: subclasses + outcome + ADCD, plus balances & deltas -
chain_feats <- c(cfg$outcome_feature, PREDICTOR_FEATURES, adcd_feats)
data_wide_chain <- data_raw %>%
  dplyr::filter(feature %in% chain_feats,
                visit_name %in% cfg$chain_visits,
                arm_name %in% cfg$maternal_arms) %>%
  dplyr::select(subject_accession, arm_name, infant_arm, feature, visit_name, log_assay_value) %>%
  dplyr::distinct(subject_accession, visit_name, feature, .keep_all = TRUE) %>%
  tidyr::pivot_wider(id_cols = c(subject_accession, arm_name, infant_arm),
                     names_from = c(feature, visit_name),
                     values_from = log_assay_value, names_sep = "_") %>%
  dplyr::mutate(arm_name = factor(arm_name, levels = cfg$maternal_arms)) %>%
  add_chain_derived(cfg) %>%                                   # gives <ANT>_ratio_<v> (pertussis core)
  add_subclass_balances(ants, cfg$chain_visits) %>%            # gives <ANT>_bal_<v> for all antigens
  add_balance_deltas(ants)                                     # gives <ANT>_baldelta_<stage>
if (adcd_available)                                            # gives <ANT>_ADCDdelta_<stage>
  data_wide_chain <- add_stage_deltas(data_wide_chain, ants, tag = adcd_tag)

oc      <- cfg$outcome_col       # WHOLE_SBA_InfMon2
prod_v  <- roles$matbirth %||% "MatBirth"
walk_nodes <- c(Maternal = roles$matbirth %||% "MatBirth",
                Cord     = roles$cord     %||% "CordBlood",
                Baseline = roles$infant   %||% "InfMon2")
STAGE_LABEL <- c(production = "Maternal production (level at delivery)",
                 transfer   = "Placental transfer (\u0394 delivery\u2192cord)",
                 decay      = "Post-natal decay (\u0394 cord\u21928 weeks)")

# small renderers reused for the balance (bal) and complement (ADCD) measures
render_within_stage <- function(tag, measure) {
  u <- fit_stage_univariate(data_wide_chain, ants, oc, tag = tag, arm = "arm_name",
                            production_visit = prod_v, nt = nt)
  if (is.null(u) || !nrow(u)) { cat("> No", measure, "stage quantities available.\n\n"); return(invisible(NULL)) }
  u <- u[order(u$stage, match(u$antigen, ants)), ]
  d <- transform(u, Stage = STAGE_LABEL[as.character(stage)],
                 `95% CI` = sprintf("%.3f to %.3f", lo, hi),
                 beta = round(beta, 3), p = signif(p, 3), sig = pvalue_to_label(p))
  d <- d[, c("antigen", "Stage", "n", "beta", "95% CI", "p", "sig")]; names(d)[1] <- "Antigen"
  print(knitr::kable(d, row.names = FALSE, align = c("l","l","r","r","r","r","c"),
    caption = paste0("Within-stage ", measure, " vs ", cfg$outcome_label,
                     " at 8 weeks (SBA ~ stage quantity + maternal group).")))
  cat("\n"); invisible(u)
}
render_forest <- function(u, measure) {
  if (is.null(u) || !nrow(u)) { cat("\n"); return(invisible(NULL)) }
  u$sig <- (u$lo > 0) | (u$hi < 0)
  u$antigen <- factor(u$antigen, levels = rev(ants))
  u$Stage <- factor(STAGE_LABEL[as.character(u$stage)], levels = STAGE_LABEL)
  print(
    ggplot(u, aes(beta, antigen)) +
      geom_vline(xintercept = 0, linetype = "dashed", colour = "grey55", linewidth = 0.7) +
      geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.22, colour = "grey25") +
      geom_point(aes(fill = sig), shape = 21, size = 3, colour = "grey15") +
      scale_fill_manual(values = c(`TRUE` = "grey15", `FALSE` = "white"), guide = "none") +
      facet_wrap(~ Stage, ncol = 1, strip.position = "right") +
      labs(x = paste0("Adjusted \u03b2 for log10 ", measure, " at 8 weeks (95% CI)"), y = NULL) +
      theme_bw(base_size = 11))
  invisible(NULL)
}
render_sequential <- function(tag, measure) {
  for (a in ants) {
    s <- fit_sequential_stages(data_wide_chain, a, oc, tag = tag, arm = "arm_name",
                               production_visit = prod_v, nt = nt)
    cat("\n**", a, "**\n\n", sep = "")
    if (is.null(s$steps)) { cat("> ", measure, " columns unavailable for ", a, ".\n\n", sep = ""); next }
    st <- transform(s$steps, Step = STAGE_LABEL[step], R2 = round(R2, 3),
                    adj_R2 = round(adj_R2, 3), `ΔR²` = round(delta_R2, 3))[,
                    c("Step","n","R2","adj_R2","\u0394R\u00b2")]
    print(knitr::kable(st, row.names = FALSE, align = c("l","r","r","r","r"),
      caption = paste0(a, " (", measure, "): incremental fit as stages are added (common n=", s$n, ").")))
    cat("\n")
    if (!is.null(s$full_model)) {
      fm <- s$full_model
      fm$Stage <- ifelse(grepl("delta_transfer", fm$term), STAGE_LABEL["transfer"],
                  ifelse(grepl("delta_decay", fm$term), STAGE_LABEL["decay"], STAGE_LABEL["production"]))
      fmd <- data.frame(Stage = fm$Stage, beta = round(fm$beta, 3),
        `95% CI` = sprintf("%.3f to %.3f", fm$lo, fm$hi),
        p = signif(fm$p, 3), sig = pvalue_to_label(fm$p), check.names = FALSE)
      print(knitr::kable(fmd, row.names = FALSE, align = c("l","r","r","r","c"),
        caption = paste0(a, " (", measure, "): all three stage terms together + arm.")))
      cat("\n")
    }
  }
}
```

# Background: what the pro-/tolerogenic ratio assumes

Our earlier analyses summarised each antigen's IgG-subclass response as a single
**pro-inflammatory-to-tolerogenic ratio**, log10[(IgG1+IgG3)/(IgG2+IgG4)], and
related it to serum bactericidal activity (SBA) at 8 weeks. That ratio is a
reasonable first pass, but it bundles together several biological and statistical
assumptions that are worth making explicit before we extend the work.

The **effector split itself is well grounded.** Across the literature the
functional hierarchy runs roughly IgG4 < IgG2 < IgG1 < IgG3 for complement
fixation and activating-Fc\u03b3-receptor engagement: IgG1 and IgG3 are the strong
mediators of complement- and Fc\u03b3R-dependent functions, IgG2 is a weak mediator
tied to polysaccharide responses, and IgG4 has minimal effector capacity and can
actively block IgG1/IgG3 complement activation while preferentially engaging the
inhibitory receptor Fc\u03b3RIIb [1\u20134]. This matters here because **our readout is
itself complement-mediated** \u2014 SBA is complement-dependent bacterial killing \u2014 so
placing the complement-fixing subclasses in the numerator and the
non-complement-fixing IgG4 in the denominator is mechanistically defensible, and
almost certainly explains why the ratio predicts SBA at all.

But the dichotomy is crude in ways that matter for this dataset:

* **IgG2 is not "tolerogenic".** IgG4 is the genuinely immunoregulatory,
  chronic-antigen "blocking" subclass; IgG2 is simply the low-effector
  anti-polysaccharide subclass. Summing them conflates "does little" with
  "actively suppresses" [1,3].
* **In pertussis the contrast is really IgG1 vs IgG4.** For pertussis antigens
  IgG1 dominates and IgG3 contributes little, while IgG4 rises specifically after
  repeated acellular vaccination and reflects Th2 skewing; whole-cell priming
  instead favours IgG1 and a Th1/Th17 profile with better opsonisation and killing
  [5\u20137]. Because IgG4 does not fix complement, a high IgG4 share is expected to
  depress a complement-dependent readout such as SBA, and low IgG1/IgG4 ratios have
  been linked to reduced protection [7]. So our "polarisation" axis is, in this
  cohort, largely an aP-vs-wP / Th2-vs-Th1 axis \u2014 which is meaningful, but should be
  named as such.
* **Cord blood adds a placental-transfer confound.** The subclasses cross the
  placenta with different efficiencies (IgG1 most efficiently, IgG2 least; the exact
  hierarchy is not settled and transfer falls as maternal titre rises) [8,9]. A cord
  ratio therefore mixes maternal production with differential transfer, and will not
  equal the maternal ratio in the same woman.
* **The arithmetic hides the potency we care about.** The original ratio sums
  *linear* concentrations, so it is dominated by the most abundant subclass; IgG3 \u2014
  the most potent per molecule \u2014 barely moves it.
* **Effector class is not fixed by subclass.** Fc glycosylation (galactosylation,
  sialylation, fucosylation) strongly re-tunes the activating/inhibitory balance of
  any given subclass [4]. A subclass-only index cannot see this.

# Why a balance, and why a complement-mediation check

Two consequences follow, and they define this analysis.

**(2) Use a compositional balance rather than a linear-sum ratio.** The four
subclasses are parts of a whole and carry *relative* information, which is exactly
the setting compositional data analysis (CoDA) was built for; naive ratios of parts
have produced spurious correlations since Pearson (1897), and the principled fix is a
log-ratio *balance* [10\u201313]. For the {IgG1,IgG3} vs {IgG2,IgG4} partition the
isometric-log-ratio balance is

$$ b \;=\; \tfrac{1}{2}\big(\log \mathrm{IgG1} + \log \mathrm{IgG3}\big) \;-\; \tfrac{1}{2}\big(\log \mathrm{IgG2} + \log \mathrm{IgG4}\big), $$

i.e. a **geometric-mean contrast** that weights each subclass equally (so IgG3
counts as much as IgG1) and is scale-invariant. This is the same move the microbiome
field made away from the crude Firmicutes/Bacteroidetes ratio toward CoDA balances
[11\u201313].

**Scale note (crucial for our arbitrary-unit assays).** Each subclass is measured in
its own arbitrary unit, so absolute balance values are not concentration-interpretable.
However, a per-subclass unit change multiplies each subclass by a constant, i.e. adds a
constant to its log \u2014 shifting $b$ by a fixed offset that is identical for every
subject. In a regression SBA ~ $b$ (+ arm) that offset is absorbed by the intercept:
the **slope, its confidence interval and p value are unchanged**. Stage **deltas**
($b_{\text{to}}-b_{\text{from}}$) cancel the offset entirely. So associations, ordering,
$\Delta R^2$ and proportion-mediated are all interpretable on ordinal data; only absolute
balance levels and individual mediation path coefficients are not.

**(6) Ask whether the balance acts through measured complement deposition, at every
stage.** Because SBA is complement killing, the sharpest mechanistic test is whether a
polarisation signal predicts SBA *through* the antibody's measured complement-deposition
capacity (antigen-specific ADCD). Now that ADCD is available at the maternal and cord
visits as well as the 8-week baseline, we can **walk complement through the chain** \u2014 at
the maternal, cord and baseline nodes in turn \u2014 fitting balance \u2192 ADCD \u2192 SBA at each,
so complement is analysed with the same stage machinery as the subclasses rather than at
a single timepoint.

# Methods

We work on the recoded, mixed-case chain frame produced by `load_serology_data()`
and compute, for each antigen and visit, the balance $b$ above
(`add_subclass_balances()`). Antigens comprise the three pertussis targets
(PT, FHA, PRN) plus the two shared vaccine-control toxoids, **tetanus toxoid (TT)** and
**diphtheria toxoid (DT)**. TT and DT serve as a specificity check: because the outcome
(SBA) is bactericidal killing of *B pertussis*, a pertussis-specific mechanism should
show the balance\u2192SBA and balance\u2192ADCD\u2192SBA signal for PT/FHA/PRN but *not* for TT/DT;
association of the toxoid balances with SBA would instead point to a nonspecific,
global IgG-polarisation effect. We then run **two measures in parallel through the same
stage machinery** \u2014 the subclass balance and antibody-dependent complement deposition
(ADCD) \u2014 decomposing the maternal\u2192cord\u2192infant chain into three **stages**: maternal
production (level at delivery), placental transfer ($\Delta$ delivery\u2192cord) and
post-natal decay ($\Delta$ cord\u21928 weeks) (`add_stage_deltas()`, generic over the
measure). All models take log10 SBA at 8 weeks (`r oc`) as the outcome and adjust for
maternal vaccine group (Tdap-IPV vs tetanus toxoid, Tdap-IPV reference).

Four questions are addressed: (i) *within each stage*, does that stage's quantity
predict SBA, for balance and for ADCD (one antigen at a time, arm-adjusted); (ii)
*sequentially*, do transfer and decay add predictive value once maternal production is
in the model (nested models on one common complete-case sample, reporting $\Delta R^2$);
(iii) a **complement walk** tracing balance \u2192 ADCD \u2192 SBA at the maternal, cord and
baseline nodes (a-path ADCD~balance, b-path SBA~ADCD, and the mediated effect; bootstrap
product-of-coefficients, and the `mediation` package where available); and (iv) a
robustness check of the geometric balance against the original linear-sum ratio.
Analyses were run in R; participant accounting for every model is tabulated at the end.

---

# Result A — within each stage, what predicts SBA?

Each row is a single arm-adjusted model, SBA ~ (that stage's quantity) + maternal group,
run in parallel for the subclass **balance** and for **complement deposition (ADCD)**.

## Subclass balance

```{r stage-uni-bal, results='asis'}
uni_bal <- render_within_stage("bal", "subclass balance")
```

```{r stage-forest-bal, fig.width=8, fig.height=4.8, fig.cap="Within-stage arm-adjusted association of each antigen's subclass balance with SBA at 8 weeks. Filled points: 95% CI excludes zero."}
render_forest(uni_bal, cfg$outcome_label)
```

## Complement deposition (ADCD)

```{r stage-uni-adcd, results='asis'}
uni_adcd <- NULL
if (!adcd_available) cat("> No ADCD measure was found in this dataset; complement stages are skipped.\n\n") else
  uni_adcd <- render_within_stage(adcd_tag, "complement (ADCD)")
```

```{r stage-forest-adcd, fig.width=8, fig.height=4.8, fig.cap="Within-stage arm-adjusted association of each antigen's ADCD with SBA at 8 weeks. Filled points: 95% CI excludes zero."}
if (adcd_available) render_forest(uni_adcd, paste0(cfg$outcome_label, " (ADCD predictor)"))
```

# Result B — do transfer and decay survive once maternal production is in the model?

For each antigen we fit three nested models on one common complete-case sample:
production alone (+ arm), then adding transfer, then adding decay, reporting the term
added at each step and the gain in variance explained ($\Delta R^2$). This is run for the
balance and, in parallel, for complement.

## Subclass balance

```{r seq-bal, results='asis'}
render_sequential("bal", "subclass balance")
```

## Complement deposition (ADCD)

```{r seq-adcd, results='asis'}
if (!adcd_available) cat("> No ADCD measure; complement sequential models skipped.\n\n") else
  render_sequential(adcd_tag, "complement (ADCD)")
```

# Result C — walking complement through the stages (maternal → cord → baseline)

At each node we fit the two mechanistic paths and the mediated effect: the **a-path**
(ADCD ~ balance + arm) asks whether polarisation drives complement deposition *at that
node*; the **b-path** (SBA ~ ADCD + arm) asks whether complement *at that node* predicts
8-week SBA; and the **indirect effect** is the balance → ADCD → SBA path, with a
dimensionless proportion mediated.

```{r walk, results='asis'}
if (!adcd_available) {
  cat("> No ADCD measure found; the complement walk is skipped.\n\n")
} else {
  walk <- stage_complement_walk(data_wide_chain, ants, oc, nodes = walk_nodes,
                                bal_tag = "bal", adcd_tag = adcd_tag,
                                arm = "arm_name", R = 2000, nt = nt)
  if (is.null(walk) || !nrow(walk)) {
    cat("> No node had both a balance and an ADCD column; walk skipped.\n\n")
  } else {
    wd <- data.frame(
      Antigen = walk$antigen,
      Node    = as.character(walk$node),
      Visit   = walk$visit,
      n       = walk$n,
      `a: ADCD~balance, β (p)` = sprintf("%.3f (%.2g)", walk$a_beta, walk$a_p),
      `b: SBA~ADCD, β (p)`     = sprintf("%.3f (%.2g)", walk$b_beta, walk$b_p),
      `indirect (95% CI)`        = sprintf("%.3f (%.3f, %.3f)", walk$indirect, walk$indirect_lo, walk$indirect_hi),
      `prop. mediated`           = ifelse(is.na(walk$prop_med), "\u2014", sprintf("%.0f%%", 100 * walk$prop_med)),
      check.names = FALSE, stringsAsFactors = FALSE)
    print(knitr::kable(wd, row.names = FALSE,
      align = c("l","l","l","r","r","r","r","r"),
      caption = paste0("Complement walk (", adcd_tag, "): balance \u2192 ADCD \u2192 SBA at 8 weeks, ",
        "by antigen and node. a-path = ADCD ~ balance + arm; b-path = SBA ~ ADCD + arm; ",
        "indirect = balance\u2192ADCD\u2192SBA (bootstrap). Proportion mediated is dimensionless.")))
    cat("\n\n*Path coefficients are on arbitrary units and are not comparable in magnitude ",
        "across nodes; signs, p values and proportion mediated are the interpretable quantities.*\n\n")
  }
}
```

# Result D — balance vs the original linear-sum ratio (robustness)

For the driver antigen at cord blood, does the CoDA balance recover the same signal
as the original linear-sum ratio?

```{r compare, results='asis'}
cmp <- compare_balance_vs_ratio(data_wide_chain, params$med_antigen,
                                params$med_x_visit, oc, arm = "arm_name")
if (is.null(cmp)) {
  cat("> Neither representation was available for the comparison.\n\n")
} else {
  cmpd <- transform(cmp, beta = round(beta, 3),
    `95% CI` = sprintf("%.3f to %.3f", lo, hi), p = signif(p, 3), R2 = round(R2, 3))[,
    c("representation","n","beta","95% CI","p","R2")]
  names(cmpd)[1] <- "Representation"
  print(knitr::kable(cmpd, row.names = FALSE, align = c("l","r","r","r","r","r"),
    caption = paste0(params$med_antigen, " at ", params$med_x_visit,
                     ": geometric balance vs linear-sum ratio as predictors of SBA at 8 weeks.")))
  cat("\n")
}
```

# Interpretation

Read together: Result A localises which *stage* of the maternal\u2192cord\u2192infant chain
carries the signal for SBA, separately for subclass polarisation and for complement
deposition; Result B asks whether transfer and decay add anything once maternal
production is accounted for (if their coefficients collapse, the 8-week profile is
essentially inherited maternal production; if they survive, transfer selectivity and/or
differential decay independently shape it), again for both measures; Result C **walks
complement through the chain** \u2014 the a-path shows where polarisation drives complement
deposition, the b-path shows which node's complement predicts eventual SBA, and the
mediated effect shows where the balance\u2192complement\u2192killing mechanism is operative
(given that SBA is complement killing, we expect the b-path and mediation to be strongest
at the cord and baseline nodes closest to the readout); and Result D confirms the
compositional balance is not an artefact of the original ratio's arithmetic. Throughout,
only associations, $\Delta R^2$ and proportion-mediated are interpreted, consistent with
the ordinal nature of the arbitrary-unit assays.

The tetanus (TT) and diphtheria (DT) toxoids are read as specificity controls: their
balances and ADCD are processed identically to the pertussis antigens, so if the
pertussis PT/FHA/PRN signals reflect a genuinely pertussis-specific,
complement-dependent mechanism we expect the TT/DT rows to be null (weak a-path, weak
b-path, little mediation), whereas broadly positive TT/DT associations would flag a
nonspecific global-polarisation or shared-arm effect that any interpretation of the
pertussis findings would need to accommodate.

# Participant accounting

```{r naccount, results='asis'}
nt_render(nt, caption = "Participant accounting for the balance, complement, sequential, and walk models.")
```

# References

1. Vidarsson G, Dekkers G, Rispens T. IgG subclasses and allotypes: from structure to effector functions. *Front Immunol* 2014;5:520.
2. IgG subclass FcγR and complement effector functions (technical review). Effector hierarchy IgG4 < IgG2 < IgG1 < IgG3.
3. Lilienthal G-M, et al. Potential of murine IgG1 and human IgG4 to inhibit the classical complement and Fcγ receptor activation pathways. *Front Immunol* 2018;9:958 (PMC5954034).
4. Harnessing IgG Fc glycosylation for clinical benefit (review), 2023 (PMC9870045).
5. Hendrikx LH, et al. Different IgG-subclass distributions after whole-cell and acellular pertussis infant primary vaccinations. *Vaccine* 2011.
6. What is wrong with pertussis vaccine immunity? Why immunological memory to pertussis is failing (review), 2017 (PMID 28289059).
7. Whole-cell or acellular pertussis vaccination in infancy determines IgG subclass profiles to DTaP booster vaccination. *Vaccine* 2017.
8. Clements T, et al. Update on transplacental transfer of IgG subclasses: impact of maternal and fetal factors. *Front Immunol* 2020;11:1920 (PMC7516031).
9. On the perplexingly low rate of transport of IgG2 across the human placenta. *PLoS One* 2014;9:e108319 (PMID 25251461).
10. Aitchison J. The statistical analysis of compositional data. *J R Stat Soc Ser B* 1982;44:139\u2013177.
11. Pearson K. On a form of spurious correlation which may arise when indices are used in the measurement of organs. *Proc R Soc Lond* 1897;60:489\u2013498.
12. Rivera-Pinto J, et al. Balances: a new perspective for microbiome analysis. *mSystems* 2018;3:e00053-18.
13. Calle ML, Susin A. coda4microbiome: compositional data analysis for microbiome studies. *BMC Bioinformatics* 2023;24:82.
14. Chung AW, et al. Dissecting polyclonal vaccine-induced humoral immunity against HIV using systems serology. *Cell* 2015;163:988\u2013998.

```{r sessioninfo}
sessionInfo()
```
