---
title: "Figure 5 (revised) — Where predictability enters, six-block ladder"
subtitle: |
  Left: cross-validated R² block ladder for the three whole-organism infant
  outcomes, with the design/clinical foundation split into design (vaccination),
  pre-vaccination clinical, and birth clinical. Right: principal maternal-
  production antigen (LMG) of each outcome.
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    theme: readable
params:
  # antigens that feed the whole-organism readouts
  anti_igg: !r c("PRN_IgG", "PT_IgG", "FHA_IgG", "FIM_IgG")
  # outcomes (assay -> display label). All three are whole-organism functional readouts.
  outcomes: !r c(WHOLE_WT_IgG = "Whole-cell binding (IgG)", WHOLE_SBA = "SBA", WHOLE_PTNA = "PTNA")
  # visits used for the maternal-production / cord-transfer / concurrent frames
  visit_outcome:   "InfMon2"
  visit_maternal:  "MatBirth"
  visit_cord:      "CordBlood"
  visit_concurrent: "InfMon2"
  # NEW: the design/clinical foundation is split into three ordered sub-blocks.
  design_vars:      !r c("maternal_arm")
  clin_prevax_vars: !r c("maternal_age", "maternal_bmi", "maternal_Hb", "parity", "gestational_age_vaccination")
  clin_birth_vars:  !r c("gestational_age_birth", "delivery_mode", "birth_weight", "infant_sex", "vaccine_birth_interval_days")
  cv_folds:         10
  seed:             1
  fig_out_dir:      "figures"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
suppressPackageStartupMessages({ library(dplyr); library(tidyr); library(ggplot2); library(knitr) })
```

```{r bootstrap, include=FALSE}
library(here)
if (!exists("shared_utils", mode = "function")) {
  .ff_cand <- c("R/shared_utils.R", "../R/shared_utils.R", "analysis/R/shared_utils.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/shared_utils.R"), here::here("analysis/R/shared_utils.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/shared_utils.R"),
                  file.path(dirname(.ff_rmd), "R/shared_utils.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/shared_utils.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE); rm(.ff_cand, .ff_rmd, .ff_hit)
}
source(shared_utils("config/endpoints.R"),           local = FALSE, encoding = "UTF-8")
source(shared_utils("config/endpoints_additions.R"), local = FALSE, encoding = "UTF-8")
source(shared_utils("R/C1_C1_serology_helpers.R"),         local = FALSE, encoding = "UTF-8")

data_raw <- load_serology_data(shared_utils("data/c_set.RData"))
```

# Data assembly

The modelling frame is identical to Results C–F, except the design/clinical base
is now assembled from **three** named variable groups so the ladder can attribute
predictability to design (vaccination), pre-vaccination clinical, and birth
clinical separately. Antibody blocks (maternal production / cord transfer /
concurrent) are unchanged.

```{r assemble}
ANTI_IGG <- params$anti_igg

pull_wide <- function(d, visit, feats, suffix) {
  d %>%
    dplyr::filter(visit_name == visit, feature %in% feats) %>%
    dplyr::distinct(subject_accession, feature, .keep_all = TRUE) %>%
    dplyr::select(subject_accession, feature, log_assay_value) %>%
    tidyr::pivot_wider(names_from = feature, values_from = log_assay_value) %>%
    dplyr::rename_with(~ paste0(.x, suffix), -subject_accession)
}

mat <- pull_wide(data_raw, params$visit_maternal,   ANTI_IGG, "_mat")
crd <- pull_wide(data_raw, params$visit_cord,       ANTI_IGG, "_cord")
inf <- pull_wide(data_raw, params$visit_concurrent, ANTI_IGG, "_inf")
arm <- dplyr::distinct(data_raw, subject_accession, maternal_arm)

# clinical covariates: pull EVERYTHING we might need for the two clinical blocks
clin <- tryCatch(load_clinical_data(shared_utils("data/clin_assess.RData")),
                 error = function(e) NULL)
want_clin <- unique(c(params$clin_prevax_vars, params$clin_birth_vars))
if (!is.null(clin)) {
  want_clin <- intersect(want_clin, names(clin))
  clind <- dplyr::select(clin, subject_accession, dplyr::all_of(want_clin))
  # light preparation so lm dummy-codes categoricals cleanly
  for (v in intersect(c("infant_sex", "delivery_mode"), names(clind)))
    clind[[v]] <- factor(clind[[v]])
  if ("parity" %in% names(clind))
    clind$parity <- factor(ifelse(clind$parity == 0, "nulliparous", "parous"))
} else { want_clin <- character(0); clind <- NULL }

# resolve the three foundation groups to variables that actually exist
DESIGN      <- intersect(params$design_vars,      c("maternal_arm", names(clind)))
CLIN_PREVAX <- intersect(params$clin_prevax_vars, names(clind))
CLIN_BIRTH  <- intersect(params$clin_birth_vars,  names(clind))
cat(sprintf("Design: %s\nPre-vaccination clinical: %s\nBirth clinical: %s\n",
            paste(DESIGN, collapse = ", "),
            paste(CLIN_PREVAX, collapse = ", "),
            paste(CLIN_BIRTH,  collapse = ", ")))
```

```{r fit-helpers}
# base-R fit utilities (identical contract to Results C–F)
r2 <- function(form, data) tryCatch({
  cc <- data[stats::complete.cases(data[, all.vars(form), drop = FALSE]), , drop = FALSE]
  if (nrow(cc) < 5) return(NA_real_)
  summary(stats::lm(form, cc))$r.squared
}, error = function(e) NA_real_)

cv_r2 <- function(form, data, K = params$cv_folds, seed = params$seed) tryCatch({
  vars <- all.vars(form)
  cc <- data[stats::complete.cases(data[, vars, drop = FALSE]), , drop = FALSE]
  if (nrow(cc) < 2 * K) return(NA_real_)
  set.seed(seed); fold <- sample(rep_len(1:K, nrow(cc)))
  pr <- numeric(nrow(cc))
  for (k in 1:K) {
    fit <- stats::lm(form, cc[fold != k, , drop = FALSE])
    pr[fold == k] <- stats::predict(fit, cc[fold == k, , drop = FALSE])
  }
  yv <- cc[[vars[1]]]; 1 - sum((yv - pr)^2) / sum((yv - mean(yv))^2)
}, error = function(e) NA_real_)

mk <- function(terms) reformulate(if (length(terms)) terms else "1", "y")

# Shapley/LMG within the maternal-production block, over the full foundation base.
shapley_lmg <- function(preds, base, data, ycol = "y") {
  vars <- c(ycol, base, preds)
  cc <- data[stats::complete.cases(data[, vars, drop = FALSE]), , drop = FALSE]
  R  <- function(terms) summary(stats::lm(reformulate(if (length(terms)) terms else "1", ycol), cc))$r.squared
  perms <- function(x) if (length(x) <= 1) list(x) else
    do.call(c, lapply(seq_along(x), function(i) lapply(perms(x[-i]), function(p) c(x[i], p))))
  P <- perms(preds); contrib <- stats::setNames(numeric(length(preds)), preds)
  for (p in P) { prev <- base
    for (v in p) { contrib[v] <- contrib[v] + (R(c(prev, v)) - R(prev)); prev <- c(prev, v) } }
  contrib / length(P)
}
```

```{r ladder-engine}
# block labels, in ladder order (design/clinical foundation split three ways)
BLOCK_LEVELS <- c("design", "+ clinical (pre-vaccination)", "+ clinical (birth)",
                  "+ maternal production", "+ cord (transfer)", "+ concurrent (ceiling)")

build_frame <- function(assay) {
  yw <- pull_wide(data_raw, params$visit_outcome, assay, "")
  names(yw)[names(yw) == assay] <- "y"
  m <- yw %>%
    dplyr::left_join(mat, "subject_accession") %>%
    dplyr::left_join(crd, "subject_accession") %>%
    dplyr::left_join(inf, "subject_accession") %>%
    dplyr::left_join(arm, "subject_accession")
  if (!is.null(clind)) m <- dplyr::left_join(m, clind, "subject_accession")
  dplyr::filter(m, !is.na(y))
}

# one outcome -> the six cumulative CV-R2 rungs + the maternal LMG breakdown
run_outcome <- function(assay, label) {
  m <- build_frame(assay)
  pres <- function(cols) cols[cols %in% names(m)]
  design <- pres(DESIGN); prevax <- pres(CLIN_PREVAX); birth <- pres(CLIN_BIRTH)
  mp <- pres(paste0(ANTI_IGG, "_mat"))
  cp <- pres(paste0(ANTI_IGG, "_cord"))
  ip <- pres(paste0(ANTI_IGG, "_inf"))

  # cumulative term sets for the six rungs
  cum <- list(design,
              c(design, prevax),
              c(design, prevax, birth),
              c(design, prevax, birth, mp),
              c(design, prevax, birth, mp, cp),
              c(design, prevax, birth, mp, cp, ip))
  ladder <- data.frame(
    outcome = label,
    block   = factor(BLOCK_LEVELS, levels = BLOCK_LEVELS),
    n       = vapply(cum, function(t) sum(stats::complete.cases(m[, all.vars(mk(t)), drop = FALSE])), integer(1)),
    cv_r2   = vapply(cum, function(t) cv_r2(mk(t), m), numeric(1)),
    stringsAsFactors = FALSE)
  ladder[["delta"]] <- c(NA, diff(ladder$cv_r2))

  # maternal-production antigen shares (LMG) over the full foundation base
  base <- c(design, prevax, birth)
  lmg  <- if (length(mp) >= 2) shapley_lmg(mp, base, m) else stats::setNames(numeric(0), character(0))
  lmg_df <- if (length(lmg))
    data.frame(outcome = label,
               antigen = sub("_IgG_mat$", "", names(lmg)),
               lmg_r2  = as.numeric(lmg), row.names = NULL) else NULL
  list(ladder = ladder, lmg = lmg_df)
}

res <- Map(run_outcome, names(params$outcomes), unname(params$outcomes))
ladder_df <- dplyr::bind_rows(lapply(res, `[[`, "ladder"))
lmg_df    <- dplyr::bind_rows(lapply(res, `[[`, "lmg"))

# fix outcome + antigen ordering for the plots
ladder_df$outcome <- factor(ladder_df$outcome, levels = unname(params$outcomes))
lmg_df$outcome    <- factor(lmg_df$outcome,    levels = unname(params$outcomes))
lmg_df$antigen    <- factor(lmg_df$antigen,    levels = sub("_IgG$", "", ANTI_IGG))

kable(tidyr::pivot_wider(ladder_df[, c("outcome","block","cv_r2")],
                         names_from = block, values_from = cv_r2),
      caption = "Six-block cross-validated R² ladder (cumulative) for each whole-organism outcome.")
```

# The figure

```{r palettes}
BLOCK_COL <- c("design"                       = "#6f6f6f",
               "+ clinical (pre-vaccination)" = "#a3a3a3",
               "+ clinical (birth)"           = "#d0d0d0",
               "+ maternal production"        = "#2b6098",
               "+ cord (transfer)"            = "#5ba3c0",
               "+ concurrent (ceiling)"       = "#d0692f")
ANTIGEN_COL <- c(PRN = "#206090", PT = "#d04050", FHA = "#308040", FIM = "#4c9a9a")

# principal antigen (largest LMG share) per outcome, for the right-panel annotation
totals <- lmg_df %>% dplyr::group_by(outcome) %>%
  dplyr::summarise(total = sum(lmg_r2), .groups = "drop")
principal <- lmg_df %>% dplyr::group_by(outcome) %>%
  dplyr::mutate(share = lmg_r2 / sum(lmg_r2)) %>%
  dplyr::slice_max(lmg_r2, n = 1, with_ties = FALSE) %>% dplyr::ungroup() %>%
  dplyr::left_join(totals, by = "outcome")
```

```{r left-panel}
p_left <- ggplot(ladder_df, aes(outcome, cv_r2, fill = block)) +
  geom_col(position = position_dodge(width = 0.82), width = 0.78,
           colour = "white", linewidth = 0.3) +
  geom_text(aes(label = sprintf("%.2f", cv_r2)),
            position = position_dodge(width = 0.82), vjust = -0.35,
            size = 2.7, colour = "grey25") +
  scale_fill_manual(values = BLOCK_COL, name = NULL) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
  labs(x = NULL, y = "Cross-validated R\u00b2",
       title = "Where predictability enters (CV-R\u00b2 block ladder)") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 12),
        legend.position = "bottom", axis.ticks = element_blank()) +
  guides(fill = guide_legend(nrow = 1, byrow = TRUE))
```

```{r right-panel}
p_right <- ggplot(lmg_df, aes(lmg_r2, outcome, fill = antigen)) +
  geom_col(colour = "white", linewidth = 0.5) +
  geom_text(data = principal, aes(x = total, y = outcome, label = sprintf("%s %.0f%%", antigen, 100 * share)),
            hjust = -0.08, size = 3, colour = "grey25", inherit.aes = FALSE) +
  scale_fill_manual(values = ANTIGEN_COL, name = "antigen") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.22))) +
  labs(x = "maternal production LMG R\u00b2", y = NULL,
       title = "Principal antigen of each outcome") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 12),
        legend.position = "bottom", axis.ticks = element_blank())
```

```{r figure5, fig.width=14.5, fig.height=5.6, fig.cap="Figure 5 (revised). Where predictability of infant whole-organism function enters. Left: cumulative cross-validated R\u00b2 for a six-block ladder \u2014 design (vaccination type), + pre-vaccination clinical, + birth clinical, + maternal production (maternal antigen at delivery), + cord transfer, + concurrent (two-month co-measured ceiling). The concurrent rung is a descriptive ceiling, not forward prediction. Right: within-outcome Shapley/LMG variance shares of the maternal-production antigens; the labelled antigen is the principal (largest-share) predictor of each outcome. Both panels exclude the concurrent block from any predictive claim."}
if (!requireNamespace("patchwork", quietly = TRUE))
  stop("Install `patchwork`: install.packages('patchwork').")
library(patchwork)
fig5 <- p_left + p_right + plot_layout(widths = c(1.55, 1.0))
fig5
```

```{r save}
dir.create(params$fig_out_dir, showWarnings = FALSE, recursive = TRUE)
png_path <- file.path(params$fig_out_dir, "figure5_block_ladder_sixway.png")
pdf_path <- file.path(params$fig_out_dir, "figure5_block_ladder_sixway.pdf")
ggsave(png_path, fig5, width = 14.5, height = 5.6, dpi = 300, bg = "white")
ggsave(pdf_path, fig5, width = 14.5, height = 5.6, bg = "white")
cat("Saved:\n- ", png_path, "\n- ", pdf_path, "\n", sep = "")
```

---

*Revised Figure 5. The design/clinical foundation of the block ladder is split
into three ordered blocks — design (vaccination type), pre-vaccination clinical
(maternal age, BMI, haemoglobin, parity, gestational age at vaccination), and
birth clinical (gestational age at birth, delivery mode, birth weight, infant sex,
vaccine-to-birth interval) — so the six rungs are design, + pre-vaccination
clinical, + birth clinical, + maternal production, + cord transfer, + concurrent
ceiling. The former "maternal (forward)" block is relabelled "maternal
production" throughout, and the right-panel axis reads "maternal production LMG
R²". Adjust the three `*_vars` params to move a covariate between the
pre-vaccination and birth clinical blocks.*
