---
title: "Predicting Infant M02 PTNA from Maternal and Cord IgG"
subtitle: |
  Six predictor blocks tracing the maternal antibody signal from vaccination
  through placental transfer to the infant at 2 months
author: "Scot Zens"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
params:
  outcome_feature: "WHOLE_PTNA"       # WHOLE_PTNA | WHOLE_SBA | WHOLE_WT_IgG
  outcome_label: "PTNA"             # used in cat() and plot titles
  outcome_y_label: "Log10 PTNA"
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo    = TRUE
)
```

---

# Analytical Framework

This document investigates how the infant's pertussis toxin neutralization
capacity at **M02** (2 months of age, immediately before first vaccination)
is shaped by the complete chain of antibody events between maternal
vaccination and early infancy. The full six-block framework is run twice,
under two independent stratifications:

- **Infant vaccination arm** (aP vs wP) — the infant's own priming history.
- **Maternal vaccination arm** (TdaP vs TT) — what the mother received
  during pregnancy.

Both stratifications use identical blocks, models, tables, and figures, so
the two can be compared directly. The infant-arm analysis is presented
first (Blocks 1–6 plus a cross-arm comparison), followed by the parallel
maternal-arm analysis.

The outcome — **`r params$outcome_label` at M02** — is a functional assay
measuring the ability of infant serum antibodies to neutralize pertussis
toxin. Unlike total IgG titres, it captures the biologically active
(neutralizing) fraction of the antibody response.

Six sequential predictor blocks are tested, each interrogating a distinct
biological mechanism:

```{r framework-table, echo=FALSE}
framework <- data.frame(
  Block = paste0("Block ", 1:6),
  Period = c(
    "P02 and P09 (maternal)",
    "P02 → P09 (maternal)",
    "P09 → M00 (maternal → cord)",
    "P09 → M00 (maternal → cord)",
    "M00 → M02 (cord → infant)",
    "M00 → M02 (cord → infant)"
  ),
  Type = c(
    "Antibody quantity",
    "Class switching",
    "Antibody quantity",
    "Class switching",
    "Antibody quantity",
    "Class switching"
  ),
  Biological_Question = c(
    "Do maternal IgG levels before and at delivery predict infant PTNA?",
    "Does gestational subclass polarisation predict infant PTNA?",
    "Does the quantity transferred across the placenta predict infant PTNA?",
    "Does subclass-selective placental transfer predict infant PTNA?",
    "Do antibody decay rates from birth to M02 predict infant PTNA?",
    "Does subclass-selective neonatal decay predict infant PTNA?"
  ),
  Expected_driver = c(
    "Higher PT IgG1 at P09 → higher infant PTNA",
    "Shift toward pro-inflammatory (IgG1+IgG3) → higher infant PTNA",
    "Greater net transfer (cord > maternal) → higher infant PTNA",
    "Preferential IgG1 transfer → higher infant PTNA",
    "Less decay (smaller negative delta) → higher infant PTNA",
    "Selective IgG1 preservation → higher infant PTNA"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(
  framework,
  col.names = c("Block", "Period", "Predictor type",
                "Biological question", "Expected direction"),
  caption   = "Table 1. Six-block predictor framework."
)
```

**Pro-inflammatory subclasses:** IgG1, IgG3  
**Tolerogenic subclasses:** IgG2, IgG4  
**Ratio:** log₁₀[(IgG1 + IgG3) / (IgG2 + IgG4)]; positive values indicate
pro-inflammatory polarisation.

---

# Setup and Parameters

```{r setup}
library(data.table)
library(tidyverse)
library(here)
library(RColorBrewer)
library(broom)
library(relaimpo)
library(knitr)

# ---- Outcome (edit here to switch to SBA or WT_IgG) ---------
outcome_feature <- params$outcome_feature   # "WHOLE_PTNA" | "WHOLE_SBA" | "WHOLE_WT_IgG"
outcome_label   <- params$outcome_label
outcome_y_label <- params$outcome_y_label
outcome_col     <- paste0(outcome_feature, "_M02")

# ---- Predictor feature lists --------------------------------
antigen_totals <- c("PT_IgG",  "FHA_IgG",  "PRN_IgG")
pt_subclasses  <- c("PT_IgG1",  "PT_IgG2",  "PT_IgG3",  "PT_IgG4")
fha_subclasses <- c("FHA_IgG1", "FHA_IgG2", "FHA_IgG3", "FHA_IgG4")
prn_subclasses <- c("PRN_IgG1", "PRN_IgG2", "PRN_IgG3", "PRN_IgG4")
all_subclasses <- c(pt_subclasses, fha_subclasses, prn_subclasses)
predictor_features <- c(antigen_totals, all_subclasses)

# ---- Visit and arm definitions ------------------------------
PRED_VISITS <- c("P02", "P09", "M00", "M02")
arms        <- c("TdaP", "TT")
infant_arms <- c("aP", "wP")
iarm_colors <- c("aP" = "dodgerblue2", "wP" = "forestgreen")
marm_colors <- c("TdaP" = "skyblue",   "TT" = "salmon")

# ---- Helper: p-value label ----------------------------------
pvalue_to_label <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE      ~ "ns"
  )
}

# ---- Helper: ratio (log10 pro-inflam / tolerogenic) ---------
ratio_col <- function(g1, g2, g3, g4) {
  log10((10^g1 + 10^g3) / (10^g2 + 10^g4))
}

# ---- Helper: fit lm, return structured result (silent) ------
run_lm <- function(outcome_col, predictors, data, label = "") {
  f <- reformulate(predictors, response = outcome_col)
  m <- lm(f, data = data, na.action = na.omit)
  s <- summary(m)

  tidy_coef <- tidy(m) %>%
    mutate(
      sig    = pvalue_to_label(p.value),
      across(c(estimate, std.error, statistic), ~round(., 4)),
      p.value = signif(p.value, 3)
    ) %>%
    rename(Predictor = term, `β` = estimate, SE = std.error,
           t = statistic, `p` = p.value, ` ` = sig)

  ri <- tryCatch(calc.relimp(m, type = "lmg"),
                 error = function(e) NULL)

ri_table <- if (!is.null(ri)) {
    data.frame(
      Predictor = names(ri@lmg),
      lmg_pct   = round(ri@lmg * 100, 2),
      stringsAsFactors = FALSE
    ) %>% arrange(desc(lmg_pct))
  } else NULL

  invisible(list(
    model     = m,
    relimp    = ri,
    label     = label,
    R2        = round(s$r.squared,     4),
    adj_R2    = round(s$adj.r.squared, 4),
    n_obs     = nobs(m),
    p_model   = pf(s$fstatistic[1], s$fstatistic[2],
                   s$fstatistic[3], lower.tail = FALSE),
    tidy_coef = tidy_coef,
    ri_table  = ri_table
  ))
}

# ---- Helper: display a run_lm result with kable -------------
show_model <- function(res, show_relimp = TRUE) {
  cat(sprintf(
    "\n**n = %d &nbsp;|&nbsp; R² = %.3f &nbsp;|&nbsp; adj-R² = %.3f &nbsp;|&nbsp; p~model~ = %.3g**\n\n",
    res$n_obs, res$R2, res$adj_R2, res$p_model
  ))
  print(knitr::kable(
    res$tidy_coef,
    caption = paste0("Coefficients — ", res$label),
    align   = c("l", rep("r", ncol(res$tidy_coef) - 1))
  ))
  cat("\n")
  if (show_relimp && !is.null(res$ri_table)) {
    print(knitr::kable(
        res$ri_table %>% rename(`lmg (%)` = lmg_pct),
        caption = paste0("Relative importance (lmg) — ", res$label),
        align   = c("l", "r")
      ))
    cat("\n")
  }
}

# ---- Helper: collect R² into a one-row data frame -----------
# `stratum` holds the level of whichever stratification is active
# (aP/wP for infant arm, or TdaP/TT for maternal arm).
r2_row <- function(res, stratum, block, model_name) {
  data.frame(
    stratum = stratum, block = block, model = model_name,
    n_obs = res$n_obs, R2 = res$R2, adj_R2 = res$adj_R2,
    stringsAsFactors = FALSE
  )
}

# ---- Helper: render every stored model for one stratum level ----
# Used by both the infant-arm and maternal-arm display sections.
show_all_models <- function(br, lvl, ids, show_relimp = TRUE) {
  for (id in ids) {
    if (!is.null(br[[lvl]][[id]]))
      show_model(br[[lvl]][[id]], show_relimp = show_relimp)
  }
}
```

---

# Data Preparation

## Load and Filter

```{r data-load}
load(file = here::here("./data/c_set.RData"))
data_raw          <- ebaa_extra
data_raw$arm_name <- data_raw$maternal_arm
data_raw$arm_name <- fct_recode(data_raw$arm_name, "TT" = "TT", "TdaP" = "TdaP")
data_raw$feature  <- factor(paste(data_raw$antigen, data_raw$analyte, sep = "_"))

all_features_needed <- c(predictor_features, outcome_feature)

data_long <- data_raw %>%
  dplyr::filter(feature    %in% all_features_needed,
                visit_name %in% PRED_VISITS,
                arm_name   %in% arms) %>%
  dplyr::select(subject_accession, visit_name, arm_name, infant_arm,
                feature, log_assay_value) %>%
  distinct(subject_accession, visit_name, feature, .keep_all = TRUE)
```

```{r data-diagnostics, results='asis'}
cat("**Feature × Visit observation counts (after filtering)**\n\n")
ft <- as.data.frame(table(data_long$feature, data_long$visit_name))
colnames(ft) <- c("Feature", "Visit", "n")
ft <- ft %>% filter(n > 0) %>%
  pivot_wider(names_from = Visit, values_from = n, values_fill = 0)
knitr::kable(ft, caption = "Table 2. Observations per feature × visit.") %>% print()
```

## Pivot to Wide Format

Each row represents one subject. Column names follow the convention
`{feature}_{visit}` — for example `PT_IgG1_P09` or `WHOLE_PTNA_M02`.

```{r data-pivot}
data_wide <- data_long %>%
  pivot_wider(
    id_cols     = c(subject_accession, arm_name, infant_arm),
    names_from  = c(feature, visit_name),
    values_from = log_assay_value,
    names_sep   = "_"
  )

cat(sprintf("Wide dataset: **%d subjects × %d columns**  \n",
            nrow(data_wide), ncol(data_wide)))
cat(sprintf("Outcome column `%s`: **%d non-missing**\n",
            outcome_col, sum(!is.na(data_wide[[outcome_col]]))))
```

## Compute Derived Quantities

Ratios and delta (Δ) variables are computed here for all six blocks.

```{r data-derive}
data_wide <- data_wide %>%
  mutate(
    # ---- Ratios at each visit ----
    PT_ratio_P02  = ratio_col(PT_IgG1_P02,  PT_IgG2_P02,  PT_IgG3_P02,  PT_IgG4_P02),
    FHA_ratio_P02 = ratio_col(FHA_IgG1_P02, FHA_IgG2_P02, FHA_IgG3_P02, FHA_IgG4_P02),
    PRN_ratio_P02 = ratio_col(PRN_IgG1_P02, PRN_IgG2_P02, PRN_IgG3_P02, PRN_IgG4_P02),
    PT_ratio_P09  = ratio_col(PT_IgG1_P09,  PT_IgG2_P09,  PT_IgG3_P09,  PT_IgG4_P09),
    FHA_ratio_P09 = ratio_col(FHA_IgG1_P09, FHA_IgG2_P09, FHA_IgG3_P09, FHA_IgG4_P09),
    PRN_ratio_P09 = ratio_col(PRN_IgG1_P09, PRN_IgG2_P09, PRN_IgG3_P09, PRN_IgG4_P09),
    PT_ratio_M00  = ratio_col(PT_IgG1_M00,  PT_IgG2_M00,  PT_IgG3_M00,  PT_IgG4_M00),
    FHA_ratio_M00 = ratio_col(FHA_IgG1_M00, FHA_IgG2_M00, FHA_IgG3_M00, FHA_IgG4_M00),
    PRN_ratio_M00 = ratio_col(PRN_IgG1_M00, PRN_IgG2_M00, PRN_IgG3_M00, PRN_IgG4_M00),
    PT_ratio_M02  = ratio_col(PT_IgG1_M02,  PT_IgG2_M02,  PT_IgG3_M02,  PT_IgG4_M02),
    FHA_ratio_M02 = ratio_col(FHA_IgG1_M02, FHA_IgG2_M02, FHA_IgG3_M02, FHA_IgG4_M02),
    PRN_ratio_M02 = ratio_col(PRN_IgG1_M02, PRN_IgG2_M02, PRN_IgG3_M02, PRN_IgG4_M02),
    # ---- Block 2: Δ ratio P02→P09 ----
    delta_PT_ratio_P02_P09  = PT_ratio_P09  - PT_ratio_P02,
    delta_FHA_ratio_P02_P09 = FHA_ratio_P09 - FHA_ratio_P02,
    delta_PRN_ratio_P02_P09 = PRN_ratio_P09 - PRN_ratio_P02,
    # ---- Block 3: Δ quantity P09→M00 ----
    delta_PT_IgG_P09_M00   = PT_IgG_M00   - PT_IgG_P09,
    delta_PT_IgG1_P09_M00  = PT_IgG1_M00  - PT_IgG1_P09,
    delta_PT_IgG2_P09_M00  = PT_IgG2_M00  - PT_IgG2_P09,
    delta_PT_IgG3_P09_M00  = PT_IgG3_M00  - PT_IgG3_P09,
    delta_PT_IgG4_P09_M00  = PT_IgG4_M00  - PT_IgG4_P09,
    delta_FHA_IgG_P09_M00  = FHA_IgG_M00  - FHA_IgG_P09,
    delta_FHA_IgG1_P09_M00 = FHA_IgG1_M00 - FHA_IgG1_P09,
    delta_FHA_IgG2_P09_M00 = FHA_IgG2_M00 - FHA_IgG2_P09,
    delta_FHA_IgG3_P09_M00 = FHA_IgG3_M00 - FHA_IgG3_P09,
    delta_FHA_IgG4_P09_M00 = FHA_IgG4_M00 - FHA_IgG4_P09,
    delta_PRN_IgG_P09_M00  = PRN_IgG_M00  - PRN_IgG_P09,
    delta_PRN_IgG1_P09_M00 = PRN_IgG1_M00 - PRN_IgG1_P09,
    delta_PRN_IgG2_P09_M00 = PRN_IgG2_M00 - PRN_IgG2_P09,
    delta_PRN_IgG3_P09_M00 = PRN_IgG3_M00 - PRN_IgG3_P09,
    delta_PRN_IgG4_P09_M00 = PRN_IgG4_M00 - PRN_IgG4_P09,
    # ---- Block 4: Δ ratio P09→M00 ----
    delta_PT_ratio_P09_M00  = PT_ratio_M00  - PT_ratio_P09,
    delta_FHA_ratio_P09_M00 = FHA_ratio_M00 - FHA_ratio_P09,
    delta_PRN_ratio_P09_M00 = PRN_ratio_M00 - PRN_ratio_P09,
    # ---- Block 5: Δ quantity M00→M02 ----
    delta_PT_IgG_M00_M02   = PT_IgG_M02   - PT_IgG_M00,
    delta_PT_IgG1_M00_M02  = PT_IgG1_M02  - PT_IgG1_M00,
    delta_PT_IgG2_M00_M02  = PT_IgG2_M02  - PT_IgG2_M00,
    delta_PT_IgG3_M00_M02  = PT_IgG3_M02  - PT_IgG3_M00,
    delta_PT_IgG4_M00_M02  = PT_IgG4_M02  - PT_IgG4_M00,
    delta_FHA_IgG_M00_M02  = FHA_IgG_M02  - FHA_IgG_M00,
    delta_FHA_IgG1_M00_M02 = FHA_IgG1_M02 - FHA_IgG1_M00,
    delta_FHA_IgG2_M00_M02 = FHA_IgG2_M02 - FHA_IgG2_M00,
    delta_FHA_IgG3_M00_M02 = FHA_IgG3_M02 - FHA_IgG3_M00,
    delta_FHA_IgG4_M00_M02 = FHA_IgG4_M02 - FHA_IgG4_M00,
    delta_PRN_IgG_M00_M02  = PRN_IgG_M02  - PRN_IgG_M00,
    delta_PRN_IgG1_M00_M02 = PRN_IgG1_M02 - PRN_IgG1_M00,
    delta_PRN_IgG2_M00_M02 = PRN_IgG2_M02 - PRN_IgG2_M00,
    delta_PRN_IgG3_M00_M02 = PRN_IgG3_M02 - PRN_IgG3_M00,
    delta_PRN_IgG4_M00_M02 = PRN_IgG4_M02 - PRN_IgG4_M00,
    # ---- Block 6: Δ ratio M00→M02 ----
    delta_PT_ratio_M00_M02  = PT_ratio_M02  - PT_ratio_M00,
    delta_FHA_ratio_M00_M02 = FHA_ratio_M02 - FHA_ratio_M00,
    delta_PRN_ratio_M00_M02 = PRN_ratio_M02 - PRN_ratio_M00
  )
```

```{r missingness-summary, results='asis'}
check_cols <- c(
  "PT_ratio_P02", "PT_ratio_P09", "PT_ratio_M00", "PT_ratio_M02",
  "delta_PT_ratio_P02_P09", "delta_PT_ratio_P09_M00", "delta_PT_ratio_M00_M02",
  "delta_PT_IgG_P09_M00",   "delta_PT_IgG_M00_M02",   outcome_col
)
miss_df <- data.frame(
  Variable  = check_cols,
  Missing_n = colSums(is.na(data_wide[, intersect(check_cols, names(data_wide))]))
)
knitr::kable(miss_df, caption = "Table 3. Missingness in key derived variables.",
             row.names = FALSE) %>% print()
```

---

# Run All Models

The model-fitting logic is wrapped in a single function, `run_all_blocks()`,
which is called twice — once stratifying by **infant arm** (aP vs wP) and
once by **maternal arm** (TdaP vs TT). This guarantees both stratifications
use byte-for-byte identical models, predictors, and storage structure. Each
call returns a named list of result objects that the display sections below
read from.

```{r run-all-models, results='hide'}
# ============================================================
# run_all_blocks(): fit all six blocks for every level of a
# stratification variable. Returns a list of storage objects.
#   d_all        : the wide data frame
#   strat_col    : "infant_arm" or "arm_name"
#   strat_levels : e.g. c("aP","wP") or c("TdaP","TT")
# ============================================================
run_all_blocks <- function(d_all, strat_col, strat_levels) {

  r2_all        <- list()
  ratio_models  <- list()
  relimp_b1     <- list()
  relimp_b3     <- list()
  relimp_b5     <- list()
  block_results <- list()

  for (lvl in strat_levels) {
    d  <- d_all %>% dplyr::filter(.data[[strat_col]] == lvl)
    br <- list()

    # ---- BLOCK 1 ----
    br[["1a"]] <- run_lm(outcome_col, paste0(antigen_totals, "_P02"), d, "Antigen totals @ P02")
    br[["1b"]] <- run_lm(outcome_col, paste0(antigen_totals, "_P09"), d, "Antigen totals @ P09")
    br[["1c"]] <- run_lm(outcome_col,
                         c(paste0(antigen_totals, "_P02"), paste0(antigen_totals, "_P09")),
                         d, "Antigen totals P02 + P09 combined")
    br[["1d"]] <- run_lm(outcome_col, paste0(pt_subclasses,  "_P02"), d, "PT subclasses @ P02")
    br[["1e"]] <- run_lm(outcome_col, paste0(pt_subclasses,  "_P09"), d, "PT subclasses @ P09")
    br[["1f"]] <- run_lm(outcome_col, paste0(fha_subclasses, "_P02"), d, "FHA subclasses @ P02")
    br[["1g"]] <- run_lm(outcome_col, paste0(fha_subclasses, "_P09"), d, "FHA subclasses @ P09")
    br[["1h"]] <- run_lm(outcome_col, paste0(prn_subclasses, "_P02"), d, "PRN subclasses @ P02")
    br[["1i"]] <- run_lm(outcome_col, paste0(prn_subclasses, "_P09"), d, "PRN subclasses @ P09")
    br[["1j"]] <- run_lm(outcome_col, paste0(all_subclasses, "_P02"), d, "Full 12-subclass @ P02")
    br[["1k"]] <- run_lm(outcome_col, paste0(all_subclasses, "_P09"), d, "Full 12-subclass @ P09")
    br[["1l"]] <- run_lm(outcome_col,
                         c(paste0(all_subclasses, "_P02"), paste0(all_subclasses, "_P09")),
                         d, "Full 24-predictor P02 + P09")
    relimp_b1[[lvl]] <- br[["1k"]]$relimp

    for (id in paste0("1", letters[1:12]))
      r2_all[[paste0(id, "_", lvl)]] <- r2_row(br[[id]], lvl, "B1_quantity", sub("1", "", id))

    # ---- BLOCK 2 ----
    br[["2"]] <- run_lm(outcome_col,
                        c("delta_PT_ratio_P02_P09","delta_FHA_ratio_P02_P09",
                          "delta_PRN_ratio_P02_P09"),
                        d, "Δ ratio P02→P09")
    r2_all[[paste0("B2_", lvl)]]    <- r2_row(br[["2"]], lvl, "B2_ratio_switch", "3-ratio")
    ratio_models[[paste0("B2_", lvl)]] <- br[["2"]]$model

    # ---- BLOCK 3 ----
    b3_ant  <- paste0("delta_", antigen_totals, "_P09_M00")
    b3_pt   <- paste0("delta_", pt_subclasses,  "_P09_M00")
    b3_fha  <- paste0("delta_", fha_subclasses, "_P09_M00")
    b3_prn  <- paste0("delta_", prn_subclasses, "_P09_M00")
    b3_full <- c(b3_pt, b3_fha, b3_prn)
    br[["3a"]] <- run_lm(outcome_col, b3_ant,  d, "Δ antigen totals P09→M00")
    br[["3b"]] <- run_lm(outcome_col, b3_pt,   d, "Δ PT subclasses P09→M00")
    br[["3c"]] <- run_lm(outcome_col, b3_fha,  d, "Δ FHA subclasses P09→M00")
    br[["3d"]] <- run_lm(outcome_col, b3_prn,  d, "Δ PRN subclasses P09→M00")
    br[["3e"]] <- run_lm(outcome_col, b3_full, d, "Δ Full 12-subclass P09→M00")
    relimp_b3[[lvl]] <- br[["3e"]]$relimp
    for (id in paste0("3", letters[1:5]))
      r2_all[[paste0(id, "_", lvl)]] <- r2_row(br[[id]], lvl, "B3_transfer_qty", sub("3", "", id))

    # ---- BLOCK 4 ----
    br[["4"]] <- run_lm(outcome_col,
                        c("delta_PT_ratio_P09_M00","delta_FHA_ratio_P09_M00",
                          "delta_PRN_ratio_P09_M00"),
                        d, "Δ ratio P09→M00")
    r2_all[[paste0("B4_", lvl)]]    <- r2_row(br[["4"]], lvl, "B4_transfer_ratio", "3-ratio")
    ratio_models[[paste0("B4_", lvl)]] <- br[["4"]]$model

    # ---- BLOCK 5 ----
    b5_ant  <- paste0("delta_", antigen_totals, "_M00_M02")
    b5_pt   <- paste0("delta_", pt_subclasses,  "_M00_M02")
    b5_fha  <- paste0("delta_", fha_subclasses, "_M00_M02")
    b5_prn  <- paste0("delta_", prn_subclasses, "_M00_M02")
    b5_full <- c(b5_pt, b5_fha, b5_prn)
    br[["5a"]] <- run_lm(outcome_col, b5_ant,  d, "Δ antigen totals M00→M02")
    br[["5b"]] <- run_lm(outcome_col, b5_pt,   d, "Δ PT subclasses M00→M02")
    br[["5c"]] <- run_lm(outcome_col, b5_fha,  d, "Δ FHA subclasses M00→M02")
    br[["5d"]] <- run_lm(outcome_col, b5_prn,  d, "Δ PRN subclasses M00→M02")
    br[["5e"]] <- run_lm(outcome_col, b5_full, d, "Δ Full 12-subclass M00→M02")
    relimp_b5[[lvl]] <- br[["5e"]]$relimp
    for (id in paste0("5", letters[1:5]))
      r2_all[[paste0(id, "_", lvl)]] <- r2_row(br[[id]], lvl, "B5_decay_qty", sub("5", "", id))

    # ---- BLOCK 6 ----
    br[["6"]] <- run_lm(outcome_col,
                        c("delta_PT_ratio_M00_M02","delta_FHA_ratio_M00_M02",
                          "delta_PRN_ratio_M00_M02"),
                        d, "Δ ratio M00→M02")
    r2_all[[paste0("B6_", lvl)]]    <- r2_row(br[["6"]], lvl, "B6_decay_ratio", "3-ratio")
    ratio_models[[paste0("B6_", lvl)]] <- br[["6"]]$model

    block_results[[lvl]] <- br
  }

  list(block_results = block_results, r2_all = r2_all,
       ratio_models = ratio_models,
       relimp_b1 = relimp_b1, relimp_b3 = relimp_b3, relimp_b5 = relimp_b5)
}

# ---- Call 1: stratify by INFANT arm (aP vs wP) -------------
inf <- run_all_blocks(data_wide, "infant_arm", infant_arms)
block_results             <- inf$block_results
r2_all                    <- inf$r2_all
ratio_models_by_block_arm <- inf$ratio_models
relimp_b1_full_by_iarm    <- inf$relimp_b1
relimp_b3_full_by_iarm    <- inf$relimp_b3
relimp_b5_full_by_iarm    <- inf$relimp_b5

# ---- Call 2: stratify by MATERNAL arm (TdaP vs TT) ---------
mat <- run_all_blocks(data_wide, "arm_name", arms)
block_results_m            <- mat$block_results
r2_all_m                   <- mat$r2_all
ratio_models_by_block_marm <- mat$ratio_models
relimp_b1_full_by_marm     <- mat$relimp_b1
relimp_b3_full_by_marm     <- mat$relimp_b3
relimp_b5_full_by_marm     <- mat$relimp_b5
```

---

# Block 1: Maternal Antibody Levels at P02 and P09

**Biological context.** Maternal vaccination with Tdap in the third
trimester induces a sharp rise in PT-specific IgG. The levels present in
maternal circulation at **P09 (delivery)** represent the pool available for
placental transfer. The levels at **P02 (pre-vaccination)** represent the
baseline. Including both visits allows us to separate the pre-existing
antibody background from the vaccine-induced boost, and to ask whether it
is absolute level at delivery or the magnitude of the gestational change
that matters most for the infant's eventual neutralisation capacity.

*Predictors:* Log₁₀ antigen-specific total IgG (PT, FHA, PRN) and
subclasses (IgG1–4) at P02 and P09. All models use `na.action = na.omit`
so effective n varies with missingness in the predictor set.

## aP infants

### Antigen totals

```{r b1-ap-antigen, results='asis'}
show_model(block_results[["aP"]][["1a"]], show_relimp = TRUE)
show_model(block_results[["aP"]][["1b"]], show_relimp = TRUE)
show_model(block_results[["aP"]][["1c"]], show_relimp = FALSE)
```

### PT subclasses

```{r b1-ap-pt, results='asis'}
show_model(block_results[["aP"]][["1d"]])
show_model(block_results[["aP"]][["1e"]])
```

### FHA subclasses

```{r b1-ap-fha, results='asis'}
show_model(block_results[["aP"]][["1f"]])
show_model(block_results[["aP"]][["1g"]])
```

### PRN subclasses

```{r b1-ap-prn, results='asis'}
show_model(block_results[["aP"]][["1h"]])
show_model(block_results[["aP"]][["1i"]])
```

### Full 12-subclass models

```{r b1-ap-full, results='asis'}
show_model(block_results[["aP"]][["1j"]])
show_model(block_results[["aP"]][["1k"]])
show_model(block_results[["aP"]][["1l"]], show_relimp = FALSE)
```

## wP infants

### Antigen totals

```{r b1-wp-antigen, results='asis'}
show_model(block_results[["wP"]][["1a"]], show_relimp = TRUE)
show_model(block_results[["wP"]][["1b"]], show_relimp = TRUE)
show_model(block_results[["wP"]][["1c"]], show_relimp = FALSE)
```

### PT subclasses

```{r b1-wp-pt, results='asis'}
show_model(block_results[["wP"]][["1d"]])
show_model(block_results[["wP"]][["1e"]])
```

### FHA subclasses

```{r b1-wp-fha, results='asis'}
show_model(block_results[["wP"]][["1f"]])
show_model(block_results[["wP"]][["1g"]])
```

### PRN subclasses

```{r b1-wp-prn, results='asis'}
show_model(block_results[["wP"]][["1h"]])
show_model(block_results[["wP"]][["1i"]])
```

### Full 12-subclass models

```{r b1-wp-full, results='asis'}
show_model(block_results[["wP"]][["1j"]])
show_model(block_results[["wP"]][["1k"]])
show_model(block_results[["wP"]][["1l"]], show_relimp = FALSE)
```

## Subclass slopes: aP vs wP (P09)

```{r b1-interaction-plot, fig.cap="Figure 1. Maternal subclass IgG at P09 vs infant PTNA at M02, by antigen and infant arm. Each panel shows log-linear slopes for IgG1 (red), IgG2 (blue), IgG3 (green), IgG4 (purple)."}
data_long_b1 <- bind_rows(
  data_wide %>% dplyr::filter(infant_arm == "aP") %>%
    dplyr::select(all_of(outcome_col), matches("IgG[1-4]_P09$")) %>%
    mutate(infant_arm = "aP"),
  data_wide %>% dplyr::filter(infant_arm == "wP") %>%
    dplyr::select(all_of(outcome_col), matches("IgG[1-4]_P09$")) %>%
    mutate(infant_arm = "wP")
) %>%
  pivot_longer(cols = matches("IgG[1-4]_P09$"),
               names_to = c("antigen","subclass","visit"),
               names_pattern = "(PT|FHA|PRN)_(IgG[1-4])_(P09)",
               values_to = "subclass_value") %>%
  drop_na()

ggplot(data_long_b1,
       aes(x = subclass_value, y = .data[[outcome_col]],
           color = subclass, linetype = infant_arm)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ infant_arm) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Log10 Maternal Subclass IgG at P09",
       y = outcome_y_label, color = "Subclass", linetype = "Infant arm",
       title = paste0("Block 1 (P09): Antigen × Subclass → ",
                      outcome_label, " at M02")) +
  theme_bw(base_size = 11)
```

---

# Block 2: Gestational Subclass Switching (P02 → P09)

**Biological context.** Beyond the absolute antibody levels, the
*direction of class switching* during pregnancy may carry independent
information. If vaccination biases the immune response toward
pro-inflammatory subclasses (IgG1, IgG3), the delta ratio (P09 − P02)
will be positive. A positive coefficient in this model means that infants
whose mothers shifted toward pro-inflammatory subclass dominance during
pregnancy have higher PTNA at M02 — consistent with the idea that
functional antibody quality, not just quantity, is imprinted by the
gestational immune environment.

*Predictors:* Δ pro-inflammatory/tolerogenic ratio (log₁₀) for each
antigen, where Δ = ratio at P09 minus ratio at P02. Positive values
indicate a shift toward IgG1+IgG3 dominance during pregnancy.

## aP infants

```{r b2-ap, results='asis'}
show_model(block_results[["aP"]][["2"]])
```

**Univariate antigen-level associations:**

```{r b2-ap-univariate, results='asis'}
univar <- lapply(c("PT","FHA","PRN"), function(ant) {
  dc <- paste0("delta_", ant, "_ratio_P02_P09")
  d  <- data_wide %>% dplyr::filter(infant_arm == "aP")
  if (!dc %in% names(d)) return(NULL)
  m  <- lm(reformulate(dc, response = outcome_col), data = d, na.action = na.omit)
  tidy(m) %>% dplyr::filter(term != "(Intercept)") %>%
    mutate(antigen = ant, R2 = summary(m)$r.squared,
           sig = pvalue_to_label(p.value),
           across(c(estimate, std.error, statistic, R2), ~round(., 4)),
           p.value = signif(p.value, 3))
}) %>% bind_rows()
knitr::kable(univar, caption = "Block 2 univariate: aP infants") %>% print()
```

## wP infants

```{r b2-wp, results='asis'}
show_model(block_results[["wP"]][["2"]])
```

```{r b2-wp-univariate, results='asis'}
univar <- lapply(c("PT","FHA","PRN"), function(ant) {
  dc <- paste0("delta_", ant, "_ratio_P02_P09")
  d  <- data_wide %>% dplyr::filter(infant_arm == "wP")
  if (!dc %in% names(d)) return(NULL)
  m  <- lm(reformulate(dc, response = outcome_col), data = d, na.action = na.omit)
  tidy(m) %>% dplyr::filter(term != "(Intercept)") %>%
    mutate(antigen = ant, R2 = summary(m)$r.squared,
           sig = pvalue_to_label(p.value),
           across(c(estimate, std.error, statistic, R2), ~round(., 4)),
           p.value = signif(p.value, 3))
}) %>% bind_rows()
knitr::kable(univar, caption = "Block 2 univariate: wP infants") %>% print()
```

---

# Block 3: Placental Transfer — Quantity (P09 → M00)

**Biological context.** Placental transfer of IgG is an active FcRn-mediated
process that concentrates maternal antibodies in fetal circulation. The
quantity transferred is captured by Δ = cordblood (M00) − maternal (P09),
measured on the log₁₀ scale. Values near zero indicate equivalent levels;
positive values indicate the infant has *more* of that subclass than the
mother (active concentration), which occurs for IgG1 in particular.
A positive coefficient here means that more efficient transfer of that
subclass is associated with higher infant PTNA at M02.

*Predictors:* Δ log₁₀ IgG for each antigen/subclass (M00 − P09). Values
are expected to cluster near zero with some positive skew for well-transferred
subclasses.

## aP infants

### Antigen totals

```{r b3-ap-antigen, results='asis'}
show_model(block_results[["aP"]][["3a"]])
```

### Per-antigen subclass models

```{r b3-ap-sub, results='asis'}
show_model(block_results[["aP"]][["3b"]])
show_model(block_results[["aP"]][["3c"]])
show_model(block_results[["aP"]][["3d"]])
```

### Full 12-subclass

```{r b3-ap-full, results='asis'}
show_model(block_results[["aP"]][["3e"]])
```

## wP infants

### Antigen totals

```{r b3-wp-antigen, results='asis'}
show_model(block_results[["wP"]][["3a"]])
```

### Per-antigen subclass models

```{r b3-wp-sub, results='asis'}
show_model(block_results[["wP"]][["3b"]])
show_model(block_results[["wP"]][["3c"]])
show_model(block_results[["wP"]][["3d"]])
```

### Full 12-subclass

```{r b3-wp-full, results='asis'}
show_model(block_results[["wP"]][["3e"]])
```

---

# Block 4: Placental Transfer — Selectivity (P09 → M00)

**Biological context.** FcRn preferentially transfers IgG1 and IgG4 over
IgG2 and IgG3 across the placenta. If this selectivity matters for
functional infant immunity, the change in pro-/tolerogenic ratio between
maternal blood (P09) and cord blood (M00) — reflecting IgG1 enrichment
or depletion during transfer — should predict infant PTNA. A positive
coefficient means that infants who received proportionally more
pro-inflammatory IgG1 relative to tolerogenic IgG2/IgG4 during transfer
have higher PTNA at M02.

*Predictors:* Δ ratio (M00 − P09) for PT, FHA, PRN.

## aP infants

```{r b4-ap, results='asis'}
show_model(block_results[["aP"]][["4"]])
```

## wP infants

```{r b4-wp, results='asis'}
show_model(block_results[["wP"]][["4"]])
```

---

# Block 5: Neonatal Antibody Decay — Quantity (M00 → M02)

**Biological context.** Maternal antibodies transferred to the infant
undergo exponential decay during the first months of life. Different IgG
subclasses have different half-lives: IgG1 and IgG4 decay more slowly than
IgG3. The decay quantity (Δ = M02 − M00, expected negative for all
subclasses) captures how much has been lost by 2 months. A positive
coefficient here means *less decay* (smaller negative or even positive
delta) is associated with higher PTNA — i.e., better preservation of
maternal antibody predicts better functional immunity.

*Predictors:* Δ log₁₀ IgG (M02 − M00). Medians should be negative,
reflecting net loss between birth and 2 months.

## aP infants

```{r b5-ap-decay-check, results='asis'}
d_ap <- data_wide %>% dplyr::filter(infant_arm == "aP")
decay_cols <- paste0("delta_", antigen_totals, "_M00_M02")
decay_present <- intersect(decay_cols, names(d_ap))
if (length(decay_present) > 0) {
  meds <- round(apply(d_ap[, decay_present], 2,
                      function(x) median(x, na.rm = TRUE)), 3)
  knitr::kable(
    data.frame(Antigen = names(meds), `Median Δ log10` = meds),
    caption = "Antibody decay medians (M02 − M00): aP infants — negative = net loss",
    row.names = FALSE
  ) %>% print()
}
```

### Antigen totals

```{r b5-ap-antigen, results='asis'}
show_model(block_results[["aP"]][["5a"]])
```

### Per-antigen subclass models

```{r b5-ap-sub, results='asis'}
show_model(block_results[["aP"]][["5b"]])
show_model(block_results[["aP"]][["5c"]])
show_model(block_results[["aP"]][["5d"]])
```

### Full 12-subclass

```{r b5-ap-full, results='asis'}
show_model(block_results[["aP"]][["5e"]])
```

## wP infants

```{r b5-wp-decay-check, results='asis'}
d_wp <- data_wide %>% dplyr::filter(infant_arm == "wP")
decay_present_wp <- intersect(decay_cols, names(d_wp))
if (length(decay_present_wp) > 0) {
  meds_wp <- round(apply(d_wp[, decay_present_wp], 2,
                         function(x) median(x, na.rm = TRUE)), 3)
  knitr::kable(
    data.frame(Antigen = names(meds_wp), `Median Δ log10` = meds_wp),
    caption = "Antibody decay medians (M02 − M00): wP infants",
    row.names = FALSE
  ) %>% print()
}
```

### Antigen totals

```{r b5-wp-antigen, results='asis'}
show_model(block_results[["wP"]][["5a"]])
```

### Per-antigen subclass models

```{r b5-wp-sub, results='asis'}
show_model(block_results[["wP"]][["5b"]])
show_model(block_results[["wP"]][["5c"]])
show_model(block_results[["wP"]][["5d"]])
```

### Full 12-subclass

```{r b5-wp-full, results='asis'}
show_model(block_results[["wP"]][["5e"]])
```

## Subclass decay slopes: aP vs wP

```{r b5-interaction-plot, fig.cap="Figure 2. Subclass decay (M02 − M00) vs infant PTNA at M02, by antigen and infant arm. Negative x-axis values indicate net loss of that subclass."}
data_long_b5 <- bind_rows(
  data_wide %>% dplyr::filter(infant_arm == "aP") %>%
    dplyr::select(all_of(outcome_col),
                  matches("^delta_(PT|FHA|PRN)_IgG[1-4]_M00_M02$")) %>%
    mutate(infant_arm = "aP"),
  data_wide %>% dplyr::filter(infant_arm == "wP") %>%
    dplyr::select(all_of(outcome_col),
                  matches("^delta_(PT|FHA|PRN)_IgG[1-4]_M00_M02$")) %>%
    mutate(infant_arm = "wP")
) %>%
  pivot_longer(cols = matches("^delta_(PT|FHA|PRN)_IgG[1-4]_M00_M02$"),
               names_to = c("antigen","subclass"),
               names_pattern = "^delta_(PT|FHA|PRN)_(IgG[1-4])_M00_M02$",
               values_to = "delta_value") %>%
  drop_na()

ggplot(data_long_b5,
       aes(x = delta_value, y = .data[[outcome_col]],
           color = subclass, linetype = infant_arm)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ infant_arm) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Δ Log10 Subclass IgG (M02 − M00, decay = negative)",
       y = outcome_y_label, color = "Subclass", linetype = "Infant arm",
       title = paste0("Block 5 (decay): Antigen × Subclass → ",
                      outcome_label, " at M02")) +
  theme_bw(base_size = 11)
```

---

# Block 6: Neonatal Antibody Decay — Selectivity (M00 → M02)

**Biological context.** If subclasses decay at different rates during the
first two months (e.g. IgG1 more slowly than IgG3), the ratio will shift
over this period. The Δ ratio (M02 − M00) captures this shift: a positive
value means pro-inflammatory subclasses are *better preserved* relative to
tolerogenic ones. A positive coefficient here means that infants in whom
the pro-inflammatory subclass balance is maintained through the neonatal
period have higher PTNA — indicating that subclass-selective decay, not
just total antibody loss, shapes functional immunity.

*Predictors:* Δ ratio (M02 − M00) for PT, FHA, PRN.

## aP infants

```{r b6-ap, results='asis'}
show_model(block_results[["aP"]][["6"]])
```

## wP infants

```{r b6-wp, results='asis'}
show_model(block_results[["wP"]][["6"]])
```

---

# Cross-Arm Comparisons (Infant Arm: aP vs wP)

## R² Summary Across All Blocks

```{r r2-table, results='asis'}
r2_table <- bind_rows(r2_all) %>%
  arrange(block, model, stratum) %>%
  dplyr::select(stratum, block, model, n_obs, R2, adj_R2)

knitr::kable(as.data.frame(r2_table),
             caption = "Table 4. R² for all models across all blocks and infant arms.",
             align = c("l","l","l","r","r","r")) %>% print()
```

```{r r2-wide, results='asis'}
r2_wide <- r2_table %>%
  pivot_wider(names_from  = stratum,
              values_from = c(R2, adj_R2, n_obs),
              names_sep   = "_") %>%
  mutate(delta_R2 = round(R2_aP - R2_wP, 4)) %>%
  arrange(block, model)

knitr::kable(as.data.frame(r2_wide),
             caption = "Table 5. R² side-by-side: aP vs wP. delta_R2 = R²(aP) − R²(wP).",
             align = c("l","l","r","r","r","r","r","r","r")) %>% print()
```

```{r r2-barplot, fig.cap="Figure 3. R² by block and infant arm. Bars show model fit for antigen-level (total IgG) models; lines connect aP and wP arms.", fig.height=5}
r2_antigen <- r2_table %>%
  dplyr::filter(grepl("antigen", model)) %>%
  mutate(block_clean = gsub("_qty|_quantity|_switch|_ratio", "", block))

ggplot(r2_antigen, aes(x = block_clean, y = R2,
                        fill = stratum, group = stratum)) +
  geom_col(position = position_dodge(0.6), width = 0.5) +
  geom_line(aes(color = stratum), position = position_dodge(0.6),
            linewidth = 0.6, linetype = "dashed") +
  scale_fill_manual(values  = iarm_colors) +
  scale_color_manual(values = iarm_colors) +
  labs(x = "Block", y = "R²", fill = "Infant arm", color = "Infant arm",
       title = paste0("Explained variance (antigen-level models) — ",
                      outcome_label)) +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Ratio-Model Coefficient Comparison Plots

```{r ratio-coef-plots, fig.cap="Figures 4–6. Ratio-model coefficients (Blocks 2, 4, 6) side-by-side for aP and wP. Error bars = 95% CI.", fig.height=5}
block_labels <- c(
  B2 = "Block 2: Δ ratio P02→P09\n(class switching during pregnancy)",
  B4 = "Block 4: Δ ratio P09→M00\n(class-selective placental transfer)",
  B6 = "Block 6: Δ ratio M00→M02\n(class-selective neonatal decay)"
)

for (blk in c("B2","B4","B6")) {
  coef_rows <- lapply(infant_arms, function(iarm) {
    key <- paste0(blk, "_", iarm)
    if (is.null(ratio_models_by_block_arm[[key]])) return(NULL)
    tidy(ratio_models_by_block_arm[[key]]) %>%
      dplyr::filter(term != "(Intercept)") %>%
      mutate(infant_arm = iarm)
  }) %>% bind_rows()

  if (nrow(coef_rows) == 0) next

  p <- ggplot(coef_rows,
              aes(x = term, y = estimate, fill = infant_arm,
                  ymin = estimate - 1.96 * std.error,
                  ymax = estimate + 1.96 * std.error)) +
    geom_col(position = position_dodge(0.6), width = 0.5) +
    geom_errorbar(position = position_dodge(0.6), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_fill_manual(values = iarm_colors) +
    labs(x     = "Pro-inflammatory/Tolerogenic Ratio",
         y     = paste0("β (→ ", outcome_label, " M02)"),
         title = paste0(block_labels[blk], "\n(outcome = ", outcome_label, ")"),
         fill  = "Infant arm") +
    theme_bw(base_size = 11)
  print(p)
}
```

## Relative Importance Comparison (Full 12-Subclass Models)

```{r relimp-comparison, results='asis'}
for (blk_nm in c("Block 1 (P09)", "Block 3 (transfer)", "Block 5 (decay)")) {
  rl <- switch(blk_nm,
    "Block 1 (P09)"       = relimp_b1_full_by_iarm,
    "Block 3 (transfer)"  = relimp_b3_full_by_iarm,
    "Block 5 (decay)"     = relimp_b5_full_by_iarm
  )
  if (is.null(rl[["aP"]]) || is.null(rl[["wP"]])) next
  ri_compare <- merge(
    data.frame(feature = names(rl[["aP"]]@lmg), lmg_aP = round(rl[["aP"]]@lmg*100, 2)),
    data.frame(feature = names(rl[["wP"]]@lmg), lmg_wP = round(rl[["wP"]]@lmg*100, 2)),
    by = "feature"
  ) %>%
    mutate(delta_lmg = round(lmg_aP - lmg_wP, 2)) %>%
    arrange(desc(abs(delta_lmg)))
  knitr::kable(
    ri_compare,
    caption = paste0("Relative importance (lmg %) — ", blk_nm,
                     " full 12-subclass model. delta_lmg = aP − wP."),
    align = c("l","r","r","r")
  ) %>% print()
  cat("\n")
}
```

---

---

# Maternal Arm Analysis (TdaP vs TT)

This section repeats the entire six-block framework, identical in every
respect, but stratifying by **maternal vaccination arm** rather than infant
arm. TdaP mothers received the acellular pertussis-containing booster during
pregnancy; TT mothers received tetanus toxoid only (the comparator).
Comparing the two isolates the contribution of maternal pertussis
vaccination itself to the antibody chain that predicts infant PTNA. All
models, predictors, captions, and statistics are produced by the same
`run_all_blocks()` function used above, so the maternal results are directly
comparable to the infant results section-for-section.

Each model's caption (e.g. "Antigen totals @ P09") identifies the predictor
set; results are grouped by block and split into TdaP and TT.

## Block 1 — Maternal Antibody Levels (P02 and P09)

### TdaP mothers

```{r b1m-tdap, results='asis'}
show_all_models(block_results_m, "TdaP", paste0("1", letters[1:12]))
```

### TT mothers

```{r b1m-tt, results='asis'}
show_all_models(block_results_m, "TT", paste0("1", letters[1:12]))
```

### Maternal subclass slopes: TdaP vs TT (P09)

```{r b1m-interaction-plot, fig.cap="Figure M1. Maternal subclass IgG at P09 vs infant PTNA at M02, by antigen and maternal arm.", fig.height=6}
data_long_b1m <- bind_rows(
  data_wide %>% dplyr::filter(arm_name == "TdaP") %>%
    dplyr::select(all_of(outcome_col), matches("IgG[1-4]_P09$")) %>%
    mutate(maternal_arm = "TdaP"),
  data_wide %>% dplyr::filter(arm_name == "TT") %>%
    dplyr::select(all_of(outcome_col), matches("IgG[1-4]_P09$")) %>%
    mutate(maternal_arm = "TT")
) %>%
  pivot_longer(cols = matches("IgG[1-4]_P09$"),
               names_to = c("antigen","subclass","visit"),
               names_pattern = "(PT|FHA|PRN)_(IgG[1-4])_(P09)",
               values_to = "subclass_value") %>%
  drop_na()

ggplot(data_long_b1m,
       aes(x = subclass_value, y = .data[[outcome_col]],
           color = subclass, linetype = maternal_arm)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ maternal_arm) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Log10 Maternal Subclass IgG at P09",
       y = outcome_y_label, color = "Subclass", linetype = "Maternal arm",
       title = paste0("Block 1 (P09): Antigen × Subclass → ",
                      outcome_label, " at M02 — TdaP vs TT")) +
  theme_bw(base_size = 11)
```

## Block 2 — Gestational Subclass Switching (P02 → P09)

### TdaP mothers

```{r b2m-tdap, results='asis'}
show_all_models(block_results_m, "TdaP", "2")
```

### TT mothers

```{r b2m-tt, results='asis'}
show_all_models(block_results_m, "TT", "2")
```

## Block 3 — Placental Transfer: Quantity (P09 → M00)

### TdaP mothers

```{r b3m-tdap, results='asis'}
show_all_models(block_results_m, "TdaP", paste0("3", letters[1:5]))
```

### TT mothers

```{r b3m-tt, results='asis'}
show_all_models(block_results_m, "TT", paste0("3", letters[1:5]))
```

## Block 4 — Placental Transfer: Selectivity (P09 → M00)

### TdaP mothers

```{r b4m-tdap, results='asis'}
show_all_models(block_results_m, "TdaP", "4")
```

### TT mothers

```{r b4m-tt, results='asis'}
show_all_models(block_results_m, "TT", "4")
```

## Block 5 — Neonatal Antibody Decay: Quantity (M00 → M02)

### TdaP mothers

```{r b5m-tdap, results='asis'}
show_all_models(block_results_m, "TdaP", paste0("5", letters[1:5]))
```

### TT mothers

```{r b5m-tt, results='asis'}
show_all_models(block_results_m, "TT", paste0("5", letters[1:5]))
```

### Maternal subclass decay slopes: TdaP vs TT

```{r b5m-interaction-plot, fig.cap="Figure M2. Subclass decay (M02 − M00) vs infant PTNA at M02, by antigen and maternal arm.", fig.height=6}
data_long_b5m <- bind_rows(
  data_wide %>% dplyr::filter(arm_name == "TdaP") %>%
    dplyr::select(all_of(outcome_col),
                  matches("^delta_(PT|FHA|PRN)_IgG[1-4]_M00_M02$")) %>%
    mutate(maternal_arm = "TdaP"),
  data_wide %>% dplyr::filter(arm_name == "TT") %>%
    dplyr::select(all_of(outcome_col),
                  matches("^delta_(PT|FHA|PRN)_IgG[1-4]_M00_M02$")) %>%
    mutate(maternal_arm = "TT")
) %>%
  pivot_longer(cols = matches("^delta_(PT|FHA|PRN)_IgG[1-4]_M00_M02$"),
               names_to = c("antigen","subclass"),
               names_pattern = "^delta_(PT|FHA|PRN)_(IgG[1-4])_M00_M02$",
               values_to = "delta_value") %>%
  drop_na()

ggplot(data_long_b5m,
       aes(x = delta_value, y = .data[[outcome_col]],
           color = subclass, linetype = maternal_arm)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(antigen ~ maternal_arm) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Δ Log10 Subclass IgG (M02 − M00, decay = negative)",
       y = outcome_y_label, color = "Subclass", linetype = "Maternal arm",
       title = paste0("Block 5 (decay): Antigen × Subclass → ",
                      outcome_label, " at M02 — TdaP vs TT")) +
  theme_bw(base_size = 11)
```

## Block 6 — Neonatal Antibody Decay: Selectivity (M00 → M02)

### TdaP mothers

```{r b6m-tdap, results='asis'}
show_all_models(block_results_m, "TdaP", "6")
```

### TT mothers

```{r b6m-tt, results='asis'}
show_all_models(block_results_m, "TT", "6")
```

---

# Maternal Arm Comparisons (TdaP vs TT)

## R² Summary Across All Blocks

```{r r2m-table, results='asis'}
r2_table_m <- bind_rows(r2_all_m) %>%
  arrange(block, model, stratum) %>%
  dplyr::select(stratum, block, model, n_obs, R2, adj_R2)

knitr::kable(as.data.frame(r2_table_m),
             caption = "Table M4. R² for all models across all blocks and maternal arms.",
             align = c("l","l","l","r","r","r")) %>% print()
```

```{r r2m-wide, results='asis'}
r2_wide_m <- r2_table_m %>%
  pivot_wider(names_from  = stratum,
              values_from = c(R2, adj_R2, n_obs),
              names_sep   = "_") %>%
  mutate(delta_R2 = round(R2_TdaP - R2_TT, 4)) %>%
  arrange(block, model)

knitr::kable(as.data.frame(r2_wide_m),
             caption = "Table M5. R² side-by-side: TdaP vs TT. delta_R2 = R²(TdaP) − R²(TT).",
             align = c("l","l","r","r","r","r","r","r","r")) %>% print()
```

```{r r2m-barplot, fig.cap="Figure M3. R² by block and maternal arm (antigen-level models).", fig.height=5}
r2_antigen_m <- r2_table_m %>%
  dplyr::filter(grepl("antigen", model)) %>%
  mutate(block_clean = gsub("_qty|_quantity|_switch|_ratio", "", block))

ggplot(r2_antigen_m, aes(x = block_clean, y = R2,
                          fill = stratum, group = stratum)) +
  geom_col(position = position_dodge(0.6), width = 0.5) +
  geom_line(aes(color = stratum), position = position_dodge(0.6),
            linewidth = 0.6, linetype = "dashed") +
  scale_fill_manual(values  = marm_colors) +
  scale_color_manual(values = marm_colors) +
  labs(x = "Block", y = "R²", fill = "Maternal arm", color = "Maternal arm",
       title = paste0("Explained variance (antigen-level models) — ",
                      outcome_label, " — maternal arm")) +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Ratio-Model Coefficient Comparison Plots

```{r ratiom-coef-plots, fig.cap="Figures M4–M6. Ratio-model coefficients (Blocks 2, 4, 6) side-by-side for TdaP and TT. Error bars = 95% CI.", fig.height=5}
block_labels_m <- c(
  B2 = "Block 2: Δ ratio P02→P09\n(class switching during pregnancy)",
  B4 = "Block 4: Δ ratio P09→M00\n(class-selective placental transfer)",
  B6 = "Block 6: Δ ratio M00→M02\n(class-selective neonatal decay)"
)

for (blk in c("B2","B4","B6")) {
  coef_rows <- lapply(arms, function(marm) {
    key <- paste0(blk, "_", marm)
    if (is.null(ratio_models_by_block_marm[[key]])) return(NULL)
    tidy(ratio_models_by_block_marm[[key]]) %>%
      dplyr::filter(term != "(Intercept)") %>%
      mutate(maternal_arm = marm)
  }) %>% bind_rows()

  if (nrow(coef_rows) == 0) next

  p <- ggplot(coef_rows,
              aes(x = term, y = estimate, fill = maternal_arm,
                  ymin = estimate - 1.96 * std.error,
                  ymax = estimate + 1.96 * std.error)) +
    geom_col(position = position_dodge(0.6), width = 0.5) +
    geom_errorbar(position = position_dodge(0.6), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_fill_manual(values = marm_colors) +
    labs(x     = "Pro-inflammatory/Tolerogenic Ratio",
         y     = paste0("β (→ ", outcome_label, " M02)"),
         title = paste0(block_labels_m[blk], "\n(outcome = ", outcome_label, ")"),
         fill  = "Maternal arm") +
    theme_bw(base_size = 11)
  print(p)
}
```

## Relative Importance Comparison (Full 12-Subclass Models)

```{r relimpm-comparison, results='asis'}
for (blk_nm in c("Block 1 (P09)", "Block 3 (transfer)", "Block 5 (decay)")) {
  rl <- switch(blk_nm,
    "Block 1 (P09)"      = relimp_b1_full_by_marm,
    "Block 3 (transfer)" = relimp_b3_full_by_marm,
    "Block 5 (decay)"    = relimp_b5_full_by_marm
  )
  if (is.null(rl[["TdaP"]]) || is.null(rl[["TT"]])) next
  ri_compare <- merge(
    data.frame(feature = names(rl[["TdaP"]]@lmg), lmg_TdaP = round(rl[["TdaP"]]@lmg*100, 2)),
    data.frame(feature = names(rl[["TT"]]@lmg),   lmg_TT   = round(rl[["TT"]]@lmg*100, 2)),
    by = "feature"
  ) %>%
    mutate(delta_lmg = round(lmg_TdaP - lmg_TT, 2)) %>%
    arrange(desc(abs(delta_lmg)))
  knitr::kable(
    ri_compare,
    caption = paste0("Relative importance (lmg %) — ", blk_nm,
                     " full 12-subclass model. delta_lmg = TdaP − TT."),
    align = c("l","r","r","r")
  ) %>% print()
  cat("\n")
}
```

---

# Synthesis and Interpretation

*This section is intended to be completed after reviewing the results above.
The scaffolding below provides the interpretive structure needed to integrate
findings across all six blocks.*

## 1. Which Block Explains the Most Variance?

> **[To complete]** Compare R² across all blocks (Table 5 above).
> Key question: Is the infant's PTNA at M02 more strongly determined by
> what the mother had (Block 1), how much was transferred (Block 3), or
> how much was preserved after birth (Block 5)? If Block 5 dominates, decay
> kinetics are the rate-limiting step; if Block 1 dominates, the maternal
> vaccination response is the primary determinant regardless of transfer or
> decay efficiency.

## 2. Dominant Predictors Within the Leading Block

> **[To complete]** Identify the subclass(es) with the highest lmg values
> in the full 12-subclass model of the leading block. Cross-reference with
> the antigen-level models: is PT antigen dominant (as in the previously
> described PTNA architecture), or does a different antigen emerge when
> the predictors are upstream measurements (maternal serum at P09)?

## 3. Are the Ratio-Based Effects (Blocks 2, 4, 6) Independent of Quantity?

> **[To complete]** The ratio models test whether subclass *balance*
> predicts PTNA independent of total level. A significant ratio predictor
> in Block 2 means gestational class switching adds information beyond raw
> IgG level. A significant predictor in Block 6 means subclass-selective
> decay shapes functional outcomes independently of how much was present
> at birth. Compare the R² of ratio models vs quantity models within each
> period.

## 4. Do aP and wP Infants Follow Different Predictive Pathways?

> **[To complete]** Use the delta_R2 column in Table 5 and the relative
> importance comparison tables to identify where the arms diverge most.
> Interpret arm differences in light of what is known about aP vs wP
> infant immune responses:
>
> - aP infants receive an acellular vaccine that induces a mixed Th1/Th2
>   profile and may interact differently with maternally derived IgG1 vs IgG2
> - wP infants receive a whole-cell vaccine that induces a stronger
>   pro-inflammatory Th1 response, which may interact more strongly with
>   maternally derived IgG1

## 5. Does Maternal Vaccine (TdaP vs TT) Change the Predictive Chain?

> **[To complete]** Compare the maternal-arm results (Tables M4–M5 and the
> maternal block sections) against the infant-arm results. TT mothers
> received no pertussis antigen, so a strong contrast is expected at the
> source: the maternal P09 PT-IgG signal that dominates Block 1 should be
> far weaker (or absent) in TT mothers, since their PT-specific response
> was not boosted. Key questions:
>
> - Is the Block 1 P09 PT-IgG → infant PTNA association present in TdaP
>   mothers but attenuated or null in TT mothers? That would confirm the
>   maternal pertussis booster is what installs the predictive signal.
> - Do the transfer and decay blocks (3–6) differ by maternal arm, or are
>   they vaccine-independent kinetic processes that behave the same in both?
> - Do the maternal-arm and infant-arm stratifications point to the same
>   dominant antigen/subclass, or do they isolate different parts of the
>   mechanism (maternal arm = supply; infant arm = downstream handling)?

## 6. Mechanistic Model

> **[To complete]** Assemble the cross-block findings into a narrative
> pathway diagram:
>
> ```
> Maternal vaccination
>       ↓ (Block 1: quantity — how much is produced?)
> Maternal IgG at delivery (P09)
>       ↓ (Blocks 3 & 4: transfer — how much/what composition crosses the placenta?)
> Cord blood IgG at birth (M00)
>       ↓ (Blocks 5 & 6: decay — how much/what composition survives to M02?)
> Infant IgG available at M02
>       ↓
> PTNA at M02 (functional outcome)
> ```
>
> At which step is the signal most amplified or most attenuated?
> Which subclass (IgG1? IgG3?) carries the signal most faithfully
> through the full chain?

## 7. Comparison with the M02/M05/M09 Concurrent Architecture

> **[To complete]** The prior PTNA analysis (within-infant-visit analysis)
> showed that PT_IgG1 is the dominant subclass predictor of PTNA when
> measuring both at the same time. Does that same subclass dominate the
> predictive pathway when the predictors are measured earlier in the causal
> chain (maternal serum → cord → infant)? Convergence would strengthen the
> mechanistic model; divergence would suggest that the maternal and
> infant-period mechanisms are distinct.

---

*Document generated by `predict_M02_PTNA_from_maternal_v1.Rmd`.  
Adapt `params$outcome_feature` in the YAML header to run for
`WHOLE_SBA` or `WHOLE_WT_IgG`.*
