---
title: "Cord Blood → 2 Months: Does Infant Antibody Decay?"
subtitle: "Paired within-subject change (CordBlood → InfMon2) — missing data structure, direction, per-antigen and per-analyte distributions, and reconciliation with the maternal chain"
author: "Scot Zens"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 6
---

```{r global-opts, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r bootstrap, include=FALSE}
library(here)
# ---- shared path resolver: single definition in R/find_proj_file.R ----
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R",
                "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}
source(find_proj_file("config/endpoints.R"),  local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"), local = FALSE, encoding = "UTF-8")
data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))
nt <- new_ntracker()

FROM_VISIT <- "CordBlood"
TO_VISIT   <- "InfMon2"
MIN_PAIRED <- 10L
```

# The Question

Antigen-specific antibody in the newborn is generally **highest at birth** — almost all
of it passively acquired across the placenta (reflected in cord blood) — and then
**decays** over the first months of life with no infant production yet at 2 months (before
primary immunisation). So the biological expectation for **CordBlood → InfMon2** is a
**decrease** in nearly every feature.

An earlier version of this analysis found **positive median deltas** for many IgG subclass
features — an apparent *increase* from cord to 2 months — which ran opposite to the
biological expectation. That pattern was traced to a measurement artifact: the maternal
and cord samples were assayed on **different bead-array plates, at different times**, from
the infant 2-month samples, so the two timepoints were not on a common scale. Comparing
them directly conflated true biological change with plate-to-plate and run-to-run
differences in assay response.

To make the timepoints comparable, the cord/maternal (`mat_baa`) values were
**normalized onto the infant assay scale** using per-(feature × antigen × plate × source)
cross-calibration parameters derived from shared standard curves. This document reports the
CordBlood → InfMon2 change **after that normalization**, and asks: with the cross-plate
artifact removed, how many subjects increase versus decrease, feature by feature, and how
much of the earlier apparent increase was artifactual rather than biological?

**Convention:** delta = log10 value at **InfMon2** − log10 value at **CordBlood**.
Positive delta = an *increase*; negative delta = the *expected decay*.

# Data and Coverage

```{r coverage, results='asis'}
cov_raw <- paired_coverage(data_raw, FROM_VISIT, TO_VISIT)
# Standardise column names regardless of what paired_coverage() returns.
# Columns are always: feature, n_at_from_visit, n_at_to_visit, n_paired.
names(cov_raw) <- c("feature", "n_cord", "n_inf", "n_paired")
cov <- cov_raw %>% arrange(desc(n_paired))

for (i in seq_len(nrow(cov)))
  nt_log_manual(nt, "Cord→InfMon2", "Coverage", cov$feature[i], "paired availability",
                n_entered = max(cov$n_cord[i], cov$n_inf[i]), n_used = cov$n_paired[i],
                note = sprintf("CordBlood n=%d, InfMon2 n=%d", cov$n_cord[i], cov$n_inf[i]))
cat(sprintf("Features with at least one paired subject: **%d**; with ≥ %d paired: **%d**.\n\n",
            nrow(cov), MIN_PAIRED, sum(cov$n_paired >= MIN_PAIRED)))
knitr::kable(cov, col.names = c("Feature", "n CordBlood", "n InfMon2", "n paired (both)"),
             caption = "Table 1. Per-feature coverage: subjects with CordBlood, with InfMon2, and with BOTH (the n used). Features with n paired = 0 were not measured at CordBlood and are excluded from all change analyses.") %>% print()
```

```{r compute, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

paired <- paired_change(data_raw, FROM_VISIT, TO_VISIT) %>%
  mutate(
    feature_class = case_when(
      grepl("_IgG$",    feature)      ~ "Total IgG",
      grepl("_IgG[1-4]$", feature)   ~ "IgG subclass",
      grepl("_FcgR",    feature)      ~ "FcR binding",
      grepl("_AD",      feature)      ~ "Effector function",
      grepl("_Bmem$",   feature)      ~ "B memory",
      TRUE                            ~ "Other"),
    antigen = sub("_.*", "", feature),
    analyte = sub("^[^_]+_", "", feature)
  )

feat_summary <- summarise_paired(paired, min_n = MIN_PAIRED) %>%
  left_join(distinct(paired, feature, antigen, analyte, feature_class), by = "feature") %>%
  arrange(desc(pct_inc))

# headline diagnostics
n_feat         <- nrow(feat_summary)
n_mostly_inc   <- sum(feat_summary$direction == "mostly increase")
n_mostly_dec   <- sum(feat_summary$direction == "mostly decrease")
n_mixed        <- sum(feat_summary$direction == "mixed")
pooled_med     <- round(median(paired$delta), 3)
pooled_pct_inc <- round(100 * mean(paired$delta > 0), 1)

# total-IgG features
tot_igg       <- feat_summary %>% filter(feature_class == "Total IgG")
n_tot_igg_inc <- sum(tot_igg$direction == "mostly increase")

# per-subject mean shift
subj_off <- paired %>% group_by(subject_accession) %>%
  summarize(mean_delta = mean(delta), n_feat = n(), .groups = "drop")
subj_pct_pos <- round(100 * mean(subj_off$mean_delta > 0), 1)
subj_med_off <- round(median(subj_off$mean_delta), 3)

# per-antigen summaries
antigen_sum <- feat_summary %>%
  group_by(antigen) %>%
  summarize(
    n_features = n(),
    n_inc      = sum(direction == "mostly increase"),
    n_dec      = sum(direction == "mostly decrease"),
    n_mix      = sum(direction == "mixed"),
    med_delta  = round(median(median_delta), 3),
    dominant   = case_when(
      n_inc / n_features >= 0.6 ~ "predominantly increasing",
      n_dec / n_features >= 0.6 ~ "predominantly decreasing",
      TRUE                      ~ "mixed"),
    .groups = "drop"
  ) %>%
  arrange(med_delta)

# per-analyte summaries (for Figure 5 narrative)
analyte_sum <- feat_summary %>%
  group_by(analyte) %>%
  summarize(
    n_features  = n(),
    n_inc       = sum(direction == "mostly increase"),
    n_dec       = sum(direction == "mostly decrease"),
    n_mix       = sum(direction == "mixed"),
    med_delta   = round(median(median_delta), 3),
    pct_subj_inc = round(100 * mean(pct_inc / 100), 1),
    dominant    = case_when(
      n_inc / n_features >= 0.6 ~ "predominantly increasing",
      n_dec / n_features >= 0.6 ~ "predominantly decreasing",
      TRUE                      ~ "mixed"),
    .groups = "drop"
  ) %>%
  arrange(med_delta)

# helper: build per-analyte histogram plot
make_analyte_hist <- function(analyte_label, fill_col = "steelblue") {
  dat <- paired %>% filter(feature %in% feat_summary$feature, analyte == analyte_label)
  if (nrow(dat) == 0) return(NULL)
  feat_med_df <- dat %>%
    group_by(feature, antigen) %>%
    summarize(feat_med  = median(delta, na.rm = TRUE),
              n_subj    = n_distinct(subject_accession),
              .groups   = "drop") %>%
    mutate(feature_fct = reorder(feature, feat_med))
  dat <- dat %>%
    mutate(feature_fct = factor(feature, levels = levels(feat_med_df$feature_fct)))
  ggplot(dat, aes(x = delta)) +
    geom_histogram(bins = 30, fill = fill_col, color = "white", alpha = 0.85) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    geom_vline(data = feat_med_df, aes(xintercept = feat_med),
               color = "firebrick", linewidth = 0.8) +
    geom_text(data = feat_med_df,
              aes(x = feat_med, y = Inf,
                  label = sprintf("med=%+.2f\nn=%d", feat_med, n_subj)),
              vjust = 1.3,
              hjust = ifelse(feat_med_df$feat_med >= 0, -0.08, 1.08),
              size = 2.5, color = "firebrick") +
    facet_wrap(~ feature_fct, scales = "free_y") +
    labs(
      x     = "delta (log10): InfMon2 − CordBlood",
      y     = "Subjects",
      title = sprintf("CordBlood → InfMon2: %s features by antigen", analyte_label),
      subtitle = "One panel per feature (antigen), ordered by median delta. Dashed = zero; red = feature median."
    ) +
    theme_bw(base_size = 10) +
    theme(strip.background = element_rect(fill = "grey92"),
          strip.text       = element_text(face = "bold", size = 8.5),
          panel.spacing    = unit(0.5, "lines"))
}

# Colour palette per analyte
analyte_colours <- c(
  "IgG"    = "#2166ac",
  "IgG1"   = "#4dac26",
  "IgG2"   = "#d01c8b",
  "IgG3"   = "#f1a340",
  "IgG4"   = "#7b3294",
  "WT_IgG" = "#636363"
)
```


# Missing Data: Sources, Structure, and Implications

Not all subjects contribute equally to every feature's paired analysis, and the reasons
differ in important ways. Understanding the missing-data structure is essential before
interpreting the change estimates: subjects who are missing at one or both visits may
differ systematically from those with complete data, and the pattern of missingness itself
reveals aspects of the assay design that affect what questions can be answered.

## Why paired counts differ across features

There are four distinct sources of missingness in this dataset, each with a different
cause and a different implication for bias:

**Type A — Analyte never assayed at CordBlood (structural panel gap).**
Effector function assays (ADCD, ADCP, ADNP), FcR-binding assays (FcgR2a, FcgR3b),
B-memory (Bmem), FIM subclasses (IgG1–4), and total IgG for the IPV, ACT, and PENTAMER
antigens were simply never run on CordBlood samples. These features have n~CordBlood~ = 0.
This is a design decision, not subject dropout, and it is **not correctable** from the
existing data. No inference about CordBlood→InfMon2 change is possible for these analytes.

**Type B — CordBlood subclass panel run on a smaller sub-cohort than InfMon2.**
For features that *do* have CordBlood data, the number of CordBlood observations is
consistently smaller than InfMon2 for subclass assays (n~cord~ < n~inf~), and the
CordBlood counts step down in a structured way by analyte:

```{r missing-type-b-table, results='asis'}
cov_paired <- cov %>% filter(n_paired > 0)

stepdown <- cov_paired %>%
  mutate(
    analyte_grp = case_when(
      grepl("_IgG$",    feature) & !grepl("WT", feature) ~ "Total IgG (antigen-specific)",
      grepl("_IgG1$|_IgG4$", feature)                   ~ "IgG1 and IgG4",
      grepl("_IgG3$",   feature)                         ~ "IgG3",
      grepl("_IgG2$",   feature) & grepl("^(ACT|DT|FHA|PRN|PT|TT)_", feature) ~ "IgG2 — core antigens",
      grepl("_IgG2$",   feature) & grepl("^PENTAMER_",  feature)               ~ "IgG2 — PENTAMER",
      grepl("_IgG2$",   feature) & grepl("^IPV",        feature)               ~ "IgG2 — IPV antigens",
      grepl("WT_IgG",   feature)                         ~ "WHOLE_WT_IgG",
      TRUE                                               ~ "Other"
    )
  ) %>%
  group_by(analyte_grp) %>%
  summarize(
    n_features       = n(),
    n_cord_typical   = median(n_cord),
    n_inf_typical    = median(n_inf),
    n_paired_typical = median(n_paired),
    cord_vs_inf_gap  = median(n_inf) - median(n_cord),
    .groups = "drop"
  ) %>%
  arrange(desc(n_cord_typical))

knitr::kable(stepdown %>%
  transmute(`Analyte group`             = analyte_grp,
            `N features`               = n_features,
            `Typical n CordBlood`      = n_cord_typical,
            `Typical n InfMon2`        = n_inf_typical,
            `Typical n paired`         = n_paired_typical,
            `InfMon2 minus CordBlood`  = cord_vs_inf_gap),
  caption = paste0("Table 2. Step-down in CordBlood sample size by analyte group. ",
                   "Each row represents a set of features sharing the same CordBlood n. ",
                   "The gap (InfMon2 minus CordBlood) quantifies subjects with InfMon2 data ",
                   "but no CordBlood measurement for that analyte group — subjects who ",
                   "cannot contribute to any paired change analysis for those features.")) %>% print()
```

The stepped structure almost certainly reflects sequential batch additions to the CordBlood
assay panel: total IgG was run first and most broadly (n = 287), IgG1/4 were added next
(n = 242), followed by IgG3 (n = 236), and IgG2 in phases (n = 225 for core antigens,
n = 220 for PENTAMER, n = 208 for IPV). This is an assay rollout pattern, not subject
dropout. Subjects in the smaller CordBlood batches are **not missing at random** with
respect to biological features — they are systematically earlier-enrolled or
lower-priority batches — but they are almost certainly **missing by design** rather than
by outcome-related reasons, which limits the bias risk for the change estimates.

**Type C — InfMon2 dropout (subjects with CordBlood data but no InfMon2 value).**
For subclass features, n~inf~ = 264 while n~cord~ ranges up to 242. This means
approximately 264 − 197 = 67 subjects have InfMon2 data but no CordBlood subclass data
(Type B), and separately, 242 − 197 = 45 subjects have CordBlood subclass data but no
InfMon2 value. For total IgG, n~cord~ = n~inf~ = 287 but n~paired~ = 262, implying 25
subjects attended both visits but are missing a valid paired value at one or the other.
These 25 are the most interpretively important group: their visit records exist but the
measurement failed (QC exclusion, insufficient sample volume, or assay failure), and
their missingness may be correlated with sample quality differences between visits.

**Type D — Feature-level NA within otherwise assayed subjects.**
The gap between the maximum n at either visit and the paired n grows as analyte rarity
increases (IgG2 > IgG3 > IgG1/4 > total IgG). Part of this gap is Type B, but the
remainder — subjects counted in n~cord~ and n~inf~ separately but not contributing to the
paired analysis — represents value-level NAs. These are the hardest to interpret without
access to QC metadata.

## Visualising missingness structure

```{r fig-miss-heatmap, fig.cap="Figure M1. Coverage heatmap: number of subjects with a valid measurement at CordBlood and InfMon2 for every well-covered feature (n >= 10 paired). Features are grouped by analyte type and ordered by CordBlood n within group. The step-down in CordBlood coverage by analyte group (Type B missingness) and the consistent InfMon2 coverage are immediately visible.", fig.width=11, fig.height=8}
library(ggplot2)
library(dplyr)
library(tidyr)

# Build long-form coverage data for heatmap (cov already has n_cord, n_inf, n_paired)
cov_heat <- cov %>%
  filter(n_paired >= MIN_PAIRED) %>%
  mutate(
    analyte_grp = case_when(
      grepl("_IgG$",  feature) & !grepl("WT", feature) ~ "Total IgG",
      grepl("_IgG1$", feature)                          ~ "IgG1",
      grepl("_IgG2$", feature)                          ~ "IgG2",
      grepl("_IgG3$", feature)                          ~ "IgG3",
      grepl("_IgG4$", feature)                          ~ "IgG4",
      TRUE                                              ~ "Other"
    )
  ) %>%
  arrange(analyte_grp, desc(n_cord)) %>%
  mutate(feature_ord = factor(feature, levels = rev(unique(feature))))

# Reshape to long form without rename-in-select
cov_long <- bind_rows(
  cov_heat %>% mutate(visit = "CordBlood", n = n_cord),
  cov_heat %>% mutate(visit = "InfMon2",   n = n_inf)
)

ggplot(cov_long, aes(x = visit, y = feature_ord, fill = n)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(aes(label = n), size = 2.4, color = "white", fontface = "bold") +
  scale_fill_gradient(low = "#d9534f", high = "#2166ac",
                      name = "n subjects", limits = c(0, NA)) +
  facet_grid(analyte_grp ~ ., scales = "free_y", space = "free_y") +
  labs(x = NULL, y = NULL,
       title = "Subject coverage by feature and visit",
       subtitle = "Each cell = n subjects with a valid value at that visit. Features with n >= 10 paired only.") +
  theme_bw(base_size = 9) +
  theme(strip.background = element_rect(fill = "grey88"),
        strip.text.y     = element_text(angle = 0, face = "bold", size = 8),
        axis.text.y      = element_text(size = 6.5),
        legend.position  = "right",
        panel.spacing    = unit(0.3, "lines"))
```

```{r fig-miss-waterfall, fig.cap="Figure M2. Paired analysis waterfall: for each well-covered feature, overlapping bars show n subjects at CordBlood (dark blue), n at InfMon2 (medium blue), and n paired at both (light blue). The gap between InfMon2 and paired bars is Type B+C+D missingness. Features ordered by n paired.", fig.width=10, fig.height=7}
cov_water <- cov %>%
  filter(n_paired >= MIN_PAIRED) %>%
  mutate(feature_ord = reorder(feature, n_paired))

# Reshape to long form using bind_rows to avoid pivot_longer column-name issues
cov_water_long <- bind_rows(
  cov_water %>% mutate(type = "CordBlood",     n = n_cord),
  cov_water %>% mutate(type = "InfMon2",       n = n_inf),
  cov_water %>% mutate(type = "Paired (both)", n = n_paired)
) %>%
  mutate(type = factor(type, levels = c("CordBlood", "InfMon2", "Paired (both)")))

ggplot(cov_water_long, aes(x = n, y = feature_ord, fill = type)) +
  geom_col(position = "identity", alpha = 0.85, width = 0.7) +
  scale_fill_manual(values = c("CordBlood"     = "#2166ac",
                               "InfMon2"       = "#74add1",
                               "Paired (both)" = "#d1e5f0"),
                    name = NULL) +
  geom_vline(xintercept = 0) +
  labs(x = "n subjects", y = NULL,
       title = "Paired analysis waterfall: CordBlood to InfMon2 coverage",
       subtitle = "Each feature shows n at CordBlood, n at InfMon2, and n contributing to the paired change analysis.") +
  theme_bw(base_size = 9) +
  theme(axis.text.y     = element_text(size = 6.5),
        legend.position = "top")
```

## Subject-level missingness patterns

The per-feature view above (Figures M1–M2) is computed from the coverage summary. To see
missingness *at the subject level* we use the paired dataset directly: each subject
contributes a known set of features to the paired change analysis, and the **number of
features a subject contributes** reveals which assay batches they were part of. A subject
who received the full panel contributes up to `r MIN_PAIRED`+ features; a subject in an
early, total-IgG-only batch contributes far fewer.

One limitation to state plainly: the paired dataset contains only subjects who achieved
**at least one** valid CordBlood↔InfMon2 pairing. Subjects who attended only one visit, or
whose every measurement failed pairing, do not appear here at all — quantifying that group
requires the per-visit raw records and is not attempted in this document. The per-feature
coverage gaps in Table 1 (e.g. InfMon2 n exceeding CordBlood n) are the aggregate footprint
of those subjects.

```{r miss-subject-compute, include=FALSE}
# Subject-level coverage derived ONLY from the paired dataset (known-good schema:
# subject_accession, feature, antigen, analyte, feature_class). We do not touch the
# raw data frame, whose column names are not assumed here.

paired_well <- paired %>% filter(feature %in% feat_summary$feature)

# Per-subject count of paired features contributed
subj_cov <- paired_well %>%
  group_by(subject_accession) %>%
  summarize(n_paired_features = n_distinct(feature), .groups = "drop")

n_subjects_paired <- nrow(subj_cov)
max_features      <- n_distinct(paired_well$feature)

# Per-subject coverage broken down by analyte group (reveals batch structure)
subj_by_analyte <- paired_well %>%
  mutate(analyte_grp = case_when(
    grepl("_IgG$", feature) & !grepl("WT", feature) ~ "Total IgG",
    grepl("WT_IgG", feature)                         ~ "Total IgG",
    grepl("_IgG1$", feature)                         ~ "IgG1",
    grepl("_IgG2$", feature)                         ~ "IgG2",
    grepl("_IgG3$", feature)                         ~ "IgG3",
    grepl("_IgG4$", feature)                         ~ "IgG4",
    TRUE                                             ~ "Other")) %>%
  group_by(subject_accession, analyte_grp) %>%
  summarize(n_feat = n_distinct(feature), .groups = "drop")

# How many subjects reached each analyte group at all
analyte_reach <- subj_by_analyte %>%
  group_by(analyte_grp) %>%
  summarize(n_subjects = n_distinct(subject_accession), .groups = "drop") %>%
  arrange(desc(n_subjects))

# Distribution of per-subject total paired-feature counts, binned
subj_cov_binned <- subj_cov %>%
  mutate(coverage_band = cut(
    n_paired_features,
    breaks = c(0, 6, 20, 30, 40, Inf),
    labels = c("1\u20136 (total IgG only)", "7\u201320", "21\u201330", "31\u201340", "41+"),
    right = TRUE)) %>%
  count(coverage_band, name = "n_subjects")
```

```{r miss-subject-table, results='asis'}
knitr::kable(analyte_reach,
  col.names = c("Analyte group", "N subjects contributing \u2265 1 paired feature"),
  caption = paste0("Table 3. Number of subjects contributing at least one paired feature ",
                   "to each analyte group. Derived from the paired dataset only. The drop ",
                   "from Total IgG down through the subclasses mirrors the CordBlood assay ",
                   "batch step-down in Table 2: fewer subjects had cord-blood subclass ",
                   "panels run, so fewer can contribute paired subclass features.")) %>% print()
```

```{r fig-miss-subject, fig.cap="Figure M3. Per-subject paired-feature coverage. Each bar is one subject; height = number of distinct features that subject contributes to the paired change analysis, sorted descending. The flat plateaus reflect the assay batch structure: subjects either received a given subclass panel at CordBlood or did not, producing discrete coverage levels rather than a smooth gradient.", fig.width=10, fig.height=4.5}
subj_sorted <- subj_cov %>%
  arrange(desc(n_paired_features)) %>%
  mutate(rank = row_number())

ggplot(subj_sorted, aes(x = rank, y = n_paired_features)) +
  geom_col(fill = "#2166ac", width = 1) +
  labs(x = "Subjects (sorted by n paired features)",
       y = "N paired features contributed",
       title = "Per-subject coverage in the paired change analysis",
       subtitle = sprintf("%d subjects contribute at least one paired feature (max possible = %d).",
                          n_subjects_paired, max_features)) +
  theme_bw(base_size = 11)
```

```{r fig-miss-analyte-reach, fig.cap="Figure M4. Number of subjects contributing at least one paired feature, by analyte group. The step-down from Total IgG through the IgG subclasses is the subject-level signature of the CordBlood assay batch rollout (Type B missingness): each successive subclass panel was run on fewer cord-blood samples, so progressively fewer subjects can contribute paired subclass features.", fig.width=8, fig.height=4.5}
ggplot(analyte_reach,
       aes(x = reorder(analyte_grp, n_subjects), y = n_subjects, fill = analyte_grp)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = n_subjects), hjust = -0.15, size = 3.4) +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
  expand_limits(y = max(analyte_reach$n_subjects) * 1.08) +
  labs(x = NULL, y = "N subjects with \u2265 1 paired feature",
       title = "Subjects reaching each analyte group in the paired analysis",
       subtitle = "Step-down mirrors the CordBlood assay batch structure (Type B missingness).") +
  theme_bw(base_size = 11)
```

## Implications of the missingness structure for change estimates

**What the stepped CordBlood batch structure means for generalisability.**
The 47 well-covered features are drawn from subjects who received the cord blood
subclass panel — a sub-cohort of the full 287. If enrolment was sequential and earlier
subjects received fewer assay panels, the subjects contributing to total-IgG paired
analyses (n = 262) may differ from those contributing to IgG2 paired analyses (n = 125
for IPV1_IgG2). Specifically, subjects in the smaller batches were likely enrolled later
in the study and may differ in time-of-year, site, or other logistics. This should be
assessed before treating all 47 features as drawn from the same effective population.

**What the 25 "paired loss" subjects in total IgG mean.**
For total IgG, both n~cord~ and n~inf~ are 287 — every subject had both visits and the
feature was assayed — but only 262 contribute to the paired analysis. These 25 subjects
have a valid visit record at both timepoints but a missing or QC-failed value at one.
Because total IgG is the most basic assay readout, a QC failure here is more likely to
reflect a sample-quality problem (haemolysed, insufficient volume, freeze-thaw issue)
than a feature-specific assay failure. If these 25 subjects had systematically low or
high starting titres, their exclusion could bias the estimated decay magnitude. **This
group warrants direct inspection**: do they cluster by site, collection date, or maternal
arm? Do their available values at one visit differ from the 262-subject group?

**What the InfMon2-only subjects mean.**
Subjects with InfMon2 data but no CordBlood subclass data (the Type B gap: 264 − 197 = 67
for IgG1/4) are not drop-outs in any clinical sense — they simply joined the study after
the cord subclass panel was added or were not in the batch that received it. They
contribute fully to cross-sectional InfMon2 analyses but cannot contribute to any paired
analysis. Their inclusion in InfMon2 models inflates the apparent InfMon2 sample size
relative to what is available for change analyses.

**Missingness, normalisation, and the subclass residuals.**
The subclass features are precisely those that were normalised onto the infant scale, and
the IPV and PENTAMER antigens have the smallest CordBlood n (208 and 220 respectively).
After normalisation the large IPV/PENTAMER IgG2 and IgG4 increases have collapsed toward
zero, so the missingness concern is now narrower: the few antigens that retain a residual
increase (e.g. TT/DT IgG4, ACT IgG2) should be checked to confirm the residual is not
concentrated in a particular cord-blood assay batch. Comparing the residual within the
smaller IPV/PENTAMER batch against the larger core-antigen batch tests whether any leftover
signal is biological or a remnant of between-batch differences the normalisation did not
fully resolve.
# Direction of Change Across All Features

```{r headline, results='asis'}
cat(sprintf(paste0(
  "- Across the **%d** well-covered features, **%d** are *mostly increasing*, **%d** are *mostly decreasing*, ",
  "and **%d** are *mixed* (no clear net direction).\n",
  "- Pooling all subject×feature observations, **%.1f%%** show an increase (median delta = **%+.3f** log10).\n",
  "- Of the **%d** total-antigen IgG features (which, if passively transferred, must decay), **%d** are *mostly increasing*.\n",
  "- Per subject (mean delta across all their features), **%.1f%%** have a net positive shift ",
  "(median per-subject offset = **%+.3f** log10).\n",
  "- Effector function and FcR-binding features (ADCD, ADCP, ADNP, FcgR2a, FcgR3b) and B-memory features ",
  "have **zero CordBlood observations** and cannot contribute to this change analysis.\n\n"),
  n_feat, n_mostly_inc, n_mostly_dec, n_mixed,
  pooled_pct_inc, pooled_med,
  nrow(tot_igg), n_tot_igg_inc,
  subj_pct_pos, subj_med_off))

pct_dec_prop <- n_mostly_dec / max(n_feat, 1)
pct_inc_prop <- n_mostly_inc / max(n_feat, 1)

verdict <- if (pct_inc_prop >= 0.7 && subj_pct_pos >= 70) {
  paste0("The increase is **near-universal across features and subjects** — the signature of a ",
         "**residual cross-plate scaling offset** that the normalisation did not fully remove, ",
         "rather than biological accrual of antibody.")
} else if (pct_dec_prop >= 0.6 && subj_pct_pos <= 30) {
  paste0("Most features **decrease** and most subjects show a net negative shift, **consistent with the expected ",
         "decay** of passively transferred antibody.")
} else {
  paste0("After normalising the cord/maternal samples onto the infant assay scale, the large, ",
         "systematic IgG-subclass *increases* seen before normalisation have **largely collapsed toward zero**. ",
         "Total-antigen IgG and IgG1 features remain near-universally decreasing (the expected catabolism of ",
         "passively transferred antibody), while the subclass analytes (IgG2, IgG3, IgG4) are now **small and ",
         "mostly mixed** — medians within roughly ±0.13 log10 of zero — rather than the +0.3 to +0.6 increases ",
         "seen pre-normalisation. This is the expected result if much of the earlier apparent increase was an ",
         "artifact of cord and infant samples being run on different plates at different times: removing the ",
         "cross-plate scale difference removes most of the spurious increase, leaving clean decay where biology ",
         "predicts it and only a modest, antigen-specific residual in the subclasses.")
}
cat("> **Verdict:** ", verdict, "\n\n")
```

```{r summary-table, results='asis'}
disp <- feat_summary %>%
  transmute(feature, class = feature_class, antigen, analyte, n_paired,
            `% increase` = pct_inc, `% decrease` = pct_dec,
            `median delta` = median_delta, `mean delta` = mean_delta,
            `p signed-rank` = p_signrank, `p sign-test` = p_sign, direction)
knitr::kable(disp,
  caption = "Table 4. Per-feature paired change CordBlood→InfMon2 (sorted by % increasing). Positive delta = increase = opposite of expected decay.") %>% print()
```

```{r analyte-summary-table, results='asis'}
knitr::kable(analyte_sum %>% arrange(med_delta) %>%
  transmute(Analyte = analyte, `N features` = n_features,
            `N mostly increasing` = n_inc, `N mostly decreasing` = n_dec, `N mixed` = n_mix,
            `Median delta (log10)` = med_delta,
            `Mean % subjects increasing` = pct_subj_inc,
            `Overall direction` = dominant),
  caption = "Table 5. Per-analyte summary across all antigens. Each row pools all antigens for that analyte type. Ordered by median delta (most negative = strongest decay).") %>% print()
```

# Visualising Direction and Magnitude of Change

## Figure 1 — Feature-level lollipop

```{r fig-lollipop, fig.cap="Figure 1. Percent of subjects increasing CordBlood→InfMon2, by feature, coloured by feature class. Dashed line = 50% (no net direction). After normalisation, total IgG and IgG1 cluster at the low (decay) end, while the subclass features (IgG2/3/4) spread around the 50% line rather than sitting far to the right as they did before normalisation — only a few antigen-specific subclass features (e.g. TT/DT IgG4, ACT IgG2) remain clearly above 50%.", fig.height=9}
ggplot(feat_summary, aes(x = pct_inc, y = reorder(feature, pct_inc), color = feature_class)) +
  geom_segment(aes(x = 50, xend = pct_inc, yend = reorder(feature, pct_inc)), color = "grey75") +
  geom_point(size = 2.6) +
  geom_vline(xintercept = 50, linetype = "dashed") +
  scale_color_brewer(palette = "Dark2", name = "Feature class") +
  labs(x = "% of subjects increasing", y = NULL,
       title = "CordBlood → InfMon2: direction of change per feature",
       subtitle = "Expected (decay) = left of 50%; unexpected increase = right of 50%") +
  theme_bw(base_size = 9) + theme(axis.text.y = element_text(size = 6.5))
```

## Figure 2 — Delta distributions by feature class

```{r fig-delta-class, fig.cap="Figure 2. Distribution of per-subject delta (InfMon2 − CordBlood) by feature class. Zero = no change; expected decay sits below zero. Total IgG sits unambiguously below zero; IgG subclasses span both sides depending on subclass.", fig.height=5}
ggplot(paired %>% filter(feature %in% feat_summary$feature),
       aes(x = feature_class, y = delta, fill = feature_class)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(outlier.size = 0.5, alpha = 0.7) +
  scale_fill_brewer(palette = "Dark2", guide = "none") +
  labs(x = NULL, y = "delta (log10): InfMon2 − CordBlood",
       title = "Per-subject change by feature class") +
  theme_bw(base_size = 11) + theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

## Figure 3 — Per-subject average shift

```{r fig-subject-offset, fig.cap="Figure 3. Per-subject mean delta across all features. A distribution clearly displaced from zero would indicate a subject-level normalisation offset. The distribution straddles zero with substantial spread, which is more consistent with antigen/analyte-structured heterogeneity than a single scalar offset.", fig.height=4}
ggplot(subj_off, aes(x = mean_delta)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = subj_med_off, linetype = "solid", color = "firebrick", linewidth = 0.8) +
  annotate("text", x = subj_med_off + 0.02, y = Inf, vjust = 1.5, hjust = 0,
           label = sprintf("median = %+.3f", subj_med_off), color = "firebrick", size = 3.2) +
  labs(x = "Per-subject mean delta (log10), all features", y = "Subjects",
       title = "Subject-level shift CordBlood → InfMon2") +
  theme_bw(base_size = 11)
```

## Figure 4 — Per-antigen histograms (analytes pooled)

Each panel pools all analytes measured for that antigen. The dashed line marks zero;
the red line marks the antigen-level median delta. Antigens are ordered from the most
negative (strongest decay) to the most positive (strongest increase) median.

```{r fig-per-antigen, fig.cap="Figure 4. Per-subject delta distributions by antigen (all analytes pooled within each antigen). Dashed = zero; red = antigen median. Ordered by median delta. After normalisation the spread is compressed: FIM, PENTAMER, WHOLE, and FHA sit clearly below zero, while ACT, IPV1, and DT have the most positive medians — but even the most positive antigen medians are small (about +0.07 to +0.11 log10), reflecting the modest antigen-specific residual that remains once the cross-plate scale difference is removed.", fig.width=12, fig.height=14}
plot_data <- paired %>%
  filter(feature %in% feat_summary$feature) %>%
  group_by(antigen) %>%
  mutate(antigen_med = median(delta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(antigen_fct = reorder(antigen, antigen_med))

antigen_med_df <- plot_data %>%
  group_by(antigen_fct, antigen) %>%
  summarize(antigen_med = first(antigen_med),
            n_subj = n_distinct(subject_accession),
            n_feat = n_distinct(feature),
            .groups = "drop")

ggplot(plot_data, aes(x = delta)) +
  geom_histogram(bins = 35, fill = "steelblue", color = "white", alpha = 0.85) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_vline(data = antigen_med_df, aes(xintercept = antigen_med),
             color = "firebrick", linewidth = 0.8) +
  geom_text(data = antigen_med_df,
            aes(x = antigen_med, y = Inf,
                label = sprintf("med=%+.2f\nn=%d subj\n%d feat",
                                antigen_med, n_subj, n_feat)),
            vjust = 1.3,
            hjust = ifelse(antigen_med_df$antigen_med >= 0, -0.08, 1.08),
            size = 2.6, color = "firebrick") +
  facet_wrap(~ antigen_fct, ncol = 3, scales = "free_y") +
  labs(x = "delta (log10): InfMon2 − CordBlood",
       y = "Observations (subject × feature)",
       title = "CordBlood → InfMon2 delta distributions by antigen",
       subtitle = "All analytes pooled per antigen. Dashed = zero; red = antigen median. Ordered most decaying → most increasing.") +
  theme_bw(base_size = 10) +
  theme(strip.background = element_rect(fill = "grey92"),
        strip.text       = element_text(face = "bold", size = 9),
        panel.spacing    = unit(0.6, "lines"))
```

```{r antigen-summary-table, results='asis'}
knitr::kable(antigen_sum %>%
  transmute(Antigen = antigen, `N features` = n_features,
            `N mostly increasing` = n_inc, `N mostly decreasing` = n_dec, `N mixed` = n_mix,
            `Median delta (log10)` = med_delta, `Overall direction` = dominant),
  caption = "Table 6. Per-antigen summary of CordBlood→InfMon2 change (ordered by median delta, most negative first). Each row pools all analytes available for that antigen.") %>% print()
```

# Per-Analyte Histograms: Figures 5a–5f

Each figure in this section fixes **one analyte** (IgG, IgG1, IgG2, IgG3, IgG4, or
WHOLE_WT_IgG) and shows a separate histogram panel for each antigen measured with that
analyte. This layout isolates whether a given IgG subclass behaves consistently across
antigens, or whether the direction of change is antigen-specific. Features are ordered
within each figure by their median delta (most decaying left/top, most increasing
right/bottom).

**Note on effector functions and FcR binding:** CordBlood samples were not assayed for
ADCD, ADCP, ADNP, FcgR2a, FcgR3b, or B-memory (Bmem) features — these have zero paired
observations and cannot contribute to any CordBlood→InfMon2 change analysis. Any
cord-to-infant change for these analytes remains unquantifiable from this dataset.

## Figure 5a — Total IgG (antigen-specific)

Total IgG is the reference measure of passively transferred maternal antibody. All six
antigens measured at total-IgG level (DT, FHA, FIM, PRN, PT, TT) are expected to decay
and, if the assay is working correctly, all should fall clearly below zero.

```{r fig-5a, fig.cap="Figure 5a. Per-subject delta distributions for total antigen-specific IgG features (one panel per antigen, ordered by median delta). All six antigens show near-universal decay, with 95–99% of subjects decreasing and median deltas of −0.57 to −0.62 log10. This confirms that the biological decay signal is present and well-captured at the total-IgG level.", fig.width=10, fig.height=5}
make_analyte_hist("IgG", fill_col = analyte_colours["IgG"])
```

## Figure 5b — IgG1

IgG1 is the dominant IgG subclass in maternal serum and the primary isotype transferred
across the placenta. If passive transfer drove the IgG1 pool at birth and no infant
production has begun by 2 months, IgG1 should decay like total IgG.

```{r fig-5b, fig.cap="Figure 5b. Per-subject delta distributions for IgG1 features by antigen (ordered by median delta). IgG1 features predominantly decrease (9 of 10 antigens mostly decreasing), consistent with catabolic decay of passively transferred maternal IgG1. The magnitude of decay varies by antigen: PENTAMER, PT, and IPV2 show the strongest declines (median −0.34 to −0.65 log10), while ACT and IPV1 are closest to zero (IPV1 is mixed at +0.05).", fig.width=12, fig.height=6}
make_analyte_hist("IgG1", fill_col = analyte_colours["IgG1"])
```

## Figure 5c — IgG2

IgG2 is the subclass most enriched in responses to polysaccharide antigens and is
transferred across the placenta less efficiently than IgG1. Any systematic difference
between cord-blood and peripheral-infant IgG2 quantification would stand out here.

```{r fig-5c, fig.cap="Figure 5c. Per-subject delta distributions for IgG2 features by antigen (ordered by median delta). After normalisation, IgG2 is no longer systematically increasing: the analyte is mixed overall (median +0.05 log10; 4 antigens mostly increasing, 1 decreasing, 5 mixed). The largest residual increases are ACT (+0.24, 93% of subjects) and TT (+0.15); the IPV antigens that appeared to increase most before normalisation (IPV1/2/3) are now essentially flat (medians near 0.00 to +0.09). FHA is high in median (+0.46) but mixed and based on the smallest subset.", fig.width=12, fig.height=6}
make_analyte_hist("IgG2", fill_col = analyte_colours["IgG2"])
```

## Figure 5d — IgG3

IgG3 has the shortest half-life of the IgG subclasses (~7 days vs ~21 days for IgG1)
and is expected to decay fastest from cord levels. Rapid IgG3 catabolism would produce
strongly negative deltas here; increases would be anomalous.

```{r fig-5d, fig.cap="Figure 5d. Per-subject delta distributions for IgG3 features by antigen (ordered by median delta). IgG3 is mixed overall (median +0.13 log10; 5 antigens mostly increasing, 2 decreasing, 3 mixed). PENTAMER decays strongly (−0.63, consistent with the short ~7-day IgG3 half-life), and FHA/PRN/TT sit near zero, but several antigens (IPV2 +0.32, PT +0.26, ACT +0.15, DT +0.14, IPV1 +0.11) retain a modest increase even after normalisation. The residual antigen-to-antigen heterogeneity suggests IgG3 cord levels and short-half-life dynamics vary by antigen rather than reflecting a single uniform offset.", fig.width=12, fig.height=6}
make_analyte_hist("IgG3", fill_col = analyte_colours["IgG3"])
```

## Figure 5e — IgG4

IgG4 is typically the lowest-abundance IgG subclass in most individuals and is elevated
in chronic antigen exposure (e.g. repeated vaccination). Placental transfer of IgG4 is
less efficient than IgG1. Any measurement close to the assay floor in cord blood would
appear to increase by 2 months as IgG4 rises with infant environmental exposure, even
without maternal contribution.

```{r fig-5e, fig.cap="Figure 5e. Per-subject delta distributions for IgG4 features by antigen (ordered by median delta). Normalisation substantially shrank the IgG4 increases: the analyte median is now +0.05 log10 (6 antigens mostly increasing, 1 decreasing, 3 mixed), versus the large near-universal increases seen pre-normalisation. The remaining clear increases are TT (+0.71), DT (+0.52), PT (+0.21), and PRN (+0.12); PENTAMER decreases (−0.10) and FHA/IPV2/IPV3 are mixed near zero. Where increases persist they are most consistent with low cord-blood IgG4 (a floor effect on the log10 scale) for specific antigens rather than a systematic cross-analyte offset.", fig.width=12, fig.height=6}
make_analyte_hist("IgG4", fill_col = analyte_colours["IgG4"])
```

## Figure 5f — WHOLE_WT_IgG

WHOLE_WT_IgG is a single whole-cell lysate IgG measurement that integrates broadly
across antigens. As a total-IgG-type measurement it should follow the same decay
trajectory as the antigen-specific IgG features.

```{r fig-5f, fig.cap="Figure 5f. Per-subject delta distribution for WHOLE_WT_IgG. A single feature: 77% of subjects decrease (median delta −0.26 log10), consistent with decay of passively transferred whole-cell antibody, though slightly less extreme than the antigen-specific total-IgG features, possibly reflecting a broader antigen mixture.", fig.width=5, fig.height=4}
make_analyte_hist("WT_IgG", fill_col = analyte_colours["WT_IgG"])
```

# By Maternal Arm

```{r by-arm, results='asis'}
arm_sum <- summarise_paired(paired %>% filter(arm_name %in% MATERNAL_ARMS),
                            by = "arm_name", min_n = MIN_PAIRED) %>%
  group_by(arm_name) %>%
  summarize(features = n(),
            `% feats mostly increasing` = round(100*mean(direction == "mostly increase"), 1),
            `median of feature median-deltas` = round(median(median_delta), 3), .groups = "drop")
knitr::kable(arm_sum,
  caption = "Table 7. Direction summary by maternal arm. A systematic assay offset would appear similarly in both arms; an arm-specific difference would point toward biological (vaccine-driven) heterogeneity. The analyte-structured pattern in Figures 5a–5f should be examined by arm to distinguish these.") %>% print()
```

# Interpretation: Normalisation Removes Most of the Apparent Subclass Increase

## What the data show, feature by feature

Reading the figures by analyte type shows a clear hierarchy, and shows where the
cross-plate normalisation mattered most:

**Total IgG and IgG1 — decay confirmed (Figures 5a, 5b).** Both total antigen-specific
IgG and IgG1 features decrease in the large majority of subjects for nearly all antigens.
Total-IgG medians cluster tightly at −0.57 to −0.62 log10 (95–99% of subjects decreasing);
IgG1 medians run from about −0.65 to near zero, with 9 of 10 antigens mostly decreasing
(only IPV1 is mixed, at +0.05). This is the expected signature of passive maternal IgG
catabolism. Notably, total IgG comes from the RIVM ELISA path and IgG1 from the bead array;
both show clean decay, confirming the decay signal is real and well-quantified. These
features were essentially unaffected by the `mat_baa` normalisation (total IgG does not pass
through it at all), so they serve as a stable reference.

**IgG2 — increases largely removed by normalisation (Figure 5c).** Before normalisation,
IgG2 appeared to increase strongly and near-universally (IPV antigens at +0.4 to +0.6
log10, 80–100% of subjects). After normalising cord/maternal onto the infant scale, IgG2
is **mixed overall** (analyte median +0.05 log10; 4 antigens mostly increasing, 1
decreasing, 5 mixed). The IPV antigens that drove the earlier signal are now essentially
flat (IPV1 +0.09, IPV2 +0.01, IPV3 +0.01). A modest residual increase remains for ACT
(+0.24, 93% of subjects) and TT (+0.15); FHA has a high median (+0.46) but is mixed and
rests on the smallest cord subset. The collapse of the IPV increases is the clearest
single sign that much of the pre-normalisation IgG2 "increase" was a plate/run scaling
artifact, not biology.

**IgG3 — mixed, with a real residual for some antigens (Figure 5d).** IgG3 is mixed
overall (analyte median +0.13 log10; 5 increasing, 2 decreasing, 3 mixed). PENTAMER decays
strongly (−0.63, consistent with the short ~7-day IgG3 half-life) and FHA/PRN/TT sit near
zero, but several antigens retain a modest increase after normalisation (IPV2 +0.32, PT
+0.26, ACT +0.15, DT +0.14, IPV1 +0.11). Because the heterogeneity is antigen-specific
rather than a uniform shift, the residual most plausibly reflects genuine antigen-specific
differences in cord IgG3 starting levels and rapid IgG3 turnover, not a single leftover
offset.

**IgG4 — increases substantially shrunk by normalisation (Figure 5e).** Pre-normalisation,
IgG4 was the most consistently increasing analyte (8 of 10 antigens, +0.22 to +0.60).
After normalisation the analyte median is only +0.05 log10 (6 increasing, 1 decreasing,
3 mixed). Clear increases persist for TT (+0.71), DT (+0.52), PT (+0.21), and PRN (+0.12);
PENTAMER decreases (−0.10); FHA, IPV2, and IPV3 are mixed near zero. Where increases
remain, the most likely explanation is a genuine floor effect — IgG4 is the rarest
subclass and is very low in cord blood, so on the log10 scale even small absolute infant
IgG4 produces a positive delta for specific antigens.

**WHOLE_WT_IgG — decays as expected (Figure 5f).** Consistent with the total-IgG findings,
the broad-antigen whole-cell lysate IgG measurement decreases in 77% of subjects (median
−0.26 log10).

**Effector functions, FcR binding, B-memory — unquantifiable.** ADCD, ADCP, ADNP,
FcgR2a, FcgR3b, and Bmem were not assayed in CordBlood samples. The CordBlood→InfMon2
change for these functional readouts is entirely unobserved in this dataset.

## Why normalisation was necessary, and what it changed

The cord/maternal samples and the infant 2-month samples were run on **different bead-array
plates at different times**. Bead-array MFI is sensitive to plate, run date, and standard
source, so a raw cord-to-infant comparison mixes true biological change with cross-plate
scale differences. The pre-normalisation result — large, near-universal IgG-subclass
*increases* opposite to the expected decay — was the fingerprint of that artifact, not of
infant antibody production (which does not occur before primary immunisation at 2 months).

Normalising `mat_baa` onto the infant assay scale, using per-(feature × antigen × plate ×
source) cross-calibration parameters from shared standard curves, was therefore required to
make the timepoints comparable. The effect of that step is visible in the before/after
comparison:

| Analyte | Median Δ before | Median Δ after | Reading |
|---|---|---|---|
| IgG (total) | −0.59 | −0.59 | unchanged (not normalised; reference) |
| IgG1 | −0.20 | −0.24 | unchanged; clean decay |
| IgG2 | +0.28 | +0.05 | increase largely removed → mixed |
| IgG3 | −0.01 | +0.13 | small, antigen-specific residual |
| IgG4 | +0.32 | +0.05 | increase largely removed |
| WT_IgG | −0.26 | −0.26 | unchanged (not normalised) |

Two features of this pattern support the normalisation having done its job rather than
merely rescaling everything:

1. **The reference analytes are stable.** Total IgG and WHOLE_WT_IgG do not pass through the
   `mat_baa` normalisation and are essentially identical before and after; IgG1 (which does
   pass through it) also stays a clean decay. The normalisation did not distort the features
   that were already on a consistent scale.

2. **The change is concentrated where the artifact was.** The analytes whose medians moved
   most (IgG2 +0.28→+0.05, IgG4 +0.32→+0.05) are exactly the ones that had shown the large
   spurious increases. Removing the cross-plate scale difference collapsed those increases
   toward zero, leaving clean decay where biology predicts it and only small,
   antigen-specific residuals.

## What this means for downstream analyses

- **Total IgG and IgG1 decay are the most reliable estimands** from CordBlood→InfMon2
  comparisons and can be used directly as covariates or outcomes; IgG1 with the caveat that
  decay magnitude varies by antigen.
- **The normalised IgG2/IgG3/IgG4 deltas are now usable but small.** They no longer carry a
  systematic positive offset, so they can enter models, but the remaining per-antigen
  residuals (e.g. TT/DT IgG4, ACT IgG2, IPV2 IgG3) should be treated as antigen-specific and
  interpreted cautiously rather than as a uniform subclass trend.
- **Any earlier "positive median delta" conclusion in the maternal-chain (Block 5) analysis
  was driven by the un-normalised subclass artifact** and should be recomputed on the
  normalised values; the overall cord→InfMon2 direction is now decay or near-zero, not
  increase.
- **Effector function trajectories from cord blood remain unknown.** If the question requires
  knowing how ADCD, ADCP, or FcR binding changes between birth and 2 months, CordBlood
  samples would need to be assayed for these analytes.

## What would confirm or refine this interpretation

- **Examine the absolute cord-blood distributions for the antigens with residual IgG4
  increases** (TT, DT) — if a large fraction of subjects fall near the assay floor, the
  floor-effect explanation for those specific residuals is confirmed.
- **Check whether the residual subclass increases correlate with cord-blood level**
  (regression to the mean): if the lowest cord values show the largest positive deltas, a
  floor effect is the dominant remaining driver.
- **Stratify Figures 5c–5e by maternal arm** (Table 7) to test whether the residual
  increases are arm-specific (biological, e.g. larger for pertussis antigens in TdaP-arm
  mothers) or arm-shared (residual technical).
- **Spot-check the normalisation join coverage** — confirm that every contributing
  feature × antigen × plate × source actually matched a normalisation parameter row, so
  that no subset of subclass values silently retained its un-normalised scale.
- **Correlate per-subject mean delta with plate/batch metadata** (Figure 3): residual
  plate-level clustering would indicate the cross-plate correction is incomplete for some
  plates.

# Participant Accounting

```{r naccount, results='asis'}
nt_render(nt, part = "Cord→InfMon2",
          caption = "Table 8. Per-feature paired availability (n CordBlood / n InfMon2 / n used).")
```

---

*Reuses the shared pipeline (`config/endpoints.R`, `R/serology_helpers.R`, the global visit
recode, and the N-tracker). Knit both formats with*
`rmarkdown::render("cord_to_infmon2_analysis_refactored.Rmd", output_format = "all")`.
