---
title: "GaPs Results B–F: maternal–infant antibody pathway and antibody correlates of infant function"
subtitle: "Figure + tables generator for the revised results section (pre-primary / two-month baseline)"
author: "Scot Zens and Anja Saso"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
    theme: readable
  word_document:
    toc: true
    toc_depth: 2
    fig_width: 11
    fig_height: 4.3
params:
  arm: "ALL"          # "ALL" (pooled, matches the rendered transition analyses) or "TdaP" / "TT"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.align = "center")
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(knitr)
})
```

```{r bootstrap, include=FALSE}
## ---- self-locating resolver (same pattern as the other analysis files) ----
library(here)
# ---- shared path resolver: single definition in R/find_proj_file.R ----
if (!exists("find_proj_file", mode = "function")) {
  .ff_cand <- c("R/find_proj_file.R", "../R/find_proj_file.R",
                "analysis/R/find_proj_file.R")
  if (requireNamespace("here", quietly = TRUE))
    .ff_cand <- c(here::here("R/find_proj_file.R"),
                  here::here("analysis/R/find_proj_file.R"), .ff_cand)
  .ff_rmd <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) NA_character_)
  if (!is.na(.ff_rmd))
    .ff_cand <- c(file.path(.ff_rmd, "R/find_proj_file.R"),
                  file.path(dirname(.ff_rmd), "R/find_proj_file.R"), .ff_cand)
  .ff_hit <- .ff_cand[file.exists(.ff_cand)]
  if (!length(.ff_hit))
    stop("Cannot locate R/find_proj_file.R — place R/ beside this file, under analysis/, or at the project root.")
  source(.ff_hit[1], local = FALSE)
  rm(.ff_cand, .ff_rmd, .ff_hit)
}
source(find_proj_file("config/endpoints.R"),           local = FALSE, encoding = "UTF-8")
source(find_proj_file("config/endpoints_additions.R"), local = FALSE, encoding = "UTF-8")
source(find_proj_file("R/serology_helpers.R"),         local = FALSE, encoding = "UTF-8")

data_raw <- load_serology_data(find_proj_file("data/c_set.RData"))   # applies VISIT_RECODE

## Optional maternal-arm restriction. The three rendered transition analyses are
## pooled across arms (params$arm = "ALL"), which is what the decay summary in the
## text (95–99%, −0.57…−0.62) reflects. Section B's prose is framed "within the
## Tdap-IPV group"; set params$arm = "TdaP" to reproduce the panels for that arm.
## The pathway figures below are rendered for the pooled cohort AND for each
## maternal arm (Figures 3, 3-TdaP, 3-TT), so params$arm no longer restricts
## the figures — it only frames the prose / transition table. Keep the full
## cohort in data_full for the per-arm panels; apply the optional restriction
## (kept for backward compatibility) only to data_raw.
data_full <- data_raw
if (!identical(params$arm, "ALL")) {
  data_raw <- dplyr::filter(data_raw, as.character(arm_name) == params$arm)
}

ACCENT   <- "#8A1538"
# Five-series palette (total IgG + four subclasses) for the expanded Figure 3.
PATH_PAL    <- c("Total IgG" = "#8A1538", "IgG1" = "#2C7A7B", "IgG2" = "#C9882B",
                 "IgG3" = "#3B6BA5", "IgG4" = "#7B5EA7")
PATH_SERIES <- c("Total IgG", "IgG1", "IgG2", "IgG3", "IgG4")
WT_ROW      <- "WT_IgG"   # whole-cell IgG decay row, pinned to the bottom of Fig 3
ANTI6  <- ANTIGEN_ORDER   # canonical order from config/endpoints.R
```

# A. Study population

```{r tab1-demographics}
clin1      <- load_clinical_data(find_proj_file("data/clin_assess.RData"))
arm_lookup <- dplyr::distinct(data_full, subject_accession, arm_name)
demo_tab   <- demographics_by_arm(clin1, arm_lookup)
kable(demo_tab,
      caption = "Table 1. Study-population characteristics, overall and by maternal arm (Tdap-IPV vs TT). Continuous variables: median [IQR]; categorical: n (%).",
      align = c("l", rep("r", ncol(demo_tab) - 1)))
```

# B. The maternal–infant antibody pathway

We trace three within-subject transitions with the shared `paired_change()` /
`summarise_paired()` machinery (delta = log10 at the later visit − log10 at the
earlier visit; Wilcoxon signed-rank vs 0):

* **Production** PregEarly → MatBirth — antibody made during pregnancy
* **Transfer** MatBirth → CordBlood — placental transfer to the infant
* **Decay** CordBlood → InfMon2 — postnatal decay of the maternal pool

```{r transitions, include=FALSE}
MIN_PAIRED <- 10L
TRANSITIONS <- list(
  c(from = "PregEarly", to = "MatBirth",  label = "Maternal production\n(PregEarly → delivery)"),
  c(from = "MatBirth",  to = "CordBlood", label = "Placental transfer\n(delivery → cord)"),
  c(from = "CordBlood", to = "InfMon2",   label = "Postnatal decay\n(cord → 2 months)")
)
tr_levels <- vapply(TRANSITIONS, `[`, "", "label")

# One transition for an arbitrary subject subset `dat`, so the same code makes
# the pooled figure and the per-arm figures.
build_transition <- function(tr, dat) {
  paired <- paired_change(dat, tr["from"], tr["to"])
  summarise_paired(paired, min_n = MIN_PAIRED) %>%
    left_join(distinct(paired, feature, antigen, analyte), by = "feature") %>%
    mutate(transition = factor(tr["label"], levels = tr_levels),
           from = tr["from"], to = tr["to"])
}
build_all_tr <- function(dat) bind_rows(lapply(TRANSITIONS, build_transition, dat = dat))

# y-axis order: antigens top->bottom = PT, FHA, PRN, FIM, TT, DT (ANTIGEN_ORDER),
# with the whole-cell IgG (WT_IgG) decay row pinned at the very bottom. Factor
# levels run bottom->top, so WT_IgG is first.
y_levels <- c(WT_ROW, rev(ANTI6))

# Symmetric within-band offsets so the five series never collide between rows.
yoff_map <- c("Total IgG" = 0.30, "IgG1" = 0.15, "IgG2" = 0.00,
              "IgG3" = -0.15, "IgG4" = -0.30)

# Build the plot frame for one subject subset. Includes total IgG + all four
# subclasses for the six antigens, plus the whole-cell IgG decay row (cord ->
# InfMon2 only; WT_IgG has no subclasses and is not measured pre-cord). FIM
# subclasses are absent before InfMon2, so the FIM row carries only total IgG in
# the production/transfer panels — handled automatically by data availability.
build_fig_df <- function(dat) {
  at <- build_all_tr(dat)
  ag <- at %>%
    filter(antigen %in% ANTI6,
           analyte %in% c("IgG", "IgG1", "IgG2", "IgG3", "IgG4")) %>%
    mutate(row_label = factor(antigen, levels = y_levels),
           series    = factor(ifelse(analyte == "IgG", "Total IgG", analyte),
                              levels = PATH_SERIES))
  wc <- at %>%
    filter(feature == "WHOLE_WT_IgG", from == "CordBlood") %>%
    mutate(row_label = factor(WT_ROW, levels = y_levels),
           series    = factor("Total IgG", levels = PATH_SERIES))
  bind_rows(ag, wc) %>%
    mutate(yoff = unname(yoff_map[as.character(series)]),
           ypos = as.numeric(row_label) + yoff,
           declabel = ifelse(series == "Total IgG" & from == "CordBlood",
                             paste0(round(pct_dec), "%↓"), NA_character_))
}

# Figure renderer shared by the pooled and per-arm panels.
plot_pathway <- function(fig_df, subtitle = NULL) {
  ggplot(fig_df, aes(x = median_delta, y = ypos,
                     colour = series, shape = series, fill = series)) +
    geom_vline(xintercept = 0, colour = "grey20", linewidth = 0.4) +
    geom_segment(aes(x = 0, xend = median_delta, yend = ypos),
                 linewidth = 0.7, alpha = 0.30) +
    geom_point(size = 2.3, stroke = 0.9) +
    geom_text(aes(label = declabel), na.rm = TRUE, nudge_y = 0.17, hjust = 1.05,
              size = 2.2, colour = ACCENT, show.legend = FALSE) +
    facet_wrap(~ transition, nrow = 1) +
    scale_y_continuous(breaks = seq_along(y_levels), labels = y_levels,
                       name = "Antigen") +
    scale_colour_manual(values = PATH_PAL, drop = FALSE) +
    scale_fill_manual(values = c("Total IgG" = unname(PATH_PAL["Total IgG"]),
                                 "IgG1" = "white", "IgG2" = "white",
                                 "IgG3" = "white", "IgG4" = "white"), drop = FALSE) +
    scale_shape_manual(values = c("Total IgG" = 21, "IgG1" = 22, "IgG2" = 23,
                                  "IgG3" = 24, "IgG4" = 25), drop = FALSE) +
    labs(x = expression("median " * Delta * "log"[10] * " (to − from)"),
         colour = NULL, shape = NULL, fill = NULL, subtitle = subtitle,
         title = "Production loads the antigen-specific pool; transfer is near-complete; decay dominates") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom",
          panel.grid.major.y = element_blank(),
          panel.grid.minor = element_blank(),
          strip.text = element_text(face = "bold", colour = ACCENT),
          plot.title = element_text(size = 10, face = "bold"))
}

# Transition table frame (respects params$arm framing).
all_tr <- build_all_tr(data_raw)
```

## Figure 3 — maternal–infant antibody pathway (expanded: total IgG + IgG1–IgG4)

```{r fig3-pooled, fig.width=11, fig.height=5.4}
plot_pathway(build_fig_df(data_full), subtitle = "All mothers (pooled)")
```

*Each point is a per-feature median within-subject change (Δlog₁₀ = later −
earlier visit). Series: total IgG (filled circle) and IgG1–IgG4 (open
square/diamond/triangles), offset within each antigen band. Rows are the six
vaccine antigens in canonical order (PT, FHA, PRN, FIM, TT, DT) plus a whole-cell
IgG (WT_IgG) decay row at the bottom (cord→2 months only). Percentages on the
total-IgG points in the decay panel are the proportion of infants declining. FIM
subclasses are not measured before InfMon2, so the FIM row shows only total IgG in
the production and transfer panels. The pertussis-toxin IgG3 point in the decay
panel sits near the assay floor (see analyst note) and is not subclass persistence.*

### Figure 3 — Tdap-IPV mothers

```{r fig3-tdap, fig.width=11, fig.height=5.4}
plot_pathway(build_fig_df(dplyr::filter(data_full, as.character(arm_name) == "TdaP")),
             subtitle = "Tdap-IPV maternal arm")
```

### Figure 3 — TT mothers

```{r fig3-tt, fig.width=11, fig.height=5.4}
plot_pathway(build_fig_df(dplyr::filter(data_full, as.character(arm_name) == "TT")),
             subtitle = "Tetanus-toxoid (TT) maternal arm")
```

*Per-arm panels resolve the toxoid question directly: both maternal vaccines
contain tetanus toxoid, so TT-antigen IgG production should rise in both arms
(subject to a pre-existing-immunity ceiling), whereas diphtheria toxoid is in the
Tdap-IPV arm only, so DT-antigen IgG production should rise mainly after Tdap-IPV.
Read the TT and DT rows of the production (left) panel in each arm figure.*

## Table 1 — transition summary (per antigen, total IgG and IgG1)

```{r tab1}
tab1 <- all_tr %>%
  filter(antigen %in% ANTI6, analyte %in% c("IgG", "IgG1")) %>%
  transmute(Transition = sub("\n.*", "", as.character(transition)),
            Antigen = antigen, Analyte = ifelse(analyte == "IgG", "Total IgG", analyte),
            n = n_paired, `% increase` = pct_inc, `% decrease` = pct_dec,
            `median Δlog10` = median_delta, `p (signed-rank)` = p_signrank,
            Direction = direction) %>%
  arrange(match(Transition, c("Maternal production", "Placental transfer", "Postnatal decay")),
          match(Antigen, ANTI6), Analyte)
kable(tab1, caption = "Table 1. Within-subject paired change for each antigen across the three pathway transitions.")
```

```{r decay-subclass, results='asis'}
dec <- build_transition(TRANSITIONS[[3]], data_raw)
igg1_med <- dec %>% filter(analyte == "IgG1", antigen %in% ANTI6) %>% summarise(m = median(median_delta)) %>% pull(m)
cat(sprintf(paste0("\n**Decay detail.** Across antigen-specific measurements the median IgG1 change was ",
                   "≈ %.2f log10; the only subclass not to decline was pertussis toxin IgG3 ",
                   "(median Δlog10 ≈ +0.26), close to the lower assay range and interpreted cautiously.\n\n"),
            igg1_med))
```

# C–F. Antibody correlates of infant function

The maternal-chain models (sections C–D) are produced by
`parts/01_maternal_chain.Rmd` inside the per-endpoint pipelines
(`SBA_analysis.Rmd`, `WT_IgG_analysis.Rmd`, `PTNA_analysis.Rmd`); the concurrent
infant models and the Fc-effector / clinical / circularity comparisons
(sections E–F and the forward-prediction table) are produced by
`06_prediction_models.Rmd`. The tables below collate the key quantities from
those rendered outputs so the whole results section knits from one file. To
regenerate them live instead of collating, knit those modules (e.g.
`rmarkdown::render("SBA_analysis.Rmd", params = list(endpoint = "SBA"))`) and read
their `selection_table()` / Block-1 coefficient outputs.

## Table 2 — maternal antigen-specific IgG at delivery → infant function (std β)

```{r tab2}
tab2 <- tibble::tribble(
  ~`Infant outcome (InfMon2)`,        ~`Principal maternal predictor (std β)`, ~`Conditional FHA (std β)`, ~`R² (PT+FHA+PRN)`, ~`R² (+DT+TT+FIM)`,
  "Serum bactericidal activity (SBA)", "PRN IgG (0.91)",                          "−0.48",                  0.681,                0.687,
  "Whole-cell binding (WT IgG)",       "PRN IgG (0.46)",                          "−0.35",                  0.351,                0.375,
  "PT-neutralising activity (PTNA)",   "PT IgG (0.56)",                           "— (ns)",                 0.244,                0.258)
kable(tab2, caption = "Table 2. Standardised β from the maternal-delivery models (parts/01_maternal_chain, Block 1). Negative FHA coefficients are conditional/collinear.")
```

## Table 3 — maternal model fit by maternal arm

```{r tab3}
tab3 <- tibble::tribble(
  ~`Infant outcome`,     ~`R² after Tdap-IPV`, ~`R² after tetanus toxoid`, ~Interpretation,
  "SBA",                 0.687,                   0.179,                          "PRN-dependent; vaccine-sensitive",
  "Whole-cell binding",  0.375,                   0.083,                          "PRN-dependent; vaccine-sensitive",
  "PTNA",                0.258,                   0.270,                          "PT-dependent; arm-independent")
kable(tab3, caption = "Table 3. Expanded maternal-delivery model fit within each maternal arm.")
```

## Table 4 — concurrent (same-visit) infant models at two months

```{r tab4}
tab4 <- tibble::tribble(
  ~Outcome,             ~`Base R²`, ~`+ matched total IgG`, ~`+ matched IgG1 (R²/adj/CV)`, ~`Dominant term (β)`, ~n,
  "PTNA",               "0.340",       "0.520",                "0.576 / 0.569 / 0.547",          "PT IgG1 (0.64)",        235L,
  "SBA",                "0.378",       "0.766",                "0.859 / 0.856 / 0.841",          "PRN IgG1 (0.95)",       217L,
  "Whole-cell binding", "—",      "0.452",                "0.536 / 0.521 / 0.498",          "PRN IgG1 (0.62)",       221L)
kable(tab4, caption = "Table 4. Concurrent infant models (06_prediction_models, Family 1). Matched IgG1 is the dominant correlate of each function.")
```

## Table 5 — Fc-effector and clinical additions

```{r tab5}
tab5 <- tibble::tribble(
  ~Outcome,             ~`CV-R²: IgG1 model`, ~`CV-R²: + FcγRIIa/ADCD`, ~`Clinical block`,
  "PTNA",               0.547,                    0.518,                            "no CV gain; age borderline",
  "SBA",                0.841,                    0.829,                            "no CV gain",
  "Whole-cell binding", 0.498,                    0.464,                            "no CV gain")
kable(tab5, caption = "Table 5. Adding Fc-effector terms lowered cross-validated performance for every outcome; the wider clinical block did not improve CV fit.")
```

## Table 6 — concurrent vs prior (maternal-at-delivery) prediction

```{r tab6}
tab6 <- tibble::tribble(
  ~Outcome,             ~`Concurrent R²/CV`, ~`Prior maternal R²/CV`, ~`Concurrent advantage (R²)`, ~n,
  "PTNA",               "0.558 / 0.523",         "0.477 / 0.437",            "≈ 0.08",                    175L,
  "SBA",                "0.863 / 0.845",         "0.678 / 0.642",            "≈ 0.19",                    163L,
  "Whole-cell binding", "0.562 / 0.529",         "0.374 / 0.321",            "≈ 0.19",                    160L)
kable(tab6, caption = "Table 6. 06_prediction_models Family 3. Concurrent = descriptive ceiling (same-sample co-measurement); prior maternal = honest forward predictor. With both entered, maternal PRN IgG1 retains β = 0.33 for SBA.")
```

# Notes for the analyst

* **Interaction tests not yet fitted.** Section D is a difference in *variance
  explained* by arm, not a slope difference. Fit maternal-group × antibody
  interactions (group × PT IgG1 for PTNA; group × PRN IgG1 for SBA and whole-cell
  binding) before claiming the antibody–function slope differs by background.
* **Deming link.** PT–PTNA and PRN–SBA are OLS here; refit as errors-in-variables
  (Deming) and test slope by maternal background to connect to the earlier GaPs
  Deming result.
* **PT IgG3 floor.** The +0.26 cord→two-month PT IgG3 change sits near the assay
  floor and shrinks to a ~+0.13 residual after cross-plate normalisation; do not
  read it as subclass persistence.
* **Provenance.** Section B (figure + table 1) is computed here from `c_set.RData`.
  Tables 2–6 collate the rendered maternal-chain and prediction modules; verify
  against the current knit before submission.

```{r sessioninfo}
sessionInfo()
```
